### 2022-06-17 paper/seleff.tex
first commit

%                                                                 aa.dem
% AA vers. 8.2, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
% according to the author-year natbib style

%
\documentclass{aa}

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
%\usepackage{url}
%\usepackage{fontawesome}

%\usepackage{xspace}
%\usepackage{subcaption}

\def\detJ{\mathrm{det}J}

\def\tein{\theta_{\mathrm{Ein}}}
\def\trad{\theta_{\mathrm{rad}}}
\def\tsis{\theta_{\mathrm{Ein}}^{\mathrm{(SIS)}}}
\def\meantsis{<\tsis>}
\def\stdtsis{\sigma(\tsis)}

\def\teinein{\theta_{\mathrm{Ein}}^*}
\def\betaein{\beta^*}

\def\toneobs{\theta_1^{\mathrm{obs}}}
\def\ttwoobs{\theta_2^{\mathrm{obs}}}

\def\moneobs{m_1^{\mathrm{obs}}}
\def\mtwoobs{m_2^{\mathrm{obs}}}

\def\asymm{\xi_{\mathrm{asymm}}}
\def\rmur{r_{\mu_r}}
\def\rmurobs{r_{\mu_r}^{(\mathrm{obs})}}
\def\mumin{\mu_{\mathrm{min}}}
\def\betamax{\beta_{\mathrm{max}}}
\def\betasl{\beta_{\mathrm{SL}}}

\def\hyperpars{\boldsymbol{\eta}}
\def\indpar{\boldsymbol{\psi}}
\def\indpari{\boldsymbol{\psi}_i}

\def\nsource{N_{\mathrm{s}}}
\def\nbkg{n_{\mathrm{bkg}}}

\def\psilens{\boldsymbol{\psi}_\mathrm{g}}
\def\psisource{\boldsymbol{\psi}_\mathrm{s}}
\def\psisourcetwo{\boldsymbol{\psi}_{\mathrm{s},2}}
\def\psisourcens{\boldsymbol{\psi}_{\mathrm{s},\nsource}}
\def\psisourcenobeta{\boldsymbol{\psi}_\mathrm{s}^{(-\boldsymbol\beta)}}

\def\Nlens{N_{\mathrm{lens}}}
\def\Nnot{N_{\mathrm{not}}}

\def\prlens{{\rm P}_\mathrm{g}}
\def\prsource{{\rm P}_\mathrm{s}}
\def\prlensone{{\rm P}_{\mathrm{g},1}}
\def\prlenstwo{{\rm P}_{\mathrm{g},2}}
\def\prsourceone{{\rm P}_{\mathrm{s},1}}
\def\prsourcetwo{{\rm P}_{\mathrm{s},2}}
\def\prsl{{\rm P}_\mathrm{{SL}}}
\def\prslone{{\rm P}_{\mathrm{SL},1}}
\def\prsltwo{{\rm P}_{\mathrm{SL},2}}

\def\prsourcenobeta{{\rm P}_\mathrm{s}^{(-\boldsymbol\beta)}}

\def\pdet{{\rm P}_\mathrm{det}}
\def\pdetone{{\rm P}_{\mathrm{det},1}}
\def\pdettwo{{\rm P}_{\mathrm{det},2}}
\def\crosssect{\sigma_\mathrm{{SL}}}

\def\data{\mathbf{d}}
\def\datai{\mathbf{d}_i}

\def\dlens{\mathbf{d}_{\mathrm{g}}}
\def\dsource{\mathbf{d}_{\mathrm{s}}}

\def\mlim{m_{\mathrm{max}}}

\def\Sref#1{Section~\ref{#1}\xspace}
\def\Fref#1{Figure~\ref{#1}\xspace}
\def\Tref#1{Table~\ref{#1}\xspace}
\def\Eref#1{Equation~\ref{#1}\xspace}

\newcommand{\ale}[1]{\textcolor{red}{\textbf{[Ale: #1]}}}

\def\pr{{\rm P}}

\defcitealias{S+C21}{Paper~I}
\defcitealias{Son21}{Paper~II}
\defcitealias{Son22}{Paper~III}
\setcitestyle{notesep={}}

\begin{document}

   \title{Strong lensing selection effects}
   \titlerunning{Strong lensing selection effects}
   \authorrunning{Sonnenfeld et al.}

   \author{Alessandro Sonnenfeld\inst{1}
          }

   \institute{Leiden Observatory, Leiden University, Niels Bohrweg 2, 2333 CA Leiden, the Netherlands\\
              \email{sonnenfeld@strw.leidenuniv.nl}
             }

   \date{}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
    {
There are selection effects. They are important.
}
  % aims heading (mandatory)
   {
How important?
} 
   % methods heading (mandatory)
   {
I simulate lenses.
}
% results heading (mandatory)
   {
Results.
}
  % conclusions heading (optional), leave it empty if necessary 
   {
I conclude.
}
   \keywords{
             Gravitational lensing: strong --
             Galaxies: fundamental parameters
               }

   \maketitle
%
%________________________________________________________________

\section{Introduction}\label{sect:intro}

The third paper of this series \citep[][, hereafter Paper III]{Son22}, showed how to take into account selection effects.

The structure of this work is the following.
I discuss the results in \Sref{sect:discuss} and draw conclusions in \Sref{sect:concl}.

The Python code used for the simulation and analysis of the lens sample can be found in a dedicated section of a GitHub repository\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools}}.

%__________________________________________________________________

\section{Individual lenses}\label{sect:indlenses}

\subsection{Axisymmetric lenses, point sources}

\subsection{Elliptical lenses, point sources}

\subsection{Elliptical lenses, extended sources}

%__________________________________________________________________

\section{Lens populations}\label{sect:lenspop}

%__________________________________________________________________

\section{Discussion and summary}\label{sect:discuss}

%\begin{acknowledgements}

%\end{acknowledgements}

\bibliographystyle{aa}
\bibliography{references}

\end{document}

### 2022-06-26 paper/seleff.tex
first plot

@@
 
+\begin{figure}
+\includegraphics[width=\columnwidth]{axisymm_pl_crosssect.eps}
+\caption{
+Strong lensing cross-section of an axisymmetric power-law lens and a point source, as a function of the power-law slope $\gamma$.
+The system is defined as a strong lens if at least two images are detected.
+Different lines correspond to the difference $\Delta m$ between the source magnitude and the survey detection limit for a point source.
+The dashed line marks the cross-section in the case in which at least two images are formed, regardless of their magnification.
+}
+\end{figure}
+
 \subsection{Elliptical lenses, point sources}
@@
 
+Discuss.
+
 %\begin{acknowledgements}

### 2022-06-26 single_lens/axisymmetric_point/get_pl_crosssect_grid.py

@@ -1,0 +1,125 @@
+import numpy as np
+from sl_profiles import nfw, deVaucouleurs as deV
+from sl_cosmology import Mpc, c, G, M_Sun, yr
+import sl_cosmology
+from scipy.interpolate import splrep, splev, splint
+from scipy.optimize import brentq
+from scipy.stats import truncnorm
+from scipy.special import erf
+import emcee
+import h5py
+import sys
+
+
+# calculates the strong lensing cross-section for an axisymmetric power-law lens
+# and a point source, with the following selection criterion: the counter-image
+# must be brighter than the detection limit.
+# the output is the cross-section for unit Einstein radius, as a function of
+# power-law slope and dms, which is the difference between the source magnitude
+# and the detection limit
+
+tein = 1.
+
+arcsec2rad = np.deg2rad(1./3600.)
+
+# defines lensing-related functions
+def alpha(x, tein, gamma):
+    return tein * x/abs(x) * (abs(x)/tein)**(2.-gamma)
+
+def kappa(x, tein, gamma): 
+    # dimensionless surface mass density
+    return (3.-gamma)/2. * (abs(x)/tein)**(1.-gamma)
+
+def mu_r(x, tein, gamma):
+    # radial magnification
+    return (1. + alpha(x, tein, gamma)/x - 2.*kappa(x, tein, gamma))**(-1)
+
+def mu_t(x, tein, gamma):
+    # tangential magnification
+    return (1. - alpha(x, tein, gamma)/x)**(-1)
+
+def pl_ycaust(gamma):
+
+    xmin = 0.01
+
+    def radial_invmag(x):
+        return 1. + alpha(x, 1., gamma)/x - 2.*kappa(x, 1., gamma)
+
+    # finds the radial caustic
+    if radial_invmag(xmin)*radial_invmag(tein) > 0.:
+        xradcrit = xmin
+    else:
+        xradcrit = brentq(radial_invmag, xmin, tein)
+
+    ycaust = -(xradcrit - alpha(xradcrit, 1., gamma))
+
+    return ycaust, xradcrit
+
+xmin = 0.01
+xmax = 100.
+
+nxB = 10001
+
+gamma_min = 1.2
+gamma_max = 2.8
+
+ngamma = 81
+
[…]

### 2022-06-26 single_lens/axisymmetric_point/plot_pl_crosssect.py

@@ -1,0 +1,72 @@
+import numpy as np
+import pylab
+import h5py
+from scipy.optimize import brentq
+from plotters import probcontour
+from labellines import labelLine, labelLines
+from matplotlib.ticker import MultipleLocator
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 18
+
+logscale = True
+
+if logscale:
+    leftm = 0.14
+else:
+    leftm = 0.11
+
+cs_file = h5py.File('pl_point_crosssect_grid.hdf5', 'r')
+
+dms_grid = cs_file['dms_grid'][()]
+nms = len(dms_grid)
+
+nind = 6
+dms_indices = 20*np.arange(nind)
+
+gamma_grid = cs_file['gamma_grid'][()]
+ngamma = len(gamma_grid)
+
+cs_grid = cs_file['crosssect_grid'][()]
+
+fig = pylab.figure()
+pylab.subplots_adjust(left=leftm, right=0.99, bottom=0.13, top=0.99, wspace=0., hspace=0.)
+
+ax = fig.add_subplot(1, 1, 1)
+
+for i in range(nind):
+    ax.plot(gamma_grid, cs_grid[:, dms_indices[i]], label='$\Delta m_s=%2.1f$'%dms_grid[dms_indices[i]], linewidth=2)
+
+lines = pylab.gca().get_lines()
+gamma_vals = 2.3 - 0.05*np.arange(6)
+labelLines(lines, xvals=gamma_vals, fontsize=fsize, backgroundcolor='white')
+
+ax.plot(gamma_grid[gamma_grid<=2.], np.pi*cs_file['beta_caust_grid'][()][gamma_grid<=2.]**2, linestyle='--', color='k', label='Multiple images')
+ax.plot([2., 2.], [np.pi*cs_file['beta_caust_grid'][()][gamma_grid<=2.][-1]**2, 100.], linestyle='--', color='k')
+
+ax.xaxis.set_major_locator(MultipleLocator(0.2))
+ax.xaxis.set_minor_locator(MultipleLocator(0.05))
+
+if logscale:
+    ax.set_yscale('log')
+    ax.set_ylim(0., 13.)
+else:
+    ax.yaxis.set_major_locator(MultipleLocator(2))
+    ax.yaxis.set_minor_locator(MultipleLocator(0.5))
+
[…]

### 2022-06-29 paper/seleff.tex
started intro

@@
 
-The third paper of this series \citep[][, hereafter Paper III]{Son22}, showed how to take into account selection effects.
+\citep{Son22} showed how to correct for selection effects in a population study of a sample of strong lenses.
+However, the practical realization of the \citep{Son22} method is challenging, as it requires 1) characterising the background source population, 2)
 
 The structure of this work is the following.
+In \Sref{sect:indlenses} we study individual lens systems, to look for trends in the lensing cross-section with various lens and source properties.
+In \Sref{sect:lenspop} we study
+
 I discuss the results in \Sref{sect:discuss} and draw conclusions in \Sref{sect:concl}.
@@
 \section{Individual lenses}\label{sect:indlenses}
+
+In this section we study how the probability of a strong lensing event varies as a function of
 

### 2022-06-29 paper/seleff.tex
intro and composite grid

@@
    \keywords{
-             Gravitational lensing: strong --
-             Galaxies: fundamental parameters
+             Gravitational lensing: strong 
                }
@@
 \section{Introduction}\label{sect:intro}
+
+Strong gravitational lensing is a powerful tool for studying galaxies and cosmology.
+%Strongly lensed images enable us to make very precise measurements of the mass 
+Strong lenses have been used to probe the mass structure of massive galaxies \citep{Aug++10, ORF14, Son++15, Sha++21}, to detect substructure \citep{Veg++12, Hez++16, Nie++20}, to carry out detailed studies of magnified star-forming galaxies \citep{Jon++13}, and to measure the expansion rate of the universe with time delays \citep[see][for a review]{T+M16}.
+
+Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
+A necessary condition for a galaxy to act as a strong lens with respect to a given source is that its projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at least one position $\boldsymbol\theta$ \citep{SEF92}.
+This condition excludes objects with a very diffuse mass distribution from the population of lenses.
+In general, galaxies with a higher concentration of mass are more likely to be strong lenses, and are therefore over-represented in lens samples.
+
+The distribution of background source properties is also affected by
+
+In order for a galaxy to act as a strong lens with respect to a given source, it is necessar
+Galaxies with a higher concentration of mass are more likely to be strong lenses
+Strong lensing is a rare phenomenon
 

### 2022-06-30 paper/seleff.tex
working on intro

@@
 Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
-A necessary condition for a galaxy to act as a strong lens with respect to a given source is that its projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at least one position $\boldsymbol\theta$ \citep{SEF92}.
+A necessary condition for a galaxy to act as a strong lens with respect to a given source is that its projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
 This condition excludes objects with a very diffuse mass distribution from the population of lenses.
@@
 
-The distribution of background source properties is also affected by
+%The distribution of background source properties is also affected by strong lensing. 
+%Both the luminosity and size distribution of lensed galaxies 
+The luminosity and size distribution of strongly lensed sources is also biased with respect to the general population of background galaxies.
+For instance, lensing magnification allows the detection of fainter objects with respect to the field, and galaxies that are very extended with respect to the Einstein radius are less likely to be classified as strongly lensed, since their magnification averaged over their size tends to be low \citep[see also the discussion in section 5.3 of][]{O+A17}.
+%Moreover, if a large overall magnification is required in order to classify a source as strongly lensed, galaxies that are very extended with respect to the Einstein radius are less likely to be part of a strong lens sample.
 
-In order for a galaxy to act as a strong lens with respect to a given source, it is necessar
-Galaxies with a higher concentration of mass are more likely to be strong lenses
-Strong lensing is a rare phenomenon
+%The details of the bias on the lens and source population properties depend also on the criteria used to define a lens.
+In general, the probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
+\begin{equation}\label{eq:one}
+\prsl(\psilens,\psisource|S) \propto \prlens(\psilens)\prsource(\psisource)\pdet(\psilens,\psisource|S).
+\end{equation}
+In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the general distribution of foreground galaxies and background sources in the absence of lensing; and $\pdet$ describes the probability of detecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
 
-\citep{Son22} showed how to correct for selection effects in a population study of a sample of strong lenses.
-However, the practical realization of the \citep{Son22} method is challenging, as it requires 1) characterising the background source population, 2)
+The left-hand side of \Eref{eq:one} is directly accessible from strong lensing observations.
+If the main goal of a lensing survey is to characterise the properties of the strong lens population, then it can be accomplished by directly analysing this term.
+For many applications of strong lensing, however, the aim is to constrain the properties of the general galaxy or source population, $\prlens$ and $\prsource$, which are coupled in a nontrivial way via the lens detection probability $\pdet$.
+%These factors are coupled in a nontrivial way via the lens detection probability $\pdet$. 
+%The factor $\pdet$ takes into account both 
+%The factor $\pdet$ varies as a function of lens mass: it is larger 
+%Therefore, in order to obtain an unbiased estimate of $\prlens$ or $\prsource$ it is necessary to account for selection effects.
+In order to obtain an unbiased estimate of either $\prlens$ or $\prsource$, then, it is necessary to invert \Eref{eq:one}.
+In principle, this can be done with a Bayesian hierarchical formalism, provided that the lens detection probability $\pdet$ is known \citep{Son22}.
+%That, however, requires knowing the lens detection probability $\pdet$ as a function of lens and source properties, which is a difficult quantity to determine.
+For most of the existing strong lensing surveys, however, characterising $\pdet$ is very challenging: the process determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
+
+%In summary, the problem of the invertibility of \Eref{eq:one} is a difficult one to solve.
+%The problem of the invertibility of \Eref{eq:one} can be simplified by working with lens samples that constructing lens samples that are complete above a well-defined observational cut.
+%The problem of the invertibility of \Eref{eq:one} can be simplified by working with samples of lenses that are constructed in a well-defined way, for instance 
+The problem of the invertibility of \Eref{eq:one} can be simplified by adopting lens selection criteria that can be easily forward-modelled: for instance by building samples that are complete above a well-defined observational cut \citep{Son22}.
+Whether that is necessary, however, depends on the severity of the strong lensing bias that needs to be corrected and on the accuracy requirements on the key quantities of interest.
+
+In this paper we aim to quantify the strength of the strong lensing bias on a series of foreground galaxy and background source parameters, under realistic assumptions for the lens detection probability $\pdet$.
+In particular, we aim to determine how strong lenses differ from the general population of foreground galaxies and background sources in terms of 1) the radial mass structure of the lenses (i.e.
+their stellar and dark matter mass density profiles); 2) the ellipticity of the lenses; 3) the size-magnitude distribution of the lensed sources.
+We consider two scenarios: one in which a population of foreground massive galaxies lenses a population of extended background sources, and a second one in which the source population consists of point-like quasars.
 

### 2022-07-01 paper/seleff.tex
more intro

@@
 Strong gravitational lensing is a powerful tool for studying galaxies and cosmology.
-%Strongly lensed images enable us to make very precise measurements of the mass 
 Strong lenses have been used to probe the mass structure of massive galaxies \citep{Aug++10, ORF14, Son++15, Sha++21}, to detect substructure \citep{Veg++12, Hez++16, Nie++20}, to carry out detailed studies of magnified star-forming galaxies \citep{Jon++13}, and to measure the expansion rate of the universe with time delays \citep[see][for a review]{T+M16}.
@@
 
-%The distribution of background source properties is also affected by strong lensing. 
-%Both the luminosity and size distribution of lensed galaxies 
 The luminosity and size distribution of strongly lensed sources is also biased with respect to the general population of background galaxies.
 For instance, lensing magnification allows the detection of fainter objects with respect to the field, and galaxies that are very extended with respect to the Einstein radius are less likely to be classified as strongly lensed, since their magnification averaged over their size tends to be low \citep[see also the discussion in section 5.3 of][]{O+A17}.
-%Moreover, if a large overall magnification is required in order to classify a source as strongly lensed, galaxies that are very extended with respect to the Einstein radius are less likely to be part of a strong lens sample.
 
-%The details of the bias on the lens and source population properties depend also on the criteria used to define a lens.
 In general, the probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
@@
 In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the general distribution of foreground galaxies and background sources in the absence of lensing; and $\pdet$ describes the probability of detecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
+This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be recognised as a strong lens.
 
@@
 
-%In summary, the problem of the invertibility of \Eref{eq:one} is a difficult one to solve.
-%The problem of the invertibility of \Eref{eq:one} can be simplified by working with lens samples that constructing lens samples that are complete above a well-defined observational cut.
-%The problem of the invertibility of \Eref{eq:one} can be simplified by working with samples of lenses that are constructed in a well-defined way, for instance 
-The problem of the invertibility of \Eref{eq:one} can be simplified by adopting lens selection criteria that can be easily forward-modelled: for instance by building samples that are complete above a well-defined observational cut \citep{Son22}.
-Whether that is necessary, however, depends on the severity of the strong lensing bias that needs to be corrected and on the accuracy requirements on the key quantities of interest.
+The problem of inverting \Eref{eq:one} is a difficult one to tackle exactly.
+Few studies have attempted to explicitly account for strong lensing selection effects, usually by making ad-hoc simplifying assumptions \citep{Son++15,O+A17,Son++19}.
+%The problem of the invertibility of \Eref{eq:one} can be simplified by adopting lens selection criteria that can be easily forward-modelled: for instance by building samples that are complete above a well-defined observational cut \citep{Son22}.
+%The problem can be simplified by adopting lens selection criteria that can be easily forward-modelled: for instance by building samples that are complete above a well-defined observational cut \citep{Son22}.
+Whether it is necessary to invert \Eref{eq:one}, however, depends on the severity of the strong lensing bias that needs to be corrected and on the accuracy requirements on the key quantities of interest.
 
-In this paper we aim to quantify the strength of the strong lensing bias on a series of foreground galaxy and background source parameters, under realistic assumptions for the lens detection probability $\pdet$.
+In this paper we aim to quantify the strength of the strong lensing bias on a series of foreground galaxy and background source parameters.
+%, under realistic assumptions for the lens detection probability $\pdet$.
 In particular, we aim to determine how strong lenses differ from the general population of foreground galaxies and background sources in terms of 1) the radial mass structure of the lenses (i.e.
 their stellar and dark matter mass density profiles); 2) the ellipticity of the lenses; 3) the size-magnitude distribution of the lensed sources.
+We use semi-analytic models to simulate large populations of strong lenses.
+We simulate images of the lenses and the lensed sources, and then simulate lensing surveys with realistic assumptions for the lens detection probability $\pdet$.
+%then use these images to determine whether they are classified as lenses by a fiducial survey, under realistic assumptions for the lens detection probability $\pdet$.
 We consider two scenarios: one in which a population of foreground massive galaxies lenses a population of extended background sources, and a second one in which the source population consists of point-like quasars.
+
+The amplitude of the lensing bias depends on two aspects: 1) how the lens detection probability $\pdet$ varies as a function of galaxy and source properties, and 2) the width of the galaxy and source parameters distribution $\plens$ and $\psource$.
+To understand this second point we can imagine the limiting case in which both $\plens$ and $\psource$ are Dirac delta functions: in this limit, $\prsl$ simply reduces to the product $\plens\psource$ up to a scaling constant, corresponding to a case in which the lensing bias is none.
+\citet{MVK09} carried out a thorough study of point 1): they quantified how the properties of a lens galaxy determine its probability of creating a lensing event.
+In this work we revisit
+
+\citet{MVK09} carried out a thorough study of how the properties of a lens galaxy determine its strong lensing cross-section with respect to point sources, which is closely related to the probability of creating a lensing event.
+The lensing cross-section depends purely on the lens detection probability $\pdet$ \citep[see section 2.1.2 of][]{Son22}, but does not determine on its own
+
+ %can be derived by marginalising $\pdet$ over the source position \citep[see Equation (4) of][]{Son22}.
+%A thorough study of how the properties of a lens galaxy affect the probability of a lensing event was carried out by \citet{MVK09}.
+%In the formalism of \Eref{eq:one}, they explored how 
+In this work we take an additional step, by
+
+%There are several differences between our work and that of \citet{MVK09}.
+In this work we revisit some of the findings of \citet{MVK09}, then take the additional step of quantifying the impact that a varying
+
+The most important one is that we take the additional step of simulating realistic populations of galaxies and quantifying the difference between
+
+%Their main result is that galaxies with a cuspier or more concentrated dark matter
 

### 2022-07-01 paper/seleff.tex
finished intro

@@
    {
-I simulate lenses.
+We simulate lenses.
 }
@@
    {
-I conclude.
+We conclude.
 }
@@
 For many applications of strong lensing, however, the aim is to constrain the properties of the general galaxy or source population, $\prlens$ and $\prsource$, which are coupled in a nontrivial way via the lens detection probability $\pdet$.
-%These factors are coupled in a nontrivial way via the lens detection probability $\pdet$. 
-%The factor $\pdet$ takes into account both 
-%The factor $\pdet$ varies as a function of lens mass: it is larger 
-%Therefore, in order to obtain an unbiased estimate of $\prlens$ or $\prsource$ it is necessary to account for selection effects.
 In order to obtain an unbiased estimate of either $\prlens$ or $\prsource$, then, it is necessary to invert \Eref{eq:one}.
 In principle, this can be done with a Bayesian hierarchical formalism, provided that the lens detection probability $\pdet$ is known \citep{Son22}.
-%That, however, requires knowing the lens detection probability $\pdet$ as a function of lens and source properties, which is a difficult quantity to determine.
 For most of the existing strong lensing surveys, however, characterising $\pdet$ is very challenging: the process determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
@@
 %, under realistic assumptions for the lens detection probability $\pdet$.
-In particular, we aim to determine how strong lenses differ from the general population of foreground galaxies and background sources in terms of 1) the radial mass structure of the lenses (i.e.
-their stellar and dark matter mass density profiles); 2) the ellipticity of the lenses; 3) the size-magnitude distribution of the lensed sources.
-We use semi-analytic models to simulate large populations of strong lenses.
-We simulate images of the lenses and the lensed sources, and then simulate lensing surveys with realistic assumptions for the lens detection probability $\pdet$.
-%then use these images to determine whether they are classified as lenses by a fiducial survey, under realistic assumptions for the lens detection probability $\pdet$.
-We consider two scenarios: one in which a population of foreground massive galaxies lenses a population of extended background sources, and a second one in which the source population consists of point-like quasars.
+In particular, we aim to determine how strong lenses differ from the general population of foreground galaxies and background sources in terms of a) the radial mass structure of the lenses (i.e.
+their stellar and dark matter mass density profiles); b) the ellipticity of the lenses; c) the size-magnitude distribution of the lensed sources.
+The answer to this question depends on 1) how the lens detection probability $\pdet$ varies as a function of galaxy and source properties, and on 2) the shape of the galaxy and source parameters distribution $\prlens$ and $\prsource$.
+To understand this second point we can imagine the limiting case in which both $\prlens$ and $\prsource$ are Dirac delta functions: in this limit, $\prsl$ simply reduces to the product $\prlens\prsource$ up to a scaling constant, corresponding to a case in which the lensing bias is none.
 
-The amplitude of the lensing bias depends on two aspects: 1) how the lens detection probability $\pdet$ varies as a function of galaxy and source properties, and 2) the width of the galaxy and source parameters distribution $\plens$ and $\psource$.
-To understand this second point we can imagine the limiting case in which both $\plens$ and $\psource$ are Dirac delta functions: in this limit, $\prsl$ simply reduces to the product $\plens\psource$ up to a scaling constant, corresponding to a case in which the lensing bias is none.
-\citet{MVK09} carried out a thorough study of point 1): they quantified how the properties of a lens galaxy determine its probability of creating a lensing event.
-In this work we revisit
+\citet{MVK09} carried out a thorough study of point 1): they quantified how the properties of a lens galaxy determine its probability of creating a lensing event with a point source.
+In this work we revisit the \citet{MVK09} study, we expand it to the extended source case and, most importantly, we address point 2) as well: we simulate large populations of strong lenses using semi-analytic models, simulate lensing surveys with realistic assumptions for the lens probability $\pdet$ and quantify the lensing bias under various scenarios.
+We explore how the results change as a function of the scatter in mass parameters at fixed light, and as a function of source type: we simulate both galaxy and quasar sources.
 
-\citet{MVK09} carried out a thorough study of how the properties of a lens galaxy determine its strong lensing cross-section with respect to point sources, which is closely related to the probability of creating a lensing event.
-The lensing cross-section depends purely on the lens detection probability $\pdet$ \citep[see section 2.1.2 of][]{Son22}, but does not determine on its own
-
- %can be derived by marginalising $\pdet$ over the source position \citep[see Equation (4) of][]{Son22}.
-%A thorough study of how the properties of a lens galaxy affect the probability of a lensing event was carried out by \citet{MVK09}.
-%In the formalism of \Eref{eq:one}, they explored how 
-In this work we take an additional step, by
-
-%There are several differences between our work and that of \citet{MVK09}.
-In this work we revisit some of the findings of \citet{MVK09}, then take the additional step of quantifying the impact that a varying
-
-The most important one is that we take the additional step of simulating realistic populations of galaxies and quantifying the difference between
-
-%Their main result is that galaxies with a cuspier or more concentrated dark matter
+%In light of the results of our experiments, we discuss to what extent can studies of galaxy structure based on strong lenses be generalised to describe non-lens galaxies.
+Finally, we address the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses.
+This last point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
+Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{B+T21}, but any difference between the two lens classes can in principle introduce biases, if not corrected for.
+With this study we aim to quantify this bias.
 
@@
 In \Sref{sect:indlenses} we study individual lens systems, to look for trends in the lensing cross-section with various lens and source properties.
-In \Sref{sect:lenspop} we study
-
-I discuss the results in \Sref{sect:discuss} and draw conclusions in \Sref{sect:concl}.
+In \Sref{sect:lenspop} we describe our simulations of lens populations.
+In \Sref{sect:results} we show the results of our analysis on the simulated lens populations.
+We discuss the results in \Sref{sect:discuss} and draw conclusions in \Sref{sect:concl}.
 
@@
 
-\section{Lens populations}\label{sect:lenspop}
+\section{Lens populations simulations}\label{sect:lenspop}
 
@@
 
-\section{Discussion and summary}\label{sect:discuss}
+\section{Results}\label{sect:results}
 
-Discuss.
+There are results.
+
+%__________________________________________________________________
+
+\section{Discussion}\label{sect:discuss}
+
+We discuss.
+
+%__________________________________________________________________
+
+\section{Conclusions}\label{sect:concl}
+
+We conclude.
 

### 2022-07-03 paper/seleff.tex
started section 2

@@
 \def\detJ{\mathrm{det}J}
+
+\def\reff{R_{\mathrm{e}}}
+\def\mstar{M_*}
 
@@
 
-In this section we study how the probability of a strong lensing event varies as a function of
+In this section we study how the probability of a strong lensing event varies as a function of lens and source properties.
+In order to do so, it is useful to introduce the concept of strong lensing cross-section.
+Given a foreground galaxy with parameters $\psilens$, a background source with parameters $\psisource$, and a criterion $S$ to define a strong lensing event, the strong lensing cross-section is defined as \citep{Son22}:
+\begin{equation}\label{eq:crosssect}
+\crosssect = \int_{\mathbb{R}^2} d\boldsymbol\beta \pdet(\psilens,\psisourcenobeta,\boldsymbol\beta|S),
+\end{equation}
+where $\boldsymbol\beta$ is the position of the background source and the integral is carried out over the whole sky.
+The definition above is valid for both a point source and an extended source: although there is no unique way of defining the position of an extended source, the integral over the sky ensures that the result is independent of how the source position is defined.
+In the limit of low density of background sources, which is satisfied in all practical cases, the probability of a lensing event is proportional to $\crosssect$.
 
-\subsection{Axisymmetric lenses, point sources}
+In this section we compute $\crosssect$ in a series of different scenarios of increasing complexity.
+The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments. We describe this in \ref{ssec:profile}.
+%In section \ref{ssec:axsymmpoint} we consider axisymmetric lenses and point sources.
+In section \ref{ssec:axsymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
+In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.
+
+\subsection{Lens density profile}\label{ssec:profile}
+
+Throughout this work we describe the mass distribution of the lens galaxies with a two-component model consisting of a S\'{e}rsic profile, describing the baryonic component of a galaxy, and a generalised Navarro, Frenk \& White (gNFW) profile, describing the dark matter halo.
+We assume that the two components are concentric.
+We also assume that the baryonic component consists entirely of stars.
+This is a good approximation for the majority of the known strong lenses, which are massive elliptical galaxies with little gas.
+Finally, we assume that the light distribution of the stellar component follows its mass distribution exactly, that is we do not allow for the presence of gradients in the stellar mass-to-light ratio.
+This assumption is probably broken in real galaxies but, as we discuss in \Sref{sect:discuss}, does not impact our investigation.
+
+For the purpose of computing lensing-related quantities, the projected massdistribution $\Sigma(R)$ is needed.
+For a S\'{e}rsic profile, this is given by
+\begin{equation}
+\Sigma(R) = \Sigma_0 \exp{\left\{-b\left(\frac{R}{\reff}\right)^{1/4}\right\}},
+\end{equation}
+with
+\begin{equation}
+\Sigma_0 = \frac{\mstar b^{2n}}{2\pi n \reff^2 \Gamma(2n)},
+\end{equation}
+where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is \citep{C+B99}
+\begin{equation}
+b(n) \approx 2n -\frac13 + \frac{4}{405n} + \frac{46}{25515n^2} + O(n^{-3}).
+\end{equation}
+
+In the axisymmetric case, the projected
+
+\subsection{Axisymmetric lenses, point sources}\label{sect:axsymmpoint}
 
@@
 
-\subsection{Elliptical lenses, point sources}
+\subsection{Elliptical lenses, point sources}\label{ssec:ellpoint}
 
-\subsection{Elliptical lenses, extended sources}
+\subsection{Elliptical lenses, extended sources}\label{ssec:ellext}
 

### 2022-07-03 single_lens/axisymmetric_point/get_composite_physical_crosssect.py

@@ -1,0 +1,201 @@
+import numpy as np
+from wl_profiles import gnfw, deVaucouleurs as deV
+from wl_cosmology import Mpc, c, G, M_Sun
+import wl_cosmology
+from scipy.interpolate import splrep, splev, splint
+from scipy.optimize import brentq
+from scipy.stats import truncnorm
+from scipy.special import erf
+import emcee
+import h5py
+import sys
+
+
+# lensing cross-section of a composite (deV + gNFW) model,
+# as a function of: half-light radius, stellar mass, dark matter 
+# fraction and inner slope.
+
+reff_fixed = 7.
+lmstar_fixed = 11.5
+lmdm5_fixed = 11.
+gammadm_fixed = 1.5
+rs_fixed = 100.
+
+zd_fixed = 0.3
+zs_fixed = 1.5
+
+kpc = Mpc/1000.
+arcsec2rad = np.deg2rad(1./3600.)
+
+dd = wl_cosmology.Dang(zd_fixed)
+ds = wl_cosmology.Dang(zs_fixed)
+dds = wl_cosmology.Dang(zs_fixed, zd_fixed)
+
+arcsec2kpc = arcsec2rad * dd * 1000.
+
+s_cr_fixed = c**2/(4.*np.pi*G)*ds/dds/dd/Mpc/M_Sun*kpc**2 # critical surface mass density, in M_Sun/kpc**2
+
+# defines lensing-related functions
+def alpha_dm(x, gnfw_norm, rs, gammadm, s_cr):
+    # deflection angle (in kpc)
+    return gnfw_norm * gnfw.fast_M2d(abs(x), rs, gammadm) / np.pi/x/s_cr
+
+def alpha_star(x, mstar, reff, s_cr): 
+    # deflection angle (in kpc)
+    return mstar * deV.M2d(abs(x), reff) / np.pi/x/s_cr
+
+def alpha(x, gnfw_norm, rs, gammadm, mstar, reff, s_cr):
+    return alpha_dm(x, gnfw_norm, rs, gammadm, s_cr) + alpha_star(x, mstar, reff, s_cr)
+
+def kappa(x, gnfw_norm, rs, gammadm, mstar, reff, s_cr): 
+    # dimensionless surface mass density
[…]

### 2022-07-03 single_lens/axisymmetric_point/plot_composite_fixedap_kappa.py

@@ -1,0 +1,113 @@
+import numpy as np
+from wl_profiles import gnfw, deVaucouleurs as deV
+from wl_cosmology import Mpc, c, G, M_Sun
+import wl_cosmology
+from scipy.interpolate import splrep, splev, splint
+from scipy.optimize import brentq
+import h5py
+import sys
+import pylab
+from labellines import labelLine, labelLines
+from matplotlib.ticker import MultipleLocator
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 18
+
+reff = 1.
+rein = reff
+rs_fixed = 10. * reff
+
+s_cr = 1. # I shouldn't need to change this
+
+gammadm_plot = [1., 1.5, 2.]
+ngammadm = len(gammadm_plot)
+fdm_fixed = 0.5
+
+fdm_plot = [0.2, 0.8]
+gammadm_fixed = 1.5
+nfdm = len(fdm_plot)
+
+# defines lensing-related functions
+def alpha_dm(x, gnfw_norm, rs, gammadm):
+    # deflection angle (in kpc)
+    return gnfw_norm * gnfw.fast_M2d(abs(x), rs, gammadm) / np.pi/x/s_cr
+
+def alpha_star(x, mstar, reff): 
+    # deflection angle (in kpc)
+    return mstar * deV.M2d(abs(x), reff) / np.pi/x/s_cr
+
+def alpha(x, gnfw_norm, rs, gammadm, mstar, reff):
+    return alpha_dm(x, gnfw_norm, rs, gammadm) + alpha_star(x, mstar, reff)
+
+def kappa(x, gnfw_norm, rs, gammadm, mstar, reff): 
+    # dimensionless surface mass density
+    return (mstar * deV.Sigma(abs(x), reff) + gnfw_norm * gnfw.fast_Sigma(abs(x), rs, gammadm))/s_cr
+   
+def mu_r(x, gnfw_norm, rs, gammadm, mstar, reff):
+    # radial magnification
+    return (1. + alpha(x, gnfw_norm, rs, gammadm, mstar, reff)/x - 2.*kappa(x, gnfw_norm, rs, gammadm, mstar, reff))**(-1)
+
+def mu_t(x, gnfw_norm, rs, gammadm, mstar, reff):
+    # tangential magnification
[…]

### 2022-07-03 single_lens/axisymmetric_point/plot_composite_fixedap_scheme.py

@@ -1,0 +1,107 @@
+import numpy as np
+import pylab
+from sl_profiles import deVaucouleurs as deV, gnfw
+from matplotlib import rc
+from matplotlib.ticker import MultipleLocator
+from scipy.optimize import brentq
+rc('text', usetex=True)
+
+
+fsize = 18
+
+reff = 1.
+rein = reff
+theta_E = rein
+rs_fixed = 10. * reff
+
+s_cr = 1. # I shouldn't need to change this
+
+gammadm_plot = [1., 1.5, 2.]
+fdm_fixed = 0.5
+gammadm_fixed = 1.5
+fdm_plot = [0.2, 0.8]
+
+Rfrac_min = gnfw.R_grid[0]
+Rfrac_max = gnfw.R_grid[-1]
+
+# defines lensing-related functions
+def alpha_dm(x, gnfw_norm, rs, gammadm):
+    # deflection angle (in kpc)
+    return gnfw_norm * gnfw.fast_M2d(abs(x), rs, gammadm) / np.pi/x/s_cr
+
+def alpha_star(x, mstar, reff): 
+    # deflection angle (in kpc)
+    return mstar * deV.M2d(abs(x), reff) / np.pi/x/s_cr
+
+def alpha(x, gnfw_norm, rs, gammadm, mstar, reff):
+    return alpha_dm(x, gnfw_norm, rs, gammadm) + alpha_star(x, mstar, reff)
+
+def kappa(x, gnfw_norm, rs, gammadm, mstar, reff): 
+    # dimensionless surface mass density
+    return (mstar * deV.Sigma(abs(x), reff) + gnfw_norm * gnfw.fast_Sigma(abs(x), rs, gammadm))/s_cr
+   
+def mu_r(x, gnfw_norm, rs, gammadm, mstar, reff):
+    # radial magnification
+    return (1. + alpha(x, gnfw_norm, rs, gammadm, mstar, reff)/x - 2.*kappa(x, gnfw_norm, rs, gammadm, mstar, reff))**(-1)
+
+def mu_t(x, gnfw_norm, rs, gammadm, mstar, reff):
+    # tangential magnification
+    return (1. - alpha(x, gnfw_norm, rs, gammadm, mstar, reff)/x)**(-1)
+
+hwidth = 3.
+t_finearr = np.linspace(-hwidth, hwidth, 1001)
[…]

### 2022-07-07 paper/seleff.tex
more text

@@
 \def\mstar{M_*}
+\def\gammadm{\gamma_{\mathrm{DM}}}
+\def\fdm{f_{\mathrm{DM}}}
+\def\mdmfive{M_{\mathrm{DM}, 5}}
 
@@
 This is a good approximation for the majority of the known strong lenses, which are massive elliptical galaxies with little gas.
-Finally, we assume that the light distribution of the stellar component follows its mass distribution exactly, that is we do not allow for the presence of gradients in the stellar mass-to-light ratio.
+Finally, we assume that the light distribution of the stellar component follows its mass distribution exactly.
+That is, we do not allow for the presence of gradients in the stellar mass-to-light ratio.
 This assumption is probably broken in real galaxies but, as we discuss in \Sref{sect:discuss}, does not impact our investigation.
 
-For the purpose of computing lensing-related quantities, the projected massdistribution $\Sigma(R)$ is needed.
+Lensing-related quantities depend on the projected mass distribution $\Sigma(R)$.
+%For the purpose of computing lensing-related quantities, the projected massdistribution $\Sigma(R)$ is needed.
 For a S\'{e}rsic profile, this is given by
@@
 \end{equation}
-where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is \citep{C+B99}
+where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is given by \citep{C+B99}
 \begin{equation}
@@
 
-In the axisymmetric case, the projected
+The dark matter component is first defined by means of its three-dimensional distribution, which for a spherically symmetric gNFW profile is
+\begin{equation}
+\rho(r) = \dfrac{\rho_0}{(r/r_s)^{\gammadm}\left(1 + r/r_s\right)^{3-\gammadm}}.
+\end{equation}
+The parameter $\gammadm$ is the inner density slope, $\rho_0$ is a normalisation parameter, while $r_s$ is the scale radius.
+The logarithmic slope of the density profile transitions from $-\gammadm$ to $-3$ at a radius $r\approx r_s$.
+
+The projected surface mass density of a gNFW profile can be expressed in terms of the following integral
+\begin{equation}
+\Sigma(R) = 2r_s\rho_0 \left(\frac{R}{r_s}\right)^{1-\gammadm}\int_0^{\pi/2} dx \sin{x}(\sin{x} + R/r_s)^{\gammadm-3}.
+\end{equation}
 

### 2022-07-08 paper/seleff.tex
lensing basics

@@
 In this section we compute $\crosssect$ in a series of different scenarios of increasing complexity.
-The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments. We describe this in \ref{ssec:profile}.
+We first introduce the basic equations of gravitational lensing in section \ref{ssec:lensbasics}.
+Then, in section \ref{ssec:profile}, we present the model family adopted to describe the radial density profile of the lenses in our experiments.
+%The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments.
+We describe this in \ref{ssec:profile}, among with the basic equations of gravitational lensing.
 %In section \ref{ssec:axsymmpoint} we consider axisymmetric lenses and point sources.
@@
 In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.
+
+\subsection{Strong lensing basics}\label{ssec:lensbasics}
+
+The lensing properties of an object with respect to a source depend solely on its dimensionless surface mass density distribution $\kappa(\boldsymbol\theta)$ (also referred to as the convergence), 
+% ratio between its surface mass density distribution and the critical surface mass density for lensing:
+\begin{equation}
+\kappa(\boldsymbol\theta) = \dfrac{\Sigma(\boldsymbol\theta)}{\Sigma_{\mathrm{cs}}},
+\end{equation}
+which is the ratio between the projected surface mass density of the lens and the critical surface mass density for lensing.
+The latter quantity is defined as
+%The quantity $\Sigma_{\matrhrm{cs}}$ is the critical surface mass density for lensing, defined as
+\begin{equation}
+\Sigma_{cr} = \dfrac{c^2D_s}{4\pi G D_d D_{ds}},
+\end{equation}
+where $c$ is the speed of light, $G$ the gravitational constant, and $D_d$, $D_s$, and $D_{ds}$ are the angular diameter distances between the observer and the lens, the observer and the source, and the lens and the source, respectively.
+
+Given a source at angular position $\boldsymbol\beta$, images of it form at the positions $\boldsymbol\theta$ that are solutions of the lens equation
+\begin{equation}\label{eq:lenseq}
+\boldsymbol\beta = \boldsymbol\theta - \boldsymbol\alpha(\boldsymbol\theta),
+\end{equation}
+where $\boldsymbol\alpha$ is the deflection angle of the lens.
+This can be expressed in terms of the dimensionless surface mass density by means of the following integral over the whole sky:
+\begin{equation}
+\boldsymbol\alpha(\boldsymbol\theta) = \frac{1}{\pi}\int_{\mathbb{R}^2} d^2\boldsymbol\theta' \kappa(\boldsymbol\theta')\dfrac{\boldsymbol\theta - \boldsymbol\theta'}{|\boldsymbol\theta - \boldsymbol\theta'|^2}.
+\end{equation}
+
+In the special case of an axisymmetric lens, the deflection angle, calculated along an arbitrary axis passing through the lens centre, reduces to
+\begin{equation}
+\alpha(\theta) = \frac{2}{\theta}\int_0^{\theta} d\theta' \kappa(\theta') \theta'.
+\end{equation}
+This can also be expressed in terms of the total projected mass enclosed within a circle with angular radius equal to $\theta$:
+\begin{equation}
+\alpha(\theta) = \frac{1}{\pi\theta}\dfrac{M(<\theta)}{\Sigma_{\mathrm{cr}}D_{\mathrm{d}}^2}.
+\end{equation}
+
+A very important quantity for describing the strength of a strong lens is the Einstein radius, $\tein$.
+For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:lenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
+It can be shown that $\tein$ satisfies the following condition,
+\begin{equation}\label{eq:criteq}
+\bar{\kappa}(<\tein) = 1,
+\end{equation}
+where $\bar{\kappa}(<\theta)$ is the average surface mass density within a radius equal to $\theta$:
+\begin{equation}
+\bar{\kappa}(<\theta) \equiv \frac{2}{\theta^2}\int_0^{\theta}d\theta' \kappa(\theta')\theta'.
+\end{equation}
+Based on \Eref{eq:criteq},
+%The average surface mass density within $\tein$ is equal to one.
+%For this reason, 
+we generalise the definition of $\tein$ to the non-axisymmetric case as the circularised radius of the isodensity contour that encloses an average surface mass density equal to $\Sigma_{\mathrm{cr}}$.
 
@@
 
-Lensing-related quantities depend on the projected mass distribution $\Sigma(R)$.
+%Lensing-related quantities depend on the projected mass distribution $\Sigma(R)$.
 %For the purpose of computing lensing-related quantities, the projected massdistribution $\Sigma(R)$ is needed.
-For a S\'{e}rsic profile, this is given by
+%For a S\'{e}rsic profile, this is given by
+The projected surface mass density of a S\'{e}rsic profile is given by
 \begin{equation}
@@
 The logarithmic slope of the density profile transitions from $-\gammadm$ to $-3$ at a radius $r\approx r_s$.
-
-The projected surface mass density of a gNFW profile can be expressed in terms of the following integral
+The projected surface mass density of a gNFW profile can be expressed in terms of the following integral \citep{WTS01}:
 \begin{equation}

### 2022-07-09 paper/seleff.tex
more section 2

@@
 
-In the special case of an axisymmetric lens, the deflection angle, calculated along an arbitrary axis passing through the lens centre, reduces to
+In the special case of an axisymmetric lens we can simplify the notation by considering a single coordinate axis with origin at the centre of the lens.
+We label the components of the image position, source position and deflection angle along this axis as $\theta$, $\beta$ and $\alpha$, respectively.
+The lens equation for an axisymmetric lens then becomes
+\begin{equation}\label{eq:1dlenseq}
+\beta = \theta - \alpha(\theta),
+\end{equation}
+and the expression for the deflection angle reduces to
+%In the special case of an axisymmetric lens, the deflection angle, calculated along an arbitrary axis passing through the lens centre, reduces to
 \begin{equation}
@@
 For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:lenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
+The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form here have infinite magnification in the tangential direction.
+
 It can be shown that $\tein$ satisfies the following condition,
@@
 
+\Fref{fig:kappa} shows the dimensionless surface mass density profile S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
+All of the profiles have a S\'{e}rsic index $n=4$, a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component.
+Two main features emerge from \Fref{fig:kappa}.
+First, the baryons generally dominate the total density in the inner regions $(\theta < \tein)$, while the dark matter is the main component at large radii.
+Second, models with different dark matter fractions and inner dark matter slopes can conspire to produce very similar total density profiles. 
+%it is possible to build models with different dark matter fractions and inner dark matter slopes, but with very similar total density profiles. 
+This is the case, for example, of the ($\fdm=0.5,\gammadm=1.5$) and the ($\fdm=0.2,\gammadm=2.0$) models (green and blue lines in \Fref{fig:kappa}).
+
+%
+\begin{figure}
+\includegraphics[width=\columnwidth]{composite_fixedap_kappa.eps}
+\caption{
+Dimensionless surface mass density profile of S\'{e}rsic + gNFW composite models.
+The S\'{e}rsic index of the baryonic component is fixed to $n=4$ and the scale radius of the dark matter component is fixed to ten times the half-light radius.
+All of the profiles are normalised in such a way that the Einstein radius is equal to the half-light radius.
+Solid lines: total density profile. Dotted lines: dark matter density profile. Dashed lines: baryonic density profile.
+The blue, orange and green dashed lines are identical, as they correspond to profiles with the same fraction of baryonic mass within the half-light radius.
+\label{fig:kappa}
+}
+\end{figure}
+%
+
 \subsection{Axisymmetric lenses, point sources}\label{sect:axsymmpoint}
+
+Axisymmetric lenses with a density profile of the kind considered in section \ref{ssec:profile} typically produce either one or three images of a point source.
+\Fref{fig:scheme} helps to visualise this property.
+%\Fref{fig:scheme}, which shows the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, helps to visualise this property. 
+Plotted in \Fref{fig:scheme} is the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, for the same lens models considered in \Fref{fig:kappa}.
+According to \Eref{eq:1dlenseq}, images form at the locations where this quantity equals the position of the source. 
+Therefore, given a source position $\beta$, the number of images and their location can be determined by drawing a horizontal line at the value $\beta$ on the vertical axis, and finding the points where this line intersects the $\theta - \alpha(\theta)$ curves.
+
+For small values of $\beta$ (for sources close to the lens centre), the number of images that are produced is three: the source is strongly lensed.
+For large values of $\beta$, instead, only one image is formed.
+The value of $\beta$ where the transition occurs is known as the radial caustic, which is marked by the horizontal dotted line in \Fref{fig:scheme}. 
+As can be seen from \Fref{fig:scheme}, a source at this location is mapped to the stationary point of the $\theta - \alpha(\theta)$ curve. 
+%this value corresponds to the stationary point of the $\theta - \alpha(\theta)$ curve. 
+That location on the image plane is known as the radial critical curve. 
+Images that form there have a formally infinite magnification in the radial direction\footnote{The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the radial magnification.
+This can be understood by taking the derivative of \Eref{eq:1dlenseq} with respect to $\theta$.}.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{composite_fixedap_scheme.eps}
+\caption{
+The lens equation of axisymmetric lenses. Solid lines: right-hand side of \Eref{eq:1dlenseq} for the five lenses with density profile plotted in \Fref{fig:kappa}. 
+Given a source at position $\beta$, its lensed images form at the values of $\theta$ where the solid line intersects a horizontal line located at $\beta$ on the vertical axis. 
+Dotted lines: positions of the radial caustics. Sources located within the radial caustic of a given lens produce three images.
+\label{fig:scheme}
+}
+\end{figure}
+%The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the magnification of an image in the radial direction. 
+%This can be understood by noting that, according to \Eref{eq:1dlenseq}, the derivative of $\theta - \alpha(\theta)$ is equal to $d\beta/d\theta$, which quantifies the ratio between an infinitesimal displacement in the radial direction in the source plane and the corresponding displacement in the image plane.
+%Stationary points of the $\theta - \alpha(\theta)$ curve, therefore, are points of infinite magnification in the radial direction.
+For a given axisymmetric lens, these points describe a circle known as the radial critical curve.
+%The corresponding curve in the source plane, the radius of which is given by the value of $\theta - \alpha(\theta)$ at the stationary point, is known as the radial caustic.
+%As can be seen in \Fref{fig:scheme}, sources that lie within the radial caustic are strongly lensed into three images, while sources that lie outside of it produce only one image.
+
+%Strong lenses also have a tangential critical curve, where the magnification along the tangential direction is formally infinite.
+For an axisymmetric lens, the tangential critical curve is a circle with radius equal to the Einstein radius, and the corresponding tangential caustic collapses to the $\beta=0$ point in the source plane.
 

### 2022-07-09 single_lens/axisymmetric_point/plot_composite_fixedap_scheme.py

@@ -9,6 +9,8 @@ rc('text', usetex=True)
 
 fsize = 18
 
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+
 reff = 1.
 rein = reff
 theta_E = rein
@@ -17,9 +19,11 @@ rs_fixed = 10. * reff
 s_cr = 1. # I shouldn't need to change this
 
 gammadm_plot = [1., 1.5, 2.]
+ngamma = len(gammadm_plot)
 fdm_fixed = 0.5
 gammadm_fixed = 1.5
 fdm_plot = [0.2, 0.8]
+nfdm = len(fdm_plot)
 
 Rfrac_min = gnfw.R_grid[0]
 Rfrac_max = gnfw.R_grid[-1]
@@ -69,23 +73,45 @@ mstar_einfrac = deV.M2d(rein, reff)
 
 xmin = max(deV.rgrid_min*reff, Rfrac_min*rs_fixed)
 
-for gammadm in gammadm_plot:
+for i in range(ngamma):
+
+    gammadm = gammadm_plot[i]
 
     mstar_here = np.pi * rein**2 / mstar_einfrac * (1. - fdm_fixed)
     mdm_ein_here = np.pi * rein**2 * fdm_fixed
 
     gnfw_norm = mdm_ein_here / gnfw.fast_M2d(rein, rs_fixed, gammadm)
 
-    ax.plot(t_finearr, t_finearr - alpha(t_finearr, gnfw_norm, rs_fixed, gammadm, mstar_here, reff), label='$f_{\mathrm{DM}} = %2.1f$, $\gamma_{\mathrm{DM}}=%2.1f$'%(fdm_fixed, gammadm))
+    ax.plot(t_finearr, t_finearr - alpha(t_finearr, gnfw_norm, rs_fixed, gammadm, mstar_here, reff), label='$f_{\mathrm{DM}} = %2.1f$, $\gamma_{\mathrm{DM}}=%2.1f$'%(fdm_fixed, gammadm), color=colseq[i])
 
-for fdm in fdm_plot:
+    # looks for radial critical curve
+    def zerofunc(x):
+        return 1./mu_r(x, gnfw_norm, rs_fixed, gammadm, mstar_here, reff)
+
+    x_rad = brentq(zerofunc, -rein, -0.01)
+    beta_rad = x_rad - alpha(x_rad, gnfw_norm, rs_fixed, gammadm, mstar_here, reff)
[…]

### 2022-07-09 paper/seleff.tex
reorganized secgtion 2

@@
 The structure of this work is the following.
+\Sref{sect:basics} introduces the basics of gravitational lensing.
 In \Sref{sect:indlenses} we study individual lens systems, to look for trends in the lensing cross-section with various lens and source properties.
@@
 
-\section{Individual lenses}\label{sect:indlenses}
-
-In this section we study how the probability of a strong lensing event varies as a function of lens and source properties.
-In order to do so, it is useful to introduce the concept of strong lensing cross-section.
-Given a foreground galaxy with parameters $\psilens$, a background source with parameters $\psisource$, and a criterion $S$ to define a strong lensing event, the strong lensing cross-section is defined as \citep{Son22}:
-\begin{equation}\label{eq:crosssect}
-\crosssect = \int_{\mathbb{R}^2} d\boldsymbol\beta \pdet(\psilens,\psisourcenobeta,\boldsymbol\beta|S),
-\end{equation}
-where $\boldsymbol\beta$ is the position of the background source and the integral is carried out over the whole sky.
-The definition above is valid for both a point source and an extended source: although there is no unique way of defining the position of an extended source, the integral over the sky ensures that the result is independent of how the source position is defined.
-In the limit of low density of background sources, which is satisfied in all practical cases, the probability of a lensing event is proportional to $\crosssect$.
-
-In this section we compute $\crosssect$ in a series of different scenarios of increasing complexity.
-We first introduce the basic equations of gravitational lensing in section \ref{ssec:lensbasics}.
-Then, in section \ref{ssec:profile}, we present the model family adopted to describe the radial density profile of the lenses in our experiments.
-%The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments.
-We describe this in \ref{ssec:profile}, among with the basic equations of gravitational lensing.
-%In section \ref{ssec:axsymmpoint} we consider axisymmetric lenses and point sources.
-In section \ref{ssec:axsymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
-In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.
-
-\subsection{Strong lensing basics}\label{ssec:lensbasics}
+\section{Strong lensing basics}\label{sect:basics}
 
@@
 \end{equation}
+
+\subsection{The axisymmetric lens}
 
@@
 we generalise the definition of $\tein$ to the non-axisymmetric case as the circularised radius of the isodensity contour that encloses an average surface mass density equal to $\Sigma_{\mathrm{cr}}$.
+
+Axisymmetric lenses with a density profile of the kind considered in section \ref{ssec:profile} typically produce either one or three images of a point source.
+\Fref{fig:scheme} helps to visualise this property.
+%\Fref{fig:scheme}, which shows the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, helps to visualise this property. 
+Plotted in \Fref{fig:scheme} is the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, for the same lens models considered in \Fref{fig:kappa}.
+According to \Eref{eq:1dlenseq}, images form at the locations where this quantity equals the position of the source. 
+Therefore, given a source position $\beta$, the number of images and their location can be determined by drawing a horizontal line at the value $\beta$ on the vertical axis, and finding the points where this line intersects the $\theta - \alpha(\theta)$ curves.
+
+For small values of $\beta$ (for sources close to the lens centre), the number of images that are produced is three: the source is strongly lensed.
+For large values of $\beta$, instead, only one image is formed.
+The value of $\beta$ where the transition occurs is known as the radial caustic, which is marked by the horizontal dotted line in \Fref{fig:scheme}. 
+As can be seen from \Fref{fig:scheme}, a source at this location is mapped to the stationary point of the $\theta - \alpha(\theta)$ curve. 
+%this value corresponds to the stationary point of the $\theta - \alpha(\theta)$ curve. 
+That location on the image plane is known as the radial critical curve. 
+Images that form there have a formally infinite magnification in the radial direction\footnote{The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the radial magnification.
+This can be understood by taking the derivative of \Eref{eq:1dlenseq} with respect to $\theta$.}.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{composite_fixedap_scheme.eps}
+\caption{
+The lens equation of axisymmetric lenses. Solid lines: right-hand side of \Eref{eq:1dlenseq} for the five lenses with density profile plotted in \Fref{fig:kappa}. 
+Given a source at position $\beta$, its lensed images form at the values of $\theta$ where the solid line intersects a horizontal line located at $\beta$ on the vertical axis. 
+Dotted lines: positions of the radial caustics. Sources located within the radial caustic of a given lens produce three images.
+\label{fig:scheme}
+}
+\end{figure}
+%The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the magnification of an image in the radial direction. 
+%This can be understood by noting that, according to \Eref{eq:1dlenseq}, the derivative of $\theta - \alpha(\theta)$ is equal to $d\beta/d\theta$, which quantifies the ratio between an infinitesimal displacement in the radial direction in the source plane and the corresponding displacement in the image plane.
+%Stationary points of the $\theta - \alpha(\theta)$ curve, therefore, are points of infinite magnification in the radial direction.
+For a given axisymmetric lens, these points describe a circle known as the radial critical curve.
+%The corresponding curve in the source plane, the radius of which is given by the value of $\theta - \alpha(\theta)$ at the stationary point, is known as the radial caustic.
+%As can be seen in \Fref{fig:scheme}, sources that lie within the radial caustic are strongly lensed into three images, while sources that lie outside of it produce only one image.
+
+%Strong lenses also have a tangential critical curve, where the magnification along the tangential direction is formally infinite.
+For an axisymmetric lens, the tangential critical curve is a circle with radius equal to the Einstein radius, and the corresponding tangential caustic collapses to the $\beta=0$ point in the source plane.
+
+\subsection{The elliptical lens}
+
+%__________________________________________________________________
+
+\section{Individual lenses}\label{sect:indlenses}
+
+In this section we study how the probability of a strong lensing event varies as a function of lens and source properties.
+In order to do so, it is useful to introduce the concept of strong lensing cross-section.
+Given a foreground galaxy with parameters $\psilens$, a background source with parameters $\psisource$, and a criterion $S$ to define a strong lensing event, the strong lensing cross-section is defined as \citep{Son22}:
+\begin{equation}\label{eq:crosssect}
+\crosssect = \int_{\mathbb{R}^2} d\boldsymbol\beta \pdet(\psilens,\psisourcenobeta,\boldsymbol\beta|S),
+\end{equation}
+where $\boldsymbol\beta$ is the position of the background source and the integral is carried out over the whole sky.
+The definition above is valid for both a point source and an extended source: although there is no unique way of defining the position of an extended source, the integral over the sky ensures that the result is independent of how the source position is defined.
+In the limit of low density of background sources, which is satisfied in all practical cases, the probability of a lensing event is proportional to $\crosssect$.
+
+In this section we compute $\crosssect$ in a series of different scenarios of increasing complexity.
+%We first introduce the basic equations of gravitational lensing in section \ref{ssec:lensbasics}.
+%Then, in section \ref{ssec:profile}, we present the model family adopted to describe the radial density profile of the lenses in our experiments.
+The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments.
+We describe this in \ref{ssec:profile}.
+%, among with the basic equations of gravitational lensing.
+%In section \ref{ssec:axsymmpoint} we consider axisymmetric lenses and point sources.
+In section \ref{ssec:axsymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
+In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.
+
+%\subsection{Strong lensing basics}\label{ssec:lensbasics}
 
@@
 
-Axisymmetric lenses with a density profile of the kind considered in section \ref{ssec:profile} typically produce either one or three images of a point source.
-\Fref{fig:scheme} helps to visualise this property.
-%\Fref{fig:scheme}, which shows the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, helps to visualise this property. 
-Plotted in \Fref{fig:scheme} is the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, for the same lens models considered in \Fref{fig:kappa}.
-According to \Eref{eq:1dlenseq}, images form at the locations where this quantity equals the position of the source. 
-Therefore, given a source position $\beta$, the number of images and their location can be determined by drawing a horizontal line at the value $\beta$ on the vertical axis, and finding the points where this line intersects the $\theta - \alpha(\theta)$ curves.
-
-For small values of $\beta$ (for sources close to the lens centre), the number of images that are produced is three: the source is strongly lensed.
-For large values of $\beta$, instead, only one image is formed.
-The value of $\beta$ where the transition occurs is known as the radial caustic, which is marked by the horizontal dotted line in \Fref{fig:scheme}. 
-As can be seen from \Fref{fig:scheme}, a source at this location is mapped to the stationary point of the $\theta - \alpha(\theta)$ curve. 
-%this value corresponds to the stationary point of the $\theta - \alpha(\theta)$ curve. 
-That location on the image plane is known as the radial critical curve. 
-Images that form there have a formally infinite magnification in the radial direction\footnote{The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the radial magnification.
-This can be understood by taking the derivative of \Eref{eq:1dlenseq} with respect to $\theta$.}.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{composite_fixedap_scheme.eps}
-\caption{
-The lens equation of axisymmetric lenses. Solid lines: right-hand side of \Eref{eq:1dlenseq} for the five lenses with density profile plotted in \Fref{fig:kappa}. 
-Given a source at position $\beta$, its lensed images form at the values of $\theta$ where the solid line intersects a horizontal line located at $\beta$ on the vertical axis. 
-Dotted lines: positions of the radial caustics. Sources located within the radial caustic of a given lens produce three images.
-\label{fig:scheme}
-}
-\end{figure}
-%The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the magnification of an image in the radial direction. 
-%This can be understood by noting that, according to \Eref{eq:1dlenseq}, the derivative of $\theta - \alpha(\theta)$ is equal to $d\beta/d\theta$, which quantifies the ratio between an infinitesimal displacement in the radial direction in the source plane and the corresponding displacement in the image plane.
-%Stationary points of the $\theta - \alpha(\theta)$ curve, therefore, are points of infinite magnification in the radial direction.
-For a given axisymmetric lens, these points describe a circle known as the radial critical curve.
-%The corresponding curve in the source plane, the radius of which is given by the value of $\theta - \alpha(\theta)$ at the stationary point, is known as the radial caustic.
-%As can be seen in \Fref{fig:scheme}, sources that lie within the radial caustic are strongly lensed into three images, while sources that lie outside of it produce only one image.
-
-%Strong lenses also have a tangential critical curve, where the magnification along the tangential direction is formally infinite.
-For an axisymmetric lens, the tangential critical curve is a circle with radius equal to the Einstein radius, and the corresponding tangential caustic collapses to the $\beta=0$ point in the source plane.
-
 \begin{figure}

### 2022-07-10 paper/seleff.tex
glafic tests. Section 2

@@
 The structure of this work is the following.
-\Sref{sect:basics} introduces the basics of gravitational lensing.
+In \Sref{sect:basics} we introduce the basics of gravitational lensing.
 In \Sref{sect:indlenses} we study individual lens systems, to look for trends in the lensing cross-section with various lens and source properties.
@@
 
-The lensing properties of an object with respect to a source depend solely on its dimensionless surface mass density distribution $\kappa(\boldsymbol\theta)$ (also referred to as the convergence), 
-% ratio between its surface mass density distribution and the critical surface mass density for lensing:
+The lensing properties of an object with respect to a source depend solely on its dimensionless surface mass density distribution $\kappa(\boldsymbol\theta)$ (also referred to as the convergence). 
+This is the ratio between the surface mass density and the critical surface mass density for lensing:
 \begin{equation}
-\kappa(\boldsymbol\theta) = \dfrac{\Sigma(\boldsymbol\theta)}{\Sigma_{\mathrm{cs}}},
+\kappa(\boldsymbol\theta) = \dfrac{\Sigma(\boldsymbol\theta)}{\Sigma_{\mathrm{cs}}}.
 \end{equation}
-which is the ratio between the projected surface mass density of the lens and the critical surface mass density for lensing.
+%which is the ratio between the projected surface mass density of the lens and the critical surface mass density for lensing.
 The latter quantity is defined as
@@
 The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form here have infinite magnification in the tangential direction.
-
+%
 It can be shown that $\tein$ satisfies the following condition,
@@
 \end{equation}
-Based on \Eref{eq:criteq},
-%The average surface mass density within $\tein$ is equal to one.
-%For this reason, 
-we generalise the definition of $\tein$ to the non-axisymmetric case as the circularised radius of the isodensity contour that encloses an average surface mass density equal to $\Sigma_{\mathrm{cr}}$.
+%Based on \Eref{eq:criteq},
+%we generalise the definition of $\tein$ to the non-axisymmetric case as the circularised radius of the isodensity contour that encloses an average surface mass density equal to $\Sigma_{\mathrm{cr}}$.
 
-Axisymmetric lenses with a density profile of the kind considered in section \ref{ssec:profile} typically produce either one or three images of a point source.
+Axisymmetric lenses of the kind considered in this work 
+%Axisymmetric lenses with a density profile of the kind considered in section \ref{ssec:profile} 
+typically produce either one or three images of a point source.
 \Fref{fig:scheme} helps to visualise this property.
 %\Fref{fig:scheme}, which shows the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, helps to visualise this property. 
-Plotted in \Fref{fig:scheme} is the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, for the same lens models considered in \Fref{fig:kappa}.
+Plotted in \Fref{fig:scheme} is the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, for a few lens models.
+%for the same lens models considered in \Fref{fig:kappa}.
 According to \Eref{eq:1dlenseq}, images form at the locations where this quantity equals the position of the source. 
-Therefore, given a source position $\beta$, the number of images and their location can be determined by drawing a horizontal line at the value $\beta$ on the vertical axis, and finding the points where this line intersects the $\theta - \alpha(\theta)$ curves.
+Therefore, given a source position $\beta$, the number of images and their location can be determined by drawing a horizontal line at the value $\beta$ on the vertical axis, and finding the points where this line intersects the $\theta - \alpha(\theta)$ curve.
 
@@
 \caption{
-The lens equation of axisymmetric lenses. Solid lines: right-hand side of \Eref{eq:1dlenseq} for the five lenses with density profile plotted in \Fref{fig:kappa}. 
+The lens equation of axisymmetric lenses. Solid lines: right-hand side of \Eref{eq:1dlenseq} for five lenses with different density profiles. 
 Given a source at position $\beta$, its lensed images form at the values of $\theta$ where the solid line intersects a horizontal line located at $\beta$ on the vertical axis. 
 Dotted lines: positions of the radial caustics. Sources located within the radial caustic of a given lens produce three images.
+The lens models used in this simulations consist of a stellar component and a dark matter halo, as described in section \ref{ssec:profile}. Their density profile is plotted in \Fref{fig:kappa}.
 \label{fig:scheme}

### 2022-07-10 single_lens/elliptical_point/test_sersic.py

@@ -1,0 +1,50 @@
+import numpy as np
+import os
+import glafic
+import h5py
+
+
+np.random.seed(0)
+
+# primary parameters
+omegaM = 0.3
+omegaL = 0.7
+weos = -1.
+hubble = 0.7
+prefix = 'sersic'
+xmin = -3.
+ymin = -3.
+xmax = 3.
+ymax = 3.
+pix_ext = 0.2
+pix_poi = 0.1
+maxlev = 5
+
+glafic.init(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0)
+
+# sets parameters of the Sersic profile
+reff = 1.
+nser = 4.
+q = 1.
+e = 1. - q
+zd = 0.3
+zs = 1.5
+Mtot = 1e11 # this should be in units of M_Sun/h
+
+glafic.startup_setnum(1, 0, 1)
+glafic.set_lens(1, 'sers', zd, Mtot, 0.0, 0.0, e, 0.0, reff, nser)
+glafic.set_point(1, 1.5, 0.5, 0.)
+
+# model_init needs to be done again whenever model parameters are changed
+glafic.model_init(verb = 0)
+
+# writes critical curves and caustics
+#glafic.writecrit(zs)
+
+# calculates the Einstein radius
+glafic.calcein2(zs, 0., 0., 1)
+
+#os.system('mv out_crit.dat sersic_crit.dat')
+
+glafic.quit()
+

### 2022-07-11 paper/seleff.tex
caustics of composite model

@@
 
+\begin{figure}
+\includegraphics[width=\columnwidth]{composite_caustics.eps}
+\caption{
+Caustics of lenses with fixed radial structure and different ellipticity.
+The outer curve is the radial caustic, while the inner diamond is the tangential caustic.
+Sources located outside the caustic are not strongly lensed.
+Sources that lie in the region enclosed between the two caustics produce three images, while sources inside the tangential caustic produce five images.
+The lens model adopted for this experiment consists of a stellar component and a dark matter halo. The two components have the same ellipticity.
+}
+\end{figure}
+
 %__________________________________________________________________

### 2022-07-11 single_lens/elliptical_point/draw_composite_caust.py

@@ -1,227 +1,0 @@
-import numpy as np
-import os
-#import glafic
-import h5py
-from wl_profiles import gnfw, deVaucouleurs as deV
-from wl_cosmology import Mpc, c, G, M_Sun
-import wl_cosmology
-from scipy.optimize import brentq
-
-
-np.random.seed(0)
-
-# primary parameters
-omegaM = 0.3
-omegaL = 0.7
-weos = -1.
-hubble = 0.7
-prefix = 'out'
-xmin = -3.
-ymin = -3.
-xmax = 3.
-ymax = 3.
-pix_ext = 0.2
-pix_poi = 0.1
-maxlev = 5
-
-#glafic.init(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0, nfw_users=1, flag_hodensity=2, hodensity=200.)
-
-# defines the lens parameters
-zd = 0.3
-zs = 1.5
-
-dd = wl_cosmology.Dang(zd)
-ds = wl_cosmology.Dang(zs)
-dds = wl_cosmology.Dang(zs, zd)
-
-kpc = Mpc/1000.
-arcsec2rad = np.deg2rad(1./3600.)
-arcsec2kpc = arcsec2rad * dd * 1000.
-
-s_cr = c**2/(4.*np.pi*G)*ds/dds/dd/Mpc/M_Sun*kpc**2 # critical surface mass density, in M_Sun/kpc**2
-
-# I want a lens with a 1" Einstein radius
-rein_arcsec = 1.
-rein_phys = rein_arcsec * arcsec2kpc
-mein_phys = np.pi * rein_phys**2 * s_cr
-
-# and with Reff = rein
-reff_arcsec = rein_arcsec
-reff_phys = reff_arcsec * arcsec2kpc
-
-# and with f_dm = 0.5
-f_dm = 0.5
-mstar_phys = (1. - f_dm) * mein_phys / deV.M2d(reff_phys, reff_phys)
-
-# and with dark matter slope of 1.5
-gammadm = 1.5
-
-# Finds the virial mass of the halo
-rs_phys = 100.
-rs_arcsec = rs_phys / arcsec2kpc
-
-mein_dm = f_dm * mein_phys
-gnfw_norm = mein_dm / gnfw.M2d(rein_phys, rs_phys, gammadm)
-
-rhoc = wl_cosmology.rhoc(zd)
-
[…]

### 2022-07-11 single_lens/elliptical_point/plot_composite_caustics.py

@@ -9,14 +9,28 @@ rc('text', usetex=True)
 
 fsize = 18
 
-e_list = [0., 0.1, 0.2, 0.3]
+e_list = [0., 0.1, 0.2, 0.3, 0.4]
 ne = len(e_list)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
 fig, ax = pylab.subplots(figsize=(5.5, 5.5))
 
-pylab.subplots_adjust(left=0., right=1., bottom=0., top=1.)
+pylab.subplots_adjust(left=0.17, right=1., bottom=0.17, top=1.)
+
+xlim = (-0.95, 0.95)
+ylim = (-0.95, 0.95)
+
+"""
+xgrid = np.arange(xlim[0], xlim[1], 0.05)
+ygrid = np.arange(ylim[0], ylim[1], 0.05)
+
+for i in range(len(xgrid)):
+    ax.axvline(xgrid[i], color='grey', linewidth=0.5)
+
+for i in range(len(ygrid)):
+    ax.axhline(ygrid[i], color='grey', linewidth=0.5)
+"""
 
 for i in range(ne):
 
@@ -42,8 +56,24 @@ for i in range(ne):
 
 #ax.set_aspect(1.)
 
-pylab.legend(loc='lower right', fontsize=fsize)
+ax.tick_params(axis='both', which='both', top=True, right=True, labelsize=fsize, direction='in')#, width=1)
+ax.set_xlabel('$\\beta_x/\\theta_{\mathrm{Ein}}$', fontsize=fsize)
+ax.set_ylabel('$\\beta_y/\\theta_{\mathrm{Ein}}$', fontsize=fsize)
 
-#pylab.axis('off')
+ax.xaxis.set_major_locator(MultipleLocator(0.2))
+ax.xaxis.set_minor_locator(MultipleLocator(0.05))
+
+ax.yaxis.set_major_locator(MultipleLocator(0.2))
+ax.yaxis.set_minor_locator(MultipleLocator(0.05))
+
+ax.set_xlim(xlim[0], xlim[1])
+ax.set_ylim(ylim[0], ylim[1])
+
+pylab.arrow(-0.75, 0.78, 0.2, 0, length_includes_head=False, head_width=0.05, color='k')
+pylab.text(-0.8, 0.82, 'Major axis', fontsize=fsize)
+
+pylab.legend(loc='lower right', framealpha=1., fontsize=fsize)
+
[…]

### 2022-07-11 single_lens/elliptical_point/set_composite_pars.py

@@ -1,0 +1,114 @@
+import numpy as np
+import os
+#import glafic
+import h5py
+from wl_profiles import gnfw, deVaucouleurs as deV
+from wl_cosmology import Mpc, c, G, M_Sun
+import wl_cosmology
+from scipy.optimize import brentq
+
+
+np.random.seed(0)
+
+# primary parameters
+omegaM = 0.3
+omegaL = 0.7
+weos = -1.
+hubble = 0.7
+prefix = 'composite'
+xmin = -3.
+ymin = -3.
+xmax = 3.
+ymax = 3.
+pix_ext = 0.2
+pix_poi = 0.1
+maxlev = 5
+
+#glafic.init(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0)#, nfw_users=1, flag_hodensity=2, hodensity=200.)
+
+#glafic.reset_par('nfw_users', 1)
+
+# defines the lens parameters
+zd = 0.3
+zs = 1.5
+
+dd = wl_cosmology.Dang(zd)
+ds = wl_cosmology.Dang(zs)
+dds = wl_cosmology.Dang(zs, zd)
+
+kpc = Mpc/1000.
+arcsec2rad = np.deg2rad(1./3600.)
+arcsec2kpc = arcsec2rad * dd * 1000.
+
+s_cr = c**2/(4.*np.pi*G)*ds/dds/dd/Mpc/M_Sun*kpc**2 # critical surface mass density, in M_Sun/kpc**2
+
+# I want a lens with a 1" Einstein radius
+rein_arcsec = 1.
+rein_phys = rein_arcsec * arcsec2kpc
+mein_phys = np.pi * rein_phys**2 * s_cr
+
+# and with Reff = rein
+reff_arcsec = rein_arcsec
+reff_phys = reff_arcsec * arcsec2kpc
+
+# and with f_dm = 0.5
+f_dm = 0.5
+mstar_phys = (1. - f_dm) * mein_phys / deV.M2d(reff_phys, reff_phys)
+
+# and with dark matter slope of 1.5
+gammadm = 1.5
+
+# Finds the virial mass of the halo
+rs_phys = 100.
+rs_arcsec = rs_phys / arcsec2kpc
+
+mein_dm = f_dm * mein_phys
+gnfw_norm = mein_dm / gnfw.M2d(rein_phys, rs_phys, gammadm)
+
[…]

### 2022-07-12 paper/seleff.tex
section 3

@@
 
-\subsection{The axisymmetric lens}
+\subsection{The axisymmetric lens}\label{ssec:axisymm}
 
@@
 
+In this paper we focus mostly on lens galaxies with elliptical isodensity contours.
+Given a surface mass density profile $\Sigma(R)$, a lens with an elliptical mass distribution can be obtained by replacing the radial coordinate with the circularised radius:
+\begin{equation}
+R \rightarrow \sqrt{qx^2 + \frac{y^2}{q}},
+\end{equation}
+where $x$ and $y$ are cartesian axes centred on the lens centre, with $x$ pointing towards the major axis, and where $q$ is the minor-to-major axis ratio.
+
+\Fref{fig:ellcaust} shows the source-plane caustics of elliptical lenses with different values of the axis ratio.
+The outermost curves are radial caustics, while the inner ones are tangential ones.
+%While in the axisymmetric case these are a circle and a point respectively (blue curves), in the elliptical case they are transformed into a 
+The most striking difference with respect to the axisymmetric case (blue curves in \Fref{fig:ellcaust}) is the fact that the tangential caustic is transformed from a point into a diamond-like curve.
+Sources located within the diamond produce five images (one of which is usually highly de-magnified).
+Sources that lie in the region enclosed between the two caustics are imaged three times.
+Sources outside the radial caustic are imaged only once, as in the axisymmetric case.
+The fact that the number of images changes by two at a caustic crossing is a general feature of gravitational lenses with no singularities \citep{SEF92}.
+
 \begin{figure}
@@
 Caustics of lenses with fixed radial structure and different ellipticity.
+Source-plane angular coordinates are in units of the Einstein radius.
 The outer curve is the radial caustic, while the inner diamond is the tangential caustic.
@@
 Sources that lie in the region enclosed between the two caustics produce three images, while sources inside the tangential caustic produce five images.
-The lens model adopted for this experiment consists of a stellar component and a dark matter halo. The two components have the same ellipticity.
+The lens model adopted for this experiment consists of a stellar component and a dark matter halo, as described in \ref{ssec:profile}. The two components have the same ellipticity.
+\label{fig:ellcaust}
 }
 \end{figure}
+
+\subsection{Lensing event definition: point sources}\label{ssec:lensdefpoint}
+
+In order to compute the probability of a lensing event we must provide an exact definition for it.
+A necessary condition for a lens-source system to qualify as a strong lens is the presence of multiple images of the source.
+As we showed above, this requires the source to lie within the region enclosed by the radial caustic.
+In order to recognise a strong lens in a real strong lensing survey, however, it is not sufficient for multiple images to exist: they must be detected.
+For this reason, given the detection limit of a survey for a point source, $m_{\mathrm{lim}}$, we define as strong lensing event any lens-source system with at least two images brighter than $m_{\mathrm{lim}}$.
+This is the same definition upon which the \citet{MVK09} work is based \citep{VMK09}.
+Labelling with $m_2$ the magnitude of the second-brightest image, then, in the absence of photometric noise the lens detection probability $\pdet$ becomes
+\begin{equation}
+\pdet(\psilens,\psisource) = \left\{\begin{array}{ll} 1 & \rm{if}\,m_2(\psilens,\psisource) < m_{\mathrm{lim}} \\
+0 & \rm{otherwise}\end{array}\right. .
+\end{equation}
+
+\subsection{Lensing event definition: extended sources}\label{ssec:lensdefext}
 
@@
 
-\subsection{Axisymmetric lenses, point sources}\label{sect:axsymmpoint}
+\subsection{Axisymmetric lenses, point sources}\label{ssec:axsymmpoint}
+
+Axisymmetric lenses with a density profile of the kind introduced in section \ref{ssec:profile} can produce either one or three images of a point source.
+This can be seen in \Fref{fig:scheme}, which shows the lens equation for various values of the dark matter fraction and the inner dark matter slope.
+All of the lenses shown in this figure have the same Einstein radius, which is equal in size to the half-mass radius of the stellar component.
+The radius of the radial caustic, marked by the dotted lines in \Fref{fig:scheme}, is a strong function of the lens properties: it is largest in lenses with a smaller dark matter fraction or steeper dark matter slopes.
+As a result, the source plane area that is subject to strong lensing is an even stronger function of these properties, since it scales with the square of the radial caustic radius.
+In order to compute the lensing cross-section, however, we must take into account the magnification of the multiple images: according to our definition of a strong lensing event, given in section \ref{ssec:lensdefpoint}, at least two images must be detected in order for a system to qualify as a strong lens.
+
+\Fref{fig:1dmag} shows the magnification of the secondary image as a function of source position.
+The secondary image is the one located in the region between the radial and tangential critical curves, opposite to the source with respect to the lens centre.
+In most practical cases this is the second brightest image.
+As \Fref{fig:1dmag} shows, the magnification is very large for sources close to the lens (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
+%In particular, for $\beta > 0.5\tein$, the magnification is smaller than unity, with the exception of a thin slice near the caustic.
+While for the model with $f_{\mathrm{DM}}=0.8$ (purple curve) the magnification is above unity everywhere, other lens models can produce highly de-magnified secondary images for large values of $\beta$.
+Depending on the intrinsic brightness of the lensed source, these images may or may not be detected.
+%This has two implications for the lensing cross-section.
+First, a larger caustic size does not necessarily translate into a larger value of $\crosssect$.
+Second, the cross-section depends on the intrinsic brightness of the lensed source: an extremely bright point source can be detected regardless of its position in the source plane, but a source with a finite brightness can go undetected depending on its value of $\beta$.
+
+%
+\begin{figure}
+\includegraphics[width=\columnwidth]{composite_fixedap_mu2.eps}
+\caption{
+Magnification of the secondary image as a function of the source position.
+The lenses are axisymmetric composite model. Their density profile is plotted in \Fref{fig:kappa}.
+Vertical dotted lines mark the position of the radial caustic.
+\label{fig:1dmag}
+}
+\end{figure}
+
+Using the definition of \Eref{eq:crosssect}, we computed the lensing cross-section of a set of axisymmetric lenses, with respect to point sources with different brightnesses.
+In particular, we considered model lenses with fixed Einstein radius, with angular half-mass radius fixed to $\tein$, and varying values of the dark matter fraction and dark matter inner slope.
+The results are shown in \Fref{fig:fixedap_cs}.
+Each line corresponds to a source with a given intrinsic (i.e. unlensed) magnitude, $m_s$. The difference between this magnitude and the limiting magnitude of the survey is indicated as follows:
+\begin{equation}
+\Delta m \equiv m_s - m_{\mathrm{lim}}.
+\end{equation}
+
+We can see a clear trend between $\Delta m$ and the lensing cross-section, in both panels of \Fref{fig:fixedap_cs}: $\crosssect$ is larger for brighter sources.
+The trend saturates below a certain $\Delta m$, for sufficiently small values of $\gammadm$ or for sufficiently large values of $\fdm$.
+%These are cases in which the source is so bright that, as long as it is located within the radial caustic, it always produces at least two detectable images. 
+In these cases the lensing cross-section coincides with the full area enclosed within the caustic, and further increasing the source brightness does not result in an increased value of $\crosssect$.
 
 \begin{figure}
-\includegraphics[width=\columnwidth]{axisymm_pl_crosssect.eps}
+\includegraphics[width=\columnwidth]{axisymm_composite_crosssect.eps}
 \caption{
-Strong lensing cross-section of an axisymmetric power-law lens and a point source, as a function of the power-law slope $\gamma$.
+Strong lensing cross-section of an axisymmetric lens and a point source.
+The lens galaxy is a composite model, introduced in section \ref{ssec:profile}, with angular half-light radius equal to the Einstein radius.
+Left panel: cross-section, in units of $\pi\tein^2$, as a function of the inner dark matter slope. The dark matter fraction is fixed at $\fdm=0.5$.
+Right panel: cross-section as a function of the fraction of dark matter within the half-light radius. The inner dark matter slope is fixed at $\gammadm=1.5$.
 The system is defined as a strong lens if at least two images are detected.
 Different lines correspond to the difference $\Delta m$ between the source magnitude and the survey detection limit for a point source.
-The dashed line marks the cross-section in the case in which at least two images are formed, regardless of their magnification.
+\label{fig:fixedap_cs}
 }
 \end{figure}
+
+At fixed source brightness, trends with the dark matter inner slope or dark matter fraction are generally weak.
+\Fref{fig:fixedap_cs}, however, shows how the lensing cross-section varies with $\gammadm$ or $\fdm$ while keeping the Einstein radius fixed.
+This is achieved by adjusting other properties of the lens, such as the total mass of the baryonic or the dark matter component (see \Fref{fig:kappa}). 
+%This means that other properties of the lens, such as the total mass of the baryonic component or the dark matter component, are adjusted to maintain a constant $\tein$ (see \Fref{fig:kappa}). 
+In practice, when varying one ingredient of the lens density profile, the Einstein radius varies in response.
+To get a complete view of how the lensing cross-section depends on different lens properties, we also computed how $\crosssect$ responds in absolute terms by varying one lens parameter at a time.
+\Fref{fig:phys_cs} shows $\crosssect$ as a function of stellar mass, half-light radius, inner dark matter slope, and projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$.
+%The reference lens is a galaxy at $z=0.3$ with $\log{\mstar}=11.5$, $\reff=7\,{\rm kpc}$, $\gammadm=1.5$, $\log{\mfive}=11.0$, and a source at $z=1.5$.
+
+\begin{figure*}
+\includegraphics[width=\textwidth]{axisymm_composite_physical_crosssect.eps}
+\caption{
+Absolute value of the strong lensing cross-section as a function of various lens properties.
+The reference lens is a galaxy at $z=0.3$ with $\log{\mstar}=11.5$, $\reff=7\,{\rm kpc}$, $\gammadm=1.5$, $\log{\mdmfive}=11.0$, and a source at $z=1.5$
+In each panel, only one property of the lens is varied, as indicated on the label of the horizontal axis.
+Each curve corresponds to a different value of $\Delta m$, in accordance with \Fref{fig:fixedap_cs}.
+The dashed line in each panel shows the quantity $\pi\tein^2$.
+\label{fig:phys_cs}
+}
+\end{figure*}
+
+The lensing cross-section increases with increasing stellar mass and dark matter mass, decreases with increasing $\reff$ for bright sources, while is only a weak function of $\gammadm$. 
+The lack of a clear trend between $\crosssect$ and $\gammadm$ appears to be in contradiction with the result of \citet{MVK09}, who found a strong positive correlation between $\crosssect$ and $\gammadm$.
+The origin of this discrepancy lies in the different ways in which $\gammadm$ is varied in the two experiments.
+While we varied $\gammadm$ at fixed $\mdmfive$, \citet{MVK09} kept fixed the virial mass of the dark matter halo.
+At fixed virial mass, increasing the inner dark matter slope results in a correspondingly larger dark matter mass in the inner regions, which naturally results in a larger lensing cross-section.
 

### 2022-07-12 single_lens/axisymmetric_point/plot_composite_crosssect.py

@@ -9,21 +9,22 @@ from matplotlib import rc
 rc('text', usetex=True)
 
 
-fsize = 18
+fsize = 22
+lsize = 18
 
-ylim = (0.03, 7.)
+ylim = (0.01, 3.)
 
 logscale = True
 
 if logscale:
-    leftm = 0.11
+    leftm = 0.14
 else:
     leftm = 0.08
 
 cs_file = h5py.File('composite_rein1.0reff_grid.hdf5', 'r')
 
 dms_grid = cs_file['dms_grid'][()]
-ndms = len(dms_grid)
+ndms = len(dms_grid)-1
 
 gammadm_grid = cs_file['gammadm_grid'][()]
 ngamma = len(gammadm_grid)
@@ -35,15 +36,20 @@ cs_gammadm = cs_file['cs_vs_gammadm'][()]
 cs_fdm = cs_file['cs_vs_fdm'][()]
 
 fig, ax = pylab.subplots(1, 2, figsize=(8, 4))
-pylab.subplots_adjust(left=leftm, right=1.00, bottom=0.14, top=0.99, wspace=0.)
+pylab.subplots_adjust(left=leftm, right=1.00, bottom=0.16, top=0.99, wspace=0.)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
 for i in range(ndms):
-    ax[0].plot(gammadm_grid, cs_gammadm[:, i], label='$\Delta m_s=%2.1f$'%dms_grid[i], linewidth=2, color=colseq[i])
-    ax[1].plot(fdm_grid, cs_fdm[:, i], label='$\Delta m_s=%2.1f$'%dms_grid[i], linewidth=2, color=colseq[i])
+    if i>1:
+        ax[0].plot(gammadm_grid, cs_gammadm[:, i+1]/np.pi, label='$\Delta m_s=%2.1f$'%dms_grid[i+1], linewidth=2, color=colseq[i])
+        ax[1].plot(fdm_grid, cs_fdm[:, i+1]/np.pi, label='$\Delta m_s=%2.1f$'%dms_grid[i+1], linewidth=2, color=colseq[i])
+    else:
+        ax[0].plot(gammadm_grid, cs_gammadm[:, i+1]/np.pi, label='$%2.1f$'%dms_grid[i+1], linewidth=2, color=colseq[i])
[…]

### 2022-07-12 single_lens/axisymmetric_point/plot_composite_fixedap_mu2.py

@@ -1,0 +1,142 @@
+import numpy as np
+from wl_profiles import gnfw, deVaucouleurs as deV
+from wl_cosmology import Mpc, c, G, M_Sun
+import wl_cosmology
+from scipy.interpolate import splrep, splev, splint
+from scipy.optimize import brentq
+import h5py
+import sys
+import pylab
+from labellines import labelLine, labelLines
+from matplotlib.ticker import MultipleLocator
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 18
+
+reff = 1.
+rein = reff
+rs_fixed = 10. * reff
+
+s_cr = 1. # I shouldn't need to change this
+
+gammadm_plot = [1., 1.5, 2.]
+ngammadm = len(gammadm_plot)
+fdm_fixed = 0.5
+
+fdm_plot = [0.2, 0.8]
+gammadm_fixed = 1.5
+nfdm = len(fdm_plot)
+
+# defines lensing-related functions
+def alpha_dm(x, gnfw_norm, rs, gammadm):
+    # deflection angle (in kpc)
+    return gnfw_norm * gnfw.fast_M2d(abs(x), rs, gammadm) / np.pi/x/s_cr
+
+def alpha_star(x, mstar, reff): 
+    # deflection angle (in kpc)
+    return mstar * deV.M2d(abs(x), reff) / np.pi/x/s_cr
+
+def alpha(x, gnfw_norm, rs, gammadm, mstar, reff):
+    return alpha_dm(x, gnfw_norm, rs, gammadm) + alpha_star(x, mstar, reff)
+
+def kappa(x, gnfw_norm, rs, gammadm, mstar, reff): 
+    # dimensionless surface mass density
+    return (mstar * deV.Sigma(abs(x), reff) + gnfw_norm * gnfw.fast_Sigma(abs(x), rs, gammadm))/s_cr
+   
+def mu_r(x, gnfw_norm, rs, gammadm, mstar, reff):
+    # radial magnification
+    return (1. + alpha(x, gnfw_norm, rs, gammadm, mstar, reff)/x - 2.*kappa(x, gnfw_norm, rs, gammadm, mstar, reff))**(-1)
+
+def mu_t(x, gnfw_norm, rs, gammadm, mstar, reff):
+    # tangential magnification
[…]

### 2022-07-12 single_lens/axisymmetric_point/plot_composite_physical_crosssect.py

@@ -22,7 +22,7 @@ else:
 
 cs_file = h5py.File('composite_physical_grid.hdf5', 'r')
 
-dms_grid = cs_file['dms_grid'][()]
+dms_grid = cs_file['dms_grid'][1:]
 ndms = len(dms_grid)
 
 #lmstar_grid = cs_file['lmstar_grid'][()]
@@ -36,10 +36,10 @@ ngamma = len(gammadm_grid)
 lmdm5_grid = cs_file['lmdm5_grid'][()]
 nmdm5 = len(lmdm5_grid)
 
-cs_gammadm = cs_file['cs_vs_gammadm'][()]
-cs_lmdm5 = cs_file['cs_vs_lmdm5'][()]
-cs_lmstar = cs_file['cs_vs_lmstar'][()]
-cs_lreff = cs_file['cs_vs_lreff'][()]
+cs_gammadm = cs_file['cs_vs_gammadm'][:, 1:]
+cs_lmdm5 = cs_file['cs_vs_lmdm5'][:, 1:]
+cs_lmstar = cs_file['cs_vs_lmstar'][:, 1:]
+cs_lreff = cs_file['cs_vs_lreff'][:, 1:]
 
 rein_gammadm = cs_file['rein_vs_gammadm'][()]
 rein_lmdm5 = cs_file['rein_vs_lmdm5'][()]

### 2022-07-13 single_lens/elliptical_ext/make_footprint_images.py
started extended source

@@ -1,0 +1,24 @@
+import numpy as np
+import pylab
+from astropy.io import fits as pyfits
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+suffix = ['x0.0_y0.0', 'x0.4_y0.0', 'x0.122_y0.105', 'x0.6_y0.0']
+
+level = 1.8
+
+leftm = 0.15
+fig, ax = pylab.subplots(1, 4, figsize=(12, 4))
+pylab.subplots_adjust(left=leftm, right=1.00, bottom=0.15, top=0.99, wspace=0.)
+
+for i in range(4):
+    img = pyfits.open('composite_%s_image.fits'%suffix[i])[0].data
+
+    ax[i].contourf(img,[level,1e8], colors='b')
+    ax[i].set_aspect(1.)
+
+pylab.show()
+
+

### 2022-07-13 single_lens/elliptical_ext/plot_caustic.py

@@ -1,0 +1,63 @@
+import numpy as np
+import pylab
+from sl_profiles import deVaucouleurs as deV, gnfw
+from matplotlib import rc
+from matplotlib.ticker import MultipleLocator
+from scipy.optimize import brentq
+rc('text', usetex=True)
+
+
+fsize = 18
+
+fig, ax = pylab.subplots(figsize=(5.5, 5.5))
+
+pylab.subplots_adjust(left=0.17, right=1., bottom=0.17, top=1.)
+
+xlim = (-0.95, 0.95)
+ylim = (-0.95, 0.95)
+
+"""
+xgrid = np.arange(xlim[0], xlim[1], 0.05)
+ygrid = np.arange(ylim[0], ylim[1], 0.05)
+
+for i in range(len(xgrid)):
+    ax.axvline(xgrid[i], color='grey', linewidth=0.5)
+
+for i in range(len(ygrid)):
+    ax.axhline(ygrid[i], color='grey', linewidth=0.5)
+"""
+
+f = open('composite_crit.dat', 'r')
+table = np.loadtxt(f)
+f.close()
+
+xs1 = table[:, 2]
+ys1 = table[:, 3]
+xs2 = table[:, 6]
+ys2 = table[:, 7]
+
+nseg = len(xs1)
+
+for j in range(nseg):
+    pylab.plot([xs1[j], xs2[j]], [ys1[j], ys2[j]], color='b')
+
+#ax.set_aspect(1.)
+
+ax.tick_params(axis='both', which='both', top=True, right=True, labelsize=fsize, direction='in')#, width=1)
+ax.set_xlabel('$\\beta_x/\\theta_{\mathrm{Ein}}$', fontsize=fsize)
+ax.set_ylabel('$\\beta_y/\\theta_{\mathrm{Ein}}$', fontsize=fsize)
+
+ax.xaxis.set_major_locator(MultipleLocator(0.2))
+ax.xaxis.set_minor_locator(MultipleLocator(0.05))
+
+ax.yaxis.set_major_locator(MultipleLocator(0.2))
+ax.yaxis.set_minor_locator(MultipleLocator(0.05))
+
+ax.set_xlim(xlim[0], xlim[1])
+ax.set_ylim(ylim[0], ylim[1])
+
+pylab.arrow(-0.75, 0.78, 0.2, 0, length_includes_head=False, head_width=0.05, color='k')
+pylab.text(-0.8, 0.82, 'Major axis', fontsize=fsize)
+
+pylab.show()
+

### 2022-07-13 paper/seleff.tex
started extended source

@@
 
+Defining a strong lensing event in the case of an extended source is less straightforward.
+In principle, we should require parts of the source to be multiply imaged.
+In practice, it is not always easy to determine whether a lensed image contains multiple images or not.
+For instance, a source located on one of the cusps of a tangential caustic produces a merging set of three magnified images.
+If the source is extended and no counter-image is detected, however, the appearence of the merging triplet is very similar to that of a single arc with no multiple images, and it can be difficult to differentiate between the two cases.
+In real strong lensing surveys, the decision of including systems with a single arc or not is often arbitrary.
+
+For the purpose of carrying out our experiment we adopt the following definition.
+First, we determine the footprint of the lensed source.
+We define this as the ensemble of pixels where the source is detected with $S/N>3$.
+The footprint is in general composed of multiple disconnected regions, corresponding to the different images.
+In order to only include images that can be clearly identified, we add the condition that the integrated $S/N$ of each separate region must be $S/N > 10$.
+This condition has the effect of removing from the source footprint any isolated region consisting only of a very small number of pixels.
+In a real-world application, it would be very hard to classify such marginal detections as images.
+%Given the source footprint thus defined, we classify the system as a strong lens if the footprint subtends an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
+Given the source footprint thus defined, we classify the system as a strong lens if we can find two points on the footprint that subtend an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
+\begin{equation}
+\theta_{\mathrm{min}} < \theta_{\mathrm{s}} \leq \pi,
+\end{equation}
+with $\theta_{\mathrm{min}}=\pi/2$.
+\Fref{fig:lensdef} shows a few examples of how this criterion can be used to classify lenses.
+
+If no counter-image is detected, however, it can be difficult to distinguish such
+
+For instance, it can be difficult to distinguish a single arc with no detected counter-image from a merging set of three images
+
+This is because it is not always
+
 %__________________________________________________________________
@@
 
-In this section we compute $\crosssect$ in a series of different scenarios of increasing complexity.
+We compute $\crosssect$ in a series of different scenarios of increasing complexity.
 %We first introduce the basic equations of gravitational lensing in section \ref{ssec:lensbasics}.
@@
 The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments.
-We describe this in \ref{ssec:profile}.
+We describe this in section \ref{ssec:profile}.
 %, among with the basic equations of gravitational lensing.

### 2022-07-13 paper/seleff.tex
new figure 3

@@
 
-If no counter-image is detected, however, it can be difficult to distinguish such
-
-For instance, it can be difficult to distinguish a single arc with no detected counter-image from a merging set of three images
-
-This is because it is not always
-
+\begin{figure*}
+\includegraphics[width=\textwidth]{footprints.eps}
+\caption{
+Criterion used to classify lensed images of extended sources.
+}
+\end{figure*}
 %__________________________________________________________________

### 2022-07-14 paper/seleff.tex
elliptical lenses

@@
 Labelling with $m_2$ the magnitude of the second-brightest image, then, in the absence of photometric noise the lens detection probability $\pdet$ becomes
-\begin{equation}
+\begin{equation}\label{eq:pointdet}
 \pdet(\psilens,\psisource) = \left\{\begin{array}{ll} 1 & \rm{if}\,m_2(\psilens,\psisource) < m_{\mathrm{lim}} \\
@@
 In practice, it is not always easy to determine whether a lensed image contains multiple images or not.
-For instance, a source located on one of the cusps of a tangential caustic produces a merging set of three magnified images.
+This is because, when the source is extended, some of the images can be blended together.
+In real strong lensing surveys, it is common to find lens candidates in which the lensed source consists only a single arc.
+In those cases it is difficult to establish whether the arc is a set of blending images or not, and the decision of including such systems in a strong lens sample is often arbitrary.
+%For instance, a source located on one of the cusps of a tangential caustic produces a merging set of three magnified images.
 If the source is extended and no counter-image is detected, however, the appearence of the merging triplet is very similar to that of a single arc with no multiple images, and it can be difficult to differentiate between the two cases.
-In real strong lensing surveys, the decision of including systems with a single arc or not is often arbitrary.
+%In real strong lensing surveys, the decision of including systems with a single arc or not is often arbitrary.
 
@@
 Given the source footprint thus defined, we classify the system as a strong lens if we can find two points on the footprint that subtend an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
-\begin{equation}
-\theta_{\mathrm{min}} < \theta_{\mathrm{s}} \leq \pi,
+\begin{equation}\label{eq:subtend}
+\theta_{\mathrm{min}} < \theta_{\mathrm{s}} \leq 180^\circ,
 \end{equation}
-with $\theta_{\mathrm{min}}=\pi/2$.
+with $\theta_{\mathrm{min}}=90^\circ$.
+A similar criterion was used in the automatic lens search campaign of \citet{Son++18a}.
+
 \Fref{fig:lensdef} shows a few examples of how this criterion can be used to classify lenses.
+As the first three examples in \Fref{fig:lensdef} show, the presence of a counter-image to the main arc always guarantees that the system is classified as a lens.
+Only if a single arc is visible, and if this arc subtends an angle smaller than $90^\circ$, a system is not considered a strong lens (fourth example).
+The choice of points used to define the angle $\theta_{\mathrm{s}}$ is not unique: any pair of points for which \Eref{eq:subtend} is satisfied is acceptable.
 
@@
 \caption{
-Criterion used to classify lensed images of extended sources.
+Criterion used to classify lensed images of extended sources. Four examples. In each panel, the blue regions are the footprint of the lensed source. 
+The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
 }
@@
 Absolute value of the strong lensing cross-section as a function of various lens properties.
-The reference lens is a galaxy at $z=0.3$ with $\log{\mstar}=11.5$, $\reff=7\,{\rm kpc}$, $\gammadm=1.5$, $\log{\mdmfive}=11.0$, and a source at $z=1.5$
+The reference lens is a galaxy at $z=0.3$ with $\log{\mstar}=11.5$, $\reff=7\,{\rm kpc}$, $\gammadm=1.5$, $\log{\mdmfive}=11.0$, $r_s=100$~kpc, and a source at $z=1.5$
 In each panel, only one property of the lens is varied, as indicated on the label of the horizontal axis.
@@
 
+\Fref{fig:fixedap_cs} and \Fref{fig:phys_cs} show that the trends between lens properties and the strong lensing cross-section can be different for sources with different brightnesses.
+The net effect in a strong lensing survey is an average over the source population, weighted by the source luminosity function. 
+This implies that surveys that target different families of sources, with different luminosity functions (for instance, galaxies or quasars), can have different strong lensing biases.
+We explore this possibility in \Sref{sect:lenspop}.
+
 \subsection{Elliptical lenses, point sources}\label{ssec:ellpoint}
+
+We measured $\crosssect$ for lenses with a fixed radial density profile and different ellipticities, with respect to point sources of diferent brightnesses.
+%In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.5$, $\reff=7$~kpc, $\log{\mdmfive}=11.0$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
+In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
+This corresponds to a lens with Einstein radius equal to the half-light radius, both equal to $1''$.
+We set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
+This is the lens model used to produce the caustics plot of \Fref{Fig:caustics}.
+That figure suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
+Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.
+
+We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used the software {\sc Glafic} \citep{Ogu21} to solve the lens equation, find the number of images and their magnification.
+We then multiplied the area over which sources are located by the fraction of those that are strongly lensed according to the criterion of \Eref{eq:pointdet}.
+%\Fref{fig:ellcaust} suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
+%Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.
+The resulting $\crosssect$ is shown in \Fref{fig:ellpoint_cs} as a function of the minor-to-major axis ratio $q$.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{ell_pnt_cs.eps}
+\caption{
+Point source strong lensing cross-section as a function of lens axis ratio.Solid lines: cross-section based on the lens event definition of \Eref{eq:pointdet}.
+Dashed lines: cross-section based on the detection of four images (quad cross-section).
+Lines of different colour correspond to sources of different intrinsic magnitude.
+The lens consists of a galaxy at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
+The baryonic and dark matter components have the same ellipticity and direction of the major axis.
+}
+\end{figure}
+
+For bright sources the strong lensing cross-section is approximately constant with axis ratio.
+This is because, as pointed out earlier, in the bright source regime the cross-section is determined by the area enclosed within the radial caustic, which does not vary much with lens ellipticity.
+For faint sources we observe a larger variation with $q$, with a factor of two difference between the largest and smallest value of $\crosssect$ at fixed source brightness.
+
+Most of the sources that result in detectable lenses produce two detectable images. These are sources that are located in the region enclosed bewteen the radial and the tangential caustic. 
+%In most cases, only two of these three images are detected: the third one, close to the centre, is usually highly de-magnified.
+If the source is located within the tangential caustic, however, four detectable images are usually created.
+Lenses with four visible images, usually referred to as quad lenses, are sometimes given a high priority in certain lensing studies, because they offer more constraints compared to double lenses.
+For instance, quads make up the majority of the lenses used so far in time-delay studies \citep{Mil++20}.
+For this reason we also computed an alternative lensing cross-section, in which the definition of a lensing event requires the detection of four images, instead of two.
+This is plotted in \Fref{fig:ellpoint_cs} with dashed lines.
+The cross-section for quads is a strong function of lens ellipticity, for bright sources.
+This is a consequence of the fact that the area enclosed within the tangential caustic, which is where a source needs to be in order to produce four images, increases with increased lens ellipticity, as \Fref{fig:ellcaust} shows.
+For sources that are intrinsically fainter than the detection limit, however, the quad cross-section is extremely small.
+
+%\Fref{fig:caustics} shows that the area enclosed within the tangential caustic is a strong function of the lens ellipticity.
+%Therefore we expect that the
 

### 2022-07-14 single_lens/elliptical_ext/make_footprint_images.py

@@ -6,7 +6,7 @@ from matplotlib import rc
 rc('text', usetex=True)
 
 
-names = ['e0.2_re0.1_x0.0_y0.0', 'e0.4_re0.1_x0.122_y0.105', 'e0.4_re0.1_x0.4_y0.0', 'e0.4_re0.1_x0.7_y0.0']
+names = ['e0.0_re0.1_x0.0_y0.0', 'e0.4_re0.1_x0.122_y0.105', 'e0.4_re0.1_x0.4_y0.0', 'e0.4_re0.1_x0.7_y0.0']
 
 sky_rms = 0.6
 thresh = 3.*sky_rms
@@ -61,16 +61,18 @@ for i in range(4):
     ax[i].text(textpos[i][0], textpos[i][1], '$\\theta_{\mathrm{s}} = %s^\circ$'%text[i], fontsize=fsize)
 
     if success[i]:
-        phi = np.pi*np.linspace(-1., 1., 1001)
-        x = 3. * np.cos(phi) + 10.
-        y = 3. * np.sin(phi) + 70.
-        ax[i].plot(x, y, color=(0., 1., 0.), linewidth=3)
+        #phi = np.pi*np.linspace(-1., 1., 1001)
+        #x = 3. * np.cos(phi) + 10.
+        #y = 3. * np.sin(phi) + 70.
+        #ax[i].plot(x, y, color=(0., 1., 0.), linewidth=3)
+        ax[i].text(10., 70., 'Lens', fontsize=fsize)
     else:
-        x = np.linspace(-3., 3.) + 10.
-        y1 = np.linspace(-3., 3.) + 70.
-        y2 = np.flipud(y1)
-        ax[i].plot(x, y1, color='r', linewidth=3)
-        ax[i].plot(x, y2, color='r', linewidth=3)
+        #x = np.linspace(-3., 3.) + 10.
+        #y1 = np.linspace(-3., 3.) + 70.
+        #y2 = np.flipud(y1)
+        #ax[i].plot(x, y1, color='r', linewidth=3)
+        #ax[i].plot(x, y2, color='r', linewidth=3)
+        ax[i].text(10., 70., 'Not a lens', fontsize=fsize)
 
     ax[i].tick_params(axis='both', which='both', top=False, bottom=False, left=False, right=False, labelleft=False, labelbottom=False)
 

### 2022-07-14 single_lens/elliptical_point/plot_cs_vs_q.py

@@ -1,0 +1,119 @@
+import numpy as np
+import pylab
+import h5py
+from scipy.optimize import brentq
+from plotters import probcontour
+from labellines import labelLine, labelLines
+from matplotlib.ticker import MultipleLocator
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 18
+
+ylim = (0.01, 3.)
+
+logscale = True
+
+if logscale:
+    leftm = 0.14
+else:
+    leftm = 0.08
+
+nq = 6
+q_arr = np.linspace(0.5, 1., nq)
+e_arr = 1. - q_arr
+
+ndms = 5
+dms_arr = np.linspace(-2., 2., ndms)
+
+cs_arr = np.zeros((nq, ndms))
+cs_quad = np.zeros((nq, ndms))
+
+fig, ax = pylab.subplots(1, 1)#, figsize=(8, 4))
+pylab.subplots_adjust(left=leftm, right=0.98, bottom=0.14, top=0.97, wspace=0.)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+
+nmocks = 100000
+
+for i in range(nq):
+
+    nmulti = np.zeros(ndms)
+    nquads = np.zeros(ndms)
+
+    f = open('composite_e%2.1f_mock.dat'%e_arr[i], 'r')
+    lines = f.readlines()[1:]
+    f.close()
+    
+    nimg = 0
+
+    ndet_list = []
+
+    sline = True
+    for line in lines:
+        if sline:
+            nimg = int(line[0])
+            nhere = 0
+            ndet = np.zeros(ndms)
+            if nimg > 0:
+                sline = False
+            else:
+                sline = True
+        elif nimg > 1:
+            line = line.split()
+            mu = float(line[2])
+            for j in range(ndms):
+                if abs(mu) > 10.**(dms_arr[j]/2.5):
+                    ndet[j] += 1
+            nhere += 1
+            if nhere == nimg:
+                sline = True
+                ndet_list.append(ndet)
+        else:
+            sline = True
+
+    ndet_arr = np.array(ndet_list)
+
+    for j in range(ndms):
[…]

### 2022-07-14 single_lens/elliptical_point/set_composite_pars.py

@@ -82,6 +82,7 @@ mstar_glafic = mstar_phys * hubble
 print('GNFW component parameters. Mvir=%4.3e, rs=%4.3f arcsec'%(m200_glafic, rs_arcsec))
 print('Sersic component parameters. Mtot=%4.3e, reff=%4.3f arcsec'%(mstar_glafic, reff_arcsec))
 
+print('Parameters in physical units. logMstar=%4.3f, reff=%4.3f kpc'%(np.log10(mstar_phys), reff_phys))
 def kappa(x, gnfw_norm, rs, gammadm, mstar, reff, s_cr): 
     # dimensionless surface mass density
     return (mstar * deV.Sigma(abs(x), reff) + gnfw_norm * gnfw.fast_Sigma(abs(x), rs, gammadm))/s_cr

### 2022-07-14 single_lens/elliptical_point/plot_cs_vs_q.py
label fixes

@@ -92,7 +92,7 @@ for i in range(ndms):
 
 lines = labelLines(ax.get_lines()[2:], xvals = [0.65, 0.9, 0.58], fontsize=fsize, backgroundcolor='white')
 ax.text(0.73, 0.40, '$\Delta m_s=-1.0$', fontsize=fsize, color=colseq[1])
-ax.text(0.51, 0.67, '$\Delta m_s=-1.0$', fontsize=fsize, color=colseq[0])
+ax.text(0.51, 0.67, '$\Delta m_s=-2.0$', fontsize=fsize, color=colseq[0])
 
 for i in range(ndms):
     ax.plot(q_arr, cs_quad[:, i]/np.pi, label='$\Delta m_s=%2.1f$'%dms_arr[i], linewidth=2, color=colseq[i], linestyle='--')

### 2022-07-15 paper/seleff.tex
label fixes
simplified lens definition

@@
 In a real-world application, it would be very hard to classify such marginal detections as images.
+If, after applying this cut, the source footprint is spread over multiple separated regions, then the system is classified as a lens.
+Additionally, if the source footprint consists of only a single region, we still classify it as strongly lensed if it subtends an angle larger than $90^\circ$ measured from the lens centre.
+This last criterion allows us to include highly magnified arcs and full rings.
+
+%Additionally, we classify as strongly lensed the following set of sources any full rings, and any single arcs that subtend an angle larger than $90^\circ$ measured from the lens centre.
+This last criterion is a
+%Additionally, we classify as a strongly lensed source any full rings, and any single arcs that subtend an angle larger than $90^\circ$ measured from the lens centre. This last criterion is a
+
 %Given the source footprint thus defined, we classify the system as a strong lens if the footprint subtends an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
-Given the source footprint thus defined, we classify the system as a strong lens if we can find two points on the footprint that subtend an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
-\begin{equation}\label{eq:subtend}
-\theta_{\mathrm{min}} < \theta_{\mathrm{s}} \leq 180^\circ,
-\end{equation}
-with $\theta_{\mathrm{min}}=90^\circ$.
-A similar criterion was used in the automatic lens search campaign of \citet{Son++18a}.
+%Given the source footprint thus defined, we classify the system as a strong lens if we can find two points on the footprint that subtend an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
+%\begin{equation}\label{eq:subtend}
+%\theta_{\mathrm{min}} < \theta_{\mathrm{s}} \leq 180^\circ,
+%\end{equation}
+%with $\theta_{\mathrm{min}}=90^\circ$.
+%A similar criterion was used in the automatic lens search campaign of \citet{Son++18a}.
 
@@
 Only if a single arc is visible, and if this arc subtends an angle smaller than $90^\circ$, a system is not considered a strong lens (fourth example).
-The choice of points used to define the angle $\theta_{\mathrm{s}}$ is not unique: any pair of points for which \Eref{eq:subtend} is satisfied is acceptable.
+%The choice of points used to define the angle $\theta_{\mathrm{s}}$ is not unique: any pair of points for which \Eref{eq:subtend} is satisfied is acceptable.
 
@@
 We set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
-This is the lens model used to produce the caustics plot of \Fref{Fig:caustics}.
+This is the lens model used to produce the caustics plot of \Fref{fig:ellcaust}.
 That figure suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
@@
 The baryonic and dark matter components have the same ellipticity and direction of the major axis.
+\label{fig:ellpoint_cs}
 }
@@
 For sources that are intrinsically fainter than the detection limit, however, the quad cross-section is extremely small.
-
-%\Fref{fig:caustics} shows that the area enclosed within the tangential caustic is a strong function of the lens ellipticity.
-%Therefore we expect that the
 

### 2022-07-15 single_lens/elliptical_ext/make_footprint_images.py

@@ -49,16 +49,18 @@ for i in range(4):
     ax[i].set_aspect(1.)
     ax[i].scatter(x0, y0, color='k')
 
-    ax[i].plot([points[i][0][0], x0], [points[i][0][1], y0], color='k')
-    ax[i].plot([points[i][1][0], x0], [points[i][1][1], y0], color='k')
-
-    # draws semicircle
-    phi = np.linspace(phis[i][0], phis[i][1], 101)
-    circ_x = r * np.cos(phi) + x0
-    circ_y = r * np.sin(phi) + y0
-
-    ax[i].plot(circ_x, circ_y, color='k', linewidth=0.5)
-    ax[i].text(textpos[i][0], textpos[i][1], '$\\theta_{\mathrm{s}} = %s^\circ$'%text[i], fontsize=fsize)
+    if i==3:
+        ax[i].plot([points[i][0][0], x0], [points[i][0][1], y0], color='k')
+        ax[i].plot([points[i][1][0], x0], [points[i][1][1], y0], color='k')
+    
+        # draws semicircle
+        phi = np.linspace(phis[i][0], phis[i][1], 101)
+        circ_x = r * np.cos(phi) + x0
+        circ_y = r * np.sin(phi) + y0
+    
+        ax[i].plot(circ_x, circ_y, color='k', linewidth=0.5)
+        #ax[i].text(textpos[i][0], textpos[i][1], '$\\theta_{\mathrm{s}} = %s^\circ$'%text[i], fontsize=fsize)
+        ax[i].text(textpos[i][0], textpos[i][1], '$%s^\circ$'%text[i], fontsize=fsize)
 
     if success[i]:
         #phi = np.pi*np.linspace(-1., 1., 1001)

### 2022-07-15 paper/seleff.tex
extended source crosssect

@@
 \Fref{fig:lensdef} shows a few examples of how this criterion can be used to classify lenses.
-As the first three examples in \Fref{fig:lensdef} show, the presence of a counter-image to the main arc always guarantees that the system is classified as a lens.
-Only if a single arc is visible, and if this arc subtends an angle smaller than $90^\circ$, a system is not considered a strong lens (fourth example).
+As the first three examples in \Fref{fig:lensdef} show, the presence of multiple images or a full ring implies that the system is classified as a lens.
+Only if the source footprint subtends an angle smaller than $90^\circ$, a system is not considered a strong lens (fourth example).
 %The choice of points used to define the angle $\theta_{\mathrm{s}}$ is not unique: any pair of points for which \Eref{eq:subtend} is satisfied is acceptable.
@@
 
+In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the surface brightness distribution, including its radial profile, shape and orientation, and the point spread function (PSF).
+For the sake of reducing the dimensionality of the problem, we neglect the PSF and focus on circularly symmetric sources.
+We also fix the surface brightness profile to an exponential disk (i.e.
+a S\'{e}rsic model with $n=1$).
+We then took a lens model with the same parameters used in section \ref{ssec:ellpoint} and a minor-to-major axis ratio of $q=0.7$.
+We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
+We carried out experiments with sources with different values of the total magnitude and varying half-light radius.
+The results are shown in \Fref{fig:extcs}.
+
 %__________________________________________________________________

### 2022-07-15 single_lens/elliptical_ext/get_I0_from_ms_Re.py

@@ -1,0 +1,15 @@
+import numpy as np
+from sl_profiles import sersic
+from scipy.special import gamma as gfunc
+
+
+ftot = 20.
+re_arr = [0.1, 0.2, 0.5, 1.]
+pix = 0.05
+nser = 1.
+
+for re in re_arr:
+    I0 = ftot/(2.*np.pi*(re/pix)**2*nser/sersic.b(nser)**(2*nser)*gfunc(2.*nser))
+    print(I0)
+
+

### 2022-07-17 paper/seleff.tex
using 2 sigma above the background instead of 3
extended source results

@@
 First, we determine the footprint of the lensed source.
-We define this as the ensemble of pixels where the source is detected with $S/N>3$.
+We define this as the ensemble of pixels where the source is detected with $S/N>2$.
 The footprint is in general composed of multiple disconnected regions, corresponding to the different images.
@@
 
-In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the surface brightness distribution, including its radial profile, shape and orientation, and the point spread function (PSF).
+In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the surface brightness distribution including its radial profile, shape and orientation, and the point spread function (PSF).
 For the sake of reducing the dimensionality of the problem, we neglect the PSF and focus on circularly symmetric sources.
@@
 We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
-We carried out experiments with sources with different values of the total magnitude and varying half-light radius.
+We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_s$.
 The results are shown in \Fref{fig:extcs}.
+The total flux of the source, indicated in the legend, is expressed in terms of the sky background fluctuation within an area equal to the square of the Einstein radius.
+We assumed that the background noise is an uncorrelated Gaussian field, for this purpose.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{ell_ext_cs.eps}
+\caption{
+Strong lensing cross-section as a function of source half-light radius.
+The lens model is the same as in \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
+Each solid line corresponds to a different value of the total flux of the source. The flux $f$ is expressed in terms of the background noise fluctuation measured over an area equal to $\tein^2$.
+The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
+Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
+\label{fig:extcs}
+}
+\end{figure}
+
+As in the point source case, the lensing cross-section increasing with increasing total flux, at fixed source size.
+At fixed flux, $\crosssect$ stays approximately constant with increasing source size until a given value, then drops rapidly for larger sizes.
+From a qualitative point of view, this behaviour can be observed also in the absence of lensing: increasing the size of a galaxy while keeping its flux fixed lowers its average surface brightness.
+If the surface brightness drops below the sky fluctuation level, then it becomes very difficult to detect it.
+In order to determine whether there are lensing-specific features in the $\crosssect-\theta_s$ relation of \Fref{fig:extcs}, we computed, for each source flux, the maximum half-light radius for which it can be detected in the absence of lensing.
+We used the same criterion as that of section \ref{ssec:lensdefext} to define a detection: we defined the source footprint as the ensemble of pixels that are $2\sigma$ above the background and required the total signal-to-noise ratio within the footprint to be larger than ten.
+The resulting limiting sizes are shown as vertical lines in \Fref{fig:extcs}.
+For each given total flux, the non-lensing size limit is larger than the value of $\theta_s$ at which the lensing cross-section drops.
+This implies that strong lensing selects preferentially more compact sources, compared to the field.
 

### 2022-07-17 single_lens/elliptical_ext/get_nolensing_detection_limit.py

@@ -1,0 +1,67 @@
+import numpy as np
+from astropy.io import fits as pyfits
+from skimage import measure
+from scipy.interpolate import splrep, splev
+import h5py
+import pylab
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+
+# this scripts looks at simulations of exponential profiles with
+# different values of the half-light radius and assuming different
+# background noise levels. Then, using the same criterion used for
+# the detection of multiple images of a strong lens, it finds the
+# largest half-light radius for which the galaxy is detected.
+
+ftot = 200.
+nre = 11
+logre_grid = np.linspace(-1., 0., nre)
+
+nfrat = 4
+lfrat_grid = np.linspace(0.8, 2., nfrat)
+# frat is the ratio between the source intrinsic flux and the sky rms over an area equal to the square of the Einstein radius.
+# In my sims, the Einstein radius is 1 arcsec. Pixels have a 0.05" size.
+# Then, the sky rms over a 1 square arcsec is 20 times that over a single pixel
+# All simulations have a source intrinsic flux of 200. Then an frat=10 
+# corresponds to a sky rms per pixel of 1.
+
+nsigma = 2.
+
+X, Y = np.meshgrid(np.arange(80), np.arange(80))
+
+for l in range(nfrat):
+    lfrat = lfrat_grid[l]
+    sky_rms = 10.**(1. - lfrat) 
+
+    sn_grid = np.zeros(nre)
+
+    for m in range(nre):
+        logre = logre_grid[m]
+
+        preamble = 'ftot200_sourceonly_logre%2.1f'%logre
+
+        img = pyfits.open('mockdir/ftot200_sourceonly/'+preamble+'_source.fits')[0].data
+        footprint = img > nsigma * sky_rms
+        npix_here = footprint.sum()
+
+        signal = img[footprint].sum()
+        noise = npix_here**0.5 * sky_rms
+
+        print(l, m, npix_here, signal/noise)
+
+        sn_grid[m] = signal/noise
+
+    good = np.isfinite(sn_grid)
+    sn_spline = splrep(np.flipud(sn_grid[good]), np.flipud(logre_grid[good]))
+
+    logre_crit = splev(10., sn_spline)
+
+    pylab.plot(10.**logre_grid, sn_grid, color=colseq[l])
[…]

### 2022-07-17 single_lens/elliptical_ext/plot_cs_vs_re.py

@@ -1,6 +1,7 @@
 import numpy as np
 import pylab
 import h5py
+from astropy.io import fits as pyfits
 from scipy.interpolate import splrep, splev
 from plotters import probcontour
 from labellines import labelLine, labelLines
@@ -11,7 +12,8 @@ rc('text', usetex=True)
 
 fsize = 18
 
-ylim = (0.1, 1.5)
+#xlim = (0.1, 1.)
+ylim = (0.085, 1.3)
 
 logscale = True
 
@@ -36,7 +38,7 @@ pylab.subplots_adjust(left=leftm, right=0.98, bottom=0.14, top=0.97, wspace=0.)
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
 for i in range(nfrat):
-    ax.loglog(10.**logre_grid, cs_grid[i, :]/np.pi, linewidth=2, color=colseq[i])
+    ax.loglog(10.**logre_grid, cs_grid[i, :]/np.pi, linewidth=2, color=colseq[i], label="$\log{\\frac{f}{\sigma_{\mathrm{sky,\\theta_{\mathrm{Ein}}^{2}}}}} = %2.1f$"%lfrat_grid[i])
     cs_spline = splrep(10.**logre_grid, cs_grid[i, :]/np.pi, k=1)
 
 nfr2 = nfrat
@@ -68,10 +70,44 @@ ax.tick_params(axis='both', which='both', top=True, right=True, labelsize=fsize,
 ax.set_xlabel('$\\theta_{\mathrm{e,s}}/\\theta_{\mathrm{Ein}}$', fontsize=fsize)
 ax.set_ylabel('$\sigma_{\mathrm{SL}}/(\pi\\theta_{\mathrm{Ein}}^2)$', fontsize=fsize)
 
-ax.set_ylim(ylim[0], ylim[1])
-#ax.set_xlim(0.5, 1.)
+# now finds the detection limit for non-lensed galaxies
+nre = 11
+logre_grid = np.linspace(-1., 0., nre)
+nsigma = 2.
 
-#pylab.savefig('../../paper/ell_ext_cs.eps')
+for l in range(nfrat):
+    lfrat = lfrat_grid[l]
+    sky_rms = 10.**(1. - lfrat) 
+
+    sn_grid = np.zeros(nre)
+
+    for m in range(nre):
+        logre = logre_grid[m]
+
+        preamble = 'ftot200_sourceonly_logre%2.1f'%logre
+
[…]

### 2022-07-17 single_lens/elliptical_ext/plot_single_footprint.py

@@ -8,7 +8,7 @@ rc('text', usetex=True)
 
 
 filename = sys.argv[1]
-sky_rms = 1.
+sky_rms = float(sys.argv[2])
 thresh = 2.*sky_rms
 
 x0 = 39.5

### 2022-07-17 single_lens/elliptical_ext/search_for_sn10_example.py

@@ -1,0 +1,51 @@
+import numpy as np
+from astropy.io import fits as pyfits
+from skimage import measure
+import h5py
+
+
+ftot = 200.
+
+nsim = 10000
+
+logre = -0.6
+lfrat = 1.
+sky_rms = 10.**(1. - lfrat) 
+nsigma = 2.
+
+# frat is the ratio between the source intrinsic flux and the sky rms over an area equal to the square of the Einstein radius.
+# In my sims, the Einstein radius is 1 arcsec. Pixels have a 0.05" size.
+# Then, the sky rms over a 1 square arcsec is 20 times that over a single pixel
+# All simulations have a source intrinsic flux of 200. Then an frat=10 
+# corresponds to a sky rms per pixel of 1.
+
+x0 = 39.5
+y0 = 39.5
+
+X, Y = np.meshgrid(np.arange(80), np.arange(80))
+
+preamble = 'ftot200_logre%2.1f'%logre
+
+for i in range(nsim):
+    img = pyfits.open('mockdir/'+preamble+'/'+preamble+'_%04d_image.fits'%i)[0].data
+    footprint = img > nsigma * sky_rms
+
+    labels = measure.label(footprint)
+    nreg = labels.max()
+    nimg = 0
+    sn_close_to_10 = False
+    for n in range(nreg):
+        npix_here = (labels==n+1).sum()
+        signal = img[labels==n+1].sum()
+        noise = npix_here**0.5 * sky_rms
+        img_sn = signal/noise
+        if img_sn < 10.:
+            img[labels==n+1] = 0.
+        else:
+            nimg += 1
+            if img_sn < 11.:
+                sn_close_to_10 = True
+
+    if nimg == 2 and sn_close_to_10:
+        print(i)
+

### 2022-07-17 single_lens/elliptical_ext/search_for_sn20_example.py

@@ -1,0 +1,51 @@
+import numpy as np
+from astropy.io import fits as pyfits
+from skimage import measure
+import h5py
+
+
+ftot = 200.
+
+nsim = 10000
+
+logre = -0.8
+lfrat = 1.
+sky_rms = 10.**(1. - lfrat) 
+nsigma = 2.
+
+# frat is the ratio between the source intrinsic flux and the sky rms over an area equal to the square of the Einstein radius.
+# In my sims, the Einstein radius is 1 arcsec. Pixels have a 0.05" size.
+# Then, the sky rms over a 1 square arcsec is 20 times that over a single pixel
+# All simulations have a source intrinsic flux of 200. Then an frat=10 
+# corresponds to a sky rms per pixel of 1.
+
+x0 = 39.5
+y0 = 39.5
+
+X, Y = np.meshgrid(np.arange(80), np.arange(80))
+
+preamble = 'ftot200_logre%2.1f'%logre
+
+for i in range(nsim):
+    img = pyfits.open('mockdir/'+preamble+'/'+preamble+'_%04d_image.fits'%i)[0].data
+    footprint = img > nsigma * sky_rms
+
+    labels = measure.label(footprint)
+    nreg = labels.max()
+    nimg = 0
+    sn_close_to_20 = False
+    for n in range(nreg):
+        npix_here = (labels==n+1).sum()
+        signal = img[labels==n+1].sum()
+        noise = npix_here**0.5 * sky_rms
+        img_sn = signal/noise
+        if img_sn < 20.:
+            img[labels==n+1] = 0.
+        else:
+            nimg += 1
+            if img_sn < 21.:
+                sn_close_to_20 = True
+
+    if nimg == 2 and sn_close_to_20:
+        print(i)
+

### 2022-07-17 single_lens/elliptical_ext/write_sourceonly_sim_input.py

@@ -1,0 +1,57 @@
+import numpy as np
+from sl_profiles import sersic
+from scipy.special import gamma as gfunc
+import os
+
+
+np.random.seed(0)
+
+# prepares a Glafic input file
+
+nre = 11
+logre_grid = np.linspace(-1., 0., nre)
+
+ftot = 200.
+pix = 0.05
+nser = 1.
+
+f = open('preamble.input', 'r')
+prelines = f.readlines()
+f.close()
+
+nsource = 10000
+    
+rmax = 1.
+
+   
+prefix = 'ftot%d_sourceonly'%(ftot)
+
+lines = prelines.copy()
+    
+lines.append('\n')
+lines.append('start_command\n')
+lines.append('\n')
+
+dirname = 'mockdir/%s/'%prefix
+    
+if not os.path.isdir(dirname):
+    os.system('mkdir %s'%dirname)
+ 
+for i in range(nre):
+
+    re = 10.**logre_grid[i]
+     
+    I0 = ftot/(2.*np.pi*(re/pix)**2*nser/sersic.b(nser)**(2*nser)*gfunc(2.*nser))
+    
+    lines.append('reset_par prefix %s\n'%(prefix + '_logre%2.1f'%logre_grid[i]))
+    lines.append('reset_extend 1 2 %f\n'%I0)
+    lines.append('reset_extend 1 7 %f\n'%10.**logre_grid[i])
+    lines.append('\n')
+    
+    lines.append('writeimage_ori\n')
+    lines.append('\n')
+    
+f = open(dirname+'/%s.input'%prefix, 'w')
+f.writelines(lines)
+f.close()
+

### 2022-07-19 paper/seleff.tex
finished section 3

@@
 \def\tein{\theta_{\mathrm{Ein}}}
+\def\teff{\theta_{\mathrm{e}}}
 \def\trad{\theta_{\mathrm{rad}}}
@@
 \def\crosssect{\sigma_\mathrm{{SL}}}
+\def\crosssectps{\sigma_\mathrm{{SL,PS}}}
 
@@
 \end{equation}
+The images of the background source are in general magnified in total flux and in size, while preserving the original surface brightness.
 
@@
 The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
+\label{fig:lensdef}
 }
@@
 \end{equation}
-where $\boldsymbol\beta$ is the position of the background source and the integral is carried out over the whole sky.
+where $\boldsymbol\beta$ is the position of the background source, $\psisourcenobeta$ is the ensemble of source parameter except for the position, and the integral is carried out over the whole sky.
 The definition above is valid for both a point source and an extended source: although there is no unique way of defining the position of an extended source, the integral over the sky ensures that the result is independent of how the source position is defined.
@@
 \end{equation}
+Throughout this paper we indicate with $\teff$ the angular size of the half-light radius.
 
@@
 
-\Fref{fig:kappa} shows the dimensionless surface mass density profile S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
+\Fref{fig:kappa} shows the dimensionless surface mass density profile of S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
 All of the profiles have a S\'{e}rsic index $n=4$, a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component.
@@
 Using the definition of \Eref{eq:crosssect}, we computed the lensing cross-section of a set of axisymmetric lenses, with respect to point sources with different brightnesses.
-In particular, we considered model lenses with fixed Einstein radius, with angular half-mass radius fixed to $\tein$, and varying values of the dark matter fraction and dark matter inner slope.
+In particular, we considered model lenses with fixed Einstein radius, with angular half-mass radius $\teff$ fixed to $\tein$, and varying values of the dark matter fraction and dark matter inner slope.
 The results are shown in \Fref{fig:fixedap_cs}.
@@
 At fixed source brightness, trends with the dark matter inner slope or dark matter fraction are generally weak.
-\Fref{fig:fixedap_cs}, however, shows how the lensing cross-section varies with $\gammadm$ or $\fdm$ while keeping the Einstein radius fixed.
+The lines of \Fref{fig:fixedap_cs}, however, have been computed by keeping the Einstein radius fixed while varying $\gammadm$ or $\fdm$.
 This is achieved by adjusting other properties of the lens, such as the total mass of the baryonic or the dark matter component (see \Fref{fig:kappa}). 
@@
 %In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.5$, $\reff=7$~kpc, $\log{\mdmfive}=11.0$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
-In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
+%In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
 This corresponds to a lens with Einstein radius equal to the half-light radius, both equal to $1''$.
-We set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
+In particular, we set $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\tein=\teff$, and set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
 This is the lens model used to produce the caustics plot of \Fref{fig:ellcaust}.
@@
 \caption{
-Point source strong lensing cross-section as a function of lens axis ratio.Solid lines: cross-section based on the lens event definition of \Eref{eq:pointdet}.
+Point source strong lensing cross-section as a function of lens axis ratio.
+Solid lines: cross-section based on the lens event definition of \Eref{eq:pointdet}.
 Dashed lines: cross-section based on the detection of four images (quad cross-section).
 Lines of different colour correspond to sources of different intrinsic magnitude.
-The lens consists of a galaxy at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
+%The lens consists of a galaxy at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
+The parameters of the lens density profile are $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\teff=\tein$.
 The baryonic and dark matter components have the same ellipticity and direction of the major axis.
@@
 
-In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the surface brightness distribution including its radial profile, shape and orientation, and the point spread function (PSF).
-For the sake of reducing the dimensionality of the problem, we neglect the PSF and focus on circularly symmetric sources.
-We also fix the surface brightness profile to an exponential disk (i.e.
+In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the surface brightness distribution with its radial profile, shape and orientation, and the point spread function (PSF).
+For the sake of reducing the dimensionality of the problem, we neglected the PSF and focus on circularly symmetric sources.
+We also fixed the surface brightness profile to an exponential disk (i.e.
 a S\'{e}rsic model with $n=1$).
@@
 We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
-We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_s$.
+We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_{\mathrm{e,s}}$.
 The results are shown in \Fref{fig:extcs}.
-The total flux of the source, indicated in the legend, is expressed in terms of the sky background fluctuation within an area equal to the square of the Einstein radius.
+The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius.
 We assumed that the background noise is an uncorrelated Gaussian field, for this purpose.
@@
 The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
-Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
+%Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
 \label{fig:extcs}
@@
 
-As in the point source case, the lensing cross-section increasing with increasing total flux, at fixed source size.
+As in the point source case, the lensing cross-section increases with increasing total flux, at fixed source size.
 At fixed flux, $\crosssect$ stays approximately constant with increasing source size until a given value, then drops rapidly for larger sizes.
@@
 If the surface brightness drops below the sky fluctuation level, then it becomes very difficult to detect it.
-In order to determine whether there are lensing-specific features in the $\crosssect-\theta_s$ relation of \Fref{fig:extcs}, we computed, for each source flux, the maximum half-light radius for which it can be detected in the absence of lensing.
+In order to determine whether there are lensing-specific features in the $\crosssect-\theta_{\mathrm{e,s}}$ relation of \Fref{fig:extcs}, we computed, for each source flux, the maximum half-light radius for which it can be detected in the absence of lensing.
 We used the same criterion as that of section \ref{ssec:lensdefext} to define a detection: we defined the source footprint as the ensemble of pixels that are $2\sigma$ above the background and required the total signal-to-noise ratio within the footprint to be larger than ten.
 The resulting limiting sizes are shown as vertical lines in \Fref{fig:extcs}.
-For each given total flux, the non-lensing size limit is larger than the value of $\theta_s$ at which the lensing cross-section drops.
-This implies that strong lensing selects preferentially more compact sources, compared to the field.
+For each given total flux, the non-lensing size limit is similar to the value of $\theta_{\mathrm{e,s}}$ at which the lensing cross-section drops.
+
+This result suggests that, to first approximation, lensing does not introduce a strong selection in source size.
+This is expected, since gravitational lensing preserves surface brightness: a source that can be detected in the absence of lensing will produce images with the same surface brightness when lensed, which can be detected as well.
+%When the source is extended, however, 
+In order to classify a source as strongly lensed, however, we require that multiple images are observed. 
+Only sources that lie within a well-defined region give rise to a strong lensing configuration.
+If part of the source extends outside of this region, then only a fraction of its flux contributes to creating a set of strongly lensed images.
+This lowers the signal-to-noise ratio of the multiple images compared to the point source case.
+%The result is the decrease in $\crosssect$ at values of the source size smaller than the 
+The result is that $\crosssect$ starts to decrease with increasing source size at values of $\theta_{\mathrm{e,s}}$ that are smaller than the no-lensing detection limit, as observed in \Fref{fig:extcs}.
+
+%At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), $\crosssect$ increases with increasing source size, before dropping to zero.
+At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (red ellipse in \Fref{fig:cs}).
+This is because, when the source is very bright, it can give rise to multiple images even while its centroid lies outside of the radial caustic, as long as its surface brightness distribution extends into it.
+This also explains why $\crosssect$ increases with increasing source size, before dropping to zero: the more extended the source, the farther away from the lens centre it can be while still producing multiple images.
+
+%The limit posed by lensing to the maximum detectable source size can be understood as follows.
+%In the point source case, at a given flux, only sources that produce at least two images with a sufficiently high magnification are classified as strongly lensed.
+%In most practical cases, the limiting factor is the magnification of the secondary image, $\mu_2$, which is a strong function of the source position (see \Fref{fig:1dmag}).
+%Only sources that lie within a region of large $\mu_2$, the area of which is equal to the strong lensing cross-section, give rise to a strong lens configuration.
+%When the source is extended, different parts of it are magnified by different amounts.
+If the source size is comparable to or larger than the region of high magnification, then its net magnification averaged over its extent will be lower than in the point-source case, and the secondary image will drop below the detection limit.
+%This effect adds to the limit on the source surface brightness in the absence of lensing to produce the trend observed in \Fref{fig:extcs}.
+
+%In the point source case, the necessary and sufficient condition for a source to be strongly lensed is that two detectable images are produced.
+%The detection limit of the survey and the intrinsic flux of the source set a requirement on the minimum magnification that two images of the source must have in order for them to be detected.
+%The limiting factor is set by the magnification of the secondary image: given a source flux, 
+%In the point source case, at a given flux, only sources that lie within a certain region give rise to at least two detectable images, which are required to classify a system as a strong lens. 
+%The extent of this region, which is equal to the strong lensing cross-section, is determined by the magnification required in order for the secondary image to be detected (see \Fref{fig:1dmag}).
+%The area of this region is equal to the strong lensing cross-section, and decreases for fainter sources.
 

### 2022-07-19 single_lens/axisymmetric_point/plot_composite_crosssect.py

@@ -66,6 +66,9 @@ ax[1].tick_params(axis='both', which='both', top=True, right=True, labelsize=fsi
 
 ax[1].set_xlabel('$f_{\mathrm{DM}}$', fontsize=fsize)
 
+ax[0].text(0.5, 1., '$f_{\mathrm{DM}}=0.5$', fontsize=fsize)
+ax[1].text(0.55, 1., '$\gamma_{\mathrm{DM}}=1.5$', fontsize=fsize)
+
 ax[0].set_yscale('log')
 ax[1].set_yscale('log')
 

### 2022-07-19 single_lens/elliptical_ext/get_crosssect.py

@@ -5,12 +5,12 @@ import h5py
 
 
 ftot = 200.
-nre = 6
+nre = 11
 logre_grid = np.linspace(-1., 0., nre)
 
 nsim = 10000
 
-rsim = 1.
+rsim = 1.5
 source_area = np.pi * rsim**2
 
 nfrat = 4

### 2022-07-19 single_lens/elliptical_ext/plot_cs_vs_re.py

@@ -13,7 +13,7 @@ rc('text', usetex=True)
 fsize = 18
 
 #xlim = (0.1, 1.)
-ylim = (0.085, 1.3)
+ylim = (0.085, 1.5)
 
 logscale = True
 
@@ -31,6 +31,9 @@ logre_grid = cs_file['logre_grid'][()]
 nre = len(logre_grid)
 
 cs_grid = cs_file['cs_grid'][()]
+cs_ps = cs_grid[:, 0] # approximation of point-source cross-section
+re_max = (cs_ps/np.pi)**0.5
+print(re_max)
 
 fig, ax = pylab.subplots(1, 1)#, figsize=(8, 4))
 pylab.subplots_adjust(left=leftm, right=0.98, bottom=0.14, top=0.97, wspace=0.)
@@ -38,9 +41,12 @@ pylab.subplots_adjust(left=leftm, right=0.98, bottom=0.14, top=0.97, wspace=0.)
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
 for i in range(nfrat):
-    ax.loglog(10.**logre_grid, cs_grid[i, :]/np.pi, linewidth=2, color=colseq[i], label="$\log{\\frac{f}{\sigma_{\mathrm{sky,\\theta_{\mathrm{Ein}}^{2}}}}} = %2.1f$"%lfrat_grid[i])
+    #ax.loglog(10.**logre_grid, cs_grid[i, :]/np.pi, linewidth=2, color=colseq[i], label="$\log{\\frac{f}{\sigma_{\mathrm{sky,\\theta_{\mathrm{Ein}}^{2}}}}} = %2.1f$"%lfrat_grid[i])
+    ax.loglog(10.**logre_grid, cs_grid[i, :]/np.pi, linewidth=2, color=colseq[i], label="$\log{f} = %2.1f$"%lfrat_grid[i])
     cs_spline = splrep(10.**logre_grid, cs_grid[i, :]/np.pi, k=1)
 
+# lines of constant surface brightness (who cares?)
+"""
 nfr2 = nfrat
 cs_fixedfr2 = np.zeros((nfr2, nfrat))
 re_fixedfr2 = np.zeros((nfr2, nfrat))
@@ -59,6 +65,7 @@ for i in range(nfr2):
         istart = 0
         iend = 4
[…]

### 2022-07-19 single_lens/elliptical_ext/plot_single_footprint.py

@@ -32,7 +32,7 @@ for n in range(nreg):
     img_sn = signal/noise
     if img_sn < 10.:
         img[labels==n+1] = 0.
-    print(n, img_sn)
+    print('image n. %d. S/N: %2.1f. Flux: %2.1e'%(n, img_sn, signal))
 
 ax.contourf(img,[thresh,1e8], colors='b')
 ax.set_aspect(1.)

### 2022-07-19 single_lens/elliptical_ext/plot_sourcepos.py

@@ -1,0 +1,108 @@
+import numpy as np
+from astropy.io import fits as pyfits
+from skimage import measure
+import h5py
+import pylab
+
+
+np.random.seed(0)
+
+ftot = 200.
+nre = 11
+logre_grid = np.linspace(-1., 0., nre)
+
+nsim = 10000
+
+# first re-generates source positions to match those of the sims
+rsim = 1.5
+source_area = np.pi * rsim**2
+
+source_x = []
+source_y = []
+
+r = np.random.rand(nsim)**0.5 * rsim
+phi = 2.*np.pi*np.random.rand(nsim)
+source_x = r*np.cos(phi)
+source_y = r*np.sin(phi)
+
+lfrat = 2.
+# frat is the ratio between the source intrinsic flux and the sky rms over an area equal to the square of the Einstein radius.
+# In my sims, the Einstein radius is 1 arcsec. Pixels have a 0.05" size.
+# Then, the sky rms over a 1 square arcsec is 20 times that over a single pixel
+# All simulations have a source intrinsic flux of 200. Then an frat=10 
+# corresponds to a sky rms per pixel of 1.
+
+nsigma = 2.
+
+x0 = 39.5
+y0 = 39.5
+
+X, Y = np.meshgrid(np.arange(80), np.arange(80))
+
+sky_rms = 10.**(1. - lfrat) 
+
+re_inds = [0]
+
+for m in re_inds:
+    logre = logre_grid[m]
+
+    preamble = 'ftot200_logre%2.1f'%logre
+    nlens = 0
+    islens_sim = np.zeros(nsim, dtype=bool)
+    nimages_sim = np.zeros(nsim, dtype=int)
+    aperture_sim = np.zeros(nsim)
+
+    for i in range(nsim):
+        img = pyfits.open('mockdir/'+preamble+'/'+preamble+'_%04d_image.fits'%i)[0].data
+        footprint = img > nsigma * sky_rms
+
+        labels = measure.label(footprint)
+        nreg = labels.max()
+        nimg = 0
+        for n in range(nreg):
+            npix_here = (labels==n+1).sum()
+            signal = img[labels==n+1].sum()
+            noise = npix_here**0.5 * sky_rms
+            img_sn = signal/noise
+            if img_sn < 10.:
+                img[labels==n+1] = 0.
+            else:
+                nimg += 1
+
+        # checks subtended angle
+
+        new_footprint = img > nsigma * sky_rms
+
+        ypix = Y[new_footprint]
[…]

### 2022-07-19 single_lens/elliptical_ext/write_sim_input.py

@@ -8,7 +8,7 @@ np.random.seed(0)
 
 # prepares a Glafic input file
 
-nre = 6
+nre = 11
 logre_grid = np.linspace(-1., 0., nre)
 
 ftot = 200.
@@ -21,7 +21,16 @@ f.close()
 
 nsource = 10000
     
-rmax = 1.
+rmax = 1.5
+
+# draws the same source positions for each value of re, so that
+# different simulations can be compared more easily
+
+# generates N sources within a circle of radius 1''
+r = np.random.rand(nsource)**0.5 * rmax
+phi = 2.*np.pi*np.random.rand(nsource)
+x = r * np.cos(phi)
+y = r * np.sin(phi)
 
 for i in range(nre):
 
@@ -39,15 +48,9 @@ for i in range(nre):
     lines.append('reset_extend 1 2 %f\n'%I0)
     lines.append('reset_extend 1 7 %f\n'%10.**logre_grid[i])
     lines.append('\n')
-    
-    # generates N sources within a circle of radius 1''
-    r = np.random.rand(nsource)**0.5 * rmax
-    phi = 2.*np.pi*np.random.rand(nsource)
-    x = r * np.cos(phi)
-    y = r * np.sin(phi)
-    
+   
     for n in range(nsource):
-        lines.append('reset_par prefix %s_%04d\n'%(prefix, n))
+        lines.append('reset_par prefix %s/%s_%04d\n'%(prefix, prefix, n))
         lines.append('reset_extend 1 3 %f\n'%(x[n]))
         lines.append('reset_extend 1 4 %f\n'%(y[n]))
         lines.append('writeimage\n')
@@ -55,12 +58,14 @@ for i in range(nre):
         #lines.append('writeimage 0. 1.\n')
         lines.append('\n')
     
+    lines.append('quit\n')
+
     dirname = 'mockdir/%s/'%prefix
     
     if not os.path.isdir(dirname):
         os.system('mkdir %s'%dirname)
     
-    f = open(dirname+'/%s.input'%prefix, 'w')
+    f = open('mockdir/%s.input'%prefix, 'w')
     f.writelines(lines)
     f.close()
 

### 2022-07-19 single_lens/elliptical_point/plot_composite_caustics.py

@@ -74,6 +74,6 @@ pylab.text(-0.8, 0.82, 'Major axis', fontsize=fsize)
 
 pylab.legend(loc='lower right', framealpha=1., fontsize=fsize)
 
-pylab.savefig('../../paper/composite_caustics.eps')
+pylab.savefig('../../paper/composite_caustics.png')
 pylab.show()
 

### 2022-07-19 single_lens/elliptical_point/set_composite_pars.py

@@ -58,8 +58,10 @@ mstar_phys = (1. - f_dm) * mein_phys / deV.M2d(reff_phys, reff_phys)
 # and with dark matter slope of 1.5
 gammadm = 1.5
 
+# and with r_s = 20 * reff
+rs_phys = 10. * reff_phys
+
 # Finds the virial mass of the halo
-rs_phys = 100.
 rs_arcsec = rs_phys / arcsec2kpc
 
 mein_dm = f_dm * mein_phys

### 2022-07-21 paper/seleff.tex
started section 4

@@
 \def\mstar{M_*}
+\def\asps{\alpha_{\mathrm{sps}}}
+
+\def\mobs{M_*^{(\mathrm{obs})}}
 \def\gammadm{\gamma_{\mathrm{DM}}}
@@
 \citet{MVK09} carried out a thorough study of point 1): they quantified how the properties of a lens galaxy determine its probability of creating a lensing event with a point source.
-In this work we revisit the \citet{MVK09} study, we expand it to the extended source case and, most importantly, we address point 2) as well: we simulate large populations of strong lenses using semi-analytic models, simulate lensing surveys with realistic assumptions for the lens probability $\pdet$ and quantify the lensing bias under various scenarios.
+In this work we revisit the \citet{MVK09} study, we expand it to the extended source case and, most importantly, we address point 2) as well: we simulate large populations of strong lenses using empirical models, simulate lensing surveys with realistic assumptions for the lens probability $\pdet$ and quantify the lensing bias under various scenarios.
 We explore how the results change as a function of the scatter in mass parameters at fixed light, and as a function of source type: we simulate both galaxy and quasar sources.
@@
 
+In the previous section we studied how the lensing cross-section, which is closely related to the term $\pdet$ in \Eref{eq:one}, varies as a function of lens and source properties.
+From here on we study the effect that those trends have on populations of lenses.
+We do this by simulating populations of foreground galaxies and background sources, selecting strong lenses among them, and comparing the properties of the strong lens sample with the general galaxy population.
+Our simulations are based on empirical models, in which existing observations of the baryonic component of galaxies are complemented with a set of assumptions on the mass distribution of the lenses.
+In section \ref{ssec:lenses} we explain how we built our foreground galaxy sample,
+while in section \ref{ssec:sources} we describe the simulation of the background source population.
+In section \ref{ssec:obs} we describe how our mock observations of lenses are generated.
+
+%As we explained in \Sref{sect:intro}, a key ingredient determining the amplitude of the strong lensing bias is the scatter in the mass properties of the foreground galaxy population.
+%Although we do not have robust measurements of the intrinsic scatter in properties such as the stellar mass-to-light ratio or the dark matter halo mass, we set upper limits on their amplitude using the
+%Our simulations explore different amplitudes of the scatter. Scatter
+
+%In section \ref{ssec:lenses} we explain how we built our foreground galaxy sample.
+%In section \ref{ssec:
+
+\subsection{Foreground galaxies}\label{ssec:lenses}
+
+Our foreground galaxy population consists of a volume-limited sample of early-type galaxies, complete above a minimum observed stellar mass of $10^{11}M_\odot$.
+We chose to focus on early-type galaxies because many strong lensing surveys have preferentially targeted this class of objects in their lens-finding phase \citep[e.g.][]{Bol++06,Gav++12,Son++18a}.
+%This, in turn, was motivated by two points:
+This, in turn, had a dual motivation:
+first, early-type galaxies are among the most massive objects in the Universe, and therefore they are more likely to be lenses;
+second, their smooth surface brightness distribution and red colour makes it easier to detect arcs from strongly lensed star forming galaxies around them.
+
+We describe lenses with elliptical versions of the two-component model introduced in section \ref{ssec:profile}.
+In the following sections we explain how their parameters are generated.
+
+\subsubsection{Stellar mass and redshift distribution}
+
+We generated lens galaxies over a finite redshift range, $0.1 < z < 0.9$.
+We chose these lower and upper limits because the value of the critical surface mass density $\Sigma_{\mathrm{cs}}$ becomes very large outside of this range, and the fraction of galaxies that act as strong lenses drops substantially as a result.
+
+We drew stellar masses from the stellar mass function of quiescent galaxies measured by \citet{Muz++13}.
+In particular, we used the following comoving number density distribution
+\begin{equation}
+\Phi(\mobs) = (\ln{10})\Phi^*\left[10^{(\mobs - \mstar^*)(1+\alpha)}\right]\times \exp{\left[-10^{(\mobs -\mstar^*)}\right]}.
+\end{equation}
+We set $\Phi^*=1.009\times10^{-3}\,{\rm Mpc}^{-3}$, $\alpha=-0.92$ and $\log{\mstar^*}=11.21$. These are the best-fit values measured by \citet{Muz++13}.
+
+We refer to the stellar masses drawn from this distribution as the observed stellar masses, $\mobs$, as opposed to the true stellar masses.
+The observed stellar mass is meant to represent an estimate of $\mstar$ based on stellar population synthesis modelling, which is the method used by \citet{Muz++13} to measure the galaxy stellar mass function.
+Stellar population synthesis measurements, however, are subject to systematic uncertainties, since they have not been calibrated on galaxies whose stellar mass is known by other means.
+We quantify the discrepancy between the observed and true stellar mass by means of the stellar population synthesis mismatch parameter $\asps$, defined as follows:
+\begin{equation}
+\asps \equiv \frac{\mstar}{\mobs}.
+\end{equation}
+
+%We randomly assign a value of $\asps$ to each galaxy in the sample, by drawing it from the following distribution:
+For each galaxy in the sample, we randomly drew a value of $\log{\asps}$ from the following distribution:
+\begin{equation}
+\pr(\log{\asps}) \sim \mathcal{N}(0.1, \sigma_\alpha^2),
+\end{equation}
+where the notation $\mathcal{N}(\mu,\sigma^2)$ indicates a Gaussian with mean $\mu$ and variance $\sigma^2$.
+We set the mean of $\log{\asps}$ to $0.1$, as this is an intermediate value among estimates of $\asps$ from the literature \citep{CvD12, Cap++13, SLC15, Son++15, Son++19}.
+For fhe scatter $\sigma_\alpha$ we adopted a few different values. We explain in \ref{ssub:scat} how these were chosen.
+
+\subsubsection{Surface mass density distribution}
+
+We described the surface mass density distribution of each galaxy as an elliptical de Vaucouleurs profile (a S\'{e}rsic profile with $n=4$).
+Given the observed stellar mass of a galaxy, we assigned a half-mass radius by drawing it from the following distribution in $\log{\reff}$:
+\begin{equation}\label{eq:masssize}
+\pr(\log{\reff}) \sim \mathcal{N}(1.20 + 0.63(\log{\mobs} - 11.4), 0.14^2).
+\end{equation}
+Then, we assigned an axis ratio $q$ by drawing it from the following distribution:
+\begin{equation}\label{eq:qdist}
+\pr(q) \sim \,{\rm boh}.
+\end{equation}
+%The choice for the distributions of \Eref{eq:masssize} and \Eref{eq:qdist}
+The choice for these distributions was motivated by observations of a sample of early-type galaxies.
+In Appendix~\ref{sect:appendixa} we explain how this sample was defined and how the coefficients of \Eref{eq:masssize} and \Eref{eq:qdist} were determined.
+
+\subsubsection{Dark matter distribution}
+
+\subsubsection{Intrinsic scatter}\label{ssub:scat}
+
+\subsection{Background sources}\label{ssec:sources}
+
+\subsection{Observations}\label{ssec:obs}
+
 %__________________________________________________________________
@@
 
+\appendix
+\section{Lens galaxy surface brightness distribution}\label{sect:appendixa}
+
+To assign half-light radii and ellipticities to the simulated lenses we relied on observations of a sample of early-type galaxies selected from the Sloan Digital Sky Survey \citep[SDSS][]{Yor++00}.
+This sample was selected as follows.
+First we defined a narrow redshift slice around $z=0.2$.
+Then we applied a selection in colour, by choosing objects with $g-r>1.2$, and on S\'{e}rsic index, by selecting only galaxies with $n>2.5$.
+We used the S\'{e}rsic fit measurements by \citet{Mee++15} for this purpose.
+These cuts produced a sample of $8078$ galaxies.
+%We focused on the $r-$band de Vaucouleurs profile measurements obtained by \citet{Mee++15}, for these objects.
+%We then took measurements of their total flux, half-light radius and axis ratio from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}. Using the 
+%We then took measurements of the total flux, half-light radius and axis ratio obtained by \citet{Mee++15} from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}.
+Using the
+%For each object, we took the measurement of its de Vaucouleurs model-based $r-band$ total flux, half-light radius and axis ratio, as provided by \citet{Mee++15}.
+We then focused on the $r-$band de Vaucouleurs model-based photometric measurements of \citet{Mee++15}.
+Using the $r-$band total flux from the de Vaucouleurs model and the stellar mass-to-light ratio estimates of \citet{Men++14}, we obtained measurements of $\mobs$.
+Finally, we fitted the stellar mass-size relation with the following model:
+\begin{equation}
+\log{\reff} \sim \mathcal{N}(\mu_{R,0} + \beta_R(\log{\mobs} - 11.4), \sigma_R^2).
+\end{equation}
+We obtained
+
 \end{document}

### 2022-07-21 lens_population/fit_mass_size.py

@@ -1,8 +1,9 @@
 import numpy as np
-from scipy.optimize import leastsq, least_squares
+from scipy.optimize import leastsq, least_squares, minimize
 from astropy.io import fits as pyfits
 from sl_cosmology import Dang
 from scipy.interpolate import splrep, splev
+import pylab
 
 
 samp_table = pyfits.open('sdss_zbin_latetype_red.fits')[1].data
@@ -26,18 +27,29 @@ dd_spline = splrep(zgrid, dd_grid)
 reff_kpc = reff_arcsec * np.deg2rad(1./3600.) * splev(samp_table['z'][cut], dd_spline) * 1000.
 
 lreff = np.log10(reff_kpc)
-lreff_arr = lreff.reshape((1, len(lreff)))
-lmstar_arr = lmstar.reshape((1, len(lmstar)))
 
-def fitfunc(p, x):
-    return p[0] + p[1]*(x - lmstar_piv)
+# I have no idea why I had to do this, but leastsq wouldn't work otherwise.
+lmstar0 = np.zeros(len(lmstar))
+lmstar0 += lmstar
+lmstar = lmstar0.copy()
 
-def errfunc(p, x, y):
-    #return lreff - fitfunc(p)
-    return y - fitfunc(p, x)
+def fitfunc(p):
+    return p[0] + p[1]*(lmstar - lmstar_piv)
 
-r = leastsq(errfunc, [1., 1.], (lmstar[:10], lreff[:10]))
-print(r)
-print(errfunc([1., 1.], lmstar[:10], lreff[:10]))
+def errfunc(p):
+    return fitfunc(p) - lreff
 
+p0 = [1., 1.]
+s = leastsq(errfunc, p0)
+print(s[0])
+
+model_lreff = s[0][0] + s[0][1] * (lmstar - lmstar_piv)
+scatter = (((model_lreff - lreff)**2).sum()/len(lreff))**0.5
+print(scatter)
+
+pylab.scatter(lmstar, lreff)
+xlim = pylab.xlim()
+xs = np.linspace(xlim[0], xlim[1])
+pylab.plot(xs, s[0][0] + s[0][1]*(xs - lmstar_piv), linestyle='--', color='k')
+pylab.show()
 

### 2022-07-21 lens_population/mcmc_mass_size.py

@@ -1,0 +1,89 @@
+import numpy as np
+import emcee
+from astropy.io import fits as pyfits
+from sl_cosmology import Dang
+from scipy.interpolate import splrep, splev
+from scipy.stats import truncnorm
+import pylab
+
+
+samp_table = pyfits.open('sdss_zbin_latetype_red.fits')[1].data
+
+cut = (samp_table['logmstar_dev'] > 11.) & (samp_table['logmstar_dev'] < 12.2)
+
+lmstar = samp_table['logmstar_dev'][cut]
+lmstar_piv = 11.4
+
+reff_arcsec = samp_table['dev_reff'][cut]
+
+zmin = 1.9
+zmax = 2.1
+zgrid = np.linspace(zmin, zmax, 5)
+dd_grid = 0.*zgrid
+for i in range(5):
+    dd_grid[i] = Dang(zgrid[i])
+
+dd_spline = splrep(zgrid, dd_grid)
+
+reff_kpc = reff_arcsec * np.deg2rad(1./3600.) * splev(samp_table['z'][cut], dd_spline) * 1000.
+
+lreff = np.log10(reff_kpc)
+
+# I have no idea why I had to do this, but leastsq wouldn't work otherwise.
+lmstar0 = np.zeros(len(lmstar))
+lmstar0 += lmstar
+lmstar = lmstar0.copy()
+
+mu_par = {'name': 'mu', 'lower': -3., 'upper': 3., 'guess': 1., 'spread': 0.1}
+beta_par = {'name': 'beta', 'lower': -3., 'upper': 3., 'guess': 1., 'spread': 0.1}
+sig_par = {'name': 'sig', 'lower': 0., 'upper': 1., 'guess': 0.2, 'spread': 0.1}
+
+pars = [mu_par, beta_par, sig_par]
+npars = len(pars)
+
+bounds = []
+for par in pars:
+    bounds.append((par['lower'], par['upper']))
+
+def logprior(p):
+    for i in range(npars):
+        if p[i] < bounds[i][0] or p[i] > bounds[i][1]:
+            return -1e300
+    return 0.
+
+def logpfunc(p):
+
+    lprior = logprior(p)
+    if lprior < 0.:
+        return -1e300
+
+    mu, beta, sig = p
+
+    lreff_model = mu + beta * (lmstar - lmstar_piv)
+    
+    sumlogp = (-0.5*(lreff_model - lreff)**2/sig**2 - np.log(sig)).sum()
+
[…]

### 2022-07-22 paper/seleff.tex
q distribution

@@
 \end{equation}
-Then, we assigned an axis ratio $q$ by drawing it from the following distribution:
+Then, we assigned an axis ratio $q$ by drawing it from the following beta distribution:
 \begin{equation}\label{eq:qdist}
-\pr(q) \sim \,{\rm boh}.
+\pr(q) \propto q^{\alpha-1}(1 - q)^{\beta-1},
 \end{equation}
-%The choice for the distributions of \Eref{eq:masssize} and \Eref{eq:qdist}
+with $\alpha=6.28$ and $\beta=2.05$.
 The choice for these distributions was motivated by observations of a sample of early-type galaxies.
@@
 \end{equation}
-We obtained
+We obtained $\mu_{R,0}=1.20$, $\beta_R=0.63$ and $\sigma_R=0.14$.
+
+We then proceeded to fit for the axis ratio distribution of the same sample of galaxies.
+\Fref{fig:qhist} shows a histogram of the observed distribution.
+We fitted this with a beta disribution:
+\begin{equation}\label{eq:betaappendix}
+\pr(q) \propto q^{\alpha-1}(1 - q)^{\beta-1}.
+\end{equation}
+We obtained $\alpha=6.28$ and $\beta=2.05$.
+\begin{figure}
+\includegraphics[width=\columnwidth]{q_dist.eps}
+\caption{
+Distribution in axis ratio of a sample of early-type galaxies.
+The sample consists of $8078$ galaxies from the SDSS, selected by means of cuts in redshift, colour and S\'{e}rsic index as explained in the text.
+Measurements of the axis ratio are taken from the de Vaucouleurs model fits of \citet{Mee++15}.
+The model curve is a beta distribution, \Eref{eq:betaappendix}, with $\alpha=6.28$ and $\beta=2.05$.
+}
+\end{figure}
 

### 2022-07-22 lens_population/fit_q_dist.py

@@ -1,0 +1,77 @@
+import numpy as np
+from scipy.optimize import leastsq, least_squares, minimize
+from scipy.stats import beta
+from astropy.io import fits as pyfits
+from sl_cosmology import Dang
+from scipy.interpolate import splrep, splev
+import pylab
+from matplotlib.ticker import MultipleLocator
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+# fits a Beta distribution to the q distribution of the sample
+
+samp_table = pyfits.open('sdss_zbin_latetype_red.fits')[1].data
+
+cut = (samp_table['logmstar_dev'] > 11.) & (samp_table['logmstar_dev'] < 12.2) & (samp_table['dev_ba'] > 0.) & (samp_table['dev_ba'] < 1.)
+
+"""
+# optional: plots q for three mass bins, to see if there's a trend
+# with stellar mass. There isn't.
+qbins = np.linspace(0., 1., 11)
+
+mbins = np.linspace(11., 12., 4)
+
+for i in range(3):
+    mbin = (samp_table['logmstar_dev'] > mbins[i]) & (samp_table['logmstar_dev'] < mbins[i+1])
+    ngal = mbin.sum()
+    weights = np.ones(ngal)/ngal
+    pylab.hist(samp_table['dev_ba'][mbin], weights=weights, bins=qbins, histtype='step')
+pylab.show()
+"""
+
+q_samp = samp_table['dev_ba'][cut]
+
+def logpfunc(p):
+    a, b = p
+    return -beta.logpdf(q_samp, a, b).sum()
+
+p0 = [2., 2.]
+print(logpfunc(p0))
+
+s = minimize(logpfunc, p0)
+print(s)
+a, b = s.x
+
+qbins = np.linspace(0., 1., 11)
+
+weights = np.ones(len(q_samp))/len(q_samp)/(qbins[1] - qbins[0])
+
+fsize = 18
+
+fig, ax = pylab.subplots()
+
+pylab.subplots_adjust(left=0.14, right=1., bottom=0.14, top=0.98)
+
+ax.hist(q_samp, bins=qbins, weights=weights, histtype='stepfilled', label='Observed')
+q_arr = np.linspace(0., 1., 101)
+
+ax.plot(q_arr, beta.pdf(q_arr, a, b), label='Model', linewidth=2)
+
+ax.legend(loc='upper left', fontsize=fsize)
+
[…]

### 2022-07-22 paper/seleff.tex
started halos

@@
 \def\mdmfive{M_{\mathrm{DM}, 5}}
+\def\mhalo{M_{\mathrm{h}}}
 
@@
 
-   \author{Alessandro Sonnenfeld\inst{1}
+   \author{Alessandro Sonnenfeld\inst{1}\and
+           Others
           }
@@
 %At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), $\crosssect$ increases with increasing source size, before dropping to zero.
-At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (red ellipse in \Fref{fig:cs}).
+At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (red ellipse in \Fref{fig:ellcaust}).
 This is because, when the source is very bright, it can give rise to multiple images even while its centroid lies outside of the radial caustic, as long as its surface brightness distribution extends into it.
@@
 
+We modelled the dark matter distribution of each lens galaxy with an elliptical gNFW halo (see section \ref{ssec:profile} for its definition).
+The parameters of the radial density profile were assigned as follows.
+Given the stellar mass of a lens, we first determined the value of its halo virial mass, $\mhalo$.
+We defined this as the mass enclosed within a spherical shell with average density equal to $200$ times the critical density of the Universe.
+We drew halo masses from the following distribution:
+\begin{equation}
+\pr(\log{\mhalo}) \sim \mathcal{N}(13.0 + 1.0(\log{\mstar} - 11.5), \sigma_h^2),
+\end{equation}
+which is a Gaussian in $\log{\mhalo}$ with a mean that scales with stellar mass and scatter $\sigma_h$.
+The values of the coefficients describing the mean of the distribution were chosen to be close to current weak lensing constraints on the halo mass of elliptical galaxies \citep{Son++22}.
+The the intrinsic scatter in halo mass is more uncertain from an observational point of view, therefore we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
+
+%which is similar to current weak lensing constraints on the halo mass distribution of elliptical galaxies \citep{Son++22}.
+
 \subsubsection{Intrinsic scatter}\label{ssub:scat}

### 2022-07-23 paper/seleff.tex
fundamental plane fit

@@
 Given a surface mass density profile $\Sigma(R)$, a lens with an elliptical mass distribution can be obtained by replacing the radial coordinate with the circularised radius:
-\begin{equation}
+\begin{equation}\label{eq:elltransf}
 R \rightarrow \sqrt{qx^2 + \frac{y^2}{q}},
@@
 
-In the previous section we studied how the lensing cross-section, which is closely related to the term $\pdet$ in \Eref{eq:one}, varies as a function of lens and source properties.
-From here on we study the effect that those trends have on populations of lenses.
-We do this by simulating populations of foreground galaxies and background sources, selecting strong lenses among them, and comparing the properties of the strong lens sample with the general galaxy population.
+In the previous section we showed how the lensing cross-section, which is closely related to the term $\pdet$ in \Eref{eq:one}, varies as a function of lens and source properties.
+From here on we focus on the effect that those trends have on populations of lenses.
+We addressed this question by simulating populations of foreground galaxies and background sources, selecting strong lenses among them, and comparing the properties of the strong lens sample with the general galaxy population.
+We simulated a lens-based search (as opposed to a source-based one), in which strongly lensed images are searched among a stellar mass-selected sample of galaxies.
 Our simulations are based on empirical models, in which existing observations of the baryonic component of galaxies are complemented with a set of assumptions on the mass distribution of the lenses.
@@
 %Our simulations explore different amplitudes of the scatter. Scatter
-
-%In section \ref{ssec:lenses} we explain how we built our foreground galaxy sample.
-%In section \ref{ssec:
 
@@
 We set the mean of $\log{\asps}$ to $0.1$, as this is an intermediate value among estimates of $\asps$ from the literature \citep{CvD12, Cap++13, SLC15, Son++15, Son++19}.
-For fhe scatter $\sigma_\alpha$ we adopted a few different values. We explain in \ref{ssub:scat} how these were chosen.
+For the scatter $\sigma_\alpha$ we adopted a few different values. We explain in \ref{ssub:scat} how these were chosen.
 
-\subsubsection{Surface mass density distribution}
+\subsubsection{Stellar mass density distribution}
 
@@
 We defined this as the mass enclosed within a spherical shell with average density equal to $200$ times the critical density of the Universe.
-We drew halo masses from the following distribution:
+%We randomly drew halo masses from the following distribution:
+Using current weak lensing constraints on the halo mass of elliptical galaxise as a reference \citep{Son++22}, we drew halo masses from the following distribution:
 \begin{equation}
@@
 which is a Gaussian in $\log{\mhalo}$ with a mean that scales with stellar mass and scatter $\sigma_h$.
-The values of the coefficients describing the mean of the distribution were chosen to be close to current weak lensing constraints on the halo mass of elliptical galaxies \citep{Son++22}.
-The the intrinsic scatter in halo mass is more uncertain from an observational point of view, therefore we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
+%The values of the coefficients describing the mean of the distribution were chosen to be close to current weak lensing constraints on the halo mass of elliptical galaxies \citep{Son++22}.
+%The intrinsic scatter in halo mass is more uncertain from an observational point of view, therefore we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
+The intrinsic scatter in halo mass is not well constrained observationally, therefore we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
 
-%which is similar to current weak lensing constraints on the halo mass distribution of elliptical galaxies \citep{Son++22}.
+To determine the density profile of the halo we relied on a theoretically motivated model that takes into account the effect of baryons on the dark matter.
+We assumed the dark matter distribution to be initially described by an NFW profile with a concentration\footnote{The concentration is the ratio between the halo virial radius and the scale radius $r_s$.} of five, then used the prescription of \citet{Cau++20} to model the response of the halo to the infall of baryons.
+% on the basis of the present stellar mass distribution.
+In the \citet{Cau++20} model the halo response is approximated with an analytical function that depends on the present stellar mass distribution, and typically results in a more concentrated and steeper density profile compared to the original NFW model.
+Finally, we fitted a gNFW profile to the surface mass density of the contracted halo.
+By doing this, we were able to fully describe the dark matter density profile with three parameters: $\mhalo$, $\gammadm$ and $r_s$.
+
+The dark matter density profile defined in this way is determined uniquely by the halo mass, the stellar mass and the half-light radius (the more concentrated the stellar distribution, the stronger the halo response, the steeper the dark matter density profile).
+However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope.
+Therefore, we applied a random scatter on $\gammadm$ of amplitude $\sigma_{\gamma}$.
+We explored different values for $\sigma_{\gamma}$, as explained in the next section.
+
+Given the radial profile of the dark matter halo, we obtained an elliptical version of it by applying a transformation of the kind of \Eref{eq:elltransf} to its projected surface mass density.
+We assumed that the axis ratio and orientation of the halo is the same as that of the stellar component.
 
 \subsubsection{Intrinsic scatter}\label{ssub:scat}
+
+The distribution in stellar mass, halo mass and dark matter inner slope of our sample of simulated foreground galaxies depends on parameters describing the intinsic scatter in these properties, namely $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_{\gamma}$.
+Direct observational constraints on these quantities are poor.
+However, we can derive upper limits on them on the basis of observed scaling relations.
+
+Early-type galaxies lie on the stellar mass fundamental plane, a scaling relation between stellar mass, half-light radius and central velocity dispersion \citep{H+B09}:
+\begin{equation}\label{eq:fpcompact}
+\sigma_{\mathrm{e}} \propto \mstar^{(\mathrm{obs})\beta_{\sigma}}\reff^{\xi_{\sigma}}.
+\end{equation}
+The existence of a fundamental plane relation is a consequence of the virial theorem: the velocity dispersion of a galaxy is directly related to the 3-dimensional mass distribution in its inner regions, which is tyically dominated by the stellar component.
+At fixed observed stellar mass distribution, however, the central velocity dispersion can vary depending on the stellar population synthesis mismatch parameter, on the dark matter mass and on the dark matter density profile.
+This gives rise to a spread in the values of the velocity dispersion given $\mobs$ and $\reff$.
+Therefore, we can use the observed scatter in velocity dispersion to put an upper limit on the intrinsic scatter in the stellar population synthesis mismatch parameter, halo mass and dark matter slope.
+
+We used Jeans modelling for this purpose.
+We generated samples of early-type galaxies with the recipes described above and with different values of $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$.
+Using the spherical Jeans equation under the assumption of isotropic orbits, we predicted the central velocity dispersion of each galaxy in the sample.
+Then, we fitted a fundamental plane relation to this mock sample and measured the predicted scatter in velocity dispersion at fixed $\mobs$ and $\reff$.
+We then varied $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$ to match the observed and the predicted scatter.
+The details of this procedures are given in Appendix~\ref{sect:appendixb}.
 
@@
 
+\section{Upper limits on the intrinsic scatter parameters}\label{sect:appendixb}
+
+We used the fundamental plane of early-type galaxies to set an upper limit on the intrinsic scatter parameters of the simulation: $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$.
+First, we measured the fundamental plane of the sample of SDSS early-type galaxies introduced in Appendix~\ref{sect:appendixa}.
+We took measurements of the line-of-sight stellar velocity dispersion within the SDSS spectroscopic aperture, $\sigma_{\mathrm{ap}}$, from the SDSS data release 16 catalogue \citep{Ahu++20}.
+Then, we fitted the following model to the distribution of $\sigma_{\mathrm{ap}}$ of the sample:
+\begin{equation}
+\pr(\sigma_{\mathrm{ap}}) \sim \mathcal{N}(\mu_{\sigma} + \beta(\log{\mobs} - 11.4) + \xi(\log{\reff} - 1.2), \sigma_{\sigma}).
+\end{equation}
+We accounted for observational uncertainties on $\sigma_{\mathrm{ap}}$ when doing the fit, therefore the parameter $\sigma_{\sigma}$ describes the intrinsic scatter in the logarithm of the velocity dispersion, deconvolved from the observational scatter.
+In principle we should also account for observational uncertainties on the stellar mass measurement, as they too contribute to the inferred scatter in $\sigma_{\mathrm{ap}}$.
+In practice, however, it is difficult to estimate observational uncertainties on stellar population synthesis measurements.
+For this reason we chose not to propagate uncertainties on $\mobs$. As a result, the inferred scatter parameter $\sigma_{\sigma}$ is slightly overestimated.
+We obtained $\mu_{\sigma}=2.36$, $\beta_{\sigma}=0.33$, $\xi_\sigma=-0.17$, and $\sigma_{\sigma}=0.035$.
+
+We then generated samples of galaxies from the model of section \ref{ssec:lenses}.
+
+We did not attempt to match the other parameters of the fundamental plane (i.e.
+the constant of proportionality of \Eref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.
+
+%First, we took the sample of SDSS early-type galaxies introduced in Appendix~\ref{sect:appendixa}
+
 \end{document}

### 2022-07-23 lens_population/fit_fundamental_plane.py

@@ -1,0 +1,57 @@
+import numpy as np
+from scipy.optimize import leastsq, least_squares, minimize
+from astropy.io import fits as pyfits
+from sl_cosmology import Dang
+from scipy.interpolate import splrep, splev
+import pylab
+
+
+samp_table = pyfits.open('sdss_zbin_latetype_red.fits')[1].data
+
+cut = (samp_table['logmstar_dev'] > 11.) & (samp_table['logmstar_dev'] < 12.2) & (samp_table['veldisp'] > 0.) & (samp_table['veldisperr'] < samp_table['veldisp'])
+
+lmstar = samp_table['logmstar_dev'][cut]
+lmstar_piv = 11.4
+lreff_piv = 1.2
+
+veldisp = samp_table['veldisp'][cut]
+veldisp_err = samp_table['veldisperr'][cut]
+
+reff_arcsec = samp_table['dev_reff'][cut]
+
+zmin = 1.9
+zmax = 2.1
+zgrid = np.linspace(zmin, zmax, 5)
+dd_grid = 0.*zgrid
+for i in range(5):
+    dd_grid[i] = Dang(zgrid[i])
+
+dd_spline = splrep(zgrid, dd_grid)
+
+reff_kpc = reff_arcsec * np.deg2rad(1./3600.) * splev(samp_table['z'][cut], dd_spline) * 1000.
+
+lreff = np.log10(reff_kpc)
+
+# I have no idea why I had to do this, but leastsq wouldn't work otherwise.
+lmstar0 = np.zeros(len(lmstar))
+lmstar0 += lmstar
+lmstar = lmstar0.copy()
+
+lsigma = np.zeros(len(lmstar))
+lsigma_err = np.zeros(len(lmstar))
+lsigma += np.log10(veldisp)
+lsigma_err += 0.5*(np.log10(veldisp+veldisp_err) - np.log10(veldisp-veldisp_err))
+
+def nlogpfunc(p):
+    lsigma_model = p[0] + p[1]*(lmstar - lmstar_piv) + p[2]*(lreff - lreff_piv)
+    logp_samp = -0.5*(lsigma - lsigma_model)**2/(lsigma_err**2 + p[3]**2) - 0.5*np.log((p[3]**2 + lsigma_err**2))
+    return -logp_samp.sum()
+
+p0 = [2.3, 0.2, -0.2, 0.1]
[…]

### 2022-07-24 paper/seleff.tex
constraining the scatter parameters

@@
 \begin{equation}
-\Phi(\mobs) = (\ln{10})\Phi^*\left[10^{(\mobs - \mstar^*)(1+\alpha)}\right]\times \exp{\left[-10^{(\mobs -\mstar^*)}\right]}.
+\Phi(\mobs) = \Phi^*\left(\frac{\mobs}{\mstar^*}\right)^{\alpha} \exp{\left\{-\frac{\mobs}{\mstar^*}\right\}}.
 \end{equation}
-We set $\Phi^*=1.009\times10^{-3}\,{\rm Mpc}^{-3}$, $\alpha=-0.92$ and $\log{\mstar^*}=11.21$. These are the best-fit values measured by \citet{Muz++13}.
+We set $\Phi^*=1.009\times10^{-3}\,{\rm Mpc}^{-3}$, $\alpha=-0.92$ and $\log{\mstar^*}=11.21$: these are the best-fit values measured by \citet{Muz++13}.
 
@@
 The model curve is a beta distribution, \Eref{eq:betaappendix}, with $\alpha=6.28$ and $\beta=2.05$.
+\label{fig:qhist}
 }
@@
 Then, we fitted the following model to the distribution of $\sigma_{\mathrm{ap}}$ of the sample:
-\begin{equation}
+\begin{equation}\label{eq:fpfull}
 \pr(\sigma_{\mathrm{ap}}) \sim \mathcal{N}(\mu_{\sigma} + \beta(\log{\mobs} - 11.4) + \xi(\log{\reff} - 1.2), \sigma_{\sigma}).
@@
 
-We then generated samples of galaxies from the model of section \ref{ssec:lenses}.
+We then generated samples of $z=0.2$ galaxies from the model of section \ref{ssec:lenses} and used the spherical Jeans equation to predict their central velocity dispersion.
+The spherical Jeans equation is \citep{B+T87}
+\begin{equation}\label{eq:jeans}
+\frac{d(\rho_*\sigma_r^2)}{dr} + \frac{\beta(r)}{r}\rho_*\sigma_r^2 = -\rho_*(r)\frac{GM(r)}{r^2},
+\end{equation}
+where $\rho_*(r)$ is the 3-dimensional distribution of dynamical tracers (i.e.
+the stars), $\sigma_r$ the radial component of the velocity dispersion, $\beta(r)$ the orbital anisotropy parameter, and $M(r)$ the mass enclosed within a spherical shell of radius $r$.
+We assumed isotropic orbits ($\beta=0$), then integrated \Eref{eq:jeans} to obtain the seeing-convolved, surface brightness-weighted line-of-sight velocity dispersion within the SDSS spectroscopic aperture\footnote{We used Python code available at \url{https://github.com/astrosonnen/spherical_jeans}, for this purpose.}.
+Finally, we fitted the fundamental plane relation of \Eref{eq:fpfull} to the observed stellar mass, half-light radius and central velocity dispersion of the mock sample.
+We repeated this procedure for different values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$, $\sigma_\mathrm{h}$ and $\sigma_\gamma$.
 
-We did not attempt to match the other parameters of the fundamental plane (i.e.
+%We did not attempt to match the other parameters of the fundamental plane (i.e.
 the constant of proportionality of \Eref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.
 
-%First, we took the sample of SDSS early-type galaxies introduced in Appendix~\ref{sect:appendixa}
-
 \end{document}

### 2022-07-24 lens_population/make_gnfwpar_grids.py

@@ -1,0 +1,123 @@
+import numpy as np
+from sl_profiles import nfw, gnfw, sersic, eaglecontr, deVaucouleurs as deV
+import sl_cosmology
+from scipy.interpolate import splrep, splev, splint, interp1d
+from scipy.optimize import minimize, leastsq
+from poppars import *
+import h5py
+
+
+# each lens consists of a de Vaucouleurs stellar component and a gNFW halo.
+# The parameters of the gNFW profile are defined by first using the Cautun 2020
+# contraction model, then fitting the resulting profile with a gNFW model.
+# It is time-consuming, therefore it's more convenient to do this on a grid
+# of values of stellar mass, halo mass and half-light radius.
+
+lmstar_min = 10.5
+lmstar_max = 12.5
+
+dlreff_min = -0.7
+dlreff_max = 0.7
+
+dlm200_min = -1.
+dlm200_max = 1.
+
+nmstar = 21
+nreff = 15
+nm200 = 21
+
+lmstar_grid = np.linspace(lmstar_min, lmstar_max, nmstar)
+dlreff_grid = np.linspace(dlreff_min, dlreff_max, nreff)
+dlm200_grid = np.linspace(dlm200_min, dlm200_max, nm200)
+
+gammadm_grid = np.zeros((nmstar, nreff, nm200))
+rs_grid = np.zeros((nmstar, nreff, nm200))
+
+redshift = 0.2
+rhoc = sl_cosmology.rhoc(redshift)
+
+nr3d = 1001
+r3d_scalefree = np.logspace(-3., 3., nr3d) # radial grid, from 1/100 to 100 times Reff
+deV_rho_scalefree = deV.rho(r3d_scalefree, 1.)
+deV_m3d_unitmass = deV.fast_M3d(r3d_scalefree)
+
+nR2d = 101
+R2d_scalefree = np.logspace(-3., 2., nR2d)
+
+R2d_fit_grid = np.logspace(0., np.log10(30.), nR2d) # radial grid over which to fit the gNFW model
+
+def get_halo_splines(m200, mstar, reff):
+    r200 = (m200*3./200./(4.*np.pi)/rhoc)**(1./3.) * 1000.
+    rs = r200/c200_0
+    r3d_grid = r3d_scalefree * reff
+    R2d_grid = R2d_scalefree * reff
+
+    xmin = 1.01*R2d_grid[0]
[…]

### 2022-07-24 lens_population/make_jeans_grids.py

@@ -1,0 +1,97 @@
+import numpy as np
+from spherical_jeans import sigma_model, tracer_profiles
+from spherical_jeans.mass_profiles import nfw
+from sl_cosmology import Mpc, Dang, G, c, M_Sun
+import sl_cosmology
+from sl_profiles import deVaucouleurs as deV, gnfw
+import h5py
+from poppars import *
+
+
+# calculates grids of sigma^2 for the stellar component
+# and the dark matter component.
+# The stellar component has only 1 non-linear dimension: the size.
+# Therefore, sigma^2 is computed on a 1D with unit mass.
+# The dark matter halo has three degrees of freedom.
+# Therefore it uses the same grid over which the contracted profile has
+# been computed.
+
+redshift = 0.2
+dd = Dang(redshift)
+rhoc = sl_cosmology.rhoc(redshift)
+arcsec2kpc = np.deg2rad(1./3600.) * dd * 1000.
+kpc = Mpc / 1000.
+
+aperture_kpc = 1.5 * arcsec2kpc
+seeing_kpc = aperture_kpc
+
+nreff = 16
+lreff_grid = np.linspace(0.5, 2., nreff)
+
+nr3d = 1001
+r3d_scalefree = np.logspace(-3., 3., nr3d) # radial grid, from 1/100 to 100 times Reff
+deV_rho_scalefree = deV.rho(r3d_scalefree, 1.)
+deV_m3d_unitmass = deV.fast_M3d(r3d_scalefree)
+
+s2_deV_grid = np.zeros(nreff)
+
+for i in range(nreff):
+    reff = 10.**lreff_grid[i]
+    s2_deV_grid[i] = sigma_model.sigma2((r3d_scalefree*reff, deV_m3d_unitmass), aperture_kpc, reff, tracer_profiles.deVaucouleurs, seeing=seeing_kpc)
+
+s2_deV_grid *= G*M_Sun/kpc/1e10
+
+deV_file = h5py.File('s2_deV_grid.hdf5', 'w')
+deV_file.create_dataset('lreff_grid', data=lreff_grid)
+deV_file.create_dataset('s2_grid', data=s2_deV_grid)
+
+# loads the gnfw parameters grid
+
+gnfwpar_file = h5py.File('gnfwpar_grid.hdf5', 'r')
+
+lmstar_grid = gnfwpar_file['lmstar_grid'][()]
[…]

### 2022-07-24 lens_population/match_fp_scatter.py

@@ -1,0 +1,87 @@
+import numpy as np
+from scipy.interpolate import splrep, splev, splint
+from scipy.optimize import minimize
+from poppars import *
+import pylab
+import h5py
+import ndinterp
+
+
+np.random.seed(0)
+
+def lmobsfunc(lmobs):
+    return 10.**((lmobs - lmstar_muz) * (alpha_muz + 1)) * np.exp(-10.**(lmobs - lmstar_muz))
+
+lmobs_min = 11.
+lmobs_max = 12.5
+
+nmobs = 101
+lmobs_grid = np.linspace(lmobs_min, lmobs_max, nmobs)
+
+lmobsfunc_spline = splrep(lmobs_grid, lmobsfunc(lmobs_grid))
+norm = splint(lmobs_min, lmobs_max, lmobsfunc_spline)
+
+cumfunc_grid = 0.*lmobs_grid
+for i in range(nmobs):
+    cumfunc_grid[i] = splint(lmobs_min, lmobs_grid[i], lmobsfunc_spline) / norm
+
+nsamp = 1000
+
+invcum_spline = splrep(cumfunc_grid, lmobs_grid)
+
+lmobs_samp = splev(np.random.rand(nsamp), invcum_spline)
+
+lreff_samp = lreff_mu + lreff_beta * (lmobs_samp - lmobs_piv) + np.random.normal(0., lreff_sig, nsamp)
+
+lasps_sig = 0.01
+lasps_samp = np.random.normal(lasps_mu, lasps_sig, nsamp)
+
+lmstar_samp = lmobs_samp + lasps_samp
+
+lm200_sig = 0.5
+lm200_samp = lm200_mu + lm200_beta * (lmstar_samp - lmstar_piv) + np.random.normal(0., lm200_sig, nsamp)
+
+lreff_mumstar_samp = lreff_mu + lreff_beta * (lmstar_samp - lasps_mu - lmobs_piv)
+dlreff_samp = lreff_samp - lreff_mumstar_samp
+
+dlm200_samp = lm200_samp - (lm200_mu + lm200_beta * (lmstar_samp - lmstar_piv))
+
+point = np.array([lmstar_samp, dlreff_samp, dlm200_samp]).T
+
+sigma_grid_file = h5py.File('deVgnfw_sigma_grid.hdf5', 'r')
+sigma_grid = sigma_grid_file['sigma_grid'][()]
+
[…]

### 2022-07-24 lens_population/poppars.py

@@ -1,0 +1,19 @@
+
+alpha_muz = -0.92
+lmstar_muz = 11.21
+
+lreff_mu = 1.20
+lreff_beta = 0.63
+lreff_sig = 0.14
+lreff_piv = 1.2
+
+lm200_mu = 13.
+lm200_beta = 1.
+
+c200_0 = 5.
+
+lasps_mu = 0.1
+
+lmobs_piv = 11.4
+lmstar_piv = 11.5
+

### 2022-07-25 paper/seleff.tex
decided scatter parameters

@@
 \end{equation}
-The existence of a fundamental plane relation is a consequence of the virial theorem: the velocity dispersion of a galaxy is directly related to the 3-dimensional mass distribution in its inner regions, which is tyically dominated by the stellar component.
+The existence of a fundamental plane relation is a consequence of the virial theorem: the velocity dispersion of a galaxy is directly related to the 3-dimensional mass distribution in its inner regions, which is typically dominated by the stellar component.
 At fixed observed stellar mass distribution, however, the central velocity dispersion can vary depending on the stellar population synthesis mismatch parameter, on the dark matter mass and on the dark matter density profile.
@@
 We used Jeans modelling for this purpose.
-We generated samples of early-type galaxies with the recipes described above and with different values of $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$.
+We generated samples of early-type galaxies with the recipes described above and with different values of $\sigma_{\mathrm{sps}}$, $\sigma_\mathrm{h}$ and $\sigma_\gamma$.
 Using the spherical Jeans equation under the assumption of isotropic orbits, we predicted the central velocity dispersion of each galaxy in the sample.
@@
 We then varied $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$ to match the observed and the predicted scatter.
-The details of this procedures are given in Appendix~\ref{sect:appendixb}.
+We did not attempt to match the other parameters of the fundamental plane (i.e.
+the constant of proportionality of \Eref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.
+The details of this procedure are given in Appendix~\ref{sect:appendixb}.
+
+%After comparing the predicted and observed fundamental plane scatter for different sets of values of $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$, we settled on three 
+%After comparing the predicted and observed fundmantal plane scatter, 
+We settled on three different set of intrinsic scatter parameters, as indicated in \Tref{tab:scatter}. We label them the fiducial, the low-scatter and the high-scatter scenarios.
+The high-scatter scenario is ruled out both by the fundamental plane and by weak lensing constraints \citep{Son++22}. 
+Moreover, our dynamical model is quite simplistic, as it neglects the effects of orbital anisotropy and departures from spherical symmetry, which are additional sources of scatter. 
+Nevertheless, we use it in our experiment in order to obtain a more conservative upper limit on the amplitude of the strong lensing bias.
+%A model with the same intrinsic scatter parameters and a more realistic 3-dimensional and orbital structure would likely predict a fundamental plane scatter that is at least as large as the observed one.
+\begin{table}
+\caption{Intrinsic scatter scenarios}
+\label{tab:scatter}
+\begin{tabular}{lccc}
+\hline
+\hline
+Model name & $\sigma_{\mathrm{sps}}$ & $\sigma_{\mathrm{h}}$ & Predicted FP \\
+ & & & scatter \\
+\hline
+Fiducial & 0.08 & 0.20 & 0.034 \\
+Low scatter & 0.05 & 0.10 & 0.021 \\
+High scatter & 0.10 & 0.30 & 0.433 \\
+\end{tabular}
+\end{table}
 

### 2022-07-25 lens_population/make_gnfwpar_grids.py

@@ -94,8 +94,8 @@ def gnfw_fit(halo_Sigma_spline, m200):
 
 for i in range(nmstar):
     mstar = 10.**lmstar_grid[i]
-    lreff_model = lreff_mu + lreff_beta * (lmstar_grid[i] - lasps_mu - lmobs_piv)
-    lm200_model = lm200_mu + lm200_beta * (lmstar_grid[i] - lmstar_piv)
+    lreff_model = mu_R + beta_R * (lmstar_grid[i] - mu_sps - lmobs_piv)
+    lm200_model = mu_h + beta_h * (lmstar_grid[i] - lmstar_piv)
 
     for j in range(nreff):
         print(i, j)

### 2022-07-25 lens_population/make_jeans_grids.py

@@ -65,8 +65,8 @@ s2_grid = np.zeros((nmstar, nreff, nm200))
 
 for i in range(nmstar):
     mstar = 10.**lmstar_grid[i]
-    lreff_model = lreff_mu + lreff_beta * (lmstar_grid[i] - lasps_mu - lmobs_piv)
-    lm200_model = lm200_mu + lm200_beta * (lmstar_grid[i] - lmstar_piv)
+    lreff_model = mu_R + beta_R * (lmstar_grid[i] - mu_sps - lmobs_piv)
+    lm200_model = mu_h + beta_h * (lmstar_grid[i] - lmstar_piv)
 
     for j in range(nreff):
         print(i, j)

### 2022-07-25 lens_population/poppars.py

@@ -2,17 +2,17 @@
 alpha_muz = -0.92
 lmstar_muz = 11.21
 
-lreff_mu = 1.20
-lreff_beta = 0.63
-lreff_sig = 0.14
+mu_R = 1.20
+beta_R = 0.63
+sigma_R = 0.14
 lreff_piv = 1.2
 
-lm200_mu = 13.
-lm200_beta = 1.
+mu_h = 13.
+beta_h = 1.
 
 c200_0 = 5.
 
-lasps_mu = 0.1
+mu_sps = 0.1
 
 lmobs_piv = 11.4
 lmstar_piv = 11.5

### 2022-07-25 lens_population/predict_fp_scatter.py

@@ -9,6 +9,9 @@ import ndinterp
 
 np.random.seed(0)
 
+sigma_h = 0.3
+sigma_sps = 0.1
+
 def lmobsfunc(lmobs):
     return 10.**((lmobs - lmstar_muz) * (alpha_muz + 1)) * np.exp(-10.**(lmobs - lmstar_muz))
 
@@ -31,20 +34,18 @@ invcum_spline = splrep(cumfunc_grid, lmobs_grid)
 
 lmobs_samp = splev(np.random.rand(nsamp), invcum_spline)
 
-lreff_samp = lreff_mu + lreff_beta * (lmobs_samp - lmobs_piv) + np.random.normal(0., lreff_sig, nsamp)
+lreff_samp = mu_R + beta_R * (lmobs_samp - lmobs_piv) + np.random.normal(0., sigma_R, nsamp)
 
-lasps_sig = 0.01
-lasps_samp = np.random.normal(lasps_mu, lasps_sig, nsamp)
+lasps_samp = np.random.normal(mu_sps, sigma_sps, nsamp)
 
 lmstar_samp = lmobs_samp + lasps_samp
 
-lm200_sig = 0.5
-lm200_samp = lm200_mu + lm200_beta * (lmstar_samp - lmstar_piv) + np.random.normal(0., lm200_sig, nsamp)
+lm200_samp = mu_h + mu_h * (lmstar_samp - lmstar_piv) + np.random.normal(0., sigma_h, nsamp)
 
-lreff_mumstar_samp = lreff_mu + lreff_beta * (lmstar_samp - lasps_mu - lmobs_piv)
-dlreff_samp = lreff_samp - lreff_mumstar_samp
+mu_Rmstar_samp = mu_R + beta_R * (lmstar_samp - mu_sps - lmobs_piv)
+dlreff_samp = lreff_samp - mu_Rmstar_samp
 
-dlm200_samp = lm200_samp - (lm200_mu + lm200_beta * (lmstar_samp - lmstar_piv))
+dlm200_samp = lm200_samp - (mu_h + mu_h * (lmstar_samp - lmstar_piv))
 
 point = np.array([lmstar_samp, dlreff_samp, dlm200_samp]).T
 

### 2022-07-25 paper/seleff.tex
lens mock code done

@@
 \def\asps{\alpha_{\mathrm{sps}}}
+
+\def\zsource{z_{\mathrm{s}}}
+\def\zlens{z_{\mathrm{g}}}
 
@@
 
-\subsubsection{Stellar mass and redshift distribution}
+\subsubsection{Stellar mass and redshift distribution}\label{ssub:mstarz}
 
-We generated lens galaxies over a finite redshift range, $0.1 < z < 0.9$.
+We generated lens galaxies over a finite redshift range, $0.1 < z < 0.7$.
 We chose these lower and upper limits because the value of the critical surface mass density $\Sigma_{\mathrm{cs}}$ becomes very large outside of this range, and the fraction of galaxies that act as strong lenses drops substantially as a result.
@@
 \end{tabular}
+\notes{
+Adopted values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$ and $\sigma_{\mathrm{h}}$ in different simulations.
+For each set of values, the fourth column indicates the scatter around the fundamental plane predicted via the spherical Jeans analysis of Appendix~\ref{sect:appendixb}.
+The observed fundamental plane scatter is $0.035$.
+}
 \end{table}
 
-\subsection{Background sources}\label{ssec:sources}
+\subsection{Background sources: galaxies}\label{ssec:sources}
 
-\subsection{Observations}\label{ssec:obs}
+\ale{To be filled by Shun-Sheng}.
+We applied a lower limit to the source redshift, by selecting only sources with $z_s > 0.8$.
+This ensures that all of the sources lie behind all of the lenses.
+A similar cut could be applied to a real survey using photometric redshift, to reduce the incidence of false positives (e.g.
+arc-like features physically associated with the lens galaxy) in the lens finding phase.
+%This selection is meant to simulate a cut on the redshift (possibly photometric) of the arcs, which would help reduce the incidence of false positives during the lens finding phase.
+The resulting number density of sources is ZZZ~arcmin$^{-2}$.
+We approximated their spatial distribution as uniform in the sky, that is we neglected clustering of the sources.
+
+\subsection{Background sources: quasars}\label{ssec:quasars}
+
+\subsection{Observations}\label{ssub:obs}
+
+For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering XXX square degrees.
+The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is YYY.
+%Then we generated a population of sources behind the lenses.
+%We approximated the foreground galaxies as isolated: for each source, we only modelled the contribution to the lensing field generated by a single lens.
+We approximated the foreground galaxies as isolated: when determining whether a galaxy acts as a strong lens, we only modelled the contribution to the lensing signal from the galaxy itself, and neglected that of the environment.
+We discuss the possible implications of this approximation in \Sref{sect:discuss}.
+For each lens, we determined its caustics relative to the highest source redshift, $\zsource=2.5$.
+We then placed sources randomly behind the lens.
+If at least one source fell within a circular region enclosing the caustics, we proceeded to compute its lensed images.
+
+For the extended sources, we produced images with a pixel size of $0.1''$.
+We applied a Gaussian PSF with a FWHM of $0.2''$.
+Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
+%This is similar to the background noise expected for the Euclid Wide survey \citep{Sca++22}.
+These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
+
+We applied the lens selection criterion of section \ref{ssec:}
 
@@
 Finally, we fitted the fundamental plane relation of \Eref{eq:fpfull} to the observed stellar mass, half-light radius and central velocity dispersion of the mock sample.
-We repeated this procedure for different values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$, $\sigma_\mathrm{h}$ and $\sigma_\gamma$.
+We repeated this procedure for different values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$, $\sigma_\mathrm{h}$ and $\sigma_\gamma$, then settled on the three scenarios listed in \Tref{tab:scatter}.
 

### 2022-07-25 lens_population/make_gnfwpar_grids.py

@@ -25,16 +25,15 @@ dlm200_max = 1.
 nmstar = 21
 nreff = 15
 nm200 = 21
+nz = 7
 
 lmstar_grid = np.linspace(lmstar_min, lmstar_max, nmstar)
 dlreff_grid = np.linspace(dlreff_min, dlreff_max, nreff)
 dlm200_grid = np.linspace(dlm200_min, dlm200_max, nm200)
+z_grid = np.linspace(zmin, zmax, nz)
 
-gammadm_grid = np.zeros((nmstar, nreff, nm200))
-rs_grid = np.zeros((nmstar, nreff, nm200))
-
-redshift = 0.2
-rhoc = sl_cosmology.rhoc(redshift)
+gammadm_grid = np.zeros((nmstar, nreff, nm200, nz))
+rs_grid = np.zeros((nmstar, nreff, nm200, nz))
 
 nr3d = 1001
 r3d_scalefree = np.logspace(-3., 3., nr3d) # radial grid, from 1/100 to 100 times Reff
@@ -46,7 +45,7 @@ R2d_scalefree = np.logspace(-3., 2., nR2d)
 
 R2d_fit_grid = np.logspace(0., np.log10(30.), nR2d) # radial grid over which to fit the gNFW model
 
-def get_halo_splines(m200, mstar, reff):
+def get_halo_splines(m200, mstar, reff, rhoc):
     r200 = (m200*3./200./(4.*np.pi)/rhoc)**(1./3.) * 1000.
     rs = r200/c200_0
     r3d_grid = r3d_scalefree * reff
@@ -73,7 +72,7 @@ def get_halo_splines(m200, mstar, reff):
 
     return contr_Sigma_spline, contr_SigmaR_spline, R2d_grid
 
-def gnfw_fit(halo_Sigma_spline, m200):
+def gnfw_fit(halo_Sigma_spline, m200, rhoc):
 
     r200_here = (m200*3./200./(4.*np.pi)/rhoc)**(1./3.) * 1000.
 
@@ -105,18 +104,21 @@ for i in range(nmstar):
         for k in range(nm200):
             m200 = 10.**(lm200_model + dlm200_grid[k])
 
-            halo_Sigma_spline, halo_SigmaR_spline, R2d_grid = get_halo_splines(m200, mstar, reff)
+            for l in range(nz):
+                rhoc = sl_cosmology.rhoc(z_grid[l])
[…]

### 2022-07-25 lens_population/mcmc_mass_size.py

@@ -1,89 +1,0 @@
-import numpy as np
-import emcee
-from astropy.io import fits as pyfits
-from sl_cosmology import Dang
-from scipy.interpolate import splrep, splev
-from scipy.stats import truncnorm
-import pylab
-
-
-samp_table = pyfits.open('sdss_zbin_latetype_red.fits')[1].data
-
-cut = (samp_table['logmstar_dev'] > 11.) & (samp_table['logmstar_dev'] < 12.2)
-
-lmstar = samp_table['logmstar_dev'][cut]
-lmstar_piv = 11.4
-
-reff_arcsec = samp_table['dev_reff'][cut]
-
-zmin = 1.9
-zmax = 2.1
-zgrid = np.linspace(zmin, zmax, 5)
-dd_grid = 0.*zgrid
-for i in range(5):
-    dd_grid[i] = Dang(zgrid[i])
-
-dd_spline = splrep(zgrid, dd_grid)
-
-reff_kpc = reff_arcsec * np.deg2rad(1./3600.) * splev(samp_table['z'][cut], dd_spline) * 1000.
-
-lreff = np.log10(reff_kpc)
-
-# I have no idea why I had to do this, but leastsq wouldn't work otherwise.
-lmstar0 = np.zeros(len(lmstar))
-lmstar0 += lmstar
-lmstar = lmstar0.copy()
-
-mu_par = {'name': 'mu', 'lower': -3., 'upper': 3., 'guess': 1., 'spread': 0.1}
-beta_par = {'name': 'beta', 'lower': -3., 'upper': 3., 'guess': 1., 'spread': 0.1}
-sig_par = {'name': 'sig', 'lower': 0., 'upper': 1., 'guess': 0.2, 'spread': 0.1}
-
-pars = [mu_par, beta_par, sig_par]
-npars = len(pars)
-
-bounds = []
-for par in pars:
-    bounds.append((par['lower'], par['upper']))
-
-def logprior(p):
-    for i in range(npars):
-        if p[i] < bounds[i][0] or p[i] > bounds[i][1]:
-            return -1e300
-    return 0.
-
-def logpfunc(p):
-
-    lprior = logprior(p)
-    if lprior < 0.:
-        return -1e300
-
-    mu, beta, sig = p
-
-    lreff_model = mu + beta * (lmstar - lmstar_piv)
-    
-    sumlogp = (-0.5*(lreff_model - lreff)**2/sig**2 - np.log(sig)).sum()
-
[…]

### 2022-07-25 lens_population/poppars.py

@@ -1,19 +1,75 @@
+import numpy as np
+from scipy.interpolate import splrep, splev, splint
+import sl_cosmology
 
+
+# stellar mass function
+phi_muz = 1.009e-3
 alpha_muz = -0.92
 lmstar_muz = 11.21
 
+# lens redshift range
+zmin = 0.1
+zmax = 0.7
+
+# mass-size relation
 mu_R = 1.20
 beta_R = 0.63
 sigma_R = 0.14
 lreff_piv = 1.2
 
+# halo mass distribution
 mu_h = 13.
 beta_h = 1.
 
+# initial concentration before contraction
 c200_0 = 5.
 
+# stellar population synthesis mismatch parameter (log of the mean)
 mu_sps = 0.1
 
 lmobs_piv = 11.4
 lmstar_piv = 11.5
+lmobs_min = 11.
+lmobs_max = 12.5
+
+# axis ratio distribution
+alpha_q = 6.28
+beta_q = 2.05
+
+fourpi_volume = 4*np.pi/3. * (sl_cosmology.comovd(zmax)**3 - sl_cosmology.comovd(zmin)**3)
+
+# observed stellar mass function
+def lmobsfunc(lmobs):
+    return phi_muz * 10.**((lmobs - lmstar_muz) * (alpha_muz + 1)) * np.exp(-10.**(lmobs - lmstar_muz))
+
+nmobs = 101
+lmobs_grid = np.linspace(lmobs_min, lmobs_max, nmobs)
+
+lmobsfunc_spline = splrep(lmobs_grid, lmobsfunc(lmobs_grid))
+# number of galaxies within a cubic comoving Mpc
+ngal_1Mpc3 = splint(lmobs_min, lmobs_max, lmobsfunc_spline)
+
+cumfunc_lmobs_grid = 0.*lmobs_grid
+for i in range(nmobs):
+    cumfunc_lmobs_grid[i] = splint(lmobs_min, lmobs_grid[i], lmobsfunc_spline) / ngal_1Mpc3
+
+invcum_lmobs_spline = splrep(cumfunc_lmobs_grid, lmobs_grid)
+
+# defines the redshift distribution function
+nz = 71
+zgrid = np.linspace(zmin, zmax, nz)
+zfunc_grid = 0.*zgrid
+for i in range(nz):
[…]

### 2022-07-25 lens_population/predict_fp_scatter.py

@@ -9,14 +9,11 @@ import ndinterp
 
 np.random.seed(0)
 
-sigma_h = 0.3
-sigma_sps = 0.1
+sigma_h = 0.2
+sigma_sps = 0.08
 
 def lmobsfunc(lmobs):
     return 10.**((lmobs - lmstar_muz) * (alpha_muz + 1)) * np.exp(-10.**(lmobs - lmstar_muz))
-
-lmobs_min = 11.
-lmobs_max = 12.5
 
 nmobs = 101
 lmobs_grid = np.linspace(lmobs_min, lmobs_max, nmobs)
@@ -40,12 +37,11 @@ lasps_samp = np.random.normal(mu_sps, sigma_sps, nsamp)
 
 lmstar_samp = lmobs_samp + lasps_samp
 
-lm200_samp = mu_h + mu_h * (lmstar_samp - lmstar_piv) + np.random.normal(0., sigma_h, nsamp)
+dlm200_samp = np.random.normal(0., sigma_h, nsamp)
+lm200_samp = mu_h + beta_h * (lmstar_samp - lmstar_piv) + dlm200_samp
 
 mu_Rmstar_samp = mu_R + beta_R * (lmstar_samp - mu_sps - lmobs_piv)
 dlreff_samp = lreff_samp - mu_Rmstar_samp
-
-dlm200_samp = lm200_samp - (mu_h + mu_h * (lmstar_samp - lmstar_piv))
 
 point = np.array([lmstar_samp, dlreff_samp, dlm200_samp]).T
 

### 2022-07-26 lens_population/assign_sources.py
mocks

@@ -1,0 +1,51 @@
+import numpy as np
+from scipy.stats import poisson
+from poppars import *
+import h5py
+from astropy.io import fits as pyfits
+
+
+np.random.seed(10)
+# places sources in a circle enclosing the radial caustic
+
+circ_caust_rat = 1.2 # ratio between circle radius and caustic radius
+
+modelname = 'fiducial_100sqdeg'
+pop = h5py.File('%s_lenses.hdf5'%modelname, 'r')
+
+nsamp = pop.attrs['nsamp']
+
+sourcecat = pyfits.open('/Users/alessandro/catalogs/skills_sourceonly_zcut.fits')[1].data
+
+nsource_tot = len(sourcecat)
+
+sourceind = np.arange(nsource_tot)
+# shuffles source catalog (it's originally ranked by redshift)
+np.random.shuffle(sourceind)
+
+outlines = []
+outlines.append('# lens_id rcirc nsources source_indices(list)\n')
+
+sourcecount = 0
+for i in range(nsamp):
+    rcirc = circ_caust_rat*pop['tcaust'][i]/pop['q'][i]**0.5
+
+    area = np.pi*rcirc**2
+
+    lam = area * nbkg / 3600. # expectation value of the number of sources in the circle
+
+    nsources = poisson.rvs(lam)
+
+    line = '%d %f %d'%(i, rcirc, nsources)
+    for n in range(nsources):
+        line += ' %d'%sourceind[sourcecount]
+        sourcecount += 1
+    line += '\n'
+
+    outlines.append(line)
+
+print('Total number of sources: %d'%sourcecount)
+f = open('%s_sources.cat'%modelname, 'w')
+f.writelines(outlines)
+f.close()
+

### 2022-07-26 lens_population/generate_lenses.py

@@ -1,0 +1,246 @@
+import numpy as np
+from scipy.interpolate import splrep, splev, splint, interp1d
+from scipy.stats import poisson, beta
+from scipy.optimize import leastsq, brentq
+import sl_cosmology
+from sl_profiles import deVaucouleurs as deV, gnfw, nfw, eaglecontr
+import h5py
+import ndinterp
+from poppars import *
+
+
+seedno = 0
+np.random.seed(seedno)
+
+sky_area_deg2 = 100.
+sky_area = sky_area_deg2 * np.deg2rad(1.)**2
+
+modelname = 'fiducial_%dsqdeg'%sky_area_deg2
+sigma_h = 0.2
+sigma_sps = 0.08
+
+print('Sky area: %d square degrees'%sky_area_deg2)
+print('Expected number of foreground galaxies: %2.1e'%(ngal_1Mpc3 * fourpi_volume * sky_area / (4.*np.pi)))
+# draws the number of galaxies
+nsamp = poisson.rvs(ngal_1Mpc3 * fourpi_volume * sky_area / (4.*np.pi), 1)
+print('%d galaxies drawn'%nsamp)
+
+lmobs_samp = splev(np.random.rand(nsamp), invcum_lmobs_spline)
+z_samp = splev(np.random.rand(nsamp), invcum_z_spline)
+
+rhoc_samp = sl_cosmology.rhoc(z_samp)
+
+lasps_samp = np.random.normal(mu_sps, sigma_sps, nsamp)
+lmstar_samp = lmobs_samp + lasps_samp
+
+lreff_samp = mu_R + beta_R * (lmobs_samp - lmobs_piv) + np.random.normal(0., sigma_R, nsamp)
+
+dlreff_samp = lreff_samp - (mu_R + beta_R * (lmstar_samp - mu_sps - lmobs_piv))
+
+dlm200_samp = np.random.normal(0., sigma_h, nsamp)
+lm200_samp = mu_h + beta_h * (lmstar_samp - lmstar_piv) + dlm200_samp
+r200_samp = (10.**lm200_samp*3./200./(4.*np.pi)/rhoc_samp)**(1./3.) * 1000.
+
+q_samp = beta.rvs(alpha_q, beta_q, size=nsamp)
+
+# uses the grid built previously to get the parameters of the gnfw profile
+
+# loads the gnfw parameters grid
[…]

### 2022-07-26 lens_population/lens_extended_sources.py

@@ -1,0 +1,62 @@
+import numpy as np
+import os
+import glafic
+import h5py
+from astropy.io import fits as pyfits
+
+
+modelname = 'fiducial_100sqdeg'
+sourcecat = pyfits.open('/Users/alessandro/catalogs/skills_sourceonly_zcut.fits')[1].data
+pop = h5py.File('%s_lenses.hdf5'%modelname, 'r')
+
+f = open('%s_sources.cat'%modelname, 'r')
+sourcelines = f.readlines()
+f.close()
+
+np.random.seed(20)
+
+# primary parameters
+omegaM = 0.3
+omegaL = 0.7
+weos = -1.
+hubble = 0.7
+prefix = 'out'
+xmin = -3.
+ymin = -3.
+xmax = 3.
+ymax = 3.
+pix_ext = 0.1
+pix_poi = 0.1
+maxlev = 5
+
+glafic.init(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0)
+glafic.set_secondary('flag_halodensity 2')
+glafic.set_secondary('nfw_users 1')
+glafic.set_secondary('halodensity 200')
+
+# sets parameters of the power-law lens
+rein = 1.
+gamma0 = 2.
+q = 1.
+e = 1. - q
+zd = 0.3
+zs = 1.5
+
+glafic.startup_setnum(2, 1, 0)
+glafic.set_lens(1, 'gnfw', zd, 1e13, 0.0, 0.0, e, 90.0, 10., 1.5)
+glafic.set_lens(2, 'sers', zd, 1e11, 0.0, 0.0, e, 90.0, 1., 4.)
+glafic.set_extend(1, 'sersic', 1.5, 0.5, 0.3, 0., 0., 0., 0.1, 1.)
+
+#glafic.set_primary('xmax 5.')
+#glafic.set_primary('xmin -5.')
+glafic.set_primary(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0)
+
+# model_init needs to be done again whenever model parameters are changed
+glafic.model_init(verb = 0)
+
+a = glafic.writeimage()
+print(np.array(a).shape)
+
+pyfits.PrimaryHDU(a).writeto('test.fits', overwrite=True)
+
+

### 2022-07-26 lens_population/poppars.py

@@ -1,7 +1,11 @@
 import numpy as np
 from scipy.interpolate import splrep, splev, splint
 import sl_cosmology
+from sl_cosmology import Mpc, M_Sun, c, G
 
+
+# number density of background sources
+nbkg = 70. # per square arcminute
 
 # stellar mass function
 phi_muz = 1.009e-3
@@ -72,4 +76,24 @@ for i in range(nz):
 
 invcum_z_spline = splrep(cumfunc_z_grid, zgrid)
 
+# makes a spline for the angular diameter distance
+dd_grid = 0.*zgrid
+for i in range(nz):
+    dd_grid[i] = sl_cosmology.Dang(zgrid[i])
+
+dd_spline = splrep(zgrid, dd_grid)
+
+# and one for the critical surface mass density at the highest source redshift
+zs_ref = 2.5
+
+ds_ref = sl_cosmology.Dang(zs_ref)
+
+dds_grid = 0.*zgrid
+for i in range(nz):
+    dds_grid[i] = sl_cosmology.Dang(zgrid[i], zs_ref)
+
+kpc = Mpc/1000.
+s_cr_grid = c**2/(4.*np.pi*G)*ds_ref/dds_grid/dd_grid/Mpc/M_Sun*kpc**2 # units: M_Sun/kpc**2
+
+s_cr_spline = splrep(zgrid, s_cr_grid)
 

### 2022-07-31 paper/seleff.tex
mocks
experimenting with different lens definitions

@@
 \def\zlens{z_{\mathrm{g}}}
+
+\def\msource{m_{\mathrm{s}}}
 
@@
 
-\begin{figure*}
-\includegraphics[width=\textwidth]{footprints.eps}
+%\begin{figure*}
+%\includegraphics[width=\textwidth]{footprints.eps}
+%\caption{
+%Criterion used to classify lensed images of extended sources. Four examples. In each panel, the blue regions are the footprint of the lensed source. 
+%The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
+%\label{fig:lensdef}
+%}
+%\end{figure*}
+\begin{figure}
+\includegraphics[width=\columnwidth]{example_detections.eps}
 \caption{
-Criterion used to classify lensed images of extended sources. Four examples. In each panel, the blue regions are the footprint of the lensed source. 
-The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
+Criterion used to classify lensed images of extended sources. Four examples.
 \label{fig:lensdef}
 }
-\end{figure*}
+\end{figure}
+
 %__________________________________________________________________
@@
 \end{tabular}
-\notes{
+\tablefoot{
 Adopted values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$ and $\sigma_{\mathrm{h}}$ in different simulations.
@@
 
-\ale{To be filled by Shun-Sheng}.
+%\ale{To be filled by Shun-Sheng}.
 We applied a lower limit to the source redshift, by selecting only sources with $z_s > 0.8$.
@@
 %This selection is meant to simulate a cut on the redshift (possibly photometric) of the arcs, which would help reduce the incidence of false positives during the lens finding phase.
-The resulting number density of sources is ZZZ~arcmin$^{-2}$.
+The resulting number density of sources is $70$~arcmin$^{-2}$.
 We approximated their spatial distribution as uniform in the sky, that is we neglected clustering of the sources.

### 2022-07-31 single_lens/elliptical_ext/get_crosssect.py

@@ -1,20 +1,27 @@
 import numpy as np
 from astropy.io import fits as pyfits
 from skimage import measure
+from scipy.special import gamma as gfunc
+from scipy.optimize import brentq
+from sl_profiles import sersic
+from lensdet import detect_lens
 import h5py
 
 
 ftot = 200.
-nre = 11
-logre_grid = np.linspace(-1., 0., nre)
+nre = 21
+logre_grid = np.linspace(-1., 1., nre)
 
 nsim = 10000
 
 rsim = 1.5
 source_area = np.pi * rsim**2
 
-nfrat = 4
-lfrat_grid = np.linspace(0.8, 2., nfrat)
+pix = 0.05
+nser = 1.
+
+nfrat = 5
+lfrat_grid = np.linspace(1., 3., nfrat)
 # frat is the ratio between the source intrinsic flux and the sky rms over an area equal to the square of the Einstein radius.
 # In my sims, the Einstein radius is 1 arcsec. Pixels have a 0.05" size.
 # Then, the sky rms over a 1 square arcsec is 20 times that over a single pixel
@@ -22,6 +29,8 @@ lfrat_grid = np.linspace(0.8, 2., nfrat)
 # corresponds to a sky rms per pixel of 1.
 
 nsigma = 2.
+
+minmag = 3.
 
 x0 = 39.5
 y0 = 39.5
@@ -33,6 +42,7 @@ cs_grid = np.zeros((nfrat, nre))
 for l in range(nfrat):
     lfrat = lfrat_grid[l]
     sky_rms = 10.**(1. - lfrat) 
+    sb_min = nsigma * sky_rms
 
     for m in range(nre):
         print(l, m)
@@ -42,56 +52,15 @@ for l in range(nfrat):
         nlens = 0
         islens_sim = np.zeros(nsim, dtype=bool)
         nimages_sim = np.zeros(nsim, dtype=int)
-        aperture_sim = np.zeros(nsim)
+
+        img_file = h5py.File('mockdir/logre%2.1f_images.hdf5'%logre, 'r')
 
         for i in range(nsim):
-            img = pyfits.open('mockdir/'+preamble+'/'+preamble+'_%04d_image.fits'%i)[0].data
-            footprint = img > nsigma * sky_rms
+        
+            img = img_file['lens_%04d'%i][()]
 
-            labels = measure.label(footprint)
-            nreg = labels.max()
-            nimg = 0
-            for n in range(nreg):
-                npix_here = (labels==n+1).sum()
[…]

### 2022-07-31 single_lens/elliptical_ext/lensdet.py

@@ -1,0 +1,82 @@
+import numpy as np
+from skimage import measure
+
+
+def detect_lens(img, sky_rms, nsigma_pixdet=2., npix_min=10):
+
+    sb_min = nsigma_pixdet * sky_rms
+
+    ny, nx = img.shape
+    x0 = nx/2. - 0.5
+    y0 = ny/2. - 0.5
+
+    X, Y = np.meshgrid(np.arange(nx), np.arange(ny))
+
+    # counts the number of images with the minimum threshold (2 sigma)
+    footprint = img > sb_min
+
+    img_detected = img.copy()
+
+    labels = measure.label(footprint)
+    nreg = labels.max()
+    nimg_std = 0
+    for n in range(nreg):
+        npix_tmp = (labels==n+1).sum()
+        signal = img[labels==n+1].sum()
+        noise = npix_tmp**0.5 * sky_rms
+        img_sn = signal/noise
+        if img_sn >= 10. and npix_tmp >= npix_min:
+            nimg_std += 1
+        else:
+            img_detected[labels==n+1] = 0.
+
+    std_footprint = img_detected > sb_min
+
+    sorted_img = img.flatten().copy()
+    sorted_img.sort()
+    sb_det = sorted_img[sorted_img > sb_min]
+    nup = len(sb_det)
+
+    i = 0
+    nimg_max = nimg_std
+    nimg_tmp = nimg_std
+    sb_maxlim = sb_min
+
+    best_footprint = std_footprint.copy()
+
+    # if there's at least one detected image, increases the threshold to see
+    # if more can be detected
+
+    while nimg_std > 0 and nimg_tmp >= nimg_max and i < nup:
+        sb_lim = sb_det[i]
+
+        img_detected = img.copy()
+        footprint_tmp = img > sb_lim
+
+        labels = measure.label(footprint_tmp)
+        nreg = labels.max()
+        nimg_tmp = 0
+        for n in range(nreg):
+            npix_tmp = (labels==n+1).sum()
+            signal = img[labels==n+1].sum()
+            noise = npix_tmp**0.5 * sky_rms
+            img_sn = signal/noise
+            if img_sn >= 10. and npix_tmp >= npix_min:
+                nimg_tmp += 1
+            else:
+                img_detected[labels==n+1] = 0.
+
+        if nimg_tmp > nimg_max:
+            nimg_max = nimg_tmp
+            sb_maxlim = sb_lim
[…]

### 2022-07-31 single_lens/elliptical_ext/plot_cs_vs_re.py

@@ -45,38 +45,12 @@ for i in range(nfrat):
     ax.loglog(10.**logre_grid, cs_grid[i, :]/np.pi, linewidth=2, color=colseq[i], label="$\log{f} = %2.1f$"%lfrat_grid[i])
     cs_spline = splrep(10.**logre_grid, cs_grid[i, :]/np.pi, k=1)
 
-# lines of constant surface brightness (who cares?)
-"""
-nfr2 = nfrat
-cs_fixedfr2 = np.zeros((nfr2, nfrat))
-re_fixedfr2 = np.zeros((nfr2, nfrat))
-
-for i in range(nfr2):
-    for j in range(nfrat):
-        if i+j>1:
-            re_fixedfr2[i, j] = 10.**logre_grid[i+j-2]
-            cs_fixedfr2[i, j] = cs_grid[j, i+j-2]
-
-for i in range(nfr2):
-    if i<2:
-        istart = 2-i
-        iend = 4
-    else:
-        istart = 0
-        iend = 4
-    ax.loglog(re_fixedfr2[i, istart:iend], cs_fixedfr2[i, istart:iend]/np.pi, color='grey', linewidth=0.5)
-"""
-
-#lines = labelLines(ax.get_lines()[2:], xvals = [0.65, 0.9, 0.58], fontsize=fsize, backgroundcolor='white')
-
-#ax.xaxis.set_major_locator(MultipleLocator(0.5))
-#ax.xaxis.set_minor_locator(MultipleLocator(0.1))
-
 ax.tick_params(axis='both', which='both', top=True, right=True, labelsize=fsize, direction='in')
 
 ax.set_xlabel('$\\theta_{\mathrm{e,s}}/\\theta_{\mathrm{Ein}}$', fontsize=fsize)
 ax.set_ylabel('$\sigma_{\mathrm{SL}}/(\pi\\theta_{\mathrm{Ein}}^2)$', fontsize=fsize)
 
+"""
 # now finds the detection limit for non-lensed galaxies
 nre = 11
 logre_grid = np.linspace(-1., 0., nre)
@@ -108,6 +82,7 @@ for l in range(nfrat):
     logre_crit = splev(10., sn_spline)
 
     ax.axvline(10.**logre_crit, linestyle='--', color=colseq[l])
+"""
 
 #ax.set_xlim(xlim[0], xlim[1])
 ax.set_ylim(ylim[0], ylim[1])
[…]

### 2022-07-31 single_lens/elliptical_ext/plot_example_detections.py

@@ -1,0 +1,156 @@
+import numpy as np
+import glafic
+from sl_profiles import sersic
+from scipy.special import gamma as gfunc
+import pylab
+from astropy.io import fits as pyfits
+from lensdet import detect_lens
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 24
+
+nexamples = 6
+
+"""
+re_list = [0.2, 0.15, 0.3, 0.4]
+srcpos_list = [(0.5, 0.), (0.75, 0.), (0.2, 0.1), (0.3, 0.25)]
+q_list = [0.7, 0.3, 0.7, 0.7]
+zs_list = [1.5, 1.5, 0.5, 0.35]
+"""
+
+re_list = [0.4, 0.4, 0.4, 0.4, 0.4, 0.4]
+srcpos_list = [(0.05, 0.), (0.1, 0.), (0.15, 0.), (0.2, 0.), (0.25, 0.), (0.3, 0.)]
+q_list = [0.7, 0.7, 0.7, 0.7, 0.7, 0.7]
+zs_list = [0.35, 0.35, 0.35, 0.35, 0.35, 0.35]
+
+# creates a custom colormap
+cm_vals = np.ones((3, 4))
+cm_vals[0, :] = np.array([1., 1., 1., 1.])
+cm_vals[1, :] = np.array([0.5, 0., 1., 1.])
+cm_vals[2, :] = np.array([1., 0., 0.5, 1.])
+my_cmap = ListedColormap(cm_vals)
+
+pix = 0.05
+sky_rms = 1. # noise per pixel
+nser = 1. # source Sersic index
+
+# primary parameters
+omegaM = 0.3
+omegaL = 0.7
+weos = -1.
+hubble = 0.7
+prefix = 'tmp'
+xmin = -2.
+ymin = -2.
+xmax = 2.
+ymax = 2.
+pix_ext = pix
+pix_poi = 0.1
+maxlev = 5
+
+glafic.init(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0)
+glafic.set_secondary('flag_halodensity 2')
+glafic.set_secondary('nfw_users 1')
+glafic.set_secondary('halodensity 200')
+
+glafic.startup_setnum(2, 1, 0)
[…]

### 2022-08-02 paper/seleff.tex
new lens definition

@@
 
-\Fref{fig:ellcaust} shows the source-plane caustics of elliptical lenses with different values of the axis ratio.
+\Fref{fig:ellcaust} shows the source-plane caustics of elliptical lenses with different values of the axis ratio. We used the software {\sc Glafic} \citep{Ogu21} to obtain the caustic curves.
 The outermost curves are radial caustics, while the inner ones are tangential ones.
@@
 This is because, when the source is extended, some of the images can be blended together.
-In real strong lensing surveys, it is common to find lens candidates in which the lensed source consists only a single arc.
+In real strong lensing surveys, it is common to find lens candidates in which the lensed source consists of only a single arc.
 In those cases it is difficult to establish whether the arc is a set of blending images or not, and the decision of including such systems in a strong lens sample is often arbitrary.
-%For instance, a source located on one of the cusps of a tangential caustic produces a merging set of three magnified images.
-If the source is extended and no counter-image is detected, however, the appearence of the merging triplet is very similar to that of a single arc with no multiple images, and it can be difficult to differentiate between the two cases.
-%In real strong lensing surveys, the decision of including systems with a single arc or not is often arbitrary.
-
-For the purpose of carrying out our experiment we adopt the following definition.
-First, we determine the footprint of the lensed source.
-We define this as the ensemble of pixels where the source is detected with $S/N>2$.
+Here we adopt the following definition: an extended source is strongly lensed if the number of surface brightness peaks that are detected is larger than its intrinsic number of peaks in the absence of lensing.
+
+We explain how this definition applies in practice with a few examples.
+%We can consider a few examples.
+For simplicity, we focus on the case of a source with a single surface brightness peak, such as a S\'{e}rsic profile.
+All of the sources considered in this work belong to this family of surface brightness models.
+We adopt the following procedure to determine the number of detected peaks.
+Given the observed image of a lensed source, we define its footprint as the ensemble of pixels where the source is detected with $S/N>2$.
 The footprint is in general composed of multiple disconnected regions, corresponding to the different images.
@@
 If, after applying this cut, the source footprint is spread over multiple separated regions, then the system is classified as a lens.
-Additionally, if the source footprint consists of only a single region, we still classify it as strongly lensed if it subtends an angle larger than $90^\circ$ measured from the lens centre.
-This last criterion allows us to include highly magnified arcs and full rings.
-
-%Additionally, we classify as strongly lensed the following set of sources any full rings, and any single arcs that subtend an angle larger than $90^\circ$ measured from the lens centre.
-This last criterion is a
-%Additionally, we classify as a strongly lensed source any full rings, and any single arcs that subtend an angle larger than $90^\circ$ measured from the lens centre. This last criterion is a
-
-%Given the source footprint thus defined, we classify the system as a strong lens if the footprint subtends an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
-%Given the source footprint thus defined, we classify the system as a strong lens if we can find two points on the footprint that subtend an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
-%\begin{equation}\label{eq:subtend}
-%\theta_{\mathrm{min}} < \theta_{\mathrm{s}} \leq 180^\circ,
-%\end{equation}
-%with $\theta_{\mathrm{min}}=90^\circ$.
-%A similar criterion was used in the automatic lens search campaign of \citet{Son++18a}.
+If the source footprint consists instead of a single region, we iteratively increase the surface brightness threshold used to define the footprint and count the number of isolated regions with $S/N>10$.
+The maximum number of detected regions defines the number of detected peaks.
 
 \Fref{fig:lensdef} shows a few examples of how this criterion can be used to classify lenses.
-As the first three examples in \Fref{fig:lensdef} show, the presence of multiple images or a full ring implies that the system is classified as a lens.
-Only if the source footprint subtends an angle smaller than $90^\circ$, a system is not considered a strong lens (fourth example).
-%The choice of points used to define the angle $\theta_{\mathrm{s}}$ is not unique: any pair of points for which \Eref{eq:subtend} is satisfied is acceptable.
-
-%\begin{figure*}
-%\includegraphics[width=\textwidth]{footprints.eps}
-%\caption{
-%Criterion used to classify lensed images of extended sources. Four examples. In each panel, the blue regions are the footprint of the lensed source. 
-%The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
-%\label{fig:lensdef}
-%}
-%\end{figure*}
+The first column shows the caustic structure and source position. The second column shows an image of the lensed source. 
+The third column shows the source footprint defined with the procedure described above.
+Coral pixels correspond to the largest footprint that maximizes the number of detected images, while purple pixels belong to the $2\sigma$ detection footprint.
+In the first, second and sixth example, these two footprints coincide.
+In the third, fourth and fifth example, instead, the $2\sigma$ detection footprint consists of a single region, but an increase of the surface brightness threshold leads to the detection of additional images.
+
+Strictly speaking, our lens definition criterion fails for a perfect Einstein ring.
+In order to cover such a scenario, we also classify as strongly lensed any sources that produce a footprint with a hole.
+
+We point out that, although our lens definition relies on peak detections, it is not necessary for the peak of the source surface brightness to lie within the caustics in order for the source to be strongly lensed.
+This can be observed in the fourth example of \Fref{fig:lensdef}: although the source centroid lies outside of the caustic (as shown in the first column), the outskirts of the source overlap with the lens, and therefore multiple images are produced.
+
 \begin{figure}
@@
 Criterion used to classify lensed images of extended sources. Four examples.
+First column: caustics (red curves) and source position (blue circle).
+The radius of the circle corresponds to the radius at which the surface brightness is equal to $2\sigma$ the sky background fluctuation for a single pixel.
+In other words, the blue circle delimits the area of the source that can be detected.
+Second column: mock image of the lensed source, with added noise.
+Third column: footprint of the source. The purple footprint is obtained with a $2\sigma$ detection criterion. The coral region is the largest footprint with the highest number of detected images.
 \label{fig:lensdef}
@@
 \end{figure}
+
+We can identify three different regimes in strong lensing of extended sources, depending on the relative size of the source and of the lens caustics.
+In the limit of a very small source, the image configurations that are produced are qualitatively similar to those that can be obtained in the point-source case.
+When the source and caustic size are comparable, the multiple images tend to be blended into arcs or rings. 
+Most, if not all, of the galaxy-scale strong lenses known belong to these two categories.
+However, there is a third regime in strong lensing, corresponding to the case in which the source size is much bigger than the caustics, such as in the fifth example of \Fref{fig:lensdef}.
+In this regime, the overall size and total flux of the source are roughly preserved, and the lens produces only a relatively minor perturbation on a localised region of the image.
+Strong lenses of this kind can be difficult to detect, especially if the region of the image subject to strong lensing overlaps with the light from the foreground galaxy. 
+%For this reason, the best known examples of strong lensing in the large source size regime are produced by objects with a relatively low brightness that act as perturbers on images that are strongly lensed by a larger mass \citep[see e.g.][]{Veg++12}.
+The ultimate limiting factor to the detection of strong lenses in the large source size regime, however, is the ability to spatially resolve the multiple images.
+This limit is set by the size of the point spread function (PSF).
 
@@
 
-We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used the software {\sc Glafic} \citep{Ogu21} to solve the lens equation, find the number of images and their magnification.
+We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification.
 We then multiplied the area over which sources are located by the fraction of those that are strongly lensed according to the criterion of \Eref{eq:pointdet}.

### 2022-08-02 single_lens/elliptical_ext/plot_example_detections.py

@@ -12,21 +12,14 @@ from matplotlib import rc
 rc('text', usetex=True)
 
 
-fsize = 24
+fsize = 28
 
 nexamples = 6
 
-"""
-re_list = [0.2, 0.15, 0.3, 0.4]
-srcpos_list = [(0.5, 0.), (0.75, 0.), (0.2, 0.1), (0.3, 0.25)]
-q_list = [0.7, 0.3, 0.7, 0.7]
-zs_list = [1.5, 1.5, 0.5, 0.35]
-"""
-
-re_list = [0.4, 0.4, 0.4, 0.4, 0.4, 0.4]
-srcpos_list = [(0.05, 0.), (0.1, 0.), (0.15, 0.), (0.2, 0.), (0.25, 0.), (0.3, 0.)]
-q_list = [0.7, 0.7, 0.7, 0.7, 0.7, 0.7]
-zs_list = [0.35, 0.35, 0.35, 0.35, 0.35, 0.35]
+re_list = [0.2, 0.2, 0.1, 0.3, 0.4, 0.4]
+srcpos_list = [(0.4, -0.2), (-0.8, 0.4), (0.6, 0.), (0.25, 0.), (-0.1, 0.), (0.3, 0.25)]
+q_list = [0.7, 0.7, 0.3, 0.7, 0.7, 0.7]
+zs_list = [1.5, 1.5, 1.5, 0.4, 0.35, 0.35]
 
 # creates a custom colormap
 cm_vals = np.ones((3, 4))
@@ -104,8 +97,8 @@ for n in range(nexamples):
 
     ax[n, 0].contour(unlensed_sb, [sb_lim], colors=['b'], extent=(-2., 2., -2., 2.))
 
-    islens, nimg_std, nimg_best, std_footprint, best_footprint, sb_maxlim = detect_lens(img, sky_rms, npix_min=5)
-    print(sb_maxlim)
+    islens, nimg_std, nimg_best, nholes_std, nholes_best, std_footprint, best_footprint, sb_maxlim = detect_lens(img, sky_rms, npix_min=5)
+    print(n+1, nholes_std, nholes_best)
 
     f = open('tmp_crit.dat', 'r')
     table = np.loadtxt(f)
@@ -144,11 +137,16 @@ for n in range(nexamples):
 
         ax[n, 2].legend(loc='upper left', fontsize=fsize)
 
-    #ax[n, 1].imshow(img_wnoise, cmap='gray', vmin = -sky_rms, vmax=5.*sky_rms)
[…]

### 2022-08-02 paper/seleff.tex
added psf

@@
 This limit is set by the size of the point spread function (PSF).
+
+All of the examples that we considered were based on sources with a single surface brightness peaks.
+In general this is not necessarily the case, and the intrinsic number of surface brightness peaks of a strongly lensed source is not known a priori.
+When dealing with a real lens candidate, applying our lens definition criterion requires showing that the observed number of surface brightness peaks can be reproduced with a lens model in which the source has a smaller number of peaks.
+In samples of lenses defined via visual inspection this process is typically done implicitly, by identifying multiply imaged blobs that belong to the same source element.
+%Applying our lens definition criterion to a real lens requires showing that the observed number of surface brightness peaks can be reproduced with a lens model in which the source has a smaller number of peaks.
 

### 2022-08-02 single_lens/elliptical_ext/make_psf.py

@@ -1,0 +1,28 @@
+import numpy as np
+from astropy.io import fits as pyfits
+
+
+npix = 7
+
+fwhm = 2.
+
+beta = 5.
+alpha = 0.5*fwhm/(2.**(1./beta) - 1.)**0.5
+
+x0 = 3.
+y0 = 3.
+
+X, Y = np.meshgrid(np.arange(npix), np.arange(npix))
+R = ((X-x0)**2 + (Y-y0)**2)**0.5
+
+psf = (beta - 1.)/(np.pi*alpha)*(1. + (R/alpha)**2)**(-beta)
+
+psf /= psf.sum()
+
+phdu = pyfits.PrimaryHDU()
+psfhdu = pyfits.ImageHDU(data=psf)
+
+hdulist = pyfits.HDUList([phdu, psfhdu])
+hdulist.writeto('psf.fits', overwrite=True)
+#pyfits.PrimaryHDU(psf).writeto('psf.fits', overwrite=True)
+

### 2022-08-02 single_lens/elliptical_ext/plot_example_detections.py

@@ -2,6 +2,7 @@ import numpy as np
 import glafic
 from sl_profiles import sersic
 from scipy.special import gamma as gfunc
+from scipy.signal import convolve2d
 import pylab
 from astropy.io import fits as pyfits
 from lensdet import detect_lens
@@ -15,6 +16,8 @@ rc('text', usetex=True)
 fsize = 28
 
 nexamples = 6
+
+psf = pyfits.open('psf.fits')[1].data
 
 re_list = [0.2, 0.2, 0.1, 0.3, 0.4, 0.4]
 srcpos_list = [(0.4, -0.2), (-0.8, 0.4), (0.6, 0.), (0.25, 0.), (-0.1, 0.), (0.3, 0.25)]
@@ -55,6 +58,7 @@ glafic.startup_setnum(2, 1, 0)
 glafic.set_lens(1, 'gnfw', 0.3, 2.021e12, 0.0, 0.0, 0.3, 90.0, 10., 1.5)
 glafic.set_lens(2, 'sers', 0.3, 1.087e11, 0.0, 0.0, 0.3, 90.0, 1., 4.)
 glafic.set_extend(1, 'sersic', 1.5, 1., 0.3, 0., 0., 0., 0.1, 1.)
+#glafic.set_psf(2.*pix, 0., 0., 5., 1., 0., 0., 1., 1.)
 
 sizeone = 4.
 fig, ax = pylab.subplots(nexamples, 3, figsize=(3*sizeone, nexamples*sizeone))
@@ -79,7 +83,10 @@ for n in range(nexamples):
 
     glafic.writecrit(zs_list[n]) # writes caustics onto 'tmp_crit.dat'
 
+    #glafic.readpsf('psf.fits')
     img = np.flipud(np.array(glafic.writeimage()))
+    img = convolve2d(img, psf, mode='same')
+
     img_wnoise = img + np.random.normal(0., sky_rms, img.shape)
 
     pyfits.PrimaryHDU(img).writeto('test.fits', overwrite=True)

### 2022-08-03 paper/seleff.tex
large extended source

@@
 We explain how this definition applies in practice with a few examples.
-%We can consider a few examples.
 For simplicity, we focus on the case of a source with a single surface brightness peak, such as a S\'{e}rsic profile.
@@
 Strong lenses of this kind can be difficult to detect, especially if the region of the image subject to strong lensing overlaps with the light from the foreground galaxy. 
-%For this reason, the best known examples of strong lensing in the large source size regime are produced by objects with a relatively low brightness that act as perturbers on images that are strongly lensed by a larger mass \citep[see e.g.][]{Veg++12}.
 The ultimate limiting factor to the detection of strong lenses in the large source size regime, however, is the ability to spatially resolve the multiple images.
@@
 We compute $\crosssect$ in a series of different scenarios of increasing complexity.
-%We first introduce the basic equations of gravitational lensing in section \ref{ssec:lensbasics}.
-%Then, in section \ref{ssec:profile}, we present the model family adopted to describe the radial density profile of the lenses in our experiments.
-The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments.
-We describe this in section \ref{ssec:profile}.
-%, among with the basic equations of gravitational lensing.
-%In section \ref{ssec:axsymmpoint} we consider axisymmetric lenses and point sources.
+The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments. We describe this in section \ref{ssec:profile}.
 In section \ref{ssec:axsymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
 In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.
-
-%\subsection{Strong lensing basics}\label{ssec:lensbasics}
 
@@
 
-%Lensing-related quantities depend on the projected mass distribution $\Sigma(R)$.
-%For the purpose of computing lensing-related quantities, the projected massdistribution $\Sigma(R)$ is needed.
-%For a S\'{e}rsic profile, this is given by
 The projected surface mass density of a S\'{e}rsic profile is given by
@@
 Second, models with different dark matter fractions and inner dark matter slopes can conspire to produce very similar total density profiles. 
-%it is possible to build models with different dark matter fractions and inner dark matter slopes, but with very similar total density profiles. 
 This is the case, for example, of the ($\fdm=0.5,\gammadm=1.5$) and the ($\fdm=0.2,\gammadm=2.0$) models (green and blue lines in \Fref{fig:kappa}).
@@
 As \Fref{fig:1dmag} shows, the magnification is very large for sources close to the lens (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
-%In particular, for $\beta > 0.5\tein$, the magnification is smaller than unity, with the exception of a thin slice near the caustic.
 While for the model with $f_{\mathrm{DM}}=0.8$ (purple curve) the magnification is above unity everywhere, other lens models can produce highly de-magnified secondary images for large values of $\beta$.
@@
 The trend saturates below a certain $\Delta m$, for sufficiently small values of $\gammadm$ or for sufficiently large values of $\fdm$.
-%These are cases in which the source is so bright that, as long as it is located within the radial caustic, it always produces at least two detectable images. 
 In these cases the lensing cross-section coincides with the full area enclosed within the caustic, and further increasing the source brightness does not result in an increased value of $\crosssect$.
@@
 We measured $\crosssect$ for lenses with a fixed radial density profile and different ellipticities, with respect to point sources of diferent brightnesses.
-%In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.5$, $\reff=7$~kpc, $\log{\mdmfive}=11.0$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
-%In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
-This corresponds to a lens with Einstein radius equal to the half-light radius, both equal to $1''$.
 In particular, we set $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\tein=\teff$, and set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
 This is the lens model used to produce the caustics plot of \Fref{fig:ellcaust}.
-That figure suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
+\Fref{fig:ellcaust} suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
 Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.
@@
 We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification.
-We then multiplied the area over which sources are located by the fraction of those that are strongly lensed according to the criterion of \Eref{eq:pointdet}.
-%\Fref{fig:ellcaust} suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
-%Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.
+We then measured the fraction of sources that are strongly lensed according to the criterion of \Eref{eq:pointdet} and multiplied this value by the area over which sources are located.
 The resulting $\crosssect$ is shown in \Fref{fig:ellpoint_cs} as a function of the minor-to-major axis ratio $q$.
@@
 Most of the sources that result in detectable lenses produce two detectable images. These are sources that are located in the region enclosed bewteen the radial and the tangential caustic. 
-%In most cases, only two of these three images are detected: the third one, close to the centre, is usually highly de-magnified.
 If the source is located within the tangential caustic, however, four detectable images are usually created.
@@
 This is a consequence of the fact that the area enclosed within the tangential caustic, which is where a source needs to be in order to produce four images, increases with increased lens ellipticity, as \Fref{fig:ellcaust} shows.
-For sources that are intrinsically fainter than the detection limit, however, the quad cross-section is extremely small.
+For sources that are intrinsically fainter than the detection limit, however, the quad cross-section is extremely small, regardless of ellipticity.
 
@@
 
-In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the surface brightness distribution with its radial profile, shape and orientation, and the point spread function (PSF).
-For the sake of reducing the dimensionality of the problem, we neglected the PSF and focus on circularly symmetric sources.
+In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the source surface brightness distribution, with its radial profile, shape and orientation, and the point spread function (PSF).
+Moreover, as we discussed in section \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
+For this reason, we split our analysis in two parts: first we explore the small source size regime, then we consider cases in which the source size is comparable to or bigger than the lens caustics.
+
+\subsubsection{Small source sizes}
+
+For the sake of reducing the dimensionality of the analysis, we focused on circularly symmetric sources.
 We also fixed the surface brightness profile to an exponential disk (i.e.
@@
 We then took a lens model with the same parameters used in section \ref{ssec:ellpoint} and a minor-to-major axis ratio of $q=0.7$.
-We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
-We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_{\mathrm{e,s}}$.
+We simulated a large number of images of extended sources with {\sc Glafic} and measured the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
+For the small source size experiment, we used pixels with a size equal to $1/20\tein$ and convolved the images with a Moffat PSF with a FWHM of two pixels.
+We also assumed that the background noise is an uncorrelated Gaussian field.
+We carried out experiments with sources with different values of the total flux, $f$, and half-light radius, $\theta_{\mathrm{e,s}}$.
 The results are shown in \Fref{fig:extcs}.
 The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius.
-We assumed that the background noise is an uncorrelated Gaussian field, for this purpose.
 
@@
 Strong lensing cross-section as a function of source half-light radius.
+Small source size regime.
 The lens model is the same as in \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
@@
 The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
+The horizontal dotted line indicates the cross-section for a very bright point source (i.e. the area enclosed within the caustics).
 %Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
@@
 \end{figure}
+
+%We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
+%We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_{\mathrm{e,s}}$.
+%The results are shown in \Fref{fig:extcs}.
+%The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius.
+We assumed that the background noise is an uncorrelated Gaussian field, for this purpose.
 
@@
 
-%The limit posed by lensing to the maximum detectable source size can be understood as follows.
-%In the point source case, at a given flux, only sources that produce at least two images with a sufficiently high magnification are classified as strongly lensed.
-%In most practical cases, the limiting factor is the magnification of the secondary image, $\mu_2$, which is a strong function of the source position (see \Fref{fig:1dmag}).
-%Only sources that lie within a region of large $\mu_2$, the area of which is equal to the strong lensing cross-section, give rise to a strong lens configuration.
-%When the source is extended, different parts of it are magnified by different amounts.
-If the source size is comparable to or larger than the region of high magnification, then its net magnification averaged over its extent will be lower than in the point-source case, and the secondary image will drop below the detection limit.
-%This effect adds to the limit on the source surface brightness in the absence of lensing to produce the trend observed in \Fref{fig:extcs}.
-
-%In the point source case, the necessary and sufficient condition for a source to be strongly lensed is that two detectable images are produced.
-%The detection limit of the survey and the intrinsic flux of the source set a requirement on the minimum magnification that two images of the source must have in order for them to be detected.
-%The limiting factor is set by the magnification of the secondary image: given a source flux, 
-%In the point source case, at a given flux, only sources that lie within a certain region give rise to at least two detectable images, which are required to classify a system as a strong lens. 
-%The extent of this region, which is equal to the strong lensing cross-section, is determined by the magnification required in order for the secondary image to be detected (see \Fref{fig:1dmag}).
-%The area of this region is equal to the strong lensing cross-section, and decreases for fainter sources.
+\subsubsection{Large source sizes}
+
+For the large source size case we fixed the surface brightness profile of the source and varied the Einstein radius of the lens.
+In particular, we set the source half-light radius to $20$ pixels and adjusted its total flux in such a way that the $2\sigma$ detection footprint in the absence of lensing extends out to the half-light radius.
+Then, starting from the lens model used in the previous section, we progressively increased the critical surface mass density to reduce the Einstein radius down to values comparable to the pixel size.
+\Fref{fig:largesourcecs} shows the resulting strong lensing cross-section as a function of the ratio between lens Einstein radius and source half-light radius.
 
@@
 By doing this, we were able to fully describe the dark matter density profile with three parameters: $\mhalo$, $\gammadm$ and $r_s$.
-
+%
 The dark matter density profile defined in this way is determined uniquely by the halo mass, the stellar mass and the half-light radius (the more concentrated the stellar distribution, the stronger the halo response, the steeper the dark matter density profile).
-However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope.
+%However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope.
 Therefore, we applied a random scatter on $\gammadm$ of amplitude $\sigma_{\gamma}$.

### 2022-08-03 single_lens/elliptical_ext/get_smallsource_crosssect.py

@@ -2,26 +2,22 @@ import numpy as np
 from astropy.io import fits as pyfits
 from skimage import measure
 from scipy.special import gamma as gfunc
+from scipy.signal import convolve2d
 from scipy.optimize import brentq
 from sl_profiles import sersic
 from lensdet import detect_lens
 import h5py
 
 
-ftot = 200.
-nre = 21
-logre_grid = np.linspace(-1., 1., nre)
+nre = 11
+logre_grid = np.linspace(-1., 0., nre)
 
 nsim = 10000
 
-rsim = 1.5
-source_area = np.pi * rsim**2
-
-pix = 0.05
 nser = 1.
 
-nfrat = 5
-lfrat_grid = np.linspace(1., 3., nfrat)
+nfrat = 3
+lfrat_grid = np.linspace(1., 2., nfrat)
 # frat is the ratio between the source intrinsic flux and the sky rms over an area equal to the square of the Einstein radius.
 # In my sims, the Einstein radius is 1 arcsec. Pixels have a 0.05" size.
 # Then, the sky rms over a 1 square arcsec is 20 times that over a single pixel
@@ -30,12 +26,8 @@ lfrat_grid = np.linspace(1., 3., nfrat)
 
 nsigma = 2.
 
-minmag = 3.
-
-x0 = 39.5
-y0 = 39.5
-
-X, Y = np.meshgrid(np.arange(80), np.arange(80))
+# reads the PSF (because Glafic can't seem to be able to deal with it)
+psf = pyfits.open('psf.fits')[0].data
 
 cs_grid = np.zeros((nfrat, nre))
 
@@ -48,19 +40,22 @@ for l in range(nfrat):
         print(l, m)
         logre = logre_grid[m]
 
-        preamble = 'ftot200_logre%2.1f'%logre
         nlens = 0
         islens_sim = np.zeros(nsim, dtype=bool)
         nimages_sim = np.zeros(nsim, dtype=int)
 
         img_file = h5py.File('mockdir/logre%2.1f_images.hdf5'%logre, 'r')
 
+        rmax = img_file.attrs['rmax']
+        source_area = np.pi*rmax**2
+
         for i in range(nsim):
         
             img = img_file['lens_%04d'%i][()]
+            img = convolve2d(img, psf, mode='same')
 
-            islens, nimg_std, nimg_max, std_footprint, best_footprint = detect_lens(img, sky_rms)
-            islens_sim[i] = islens
[…]

### 2022-08-03 single_lens/elliptical_ext/make_psf.py

@@ -19,10 +19,5 @@ psf = (beta - 1.)/(np.pi*alpha)*(1. + (R/alpha)**2)**(-beta)
 
 psf /= psf.sum()
 
-phdu = pyfits.PrimaryHDU()
-psfhdu = pyfits.ImageHDU(data=psf)
-
-hdulist = pyfits.HDUList([phdu, psfhdu])
-hdulist.writeto('psf.fits', overwrite=True)
-#pyfits.PrimaryHDU(psf).writeto('psf.fits', overwrite=True)
+pyfits.PrimaryHDU(psf).writeto('psf.fits', overwrite=True)
 

### 2022-08-03 paper/seleff.tex
added cut in Einstein radius

--- 2022-08-02 paper/seleff.tex
@@
 
-\def\msource{m_{\mathrm{s}}}
-
 \def\mobs{M_*^{(\mathrm{obs})}}
@@
 
-\Fref{fig:ellcaust} shows the source-plane caustics of elliptical lenses with different values of the axis ratio. We used the software {\sc Glafic} \citep{Ogu21} to obtain the caustic curves.
+\Fref{fig:ellcaust} shows the source-plane caustics of elliptical lenses with different values of the axis ratio.
 The outermost curves are radial caustics, while the inner ones are tangential ones.
@@
 This is because, when the source is extended, some of the images can be blended together.
-In real strong lensing surveys, it is common to find lens candidates in which the lensed source consists of only a single arc.
+In real strong lensing surveys, it is common to find lens candidates in which the lensed source consists only a single arc.
 In those cases it is difficult to establish whether the arc is a set of blending images or not, and the decision of including such systems in a strong lens sample is often arbitrary.
-Here we adopt the following definition: an extended source is strongly lensed if the number of surface brightness peaks that are detected is larger than its intrinsic number of peaks in the absence of lensing.
-
-We explain how this definition applies in practice with a few examples.
-%We can consider a few examples.
-For simplicity, we focus on the case of a source with a single surface brightness peak, such as a S\'{e}rsic profile.
-All of the sources considered in this work belong to this family of surface brightness models.
-We adopt the following procedure to determine the number of detected peaks.
-Given the observed image of a lensed source, we define its footprint as the ensemble of pixels where the source is detected with $S/N>2$.
+%For instance, a source located on one of the cusps of a tangential caustic produces a merging set of three magnified images.
+If the source is extended and no counter-image is detected, however, the appearence of the merging triplet is very similar to that of a single arc with no multiple images, and it can be difficult to differentiate between the two cases.
+%In real strong lensing surveys, the decision of including systems with a single arc or not is often arbitrary.
+
+For the purpose of carrying out our experiment we adopt the following definition.
+First, we determine the footprint of the lensed source.
+We define this as the ensemble of pixels where the source is detected with $S/N>2$.
 The footprint is in general composed of multiple disconnected regions, corresponding to the different images.
@@
 If, after applying this cut, the source footprint is spread over multiple separated regions, then the system is classified as a lens.
-If the source footprint consists instead of a single region, we iteratively increase the surface brightness threshold used to define the footprint and count the number of isolated regions with $S/N>10$.
-The maximum number of detected regions defines the number of detected peaks.
+Additionally, if the source footprint consists of only a single region, we still classify it as strongly lensed if it subtends an angle larger than $90^\circ$ measured from the lens centre.
+This last criterion allows us to include highly magnified arcs and full rings.
+
+%Additionally, we classify as strongly lensed the following set of sources any full rings, and any single arcs that subtend an angle larger than $90^\circ$ measured from the lens centre.
+This last criterion is a
+%Additionally, we classify as a strongly lensed source any full rings, and any single arcs that subtend an angle larger than $90^\circ$ measured from the lens centre. This last criterion is a
+
+%Given the source footprint thus defined, we classify the system as a strong lens if the footprint subtends an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
+%Given the source footprint thus defined, we classify the system as a strong lens if we can find two points on the footprint that subtend an angle $\theta_{\mathrm{s}}$ with respect to the lens centre that satisfies the following condition:
+%\begin{equation}\label{eq:subtend}
+%\theta_{\mathrm{min}} < \theta_{\mathrm{s}} \leq 180^\circ,
+%\end{equation}
+%with $\theta_{\mathrm{min}}=90^\circ$.
+%A similar criterion was used in the automatic lens search campaign of \citet{Son++18a}.
 
 \Fref{fig:lensdef} shows a few examples of how this criterion can be used to classify lenses.
-The first column shows the caustic structure and source position. The second column shows an image of the lensed source. 
-The third column shows the source footprint defined with the procedure described above.
-Coral pixels correspond to the largest footprint that maximizes the number of detected images, while purple pixels belong to the $2\sigma$ detection footprint.
-In the first, second and sixth example, these two footprints coincide.
-In the third, fourth and fifth example, instead, the $2\sigma$ detection footprint consists of a single region, but an increase of the surface brightness threshold leads to the detection of additional images.
-
-Strictly speaking, our lens definition criterion fails for a perfect Einstein ring.
-In order to cover such a scenario, we also classify as strongly lensed any sources that produce a footprint with a hole.
-
-We point out that, although our lens definition relies on peak detections, it is not necessary for the peak of the source surface brightness to lie within the caustics in order for the source to be strongly lensed.
-This can be observed in the fourth example of \Fref{fig:lensdef}: although the source centroid lies outside of the caustic (as shown in the first column), the outskirts of the source overlap with the lens, and therefore multiple images are produced.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{example_detections.eps}
+As the first three examples in \Fref{fig:lensdef} show, the presence of multiple images or a full ring implies that the system is classified as a lens.
+Only if the source footprint subtends an angle smaller than $90^\circ$, a system is not considered a strong lens (fourth example).
+%The choice of points used to define the angle $\theta_{\mathrm{s}}$ is not unique: any pair of points for which \Eref{eq:subtend} is satisfied is acceptable.
+
+\begin{figure*}
+\includegraphics[width=\textwidth]{footprints.eps}
 \caption{
-Criterion used to classify lensed images of extended sources. Four examples.
-First column: caustics (red curves) and source position (blue circle).
-The radius of the circle corresponds to the radius at which the surface brightness is equal to $2\sigma$ the sky background fluctuation for a single pixel.
-In other words, the blue circle delimits the area of the source that can be detected.
-Second column: mock image of the lensed source, with added noise.
-Third column: footprint of the source. The purple footprint is obtained with a $2\sigma$ detection criterion. The coral region is the largest footprint with the highest number of detected images.
+Criterion used to classify lensed images of extended sources. Four examples. In each panel, the blue regions are the footprint of the lensed source. 
+The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
 \label{fig:lensdef}
 }
-\end{figure}
-
-We can identify three different regimes in strong lensing of extended sources, depending on the relative size of the source and of the lens caustics.
-In the limit of a very small source, the image configurations that are produced are qualitatively similar to those that can be obtained in the point-source case.
-When the source and caustic size are comparable, the multiple images tend to be blended into arcs or rings. 
-Most, if not all, of the galaxy-scale strong lenses known belong to these two categories.
-However, there is a third regime in strong lensing, corresponding to the case in which the source size is much bigger than the caustics, such as in the fifth example of \Fref{fig:lensdef}.
-In this regime, the overall size and total flux of the source are roughly preserved, and the lens produces only a relatively minor perturbation on a localised region of the image.
-Strong lenses of this kind can be difficult to detect, especially if the region of the image subject to strong lensing overlaps with the light from the foreground galaxy. 
-%For this reason, the best known examples of strong lensing in the large source size regime are produced by objects with a relatively low brightness that act as perturbers on images that are strongly lensed by a larger mass \citep[see e.g.][]{Veg++12}.
-The ultimate limiting factor to the detection of strong lenses in the large source size regime, however, is the ability to spatially resolve the multiple images.
-This limit is set by the size of the point spread function (PSF).
-
+\end{figure*}
 %__________________________________________________________________
@@
 
-We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification.
+We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used the software {\sc Glafic} \citep{Ogu21} to solve the lens equation, find the number of images and their magnification.
 We then multiplied the area over which sources are located by the fraction of those that are strongly lensed according to the criterion of \Eref{eq:pointdet}.
@@
 \end{tabular}
-\tablefoot{
+\notes{
 Adopted values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$ and $\sigma_{\mathrm{h}}$ in different simulations.
@@
 
-%\ale{To be filled by Shun-Sheng}.
+\ale{To be filled by Shun-Sheng}.
 We applied a lower limit to the source redshift, by selecting only sources with $z_s > 0.8$.
@@
 For the extended sources, we produced images with a pixel size of $0.1''$.
-We applied a Gaussian PSF with a FWHM of $0.2''$.
+We applied a Moffat PSF with an FWHM of $0.2''$.
 Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
@@
 
-We applied the lens selection criterion of section \ref{ssec:}
+We applied two different lens selection criteria.
+First, we used the criterion described in section \ref{ssec:lensdetext}.
+This resulted in a sample of XXX lenses.
+Then, we added a cut on the Einstein radius of the lens: we only included in the lens sample those with $\tein > 0.5''$.
+This additional cut is meant to approximate the difficulty in detecting strongly lensed images that lie very close to the lens centre: the number of known lenses with Einstein radius smaller than that cutoff value is very small, even in lens surveys that are based on high-resolution data \citep[see e.g.][]{Gar++22}.
+With this more restrictive criterion, the number of strong lenses in the simulated sample is YYY.
+
+\Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the general population.
+\Fref{fig:sourcepars} shows the distribution in source properties.
 
@@
 This sample was selected as follows.
-First we defined a narrow redshift slice around $z=0.2$.
+Starting from the SDSS spectroscopic sample, we defined a narrow redshift slice around $z=0.2$.
 Then we applied a selection in colour, by choosing objects with $g-r>1.2$, and on S\'{e}rsic index, by selecting only galaxies with $n>2.5$.

### 2022-08-03 paper/seleff.tex
stuff

@@
 %At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), $\crosssect$ increases with increasing source size, before dropping to zero.
-At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (red ellipse in \Fref{fig:ellcaust}).
+At the brightest flux explored in the experiment (green line in \Fref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (black dotted line in \Fref{fig:extcs}).
 This is because, when the source is very bright, it can give rise to multiple images even while its centroid lies outside of the radial caustic, as long as its surface brightness distribution extends into it.

### 2022-08-03 lens_population/lens_extended_sources.py

@@ -41,9 +41,9 @@ pix_poi = 0.1
 maxlev = 5
 
 glafic.init(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0)
-glafic.set_secondary('flag_halodensity 2')
+glafic.set_secondary('flag_hodensity 2')
 glafic.set_secondary('nfw_users 1')
-glafic.set_secondary('halodensity 200')
+glafic.set_secondary('hodensity 200')
 
 glafic.startup_setnum(2, 1, 0)
 glafic.set_lens(1, 'gnfw', 0.3, 1e13, 0.0, 0.0, 0., 90.0, 10., 1.5)

### 2022-08-03 single_lens/elliptical_ext/get_largesource_crosssect.py

@@ -1,0 +1,53 @@
+import numpy as np
+from astropy.io import fits as pyfits
+from skimage import measure
+from scipy.special import gamma as gfunc
+from scipy.signal import convolve2d
+from scipy.optimize import brentq
+from sl_profiles import sersic
+from lensdet import detect_lens
+import h5py
+
+
+ntein = 5
+ltein_grid = np.linspace(-0.9, -0.1, ntein)
+
+nsim = 10000
+
+nser = 1.
+
+sky_rms = 1.
+sb_min = 2. * sky_rms
+
+cs_grid = np.zeros(ntein)
+
+#for m in range(ntein):
+for m in range(3, 4):
+
+    print(m)
+    ltein = ltein_grid[m]
+
+    nlens = 0
+    islens_sim = np.zeros(nsim, dtype=bool)
+    nimages_sim = np.zeros(nsim, dtype=int)
+
+    img_file = h5py.File('mockdir/ltein%2.1f_images.hdf5'%ltein, 'r')
+
+    rmax = img_file.attrs['rmax']
+    source_area = np.pi*rmax**2
+
+    for i in range(nsim):
+
+        img = img_file['lens_%04d_wseeing'%i][()]
+
+        res = detect_lens(img, sky_rms, npix_min=1)
+        islens_sim[i] = res[0]
+        print(i, res[0])
+
+    cs_grid[m] = islens_sim.sum()/float(nsim) * source_area
+
+output = h5py.File('largesource_crosssect.hdf5', 'w')
+
+output.create_dataset('ltein_grid', data=ltein_grid)
+output.create_dataset('cs_grid', data=cs_grid)
+

### 2022-08-03 single_lens/elliptical_ext/lensdet.py

@@ -19,7 +19,9 @@ def detect_lens(img, sky_rms, nsigma_pixdet=2., npix_min=10):
 
     labels = measure.label(footprint)
     nreg = labels.max()
+    euler = measure.euler_number(footprint)
     nimg_std = 0
+    nholes_std = nreg - euler
     for n in range(nreg):
         npix_tmp = (labels==n+1).sum()
         signal = img[labels==n+1].sum()
@@ -42,6 +44,8 @@ def detect_lens(img, sky_rms, nsigma_pixdet=2., npix_min=10):
     nimg_tmp = nimg_std
     sb_maxlim = sb_min
 
+    nholes_max = 0
+
     best_footprint = std_footprint.copy()
 
     # if there's at least one detected image, increases the threshold to see
@@ -55,7 +59,9 @@ def detect_lens(img, sky_rms, nsigma_pixdet=2., npix_min=10):
 
         labels = measure.label(footprint_tmp)
         nreg = labels.max()
+
         nimg_tmp = 0
+        nholes_tmp = 0
         for n in range(nreg):
             npix_tmp = (labels==n+1).sum()
             signal = img[labels==n+1].sum()
@@ -63,6 +69,9 @@ def detect_lens(img, sky_rms, nsigma_pixdet=2., npix_min=10):
             img_sn = signal/noise
             if img_sn >= 10. and npix_tmp >= npix_min:
                 nimg_tmp += 1
+                # checks if the image has a hole
+                euler = measure.euler_number(labels==n+1)
+                nholes_tmp += 1 - euler
             else:
                 img_detected[labels==n+1] = 0.
 
@@ -71,6 +80,9 @@ def detect_lens(img, sky_rms, nsigma_pixdet=2., npix_min=10):
             sb_maxlim = sb_lim
             best_footprint = img_detected > sb_lim
 
+        if nholes_tmp > nholes_max:
+            nholes_max = nholes_tmp
+
         i += 1
 
     if nimg_max > 1:
@@ -78,5 +90,5 @@ def detect_lens(img, sky_rms, nsigma_pixdet=2., npix_min=10):
     else:
         islens = False
 
-    return islens, nimg_std, nimg_max, std_footprint, best_footprint, sb_maxlim
+    return islens, nimg_std, nimg_max, nholes_std, nholes_max, std_footprint, best_footprint, sb_maxlim
 

### 2022-08-03 single_lens/elliptical_ext/make_largesource_images.py

@@ -16,8 +16,8 @@ from scipy.optimize import brentq
 # keeps the source surface brightness fixed and varies the redshift
 np.random.seed(0)
 
-nrein = 8
-ltein_grid = np.linspace(-0.4, 1., nrein)
+nrein = 5
+ltein_grid = np.linspace(-0.9, -0.1, nrein)
 
 pix = 0.05
 nser = 1. # source Sersic index

### 2022-08-03 single_lens/elliptical_ext/plot_example_detections.py

@@ -17,7 +17,7 @@ fsize = 28
 
 nexamples = 6
 
-psf = pyfits.open('psf.fits')[1].data
+psf = pyfits.open('psf.fits')[0].data
 
 re_list = [0.2, 0.2, 0.1, 0.3, 0.4, 0.4]
 srcpos_list = [(0.4, -0.2), (-0.8, 0.4), (0.6, 0.), (0.25, 0.), (-0.1, 0.), (0.3, 0.25)]

### 2022-08-05 paper/seleff.tex
large source case

@@
 
-All of the examples that we considered were based on sources with a single surface brightness peaks.
+All of the examples that we considered were based on sources with a single surface brightness peak.
 In general this is not necessarily the case, and the intrinsic number of surface brightness peaks of a strongly lensed source is not known a priori.
@@
 %Applying our lens definition criterion to a real lens requires showing that the observed number of surface brightness peaks can be reproduced with a lens model in which the source has a smaller number of peaks.
+
+To our knowledge, we are the first to propose a surface brightness peak-based definition of a strong lensing event.
+A popular alternative definition of a strong lens is one based on magnification: only images that are magnified by more than a given threshold are considered as strongly lensed \citep[see e.g.][]{Hil++07}.
+The problem with a definition of this kind is that magnification cannot be determined unambiguously from observations, unless the intrinsic properties of the lensed source are known from other means (e.g.
+if the source is a standard candle or a standard ruler).
+Because of the mass-sheet degeneracy \citep{FGS85}, it is possible to vary the magnification of a lensed image while keeping its observed properties fixed: this could lead to the paradox of two identical lenses that are classified differently on the basis of the underlying magnification.
+Although we could still use a magnification-based definition for the sake of carrying out our experiments, it would then be difficult to apply our results to real data.
+%Because of the mass-sheet degeneracy
 
@@
 \caption{
-Strong lensing cross-section as a function of source half-light radius.
+Strong lensing cross-section of an extended source.
 Small source size regime.
-The lens model is the same as in \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
+The cross-section is plotted as a function of the ratio between the half-light radius of the source and the Einstein radius of the lens.
+The lens model is fixed to be the same of \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
 Each solid line corresponds to a different value of the total flux of the source. The flux $f$ is expressed in terms of the background noise fluctuation measured over an area equal to $\tein^2$.
@@
 \Fref{fig:largesourcecs} shows the resulting strong lensing cross-section as a function of the ratio between lens Einstein radius and source half-light radius.
+For values of the Einstein radius close to the size of the PSF, the lensing cross-section (blue solid line) is very small: this is because the perturbations caused by lensing are not well resolved.
+For larger values, $\crosssect$ stays approximately constant, around values that are comparable to the source size (horizontal dotted line) and much larger than the area enclosed within the caustics (red dashed line).
+We conclude that, in the large source size regime, the main factor that determines the lensing cross-section is the area of the background source, provided that the Einstein radius of the lens is larger than the PSF.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{largesource_cs.eps}
+\caption{
+Strong lensing cross-section of an extended source. Large source size regime.
+The cross-section is plotted as a function of the ratio between the Einstein radius of the lens and the half-light radius of the source.
+The source model is fixed to a circular S\'{e}rsic profile with $n=1$, detected out to the half-light radius.
+The source half-light radius measures $20$ pixels, and the PSF has an FWHM of two pixels.
+The lens model is varied: starting from the model of \Fref{fig:extcs}, the critical surface mass density is increased to produce progressively smaller values of $\tein$.
+Solid line: lensing cross-section, in units of the footprint area of the source in the absence of lensing, as a function of the ratio between the lens Einstein radius and the source half-light radius.
+Dashed line: area enclosed within the lens caustics.
+Dotted line: value of the cross-section corresponding to the area of the source.
+Dash-dotted line: FWHM of the PSF.
+\label{fig:largesourcecs}
+}
+\end{figure}
 

### 2022-08-05 single_lens/elliptical_ext/get_largesource_crosssect.py

@@ -9,8 +9,8 @@ from lensdet import detect_lens
 import h5py
 
 
-ntein = 5
-ltein_grid = np.linspace(-0.9, -0.1, ntein)
+ntein = 11
+ltein_grid = np.linspace(-1., 0., ntein)
 
 nsim = 10000
 
@@ -21,8 +21,11 @@ sb_min = 2. * sky_rms
 
 cs_grid = np.zeros(ntein)
 
-#for m in range(ntein):
-for m in range(3, 4):
+islens_grid = np.zeros((ntein, nsim), dtype=bool)
+x_grid = np.zeros((ntein, nsim))
+y_grid = np.zeros((ntein, nsim))
+
+for m in range(ntein):
 
     print(m)
     ltein = ltein_grid[m]
@@ -36,13 +39,16 @@ for m in range(3, 4):
     rmax = img_file.attrs['rmax']
     source_area = np.pi*rmax**2
 
+    x_grid[m, :] = img_file['x'][()]
+    y_grid[m, :] = img_file['y'][()]
+
     for i in range(nsim):
 
         img = img_file['lens_%04d_wseeing'%i][()]
 
         res = detect_lens(img, sky_rms, npix_min=1)
         islens_sim[i] = res[0]
-        print(i, res[0])
+        islens_grid[m, i] = res[0]
 
     cs_grid[m] = islens_sim.sum()/float(nsim) * source_area
 
@@ -50,4 +56,8 @@ output = h5py.File('largesource_crosssect.hdf5', 'w')
 
 output.create_dataset('ltein_grid', data=ltein_grid)
 output.create_dataset('cs_grid', data=cs_grid)
+output.create_dataset('islens_grid', data=islens_grid)
+output.create_dataset('x_grid', data=x_grid)
+output.create_dataset('y_grid', data=y_grid)
+
 

### 2022-08-05 single_lens/elliptical_ext/plot_largesource_cs.py

@@ -12,13 +12,13 @@ rc('text', usetex=True)
 
 fsize = 18
 
-#xlim = (0.1, 1.)
+xlim = (0.09, 1.)
 #ylim = (0.085, 1.5)
 
 logscale = True
 
 if logscale:
-    leftm = 0.14
+    leftm = 0.15
 else:
     leftm = 0.08
 
@@ -38,19 +38,22 @@ pylab.subplots_adjust(left=leftm, right=0.98, bottom=0.14, top=0.97, wspace=0.)
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
 ax.loglog(10.**ltein_grid, cs_grid[()]/np.pi, linewidth=2, label='Cross-section')
-ax.loglog(10.**ltein_grid, caust_grid[()]/np.pi, linestyle='--', color='k', linewidth=2, label='Caustic area')
+ax.loglog(10.**ltein_grid, caust_grid[()]/np.pi, linestyle='--', color='r', linewidth=2, label='Caustic area')
+ax.axhline(1., linestyle=':', color='k', label='Source area')
 
 ax.tick_params(axis='both', which='both', top=True, right=True, labelsize=fsize, direction='in')
 
 ax.set_xlabel('$\\theta_{\mathrm{Ein}}/\\theta_{\mathrm{s, e}}$', fontsize=fsize)
 ax.set_ylabel('$\sigma_{\mathrm{SL}}/(\pi\\theta_{\mathrm{s, e}}^2)$', fontsize=fsize)
 
+ax.axvline(0.1, linestyle='-.', color='k', label='PSF FWHM')
+
 #ax.set_xlim(xlim[0], xlim[1])
 #ax.set_ylim(ylim[0], ylim[1])
 
 #ax.loglog(re_max, cs_ps/np.pi, linestyle=':', color='k')
 
-ax.legend(loc='upper left', fontsize=fsize)
+ax.legend(loc='lower right', fontsize=fsize)
 
 pylab.savefig('../../paper/largesource_cs.eps')
 pylab.show()

### 2022-08-07 paper/seleff.tex
started section 5

@@
 In this work we revisit the \citet{MVK09} study, we expand it to the extended source case and, most importantly, we address point 2) as well: we simulate large populations of strong lenses using empirical models, simulate lensing surveys with realistic assumptions for the lens probability $\pdet$ and quantify the lensing bias under various scenarios.
-We explore how the results change as a function of the scatter in mass parameters at fixed light, and as a function of source type: we simulate both galaxy and quasar sources.
+%We explore how the results change as a function of the scatter in mass parameters at fixed light, of the efficiency of finding lenses with small image separation, and of source type: we simulate both galaxy and quasar sources.
+%We explore how the results change as a function of the scatter in mass parameters at fixed light, and as a function of the efficiency of a lens survey at discovering small image separation lenses.
+We explore how the results change as a function of the efficiency of a lens survey at discovering small image separation lenses, and of the scatter in mass parameters at fixed light.
 
@@
 
-For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering XXX square degrees.
+For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering $1000$ square degrees.
 The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is YYY.
@@
 
-For the extended sources, we produced images with a pixel size of $0.1''$.
+%\subsubsection{Lensed galaxies}
+
+For the simulation with extended sources, we produced images with a pixel size of $0.1''$.
 We applied a Moffat PSF with an FWHM of $0.2''$.
@@
 These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
-
-We applied two different lens selection criteria.
-First, we used the criterion described in section \ref{ssec:lensdetext}.
-This resulted in a sample of XXX lenses.
-Then, we added a cut on the Einstein radius of the lens: we only included in the lens sample those with $\tein > 0.5''$.
-This additional cut is meant to approximate the difficulty in detecting strongly lensed images that lie very close to the lens centre: the number of known lenses with Einstein radius smaller than that cutoff value is very small, even in lens surveys that are based on high-resolution data \citep[see e.g.][]{Gar++22}.
-With this more restrictive criterion, the number of strong lenses in the simulated sample is YYY.
-
-\Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the general population.
+%
+We then applied the peak detection-based lens selection criterion introduced in section \ref{ssec:lensdefext} to find the strong lenses.
+
+%First, we used the criterion described in section \ref{ssec:lensdefext}.
+%This resulted in a sample of XXX lenses.
+%Then, we added a cut on the Einstein radius of the lens: we only included in the lens sample those with $\tein > 0.5''$.
+%This additional cut is meant to approximate the difficulty in detecting strongly lensed images that lie very close to the lens centre: the number of known lenses with Einstein radius smaller than that cutoff value is very small, even in lens surveys that are based on high-resolution data \citep[see e.g.][]{Gar++22}.
+%With this more restrictive criterion, the number of strong lenses in the simulated sample is YYY.
+
+%\subsubsection{Lensed quasars}
+
+For the lensed quasars we did not simulate pixel-level data, but simply computed the observed magnitudes of the multiple images.
+Following the definition of section \ref{ssec:lensdefpoint}, we included in the sample of strong lenses only systems with at least two images brighter than a limiting magnitude $m_{\mathrm{lim}}$.
+We set $m_{\mathrm{lim}}=23.3$, which corresponds to the $10\sigma$ detection limit of the LSST in a single visit \citep{O+M10}.
+This is motivated by the fact that the quasar lenses are meant to simulate a sample assembled for the purpose of carrying out time delay measurements, which in turn require combining single-visit detections over many epochs.
+
+%Both for the lensed galaxies and lensed quasars, we apply an additional selection based on the angular size of the Einstein radius.
+Both for the lensed galaxies and lensed quasars, we apply an additional cut, by selecting only lenses with an angular Einstein radius larger than a limiting value $\theta_{\mathrm{Ein,min}}$.
+This additional cut is meant to approximate the difficulty in detecting strongly lensed images with small separation, because of the contamination from the light of the lens galaxy.
+We explored scenarios with different values of $\theta_{\mathrm{Ein,min}}$.
+
+%__________________________________________________________________
+
+\section{Results}\label{sect:results}
+
+\subsection{Galaxy-galaxy lenses}
+
+\Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the general population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
+A minimum Einstein radius of $0.5''$ is close to what can be currently be achieved via visual inspection of high-resolution space-based images \cite{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
+\begin{figure*}
+\includegraphics[width=\textwidth]{lens_cornerplot.eps}
+\caption{
+Comparison bewteen the properties of lens samples and the general galaxy population. Distribution in foreground galaxy parameters.
+\label{fig:lenspars}
+}
+\end{figure*}
+
+The most striking difference between the samples is in the stellar mass, which is much larger in the lenses than in the general population.
+Lenses tend to also have a larger halo mass, a smaller half-light radius at fixed stellar mass, and a larger stellar population synthesis mismatch parameter.
+The distribution in ellipticity and inner dark matter slope of the lenses instead look very similar to that of the parent population.
+%This outcome was expected, given the results of the analysis of \Sref{sect:indlenses}.
+Additionally, we can see that the amplitude of the strong lensing bias appears to be always larger in the lens sample with the more restrictive selection on Einstein radius.
+
+Qualitatively, the result shown in \Fref{fig:lenspars} matches our expectations from \Sref{sect:indlenses}.
+In order to quantitatively assess the amplitude of the strong lensing bias in various lensing properties, we proceeded as follows.
+
 \Fref{fig:sourcepars} shows the distribution in source properties.
 
-%__________________________________________________________________
-
-\section{Results}\label{sect:results}
-
-There are results.
+\subsection{Galaxy-quasar lenses}
 

### 2022-08-09 paper/seleff.tex
unlensed detection

@@
 
-The most striking difference between the samples is in the stellar mass, which is much larger in the lenses than in the general population.
+The most striking difference between the samples is in the stellar mass: strong lensing selects preferentially galaxies with larger values of $\mobs$.
+%, which is much larger in the lenses than in the general population. 
 Lenses tend to also have a larger halo mass, a smaller half-light radius at fixed stellar mass, and a larger stellar population synthesis mismatch parameter.
@@
 Additionally, we can see that the amplitude of the strong lensing bias appears to be always larger in the lens sample with the more restrictive selection on Einstein radius.
+
+\Fref{fig:sourcepars} shows the distribution in parameters describing the background source population, for the same simulations of \Fref{fig:lenspars}. 
+The unlensed distribution (blue filled contours) was obtained by taking into account the detectability of the sources: we drew a large number of sources from the distribution described in section \ref{ssec:sources}, simulated images of them with the same properties as the images of the lenses, and selected only those galaxies detected with $S/N>10$.
+Therefore, the resulting distribution is different to the parent distribution from which the simulation was drawn.
+
+Since we want to quantify the effect of lensing with respect to the case with no lensing for a given survey, we took into account the detectability of the sources when determining the unlensed background galaxy distribution.
+
+Strong lensing tends to preferentially select sources at higher redshift, especially in the more restrictive case with $\theta_{\mathrm{Ein,min}} > 1.0’’$.
+This is because, at fixed lens properties, increasing the source redshift lowers the critical surface mass density and, consequently, increases the size of the caustics and the Einstein radius.
+The distribution in source magnitude is also significantly affected by the strong lensing selection.
+In a somewhat counter-intuitive way, the $\msource$ distribution of the lenses peaks at brighter magnitudes than the parent distribution.
+The parent sample, however, is simply the input of the simulations, and does not take into account whether sources are detectable or not.
+The detection limit that we assumed in our simulations is at $\msource\approx25$, and therefore the source magnitude distribution in the absence of lensing would drop rapidly.
+In the absence of lensing, this distribution would drop rapidly beyond the detection limit $\msource\approx25$.
+Lensing, however, allows the detection of sources at fainter magnitudes, in virtue of the magnification, and therefore the $\msource$ distribution shows a faint-end tail beyond $\msource=25$.
+We can also see a bias in favour of sources with a very low S\'{e}rsic index.
+This is a consequence of the selection in magnitude: our source sample shows a strong correlation between $\msource$ and $n$: the brightest sources tend to have the smallest values of the S\'{e}rsic index.
 

### 2022-08-09 lens_population/draw_detectable_sources.py

@@ -1,0 +1,96 @@
+import numpy as np
+from scipy.stats import poisson
+from simpars import *
+from sc_profiles import sersic
+import h5py
+import os
+from scipy.signal import convolve2d
+from astropy.io import fits as pyfits
+from skimage import measure
+import sys
+
+
+np.random.seed(20)
+
+ndraw = 100000
+sourcecat = pyfits.open('/Users/alessandro/catalogs/skills_sourceonly_zcut.fits')[1].data
+
+psf = pyfits.open('psf.fits')[0].data
+pix_arcsec = 0.1
+sb_min = nsigma_pixdet * sky_rms
+
+nsource_tot = len(sourcecat)
+
+sourceind = np.arange(nsource_tot)
+# shuffles source catalog (it's originally ranked by redshift)
+np.random.shuffle(sourceind)
+
+sreff_draw = sourcecat['Re_arcsec_CM'][:ndraw]
+nser_draw = sourcecat['sersic_n_CM'][:ndraw]
+sq_draw = sourcecat['axis_ratio_CM'][:ndraw]
+smag_draw = sourcecat['g_SDSS_apparent_corr'][:ndraw]
+zs_draw = sourcecat['zobs'][:ndraw]
+pa_draw = sourcecat['PA_random'][:ndraw]
+
+f = open('preamble.input', 'r')
+prelines = f.readlines()
+f.close()
+
+outlines = prelines.copy()
+
+for i in range(ndraw):
+    ind = sourceind[i]
+    outlines.append('reset_par simdir/source_%07d\n'%ind)
+    outlines.append('reset_extend 1 1 %f\n'%zs_draw[i])
+
+    ftot = 10.**(-2./5.*(smag_draw[i] - zeropoint))
+    I0 = ftot/(2.*np.pi*(sreff/pix_arcsec)**2*nser/sersic.b(nser)**(2*nser)*gfunc(2.*nser))
+
+    outlines.append('reset_extend 1 2 %f\n'%I0)
+    outlines.append('reset_extend 1 5 %f\n'%(1. - sq_draw[i]))
+    outlines.append('reset_extend 1 6 %f\n'%pa_draw[i])
+    outlines.append('reset_extend 1 7 %f\n'%sreff_draw[i])
+    outlines.append('reset_extend 1 8 %f\n'%nser_draw[i])
+
+    outlines.append('writeimage_ori\n')
+    outlines.append('\n')
+
+outlines.append('quit\n')
+
[…]

### 2022-08-14 paper/seleff.tex
detectable sources discussion

--- 2022-08-07 paper/seleff.tex
@@
 
+%In this section we present the results of the lens population simulations. 
+%Section \ref{ssec:resgal} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:resquasar} focuses on the population of galaxy-quasar lenses.
+%In each case, we first offer
+
 \subsection{Galaxy-galaxy lenses}
 
-\Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the general population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
+In this section we show the results of the experiments with populations of galaxy-galaxy lenses.
+We first present the results in a qualitative way, then proceed to quantify the amplitude of the strong lensing bias in various quantities of interest.
+
+\Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the parent population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
 A minimum Einstein radius of $0.5''$ is close to what can be currently be achieved via visual inspection of high-resolution space-based images \cite{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
@@
 \caption{
-Comparison bewteen the properties of lens samples and the general galaxy population. Distribution in foreground galaxy parameters.
+Comparison bewteen the properties of lens samples and the general population. Distribution in foreground galaxy parameters.
+Filled contours: distribution of the parent sample.
+Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
+Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
 \label{fig:lenspars}
@@
 Additionally, we can see that the amplitude of the strong lensing bias appears to be always larger in the lens sample with the more restrictive selection on Einstein radius.
-
-Qualitatively, the result shown in \Fref{fig:lenspars} matches our expectations from \Sref{sect:indlenses}.
-In order to quantitatively assess the amplitude of the strong lensing bias in various lensing properties, we proceeded as follows.
-
-\Fref{fig:sourcepars} shows the distribution in source properties.
+We quantify the amplitude of these biases later in this section.
+
+\Fref{fig:sourcepars} shows the distribution in parameters describing the background source population, for the same simulations of \Fref{fig:lenspars}. 
+Additionally, \Fref{fig:sourcepars} shows the subset of the parent population that consists of detectable sources.
+These are background galaxies that, in the absence of lensing, can be detected according to the same criterion used for the lensed sources (i.e.
+the $S/N$ over their $2\sigma$ footprint is larger than ten).
+The detection limit of the survey is at $\msource\approx25$ (the actual limit varies depending on the surface brightness distribution parameters).
+Because our simulated background source population extends to much fainter magnitudes, the distribution of detectable sources differs substantially from that of the parent population.
+We can then consider two different strong lensing bias definitions: one that quantifies the difference in lensed source properties with respect to the parent population, and one that describes the difference with respect to the detectable source population.
+We are mostly interested in the second definition.
+
+\begin{figure*}
+\includegraphics[width=\textwidth]{source_cornerplot.eps}
+\caption{
+Comparison bewteen the properties of lens samples and the general population. Distribution in background source parameters.
+Filled contours: distribution of the parent sample.
+Black solid lines: distribution of the detectable sources.
+Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
+Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
+\label{fig:sourcepars}
+}
+\end{figure*}
+
+Strong lensing tends to preferentially select sources at higher redshift, especially in the more restrictive case with $\theta_{\mathrm{Ein,min}} > 1.0''$.
+This is because, at fixed lens properties, increasing the source redshift lowers the critical surface mass density and, consequently, increases the size of the caustics and the Einstein radius.
+The distribution in the magnitude of the lensed sources is also very different from that of the parent distribution, as it drops rapidly for values larger than $\msource\approx25$.
+Interestingly, however, there does not seem to be a large difference with respect to the distribution of detectable sources in the absence of lensing.
+This result can appear to be somewhat counter-intuitive: lensing magnification should allow the detection of sources that are intrinsically fainter than the detection limit.
+To some extent, this is the case: the distribution of lensed sources shows a slight excess of fainter galaxies compared to the unlensed case.
+However, as we quantify later in this section, the difference is far from large.
+The reason for this behaviour lies in the fact that, in the detection of both lensed and unlensed sources, the most important quantity is surface brightness, which is preserved by gravitational lensing.
+As we showed in section \ref{ssec:ellext}, the lensing cross-section drops to zero once the surface brightness of the source reaches a value that would make it undetectable in the absence of lensing.
+For this reason, also the distribution in half-light radius is very similar between the lensed sources and the detectable source population.
+%We can also see a bias in favour of sources with a very low S\'{e}rsic index.
+This is a consequence of the selection in magnitude: our source sample shows a strong correlation between $\msource$ and $n$: the brightest sources tend to have the smallest values of the S\'{e}rsic index.
+
+Qualitatively, the results shown in \Fref{fig:lenspars} and \Fref{fig:sourcepars} match our expectations from \Sref{sect:indlenses}.
+In the rest of this section we quantify the lensing bias.
+We present the results in three different parts.
+In the first part, we focus on the properties of the lens galaxies that can be observed directly.
+These are quantities that can be derived from photometry and spectroscopy with minimal assumptions: the lens redshift, the observed stellar mass, the half-light radius and the axis ratio.
+Any bias in these quantities can be determined relatively easily in a real strong lens survey.
+In the second part, we consider the parameters related to the mass distribution: the stellar population synthesis mismatch parameter and the dark matter distribution parameters.
+Determining the bias on these parameters is much more difficult, but these are quantities of great interest from a galaxy science point of view.
+In the third part, we focus on source distribution parameters.
+
+\subsubsection{Bias in observable lens parameters}
+
+\subsubsection{Bias in lens mass parameters}
+
+\subsubsection{Bias in source parameters}
 

### 2022-08-15 paper/seleff.tex
Using i-band instead of g-band. Started bias plots

@@
 
+\begin{figure}
+\includegraphics[width=\columnwidth]{lens_observable_bias.eps}
+\caption{
+Bias on lens observable properties as a function of the minimum Einstein radius.
+First panel: median redshift.
+Second panel: median observed stellar mass.
+Third panel: median half-light radius at an observed mass of $\log{\mobs}=11.4$.
+Fourth panel: median axis ratio.
+In each panel, the black dashed line indicates the value of the parent population.
+Error bars indicate the standard deviation of the mean of the lens sample.
+}
+\end{figure}
+
 \subsubsection{Bias in observable lens parameters}

### 2022-08-15 lens_population/draw_detectable_sources.py

@@ -14,7 +14,7 @@ import sys
 np.random.seed(20)
 
 ndraw = 40000
-sourcecat = pyfits.open('/data2/sonnenfeld/skills_sourceonly_zcut.fits')[1].data
+sourcecat = pyfits.open('/Users/alessandro/catalogs/skills_sourceonly_zcut.fits')[1].data
 
 psf = pyfits.open('psf.fits')[0].data
 pix_arcsec = 0.1
@@ -29,7 +29,7 @@ np.random.shuffle(sourceind)
 sreff_draw = sourcecat['Re_arcsec_CM'][sourceind[:ndraw]]
 nser_draw = sourcecat['sersic_n_CM'][sourceind[:ndraw]]
 sq_draw = sourcecat['axis_ratio_CM'][sourceind[:ndraw]]
-smag_draw = sourcecat['g_SDSS_apparent_corr'][sourceind[:ndraw]]
+smag_draw = sourcecat['i_SDSS_apparent_corr'][sourceind[:ndraw]]
 zs_draw = sourcecat['zobs'][sourceind[:ndraw]]
 spa_draw = sourcecat['PA_random'][sourceind[:ndraw]]
 
@@ -63,7 +63,7 @@ f = open('unlensed.input', 'w')
 f.writelines(outlines)
 f.close()
 
-os.system('glafic unlensed.input')
+os.system('/Users/alessandro/glafic unlensed.input')
 
 detected = np.zeros(ndraw, dtype=bool)
 

### 2022-08-15 lens_population/lens_extended_sources.py

@@ -115,7 +115,7 @@ for i in range(nsamp):
                 sq = -sq
             spa = sourcecat['PA_random'][sourceind]
             zs = sourcecat['zobs'][sourceind]
-            smag = sourcecat['g_SDSS_apparent_corr'][sourceind]
+            smag = sourcecat['i_SDSS_apparent_corr'][sourceind]
 
             ftot = 10.**(-2./5.*(smag - zeropoint))
             I0 = ftot/(2.*np.pi*(sreff/pix_arcsec)**2*nser/sersic.b(nser)**(2*nser)*gfunc(2.*nser))

### 2022-08-15 lens_population/make_lens_cornerplot.py

@@ -65,7 +65,7 @@ for i in range(npars):
     pylab.hist(lens_samp[pars[i]], bins=bins, color=lens_color, histtype='step', weights=lweights, linewidth=2, label="Lenses, $\\theta_{\mathrm{Ein}} > 0.5''$")
 
     tweights = np.ones(ntein1)/float(ntein1)/(bins[1] - bins[0])
-    pylab.hist(tein1_samp[pars[i]], bins=bins, color=tein1_color, histtype='step', weights=tweights, linewidth=2, label="Lenses, $\\theta_{\mathrm{Ein}} > 1.0''$")
+    pylab.hist(tein1_samp[pars[i]], bins=bins, color=tein1_color, histtype='step', weights=tweights, linewidth=2, linestyle='--', label="Lenses, $\\theta_{\mathrm{Ein}} > 1.0''$")
 
     if i==0:
         ylim = pylab.ylim()
@@ -97,7 +97,7 @@ for j in range(1, npars): # loops over rows
 
         probcontour(gal_samp[pars[i]], gal_samp[pars[j]], color=gal_color, style='filled', linewidths=2)
         probcontour(lens_samp[pars[i]], lens_samp[pars[j]], color=lens_color, style='lines', linewidths=2, smooth=5)
-        probcontour(tein1_samp[pars[i]], tein1_samp[pars[j]], color=tein1_color, style='lines', linewidths=2, smooth=7)
+        probcontour(tein1_samp[pars[i]], tein1_samp[pars[j]], color=tein1_color, style='lines', linewidths=2, linestyles='--', smooth=7)
 
         ax.set_xlim(lims[i])
         ax.set_ylim(lims[j])

### 2022-08-15 lens_population/make_source_cornerplot.py

@@ -22,14 +22,14 @@ modelname = 'fiducial_1000sqdeg'
 
 lenspop = h5py.File('%s_lenses.hdf5'%modelname, 'r')
 #sourcecat = pyfits.open('/Users/alessandro/catalogs/skills_sourceonly_zcut.fits')[1].data
-sourcecat = pyfits.open('/data2/sonnenfeld/skills_sourceonly_zcut.fits')[1].data
+sourcecat = pyfits.open('/Users/alessandro/catalogs/skills_sourceonly_zcut.fits')[1].data
 
 detect_file = h5py.File('detectable_sources.hdf5', 'r')
 
 tein05 = lenspop['tein_zs'][()] > 0.5
 tein1 = lenspop['tein_zs'][()] > 1.
 
-catpars = ['zobs', 'g_SDSS_apparent_corr', 'Re_arcsec_CM', 'sersic_n_CM', 'axis_ratio_CM']
+catpars = ['zobs', 'i_SDSS_apparent_corr', 'Re_arcsec_CM', 'sersic_n_CM', 'axis_ratio_CM']
 pars = ['zs', 'smag', 'sreff', 'nser', 'sq']
 npars = len(pars)
 
@@ -138,5 +138,5 @@ for j in range(1, npars): # loops over rows
             ax.tick_params(axis='x', labelbottom=False)
 
 pylab.savefig('../paper/source_cornerplot.eps')
-#pylab.show()
+pylab.show()
 

### 2022-08-15 lens_population/plot_lens_mass_bias.py

@@ -1,0 +1,112 @@
+import numpy as np
+import pylab
+from astropy.io import fits as pyfits
+import h5py
+from simpars import *
+from scipy.optimize import leastsq
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 20
+
+sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg']
+labels = ['Fiducial', 'High scatter']
+nsims = len(sims)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+
+fig, ax = pylab.subplots(4, 1, figsize=(6, 12))
+
+pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
+
+ntein = 21
+tein_arr = np.linspace(0., 2., ntein)
+
+for n in range(nsims):
+
+    galpop = h5py.File('%s_galaxies.hdf5'%sims[n], 'r')
+
+    laspsmed_gal = np.median(galpop['lasps'][()])
+    laspsmed_arr = np.zeros(ntein)
+    laspserr_arr = np.zeros(ntein)
+
+    lm200med_gal = mu_h
+    lm200med_arr = np.zeros(ntein)
+    lm200err_arr = np.zeros(ntein)
+
+    for i in range(ntein):
+        lenscut = galpop['islens'][()] & (galpop['tein_zs'][()] > tein_arr[i])
+        nlens = lenscut.sum()
+        print('Theta_Ein > %2.1f. %d lenses'%(tein_arr[i], lenscut.sum()))
+        laspsmed_arr[i] = np.median(galpop['lasps'][lenscut])
+        laspserr_arr[i] = np.std(galpop['lasps'][lenscut])/float(nlens)**0.5
+
+        # fits for the stellar-halo mass relation
+
+        lm200_here = galpop['lm200'][lenscut]
+        lmstar_here = galpop['lmstar'][lenscut]
+
+        def fitfunc(p):
+            return p[0] + p[1] * (lmstar_here - lmstar_piv)
+
+        def errfunc(p):
+            return fitfunc(p) - lm200_here
+
+        pfit = leastsq(errfunc, [13., 1.])
+
[…]

### 2022-08-15 lens_population/plot_lens_observable_bias.py

@@ -1,0 +1,125 @@
+import numpy as np
+import pylab
+from astropy.io import fits as pyfits
+import h5py
+from simpars import *
+from scipy.optimize import leastsq
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 20
+
+sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg']
+labels = ['Fiducial', 'High scatter']
+nsims = len(sims)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+
+fig, ax = pylab.subplots(4, 1, figsize=(6, 12))
+
+pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
+
+ntein = 21
+tein_arr = np.linspace(0., 2., ntein)
+
+for n in range(nsims):
+
+    galpop = h5py.File('%s_galaxies.hdf5'%sims[n], 'r')
+
+    zmed_gal = np.median(galpop['z'][()])
+    zmed_arr = np.zeros(ntein)
+    zerr_arr = np.zeros(ntein)
+
+    lmobsmed_gal = np.median(galpop['lmobs'][()])
+    lmobsmed_arr = np.zeros(ntein)
+    lmobserr_arr = np.zeros(ntein)
+
+    mbin_gal = (galpop['lmobs'][()] > 11.38) & (galpop['lmobs'][()] < 11.42)
+    lreffmed_gal = np.median(galpop['lreff'][mbin_gal])
+    lreffmed_arr = np.zeros(ntein)
+    lrefferr_arr = np.zeros(ntein)
+
+    qmed_gal = np.median(galpop['q'][()])
+    qmed_arr = np.zeros(ntein)
+    qerr_arr = np.zeros(ntein)
+
+    for i in range(ntein):
+        lenscut = galpop['islens'][()] & (galpop['tein_zs'][()] > tein_arr[i])
+        nlens = lenscut.sum()
+        print('Theta_Ein > %2.1f. %d lenses'%(tein_arr[i], lenscut.sum()))
+        zmed_arr[i] = np.median(galpop['z'][lenscut])
+        zerr_arr[i] = np.std(galpop['z'][lenscut])/float(nlens)**0.5
+
+        lmobsmed_arr[i] = np.median(galpop['lmobs'][lenscut])
[…]

### 2022-08-16 paper/seleff.tex
finished galaxy lenses results

@@
 
-The luminosity and size distribution of strongly lensed sources is also biased with respect to the general population of background galaxies.
-For instance, lensing magnification allows the detection of fainter objects with respect to the field, and galaxies that are very extended with respect to the Einstein radius are less likely to be classified as strongly lensed, since their magnification averaged over their size tends to be low \citep[see also the discussion in section 5.3 of][]{O+A17}.
+The luminosity and size distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies.
+For instance, lensing magnification allows the detection of fainter objects with respect to the field, and, in case lenses are selected on the basis of the magnification of the source, galaxies that are very extended with respect to the Einstein radius are less likely to be classified as strongly lensed.
+This is because their magnification averaged over their size tends to be low \citep[see the discussion in section 5.3 of][]{O+A17}.
 
@@
 \end{equation}
-In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the general distribution of foreground galaxies and background sources in the absence of lensing; and $\pdet$ describes the probability of detecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
+In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\pdet$ describes the probability of detecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
 This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be recognised as a strong lens.
@@
 %, under realistic assumptions for the lens detection probability $\pdet$.
-In particular, we aim to determine how strong lenses differ from the general population of foreground galaxies and background sources in terms of a) the radial mass structure of the lenses (i.e.
+In particular, we aim to determine how strong lenses differ from the parent population of foreground galaxies and background sources in terms of a) the radial mass structure of the lenses (i.e.
 their stellar and dark matter mass density profiles); b) the ellipticity of the lenses; c) the size-magnitude distribution of the lensed sources.
@@
 From here on we focus on the effect that those trends have on populations of lenses.
-We addressed this question by simulating populations of foreground galaxies and background sources, selecting strong lenses among them, and comparing the properties of the strong lens sample with the general galaxy population.
+We addressed this question by simulating populations of foreground galaxies and background sources, selecting strong lenses among them, and comparing the properties of the strong lens sample with the parent galaxy population.
 We simulated a lens-based search (as opposed to a source-based one), in which strongly lensed images are searched among a stellar mass-selected sample of galaxies.
@@
 
-\subsubsection{Dark matter distribution}
+\subsubsection{Dark matter distribution}\label{ssec:dmprofile}
 
@@
 
-%Both for the lensed galaxies and lensed quasars, we apply an additional selection based on the angular size of the Einstein radius.
-Both for the lensed galaxies and lensed quasars, we apply an additional cut, by selecting only lenses with an angular Einstein radius larger than a limiting value $\theta_{\mathrm{Ein,min}}$.
-This additional cut is meant to approximate the difficulty in detecting strongly lensed images with small separation, because of the contamination from the light of the lens galaxy.
+\subsection{Einstein radius cuts}
+
+\Fref{fig:teinhist} shows the distribution in Einstein radius of the simulated lens samples.
+In all cases, the distribution peaks at $\tein\approx0.7''$.
+Although all of the lenses in these samples are detected, it does not necessarily follow that they would all be included in a strong lensing study.
+There can be a few reasons for escluding certain lenses from a sample. 
+One is the low accuracy of lens finders: current automated lens finding algorithms tend to produce lens candidates samples with low purity \citep[see e.g.][]{Son++18a,Pet++19,Sav++22}.
+Such samples are then visually inspected, and only those candidates that can be clearly distinguished from false-positives are kept.
+This visual inspection step tends to disfavour lenses with a small image separation, because of the contamination from the light of the lens galaxy.
+\begin{figure}
+\includegraphics[width=\columnwidth]{tein_hist.eps}
+\caption{
+Einstein radius distribution of the simulated lens samples.
+\label{fig:teinhist}
+}
+\end{figure}
+
+Another reason for refining a sample of lens candidates is the availability of redshift measurements.
+Redshifts of both the lens and the source are needed in order to convert a lens model into a measurement of mass.
+When working with large samples of lenses, obtaining spectroscopic measurements is not a viable option, and photometric redshifts are a necessity.
+Measuring the photometric redshift of a strongly lensed source, however, is challenging, especially when the Einstein radius is small and the source light is blended with the light from the lens \citep{Lan++22}.
+
+Both of these scenarios can result in samples in which lenses with a larger Einstein radius are preferentially selected.
+In order to simulate these situations, we apply an additional cut to lens samples, by selecting only lenses with an angular Einstein radius larger than a limiting value $\theta_{\mathrm{Ein,min}}$.
+This is equivalent to multiplying $\pdet$ by a step function in the Einstein radius, which is zero for $\tein < \theta_{\mathrm{Ein,min}}$ and one otherwise.
+%Both for the lensed galaxies and lensed quasars, we apply an additional cut, by selecting only lenses with an angular Einstein radius larger than a limiting value $\theta_{\mathrm{Ein,min}}$.
+%This additional cut is meant to approximate the difficulty in detecting strongly lensed images with small separation, because of the contamination from the light of the lens galaxy.
 We explored scenarios with different values of $\theta_{\mathrm{Ein,min}}$.
+As the next section shows, the larger the minimum Einstein radius, the higher the strong lensing bias.
 
@@
 \caption{
-Comparison bewteen the properties of lens samples and the general population. Distribution in foreground galaxy parameters.
+Comparison bewteen the properties of lens samples and the parent population. Distribution in foreground galaxy parameters.
 Filled contours: distribution of the parent sample.
@@
 \caption{
-Comparison bewteen the properties of lens samples and the general population. Distribution in background source parameters.
+Comparison bewteen the properties of lens samples and the parent population. Distribution in background source parameters.
 Filled contours: distribution of the parent sample.
@@
 We present the results in three different parts.
-In the first part, we focus on the properties of the lens galaxies that can be observed directly.
+First, we focus on the properties of the lens galaxies that can be observed directly.
 These are quantities that can be derived from photometry and spectroscopy with minimal assumptions: the lens redshift, the observed stellar mass, the half-light radius and the axis ratio.
@@
 In the third part, we focus on source distribution parameters.
+
+\subsubsection{Bias in observable lens parameters}
+
+\Fref{fig:lensobsbias} shows the median redshift, median $\mobs$, median size for a given $\mobs$, and median axis ratio of various lens population simulations, as a function of the minimum Einstein radius.
+We defined the median size at fixed $\mobs$ by fitting the following mass-size relation to the $\reff-\mobs$ distribution:
+\begin{equation}
+\log{\reff} \sim \mu_{R,0} + \beta_R(\log{\mobs} - 11.4).
+\end{equation}
+The quantity shown in the third panel of \Fref{fig:lensobsbias} is the parameter $\mu_{R,0}$, which is the average $\log{\reff}$ at an observed stellar mass of $\log{\mobs}=11.4$.
+The horizontal dashed line in each panel shows the value of the parent population: the larger the distance between the curve of a simulation and this line, the higher the strong lensing bias.
+
+%\Fref{fig:lensobsbias} shows the median of various lens galaxy observable quantities as a function of the minimum Einstein radius, for the three lens population simulations
 
@@
 Error bars indicate the standard deviation of the mean of the lens sample.
-}
-\end{figure}
-
-\subsubsection{Bias in observable lens parameters}
+\label{fig:lensobsbias}
+}
+\end{figure}
+
+As was already visible in \Fref{fig:lenspars}, there is a clear bias towards lower redshift, higher stellar mass and smaller sizes, with the bias becoming stronger for more restrictive cuts in Einstein radius.
+Biases in stellar mass and size are stronger for the simulation with low scatter in $\asps$ and $\mhalo$.
+This can be explained as follows.
+For a given value of the observed stellar mass, strong lensing selection favors galaxies with a larger $\asps$ or $\mhalo$, or with a smaller half-light radius.
+Since the galaxy stellar mass function of the parent population is steep, lenses tend to have a relatively small $\mobs$ and large values of $\asps$ or $\mhalo$ for their observed stellar mass (this is shown more clearly in the next section).
+If the intrinsic scatter in $\asps$ and $\mhalo$ is low, however, the number of galaxies with a small $\mobs$ and a large stellar or halo mass is greatly reduced.
+Only galaxies with a large $\mobs$ or a small size can therefore act as strong lenses.
 
@@
 
+\Fref{fig:lensmassbias} shows the median of the distribution in various mass-related quantities, as a function of the minimum Einstein radius.
+The top panel is the median stellar population synthesis mismatch parameter.
+In all simulations, strong lenses are biased towards larger values than the parent population.
+The bias is larger the higher the intrinsic scatter in $\asps$ and $\mhalo$, and increases with increasing $\theta_{\mathrm{Ein,min}}$.
+For the fiducial model, the bias on $\asps$ can be as small as $0.03$~dex (7\%), if no cut on Einstein radius is applied.
+However, it can rise up to $0.09$~dex (23\%) when considering only lenses with $\tein > 2''$.
+
+The second panel of \Fref{fig:lensmassbias} shows the median halo mass of galaxies with a stellar mass of $\log{\mstar}=11.5$.
+We measured this quantity by fitting the following relation to the $\mhalo-\mstar$ distribution,
+\begin{equation}
+\log{\mhalo} \sim \mu_{\mathrm{h},0} + \beta_{\mathrm{h}}(\log{\mstar} - 11.5),
+\end{equation}
+and taking the resulting value of the parameter $\mu_{\mathrm{h},0}$.
+Similarly to the $\asps$ case, the strong lenses are biased towards larger values of the halo mass, with the bias being larger for higher-scatter simulations and more restrictive cuts on $\tein$.
+In the fiducial scatter scenario with a minimum Einstein radius of $\theta_{\mathrm{Ein,min}}=1.0''$, the halo masses of lenses are on average $0.1$~dex larger than those of their parent population.
+
+The third panel of \Fref{fig:lensmassbias} shows the median projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$, at fixed stellar mass and half-light radius.
+We obtained this quantity by fitting the following relation to the $\mdmfive-\mstar-\reff$ distribution:
+\begin{equation}\label{eq:dmscaling}
+\log{\mdmfive} \sim \mu_{\mathrm{DM},0} + \beta_{\mathrm{DM}}(\log{\mstar} - 11.5) + \xi_{\mathrm{DM}}(\log{\reff} - 1.2).
+\end{equation}
+\Fref{fig:lensmassbias} shows the value of $\mu_{\mathrm{DM},0}$.
+The bias on this quantity is qualitatively similar to that on the total halo mass, but much smaller in amplitude.
+There are two reasons for this.
+First, the dependence of $\mdmfive$ on $\mhalo$ is shallower than linear.
+This is because, as the virial mass increases, the virial radius increases as well: the extra mass is spread over a larger volume, and therefore the mass within the inner region does not increase proportionally.
+Second, while $\mu_{\mathrm{h},0}$ is the halo mass at fixed stellar mass, $\mu_{\mathrm{DM},0}$ is measured at fixed half-light radius as well.
+The density profile of the dark matter halos in our simulation have a dependence on galaxy size: the response of dark matter to baryons is stronger for more concentrated stellar distributions (see section \ref{ssec:dmprofile}).
+By capturing this dependence in the model of \Eref{eq:dmscaling}, the residual scatter in $\mdmfive$ around the mean is reduced, and so is the strong lensing bias.
+
+In the fiducial scenario, the bias on $\mdmfive$ is negligible for $\theta_{\mathrm{Ein,min}} < 0.5''$.
+Interestingly, the bias is negative in the low-scatter scenario.
+This is a counter-intuitive result that follows from the fact that the model of \Eref{eq:dmscaling} is not a perfect description of the distribution in $\mdmfive$.
+Nevertheless, the amplitude of the bias is only $0.01$~dex, which for all practical purposes is negligible.
+
+In the fourth panel of \Fref{fig:lensmassbias} we show the average inner dark matter slope at fixed stellar mass and half-light radius, which we measured by fitting the following model
+\begin{equation}\label{eq:dmscaling}
+\gammadm \sim \mu_{\gamma,0} + \beta_{\gamma}(\log{\mstar} - 11.5) + \xi_{\gamma}(\log{\reff} - 1.2),
+\end{equation}
+and taking the resulting value of $\mu_{\gamma}$.
+The trends in the avereage $\gammadm$ reflect very closely those in $\mdmfive$, but with flipped sign.
+This is because, in our model for the dark matter density profile, inner dark matter mass and density slope are perfectly correlated: both $\mdmfive$ and $\gammadm$ are determined univocally once the halo mass, stellar mass and half-light radius are specified.
+The fact that the bias is negative is somewhat counter-intuitive: lenses with a stronger response to baryons have both a larger value of $\mdmfive$ and a steeper inner dark matter slope.
+However, $\gammadm$ is also anti-correlated with total halo mass (see $\gammadm-\mhalo$ panel in \Fref{fig:lenspars}): since strong lenses tend to have a preferentially larger halo mass, they also have a shallower dark matter slope.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{lens_mass_bias.eps}
+\caption{
+Bias on lens mass properties as a function of the minimum Einstein radius.
+First panel: median $\log{\asps}$.
+Second panel: median $\mhalo$ at $\log{\mstar}=11.5$.
+Third panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+Fourth panel: median $\gammadm$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+In each panel, the dashed line indicates the value of the parent population.
+Error bars indicate the standard deviation of the mean of the lens sample.
+\label{fig:lensmassbias}
+}
+\end{figure}
+
 \subsubsection{Bias in source parameters}
+
+\Fref{fig:sourcebias} shows the lensing bias in source-related parameters, with respect to the population of sources that are detectable without lensing.
+As previously seen in \Fref{fig:sourcepars}, lensed sources are biased towards higher redshift (top panel of \Fref{fig:sourcebias}), with the trend being larger for more restrictive cuts on the Einstein radius.
+The second panel shows the bias in the source magnitude.
+This bias is negative for small values of $\theta_{\mathrm{Ein,min}}$, that is, lensed sources tend to be brighter than their field counterparts.
+Naively we would have expected the opposite trend, as strong lensing magnification allows the detection of sources that are intrinsically fainter than the detection limit.
+However, as the analysis of \Sref{sect:indlenses} clearly shows, the lensing cross-section is always larger for brighter sources, and that explains the sign of the bias.
+At the same time, the median does not capture the whole picture of the bias in $\msource$.
+For instance, in the fiducial simulation with no cut on the Einstein radius, the $90\%$-ile of the $\msource$ distribution is $26.05$, while that of the population of detectable sources is $0.14$~mag brighter: indeed, strong lensing allows the detection of fainter sources.
+
+The third and fourth panels of \Fref{fig:sourcepars} show the bias in half-light radius and S\'{e}rsic index, respectively, for sources within a magnitude bin centred on $\msource=25$ and with a $0.4$~mag width.
+We detected no clear sign of bias.
+Finally, the fifth panel shows the bias in source axis ratio. Also in this case no obvious sign of bias was detected, except in the largest values of $\theta_{\mathrm{Ein,min}}$.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{source_bias.eps}
+\caption{
+Bias on source properties as a function of the minimum Einstein radius.
+First panel: median source redshift.
+Second panel: median source magnitude.
+Third panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
+Fourth panel: median S\'{e}rsic index in the magnitude bin $24.8 < \msource < 25.2$.
+Fifth panel: median axis ratio.
+In each panel, the dashed line indicates the value of the population of detectable sources.
+Error bars indicate the standard deviation of the mean of either the full sample (first, second and fifth panel) or the bin (third and fourth panel).
+\label{fig:sourcebias}
+}
+\end{figure}
 

### 2022-08-16 lens_population/plot_lens_mass_bias.py

@@ -13,8 +13,8 @@ rc('text', usetex=True)
 
 fsize = 20
 
-sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg']
-labels = ['Fiducial', 'High scatter']
+sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg', 'lowscatter_1000sqdeg']
+labels = ['Fiducial', 'High scatter', 'Low scatter']
 nsims = len(sims)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
@@ -38,6 +38,32 @@ for n in range(nsims):
     lm200med_arr = np.zeros(ntein)
     lm200err_arr = np.zeros(ntein)
 
+    lmdm5med_arr = np.zeros(ntein)
+    lmdm5err_arr = np.zeros(ntein)
+
+    def mdm5fitfunc(p):
+        return p[0] + p[1] * (galpop['lmstar'][()] - lmstar_piv) + p[2] * (galpop['lreff'][()] - lreff_piv)
+
+    def mdm5errfunc(p):
+        return mdm5fitfunc(p) - galpop['lmdm5'][()]
+
+    pmdm5fit = leastsq(mdm5errfunc, [11., 0., 0.])
+
+    lmdm5med_gal = pmdm5fit[0][0]
+
+    gammamed_arr = np.zeros(ntein)
+    gammaerr_arr = np.zeros(ntein)
+
+    def gammafitfunc(p):
+        return p[0] + p[1] * (galpop['lmstar'][()] - lmstar_piv) + p[2] * (galpop['lreff'][()] - lreff_piv)
+
+    def gammaerrfunc(p):
+        return gammafitfunc(p) - galpop['gammadm'][()]
+
+    pgammafit = leastsq(gammaerrfunc, [1.5, 0., 0.])
+
+    gammamed_gal = pgammafit[0][0]
+
     for i in range(ntein):
         lenscut = galpop['islens'][()] & (galpop['tein_zs'][()] > tein_arr[i])
         nlens = lenscut.sum()
@@ -49,6 +75,9 @@ for n in range(nsims):
 
         lm200_here = galpop['lm200'][lenscut]
         lmstar_here = galpop['lmstar'][lenscut]
+        lreff_here = galpop['lreff'][lenscut]
+        lmdm5_here = galpop['lmdm5'][lenscut]
+        gammadm_here = galpop['gammadm'][lenscut]
 
[…]

### 2022-08-16 lens_population/plot_lens_observable_bias.py

@@ -13,8 +13,8 @@ rc('text', usetex=True)
 
 fsize = 20
 
-sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg']
-labels = ['Fiducial', 'High scatter']
+sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg', 'lowscatter_1000sqdeg']
+labels = ['Fiducial', 'High scatter', 'Low scatter']
 nsims = len(sims)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
@@ -80,44 +80,46 @@ for n in range(nsims):
         qerr_arr[i] = np.std(galpop['q'][lenscut])/float(nlens)**0.5
 
     ax[0].errorbar(tein_arr, zmed_arr, yerr=zerr_arr, color=colseq[n], label=labels[n])
-    ax[0].axhline(zmed_gal, color='k', linestyle='--', label='Parent population')
-    ax[0].set_ylabel('Median $z$', fontsize=fsize)
-
-    ax[0].yaxis.set_major_locator(MultipleLocator(0.05))
-    ax[0].yaxis.set_minor_locator(MultipleLocator(0.01))
-
     ax[1].errorbar(tein_arr, lmobsmed_arr, yerr=lmobserr_arr, color=colseq[n])
-    ax[1].axhline(lmobsmed_gal, color='k', linestyle='--')
-    ax[1].set_ylabel('Median $\log{M_*^{(\mathrm{obs})}}$', fontsize=fsize)
-
-    ax[1].yaxis.set_major_locator(MultipleLocator(0.2))
-    ax[1].yaxis.set_minor_locator(MultipleLocator(0.05))
-
     ax[2].errorbar(tein_arr, lreffmed_arr, yerr=lrefferr_arr, color=colseq[n])
-    ax[2].axhline(lreffmed_gal, color='k', linestyle='--')
-    ax[2].set_ylabel('$\mu_{\mathrm{R}}$', fontsize=fsize)
-
-    ax[2].yaxis.set_major_locator(MultipleLocator(0.05))
-    ax[2].yaxis.set_minor_locator(MultipleLocator(0.01))
-
     ax[3].errorbar(tein_arr, qmed_arr, yerr=qerr_arr, color=colseq[n])
-    ax[3].axhline(qmed_gal, color='k', linestyle='--')
[…]

### 2022-08-16 lens_population/plot_source_bias.py

@@ -1,0 +1,138 @@
+import numpy as np
+import pylab
+from astropy.io import fits as pyfits
+import h5py
+from simpars import *
+from scipy.optimize import leastsq
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 20
+
+smag_min = 24.8
+smag_max = 25.2
+
+sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg', 'lowscatter_1000sqdeg']
+labels = ['Fiducial', 'High scatter', 'Low scatter']
+nsims = len(sims)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+
+fig, ax = pylab.subplots(5, 1, figsize=(6, 12))
+
+pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
+
+ntein = 21
+tein_arr = np.linspace(0., 2., ntein)
+
+detectable = h5py.File('detectable_sources.hdf5', 'r')
+
+zmed_det = np.median(detectable['zs'][()])
+smagmed_det = np.median(detectable['smag'][()])
+
+smagbin_det = (detectable['smag'][()] > smag_min) & (detectable['smag'][()] < smag_max)
+
+nsermed_det = np.median(detectable['nser'][smagbin_det])
+sreffmed_det = np.median(detectable['sreff'][smagbin_det])
+
+qmed_det = np.median(detectable['sq'][()])
+
+for n in range(nsims):
+
+    lenspop = h5py.File('%s_lenses.hdf5'%sims[n], 'r')
+
+    zmed_arr = np.zeros(ntein)
+    zerr_arr = np.zeros(ntein)
+
+    smagmed_arr = np.zeros(ntein)
+    smagerr_arr = np.zeros(ntein)
+
+    sreffmed_arr = np.zeros(ntein)
+    srefferr_arr = np.zeros(ntein)
+
+    nsermed_arr = np.zeros(ntein)
+    nsererr_arr = np.zeros(ntein)
+
+    qmed_arr = np.zeros(ntein)
+    qerr_arr = np.zeros(ntein)
+
+    for i in range(ntein):
+        teincut = (lenspop['tein_zs'][()] > tein_arr[i])
+        nlens = teincut.sum()
[…]

### 2022-08-18 paper/seleff.tex
quasar lenses

@@
 \def\zlens{z_{\mathrm{g}}}
+\def\zqso{z_{\mathrm{qso}}}
 
 \def\msource{m_{\mathrm{s}}}
+\def\mqso{m_{\mathrm{qso}}}
 
@@
 
+We describe the population of background quasars with the following double power-law luminosity function in the rest-frame UV absolute magnitude $M$:
+\begin{equation}
+\Phi(M,\zqso) = \dfrac{\Phi(M^*)}{10^{0.4(\alpha_Q+1)(M - M^*)} + 10^{0.4(\beta_Q+1)(M - M^*)}}.
+\end{equation}
+Following \citet{Man++17}, we set $\alpha_Q=-1.35$ (faint-end slope), $\beta_Q=-3.23$ (bright-end slope), and adopted a redshift-evolving normalisation
+\begin{equation}
+\log{\Phi^*} = -6.0991 + 0.0209\zqso + 0.0171\zqso^2,
+\end{equation}
+and charachteristic magnitude
+\begin{equation}
+M^* = -22.5216 - 1.6510\zqso + 0.2869\zqso^2.
+\end{equation}
+Given the redshift and rest-frame UV luminosity of a quasar, we then computed the apparent magnitude in the observed $i-$band, $\mqso$, using a quasar spectral template\footprint{\url{https://archive.stsci.edu/hlsps/reference-atlases/cdbs/grid/comp_qso/}} built from optical and near-infrared spectra obtained by \citet{vdB++01,Gli++06}.
+
+For the sake of consistency with the population of background extended sources, we limited the redshift distribution of quasars to the range $0.8 < \zqso < 2.5$.
+We then truncated the distribution in $\mqso$ at two magnitudes fainter than the detection limit (which is specified in section \ref{ssub:obs}).
+Finally, we randomly placed quasars in the source plane with a projected number density of $70$~arcmin$^{-2}$.
+This is a much larger number density than observed in the real universe.
+Nevertheless, we are not interested in predicting the absolute number of strong lenses, so this is a viable choice.
+The advantage of boosting the number density of quasars is that it allows us to produce a large number of lenses without the need for generating too big a population of foreground galaxies.
+
 \subsection{Observations}\label{ssub:obs}
@@
 For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering $1000$ square degrees.
-The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is YYY.
+The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around $300000$.
 %Then we generated a population of sources behind the lenses.
@@
 Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
-%This is similar to the background noise expected for the Euclid Wide survey \citep{Sca++22}.
 These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
@@
 
-%First, we used the criterion described in section \ref{ssec:lensdefext}.
-%This resulted in a sample of XXX lenses.
-%Then, we added a cut on the Einstein radius of the lens: we only included in the lens sample those with $\tein > 0.5''$.
-%This additional cut is meant to approximate the difficulty in detecting strongly lensed images that lie very close to the lens centre: the number of known lenses with Einstein radius smaller than that cutoff value is very small, even in lens surveys that are based on high-resolution data \citep[see e.g.][]{Gar++22}.
-%With this more restrictive criterion, the number of strong lenses in the simulated sample is YYY.
-
-%\subsubsection{Lensed quasars}
+The fiducial scatter simulation with extended background sources produced a sample of $2113$ lenses, corresponding to a number density of $2.1$~deg$^{-2}$.
+This is about a factor of five smaller than the number density predicted by \citet{Col15} for the Euclid survey.
+We believe that this is a result of differences in the description of the source population and the criteria used to define a strong lensing event.
 
@@
 This bias is negative for small values of $\theta_{\mathrm{Ein,min}}$, that is, lensed sources tend to be brighter than their field counterparts.
-Naively we would have expected the opposite trend, as strong lensing magnification allows the detection of sources that are intrinsically fainter than the detection limit.
+Naively one might have expected the opposite trend, as strong lensing magnification allows the detection of sources that are intrinsically fainter than the detection limit.
 However, as the analysis of \Sref{sect:indlenses} clearly shows, the lensing cross-section is always larger for brighter sources, and that explains the sign of the bias.

### 2022-08-18 lens_population/generate_lenses.py

@@ -15,9 +15,9 @@ np.random.seed(seedno)
 sky_area_deg2 = 1000.
 sky_area = sky_area_deg2 * np.deg2rad(1.)**2
 
-modelname = 'fiducial_%dsqdeg'%sky_area_deg2
-sigma_h = 0.2
-sigma_sps = 0.08
+modelname = 'lowscatter_%dsqdeg'%sky_area_deg2
+sigma_h = 0.1
+sigma_sps = 0.05
 
 print('Sky area: %d square degrees'%sky_area_deg2)
 print('Expected number of foreground galaxies: %2.1e'%(ngal_1Mpc3 * fourpi_volume * sky_area / (4.*np.pi)))

### 2022-08-18 lens_population/lens_quasars.py

@@ -3,15 +3,14 @@ import os
 import glafic
 import h5py
 from simpars import *
+from qsopars import *
 from scipy.signal import convolve2d
-from astropy.io import fits as pyfits
 from scipy.interpolate import splev
 from scipy.optimize import brentq
 from sl_profiles import sersic, gnfw, deVaucouleurs as deV
 import sl_cosmology
 from sl_cosmology import G, M_Sun, Mpc, c
 from scipy.special import gamma as gfunc
-from lensdet import detect_lens
 import sys
 
 
@@ -24,13 +23,22 @@ tein_zs_samp = np.zeros(nsamp)
 
 # reads the number of sources
 f = open('%s_sources.cat'%modelname, 'r')
-nsources = np.loadtxt(f, usecols=(1, ), dtype=int)
+nqso = np.loadtxt(f, usecols=(2, ), dtype=int)
 f.close()
 
-nsource_tot = nsources.sum()
+nqso_tot = nqso.sum()
 
-# generates redshift and magnitude
+# generates redshift and magnitudes
 
+zqso_samp = splev(np.random.rand(nqso_tot), invcum_zqso_spline)
+t_samp = np.random.rand(nqso_tot)
+
+qsomag_samp = np.zeros(nqso_tot)
+
+for i in range(nqso_tot):
+    ind = ztoind(zqso_samp[i])
+    qsomag_here = splev(t_samp[i], invcum_phiqso_splines[ind])
+    qsomag_samp[i] = qsomag_here
 
 # primary parameters
 omegaM = 0.3
@@ -42,7 +50,7 @@ xmin = -5.
 ymin = -5.
 xmax = 5.
 ymax = 5.
-pix_ext = pix_arcsec
+pix_ext = 0.1
 pix_poi = 0.1
 maxlev = 5
 
@@ -56,10 +64,10 @@ glafic.set_lens(1, 'gnfw', 0.3, 1e13, 0.0, 0.0, 0., 90.0, 10., 1.5)
 glafic.set_lens(2, 'sers', 0.3, 1e11, 0.0, 0.0, 0., 90.0, 1., 4.)
 glafic.set_point(1, zs_ref, 0., 0.)
 
-zs_list = []
+zqso_list = []
 xpos_list = []
 ypos_list = []
-smag_list = []
+qsomag_list = []
 nimg_list = []
 tein_zs_list = []
 
[…]

### 2022-08-18 lens_population/plot_lens_mass_bias.py

@@ -1,6 +1,5 @@
 import numpy as np
 import pylab
-from astropy.io import fits as pyfits
 import h5py
 from simpars import *
 from scipy.optimize import leastsq
@@ -13,8 +12,10 @@ rc('text', usetex=True)
 
 fsize = 20
 
-sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg', 'lowscatter_1000sqdeg']
-labels = ['Fiducial', 'High scatter', 'Low scatter']
+sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg', 'lowscatter_1000sqdeg', 'fiducial_1000sqdeg']
+sources = ['gal', 'gal', 'gal', 'qso']
+
+labels = ['Fiducial', 'High scatter', 'Low scatter', 'Quasars']
 nsims = len(sims)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
@@ -65,7 +66,11 @@ for n in range(nsims):
     gammamed_gal = pgammafit[0][0]
 
     for i in range(ntein):
-        lenscut = galpop['islens'][()] & (galpop['tein_zs'][()] > tein_arr[i])
+        if sources[n] == 'gal':
+            lenscut = galpop['islens'][()] & (galpop['tein_zs'][()] > tein_arr[i])
+        elif sources[n] == 'qso':
+            lenscut = galpop['qsolens'][()] & (galpop['tein_zqso'][()] > tein_arr[i])
+
         nlens = lenscut.sum()
         print('Theta_Ein > %2.1f. %d lenses'%(tein_arr[i], lenscut.sum()))
         laspsmed_arr[i] = np.median(galpop['lasps'][lenscut])

### 2022-08-18 lens_population/plot_source_bias.py

@@ -22,7 +22,7 @@ nsims = len(sims)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
-fig, ax = pylab.subplots(5, 1, figsize=(6, 12))
+fig, ax = pylab.subplots(5, 1, figsize=(6, 14))
 
 pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
 
@@ -121,7 +121,7 @@ ax[4].yaxis.set_major_locator(MultipleLocator(0.02))
 ax[4].yaxis.set_minor_locator(MultipleLocator(0.005))
 
 ax[0].tick_params(axis='both', which='both', direction='in', labelbottom=False, labelsize=fsize, right=True, top=True)
-ax[1].legend(loc='upper left', fontsize=fsize)
+ax[0].legend(loc='upper left', fontsize=fsize)
 
 ax[1].tick_params(axis='both', which='both', direction='in', labelbottom=False, labelsize=fsize, right=True, top=True)
 ax[2].tick_params(axis='both', which='both', direction='in', labelbottom=False, labelsize=fsize, right=True, top=True)

### 2022-08-19 paper/seleff.tex
paragraph on lensed quasars

@@
 \end{equation}
-Given the redshift and rest-frame UV luminosity of a quasar, we then computed the apparent magnitude in the observed $i-$band, $\mqso$, using a quasar spectral template\footprint{\url{https://archive.stsci.edu/hlsps/reference-atlases/cdbs/grid/comp_qso/}} built from optical and near-infrared spectra obtained by \citet{vdB++01,Gli++06}.
+Given the redshift and rest-frame UV luminosity of a quasar, we then computed the apparent magnitude in the observed $i-$band, $\mqso$, using a quasar spectral template\footnote{\url{https://archive.stsci.edu/hlsps/reference-atlases/cdbs/grid/comp_qso/}} built from optical and near-infrared spectra obtained by \citet{vdB++01,Gli++06}.
 
@@
 Einstein radius distribution of the simulated lens samples.
+These are: galaxy-galaxy lenses in the fiducial, high scatter and low scatter scenarios; galaxy-quasar lenses with fiducial scatter, considering all lenses or only lenses that produce four images.
 \label{fig:teinhist}
@@
 
-Another reason for refining a sample of lens candidates is the availability of redshift measurements.
+Another possible reason for refining a sample of lens candidates is the availability of redshift measurements.
 Redshifts of both the lens and the source are needed in order to convert a lens model into a measurement of mass.
@@
 First, we focus on the properties of the lens galaxies that can be observed directly.
-These are quantities that can be derived from photometry and spectroscopy with minimal assumptions: the lens redshift, the observed stellar mass, the half-light radius and the axis ratio.
+These are quantities that can be derived from photometry and spectroscopy with minimal assumptions: the lens redshift, the observed stellar mass, the half-light radius and the axis ratio\footnote{Strictly speaking, the axis ratio in mass is not necessarily observable, but in the context of our simulations it is, since light and mass have the same ellipticity.}.
 Any bias in these quantities can be determined relatively easily in a real strong lens survey.
@@
 \subsection{Galaxy-quasar lenses}
+
+The main goal of the experiment with lensed quasars is to check whether there are any differences in the strong lensing bias with respect to the extended source case, at fixed properties of the foreground galaxy population.
+For this reason, we only ran simulations with lensed quasars in the fiducial scatter scenario and compared the results with those from the extended source simulation.
+%We are also interested in understanding how the entire population of quasar lenses differs from the subset of quad lenses, therefore we computed the bias for 
+We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we computed the bias for this sample as well.
+%\Fref{fig:quasarbias} shows the strong lensing bias obtained with the lensed quasars simulation, compared with that 
+\Fref{fig:quasarbias} shows the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{quasar_bias.eps}
+\caption{
+
+}
+\end{figure}
 

### 2022-08-19 lens_population/plot_lens_mass_bias.py
refined plots

@@ -12,15 +12,14 @@ rc('text', usetex=True)
 
 fsize = 20
 
-sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg', 'lowscatter_1000sqdeg', 'fiducial_1000sqdeg']
-sources = ['gal', 'gal', 'gal', 'qso']
+sims = ['fiducial_1000sqdeg', 'highscatter_1000sqdeg', 'lowscatter_1000sqdeg']
 
-labels = ['Fiducial', 'High scatter', 'Low scatter', 'Quasars']
+labels = ['Fiducial', 'High scatter', 'Low scatter']
 nsims = len(sims)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
-fig, ax = pylab.subplots(4, 1, figsize=(6, 12))
+fig, ax = pylab.subplots(4, 1, figsize=(6, 13))
 
 pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
 
@@ -66,10 +65,7 @@ for n in range(nsims):
     gammamed_gal = pgammafit[0][0]
 
     for i in range(ntein):
-        if sources[n] == 'gal':
-            lenscut = galpop['islens'][()] & (galpop['tein_zs'][()] > tein_arr[i])
-        elif sources[n] == 'qso':
-            lenscut = galpop['qsolens'][()] & (galpop['tein_zqso'][()] > tein_arr[i])
+        lenscut = galpop['islens'][()] & (galpop['tein_zs'][()] > tein_arr[i])
 
         nlens = lenscut.sum()
         print('Theta_Ein > %2.1f. %d lenses'%(tein_arr[i], lenscut.sum()))
@@ -131,32 +127,45 @@ for n in range(nsims):
     ax[1].errorbar(tein_arr, lm200med_arr, yerr=lm200err_arr, color=colseq[n], label=labels[n], linewidth=2)
     ax[2].errorbar(tein_arr, lmdm5med_arr, yerr=lmdm5err_arr, color=colseq[n], linewidth=2)
     ax[2].axhline(lmdm5med_gal, color=colseq[n], linestyle='--')
[…]

### 2022-08-19 lens_population/plot_lens_observable_bias.py

@@ -18,7 +18,7 @@ nsims = len(sims)
 
 colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
 
-fig, ax = pylab.subplots(4, 1, figsize=(6, 12))
+fig, ax = pylab.subplots(4, 1, figsize=(6, 13))
 
 pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
 
@@ -80,12 +80,12 @@ for n in range(nsims):
 
     ax[0].errorbar(tein_arr, zmed_arr, yerr=zerr_arr, color=colseq[n], label=labels[n], linewidth=2)
     ax[1].errorbar(tein_arr, lmobsmed_arr, yerr=lmobserr_arr, color=colseq[n], label=labels[n], linewidth=2)
-    ax[2].errorbar(tein_arr, lreffmed_arr, yerr=lrefferr_arr, color=colseq[n], linewidth=2)
+    ax[2].errorbar(tein_arr, lreffmed_arr, yerr=lrefferr_arr, color=colseq[n], linewidth=2, label=labels[n])
     ax[3].errorbar(tein_arr, qmed_arr, yerr=qerr_arr, color=colseq[n], linewidth=2)
 
 # NEED TO REMOVE INDENTATION
 ax[0].axhline(zmed_gal, color='k', linestyle='--')#, label='Parent population')
-ax[0].set_ylabel('Median $z$', fontsize=fsize)
+ax[0].set_ylabel('Median $z_{\mathrm{g}}$', fontsize=fsize)
 
 ax[0].yaxis.set_major_locator(MultipleLocator(0.05))
 ax[0].yaxis.set_minor_locator(MultipleLocator(0.01))
@@ -96,21 +96,22 @@ ax[1].set_ylabel('Median $\log{M_*^{(\mathrm{obs})}}$', fontsize=fsize)
 ax[1].yaxis.set_major_locator(MultipleLocator(0.2))
 ax[1].yaxis.set_minor_locator(MultipleLocator(0.05))
 
-ax[2].axhline(lreffmed_gal, color='k', linestyle='--')
-ax[2].set_ylabel('$\mu_{\mathrm{R},0}$', fontsize=fsize)
+ax[2].axhline(lreffmed_gal, color='k', linestyle='--', label='Parent pop.')
+#ax[2].set_ylabel('$\mu_{\mathrm{R},0}$', fontsize=fsize)
[…]

### 2022-08-19 lens_population/plot_quasar_bias.py

@@ -1,0 +1,170 @@
+import numpy as np
+import pylab
+import h5py
+from simpars import *
+from scipy.optimize import leastsq
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 20
+
+sims = ['fiducial_1000sqdeg', 'fiducial_1000sqdeg', 'fiducial_1000sqdeg']
+sources = ['gal', 'qso', 'quads']
+
+labels = ['Extended sources', 'Quasars (all)', 'Quasars (quads)']
+nsims = len(sims)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+colors = [colseq[0], colseq[3], colseq[4]]
+
+fig, ax = pylab.subplots(4, 1, figsize=(6, 13))
+
+pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
+
+ntein = 21
+tein_arr = np.linspace(0., 2., ntein)
+
+galpop = h5py.File('fiducial_1000sqdeg_galaxies.hdf5', 'r')
+laspsmed_gal = np.median(galpop['lasps'][()])
+lm200med_gal = mu_h
+qmed_gal = np.median(galpop['q'][()])
+
+def mdm5fitfunc(p):
+    return p[0] + p[1] * (galpop['lmstar'][()] - lmstar_piv) + p[2] * (galpop['lreff'][()] - lreff_piv)
+
+def mdm5errfunc(p):
+    return mdm5fitfunc(p) - galpop['lmdm5'][()]
+
+pmdm5fit = leastsq(mdm5errfunc, [11., 0., 0.])
+
+lmdm5med_gal = pmdm5fit[0][0]
+
+for n in range(nsims):
+
+    if sources[n] == 'gal':
+        lenspop = h5py.File('%s_lenses.hdf5'%sims[n], 'r')
+
+    elif sources[n] == 'qso':
+        lenspop = h5py.File('%s_qsolenses.hdf5'%sims[n], 'r')
+
+    elif sources[n] == 'quads':
+        lenspop = h5py.File('%s_qsolenses.hdf5'%sims[n], 'r')
+
+    laspsmed_arr = np.zeros(ntein)
+    laspserr_arr = np.zeros(ntein)
+
+    lm200med_arr = np.zeros(ntein)
[…]

### 2022-08-19 lens_population/plot_source_bias.py

@@ -84,9 +84,9 @@ for n in range(nsims):
 
     ax[0].errorbar(tein_arr, zmed_arr, yerr=zerr_arr, color=colseq[n], label=labels[n])
     ax[1].errorbar(tein_arr, smagmed_arr, yerr=smagerr_arr, color=colseq[n], label=labels[n])
-    ax[2].errorbar(tein_arr, sreffmed_arr, yerr=srefferr_arr, color=colseq[n])
-    ax[3].errorbar(tein_arr, nsermed_arr, yerr=nsererr_arr, color=colseq[n])
-    ax[4].errorbar(tein_arr, qmed_arr, yerr=qerr_arr, color=colseq[n])
+    ax[2].errorbar(tein_arr, sreffmed_arr, yerr=srefferr_arr, color=colseq[n], label=labels[n])
+    ax[3].errorbar(tein_arr, nsermed_arr, yerr=nsererr_arr, color=colseq[n], label=labels[n])
+    ax[4].errorbar(tein_arr, qmed_arr, yerr=qerr_arr, color=colseq[n], label=labels[n])
 
 # NEED TO REMOVE INDENTATION
 ax[0].axhline(zmed_det, color='k', linestyle='--')#, label='Parent population')
@@ -95,7 +95,7 @@ ax[0].set_ylabel('Median $z$', fontsize=fsize)
 ax[0].yaxis.set_major_locator(MultipleLocator(0.2))
 ax[0].yaxis.set_minor_locator(MultipleLocator(0.05))
 
-ax[1].axhline(smagmed_det, color='k', linestyle='--')
+ax[1].axhline(smagmed_det, color='k', linestyle='--', label='Detectable pop.')
 ax[1].set_ylabel('Median $m_{\mathrm{s}}$', fontsize=fsize)
 
 ax[1].yaxis.set_major_locator(MultipleLocator(0.2))
@@ -107,7 +107,7 @@ ax[2].set_ylabel("Median $\\theta_{\mathrm{s}}$\n at $m_{\mathrm{s}}=25\, ('')$"
 ax[2].yaxis.set_major_locator(MultipleLocator(0.1))
 ax[2].yaxis.set_minor_locator(MultipleLocator(0.02))
 
-ax[3].axhline(nsermed_det, color='k', linestyle='--')
+ax[3].axhline(nsermed_det, color='k', linestyle='--', label='Detectable pop.')
 ax[3].set_ylabel("Median $n$\n at $m_{\mathrm{s}}=25$", fontsize=fsize)
 
[…]

### 2022-08-19 paper/seleff.tex
refined plots
IMF

@@
 Although we could still use a magnification-based definition for the sake of carrying out our experiments, it would then be difficult to apply our results to real data.
+Our definition of a strong lensing event, instead, is robust with respect to the mass-sheet degeneracy.
 %Because of the mass-sheet degeneracy
@@
 \end{equation}
+The most important source of systematic uncertainty in the measurement of the stellar mass is the assumption of the stellar initial mass function (IMF).
+\citet{Muz++13} assumed a Kroupa IMF \citep{Kro01} to derive their measurements.
+This means that, in the absence of other systematic effects, a value of $\asps=1$ corresponds to a Kroupa IMF.
 
@@
 This is about a factor of five smaller than the number density predicted by \citet{Col15} for the Euclid survey.
-We believe that this is a result of differences in the description of the source population and the criteria used to define a strong lensing event.
+This is a result of differences in the description of the source population, in the criteria used to define a strong lensing event, and in the redshift and stellar mass cuts that we applied to define the foreground galaxy population.
 
@@
 
-%In this section we present the results of the lens population simulations. 
-%Section \ref{ssec:resgal} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:resquasar} focuses on the population of galaxy-quasar lenses.
-%In each case, we first offer
+In this section we present the results of the lens population simulations. 
+Section \ref{ssec:resgal} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:resquasar} focuses on the population of galaxy-quasar lenses.
+Given the number of parameters that are needed to describe our model, providing a complete characterisation of the strong lensing bias is a problem with relatively high dimensionality, and is beyond the scope of this paper.
+For the sake of conciseness, we focus instead on the quantities that we consider most important.
+Nevertheless, the output of our simulations is available online\footnote{\url{https://github.com/astrosonnen/sl_selection_effects}}.
+We encourage readers who are interested in studying aspects of the strong lensing bias that are not covered in this section to download our data and analyse them directly.
 
@@
 However, it can rise up to $0.09$~dex (23\%) when considering only lenses with $\tein > 2''$.
+For comparison, we also show two reference values of $\asps$, corresponding to a Chabrier IMF \citep{Cha03} and a Salpeter IMF \citep{Sal55}.
+These values roughly bracket the current systematic uncertainty on $\asps$.
 
@@
 Bias on lens mass properties as a function of the minimum Einstein radius.
-First panel: median $\log{\asps}$.
+First panel: median $\log{\asps}$. Dotted lines indicate values of $\asps$ corresponding to a Chabrier and a Salpeter IMF ($\asps=1$ corresponds to a Kroupa IMF).
 Second panel: median $\mhalo$ at $\log{\mstar}=11.5$.
@@
 For this reason, we only ran simulations with lensed quasars in the fiducial scatter scenario and compared the results with those from the extended source simulation.
-%We are also interested in understanding how the entire population of quasar lenses differs from the subset of quad lenses, therefore we computed the bias for 
-We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we computed the bias for this sample as well.
+%We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we computed the bias for this sample as well.
+We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we also analysed this subsample on its own.
+%\Fref{fig:quasarpars} shows the distribution in a 
 %\Fref{fig:quasarbias} shows the strong lensing bias obtained with the lensed quasars simulation, compared with that
+
 \Fref{fig:quasarbias} shows the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
+The biases of the population of lensed quasars (red curves) are very similar to that of lensed galaxies (blue curves), especially for values of $\theta_{\mathrm{Ein},min}<1''$, 
+When considering only quad lenses, however, there are some differences, the most remarkable of which is the bias in the axis ratio: quad lenses tend to be on average galaxies with a higher ellipticity.
+This trend was expected, given the results of section \ref{ssec:axsymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
+%As we discuss in \Sref{sect:discuss}, the difference in ellipticity between quad lenses and 
+Quad lenses tend to also have a slightly larger halo mass at fixed stellar mass, for $\theta_{\mathrm{Ein},min}<1''$.
 
@@
 \caption{
-
-}
-\end{figure}
+Bias in the population of quasar lenses (all and quads only), compared to the extended source simulation.
+%Bias on lens mass properties as a function of the minimum Einstein radius.
+First panel: median $\log{\asps}$.
+Second panel: median $\mhalo$ at $\log{\mstar}=11.5$.
+Third panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+Fourth panel: median lens axis ratio.
+In each panel, the dashed line indicates the value of the parent population.
+Error bars indicate the standard deviation of the mean of the lens sample.
+\label{fig:quasarbias}
+}
+\end{figure}
+
+%As a last step, we also analysed the bias in the source-related parameters of the lensed quasars simulation.
+%\Fref{fig:quasar_cp} shows the distribution in source redshift and magnitude of the lensed quasars, in the $\theta_{\mathrm{Ein,min}}>0.5''$ case, compared to that of the parent and detectable populations.
 
@@
 
-We discuss.
+In order for a galaxy-source pair to be included as a strong lens in a survey, three condition must be met.
+First of all, part of the source must be multiply imaged.
+Second, the multiple images must be detectable.
+Third, the lens must be recognised as such.
+Each one of these conditions introduces a bias with respect to the parent population of foreground galaxies and background sources.
+Together, they define the lens detection probability term $\pdet$ of \Eref{eq:one}.
+
+There are two sources of strong lensing bias. 
+The first one is related to the detectability of
 

### 2022-08-19 paper/seleff.tex
started discussion

@@
 Together, they define the lens detection probability term $\pdet$ of \Eref{eq:one}.
+%In the best case scenario, the only sources of bias are the first two points.
+The first two points are intrinsic to a strong lensing survey and constitute an unavoidable source of bias. 
+The best case scenario occurs when the efficiency of including a detected strong lens in a survey is always one.
+In this case, the third condition does not introduce any further selection and the strong lensing bias is minimised.
+We explored this scenario in \Sref{sect:results} when setting the minimum Einstein radius to zero.
 

### 2022-08-19 paper/seleff.tex
discussion

@@
 
-\subsection{Galaxy-quasar lenses}
+\subsection{Galaxy-quasar lenses}\label{ssec:quasarbias}
 
@@
 
-There are two sources of strong lensing bias. 
-The first one is related to the detectability of
+Strictly speaking, the results shown in \Sref{sect:results} apply only to samples of lenses and background sources with the same properties as our simulations.
+However, the comparison of section \ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
+This result suggests that the dependence of the strong lensing bias on the deails of the background source population is very weak, as long as no additional selection on the image configuration is applied (e.g.
+by selecting only quad lenses).
+This is not a coincidence, but is a consequence of the fact that the dependence of the strong lensing cross-section on the lens parameters is a weak function of source magnitude\footnote{The cross-section itself is a strong function of source magnitude, but the trends between $\crosssect$ and the lens parameters are not.} (see section \ref{ssec:axisymmpoint}).
+We can then conclude that the details of the properties of the background source population play a secondary role in determining the strong lensing bias.
 
@@
 We conclude.
+
+We also introduced a new definition for a strong lensing event, in the case of an extended source.
+This definition, which is based on the number of detected surface brightness peaks, is robust with respect to the mass-sheet degeneracy, unlike all magnification-based definitions.
 

### 2022-08-20 paper/seleff.tex
big chunk of discussion

@@
 In general, galaxies with a higher concentration of mass are more likely to be strong lenses, and are therefore over-represented in lens samples.
-
-The luminosity and size distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies.
-For instance, lensing magnification allows the detection of fainter objects with respect to the field, and, in case lenses are selected on the basis of the magnification of the source, galaxies that are very extended with respect to the Einstein radius are less likely to be classified as strongly lensed.
+The distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies: for instance, lensing magnification allows the detection of fainter objects with respect to the field.
+
+%The luminosity and size distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies.
+%For instance, lensing magnification allows the detection of fainter objects with respect to the field, and, in case lenses are selected on the basis of the magnification of the source, galaxies that are very extended with respect to the Einstein radius are less likely to be classified as strongly lensed.
 This is because their magnification averaged over their size tends to be low \citep[see the discussion in section 5.3 of][]{O+A17}.
@@
 The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments. We describe this in section \ref{ssec:profile}.
-In section \ref{ssec:axsymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
+In section \ref{ssec:axisymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
 In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.
@@
 
-\subsection{Axisymmetric lenses, point sources}\label{ssec:axsymmpoint}
+\subsection{Axisymmetric lenses, point sources}\label{ssec:axisymmpoint}
 
@@
 In section \ref{ssec:obs} we describe how our mock observations of lenses are generated.
+In section \ref{ssec:teincuts} we apply a further selection step, based on the angular size of the Einstein radius.
 
@@
 \citet{Muz++13} assumed a Kroupa IMF \citep{Kro01} to derive their measurements.
-This means that, in the absence of other systematic effects, a value of $\asps=1$ corresponds to a Kroupa IMF.
+This means that, in the absence of other systematic effects, a galaxy with a Kroupa IMF has a value of $\asps=1$.
 
@@
 
-\subsubsection{Dark matter distribution}\label{ssec:dmprofile}
+\subsubsection{Dark matter distribution}\label{ssub:dmprofile}
 
@@
 
-\subsection{Background sources: galaxies}\label{ssec:sources}
+\subsection{Background sources}\label{ssec:sources}
+
+\subsubsection{Galaxies}\label{ssub:extsources}
 
@@
 
-\subsection{Background sources: quasars}\label{ssec:quasars}
+\subsubsection{Quasars}\label{ssub:quasars}
 
@@
 
-\subsection{Observations}\label{ssub:obs}
+\subsection{Observations}\label{ssec:obs}
 
@@
 
-\subsection{Einstein radius cuts}
+The resulting number of quasar lenses in our simulation with the fiducial scatter is $1621$.
+This number is meaningless, given that the simulation was created with an unrealistically large number density of quasars.
+More interesting is the relative number of quad lenses with respect to the total, which is about $9\%$. This is a slightly smaller value than the fraction of quads predicted by \citet{O+M10}.
+The reason for this discrepancy lies in the differences bewteen the lens mass models in the two simulations.
+
+\subsection{Einstein radius cuts}\label{ssec:teincuts}
 
@@
 In this section we present the results of the lens population simulations. 
-Section \ref{ssec:resgal} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:resquasar} focuses on the population of galaxy-quasar lenses.
+Section \ref{ssec:galbias} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:quasarbias} focuses on the population of galaxy-quasar lenses.
 Given the number of parameters that are needed to describe our model, providing a complete characterisation of the strong lensing bias is a problem with relatively high dimensionality, and is beyond the scope of this paper.
@@
 
-\subsection{Galaxy-galaxy lenses}
+\subsection{Galaxy-galaxy lenses}\label{ssec:galbias}
 
@@
 Second, while $\mu_{\mathrm{h},0}$ is the halo mass at fixed stellar mass, $\mu_{\mathrm{DM},0}$ is measured at fixed half-light radius as well.
-The density profile of the dark matter halos in our simulation have a dependence on galaxy size: the response of dark matter to baryons is stronger for more concentrated stellar distributions (see section \ref{ssec:dmprofile}).
+The density profile of the dark matter halos in our simulation have a dependence on galaxy size: the response of dark matter to baryons is stronger for more concentrated stellar distributions (see section \ref{ssub:dmprofile}).
 By capturing this dependence in the model of \Eref{eq:dmscaling}, the residual scatter in $\mdmfive$ around the mean is reduced, and so is the strong lensing bias.
@@
 When considering only quad lenses, however, there are some differences, the most remarkable of which is the bias in the axis ratio: quad lenses tend to be on average galaxies with a higher ellipticity.
-This trend was expected, given the results of section \ref{ssec:axsymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
+This trend was expected, given the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
 %As we discuss in \Sref{sect:discuss}, the difference in ellipticity between quad lenses and 
@@
 \section{Discussion}\label{sect:discuss}
+
+\subsection{Key results}
 
@@
 We explored this scenario in \Sref{sect:results} when setting the minimum Einstein radius to zero.
+If only lenses with Einstein radius larger than a given threshold are selected, however, the bias generally increases.
+
+The strong lensing bias affects all quantities that are related to the mass distribution of the lens, as well as the redshifts of lens and source and the source surface brightness parameters.
+Some of these quantities, such as the observed stellar mass and half-light radius of the lenses, can be directly measured, and it is straightforward to quantify their lensing bias.
+Other quantities, however, such as the stellar population synthesis mismatch parameter or the dark matter content, are difficult to obtain via traditional, non-lensing, observations.
+Our simulations are particularly useful to quantify the bias on these properties.
+
+In the fiducial scatter scenario, the one consistent with the observed scatter around the fundamental plane, the bias on $\asps$ varies from $0.03$~dex, when no restrictions on the lens Einstein radius are applied, to $0.08$~dex, corresponding to the case in which only lenses with $\tein > 2.0''$ are selected (see \Fref{fig:lensmassbias}).
+The current systematic uncertainty on $\asps$ is $0.2-0.3$~dex: this is roughly the difference in mesaurements of the stellar mass of a galaxy obtained with a Chabrier or a Salpeter IMF.
+Compared with this uncertainty, the amplitude of the strong lensing bias on $\asps$ is small: strong lensing observations can be used directly to discriminate between these two alternative choices of IMF, without the need to correct for selection effects.
+This is especially the case when working with samples of lenses that probe the sub-arcsecond Einstein radius regime.
+Such a goal could be reached with a sample size of a thousand lenses and a statistical study of the kind proposed by \citet{S+C21}.
+
+Strong lenses are also biased towards larger halo masses.
+Nevertheless, when focusing on the dark matter content in the inner regions, the amplitude of the strong lensing bias is relatively small, especially when controlling for the stellar distribution of the lens galaxies.
+For instance, at fixed stellar mass and half-light radius, the bias on the projected dark matter mass enclosed within $5$~kpc, $\mdmfive$, is only a few percent in the fiducial scatter scenario with $\theta_{\mathrm{Ein,min}} < 1.0''$.
+This means that strong lenses can be used to understand the inner dark matter distribution of galaxies, as long as the dependence of the dark matter distribution on the properties of the baryonic component is accurately modelled \citep[for example, by following the approach of][]{S+C21}.
+
+We also looked at the bias on source-related parameters.
+Strong lensing causes the luminosity function of background sources to be broadened, compared to the distribution of detectable sources.
+On the one hand, it preferentially selects brighter sources, because the lensing cross-section increases with increasing source brightness.
+On the other hand, it allows the detection of sources that are intrinsically fainter than the detection limit in the absence of lensing.
+Interestingly, we did not detect any significant bias on the source size, at fixed magnitude.
+This result follows from the fact that our lens detection criterion relies on a surface brightness threshold, and surface brightness is preserved by lensing.
+However, it appears to be in contradiction with the work of \citet{O+A17}, who argued that their strong lens sample selected preferentially compact sources.
+%However, \citet{O+A17} showed that their strong lens sample selected preferentially compact sources. 
+The origin of this discrepancy probably lies in the differences between the criteria used to define a strong lens in the two studies.
+The \citet{O+A17} sample was selected primarily via spectroscopy, by looking for signatures of two galaxies at different redshifts in the Sloan Digital Sky Survey \citep[SDSS][]{Yor++00} data.
+SDSS spectra were taken in fibers with a $1.5''$ radius.
+Lensed galaxies that are comparable in size to this scale, or larger, are less likely to be detected, because part of their flux extends outside of the fiber.
+Compact galaxies with a large overall magnification, instead, are more likely to be detected \citep[see the discussion in section 5.3 of][]{O+A17}.
+
+%Our analysis based on photometric detection showed that the tendency to select compact sources is not a general feature of strong lens samples.
+There is also an indirect way in which strong lensing could preferentially probe more compact sources than possible in the field.
+Sources that are smaller than the size of the PSF can be easily confused as stars in our own Galaxy.
+Since these stars cannot be strongly lensed into multiple images, the detection of a strong lens automatically confirms the extragalactic nature of a lensed source.
+This type of selection effect could explain, for example, the detection of an apparent outlier in the magnitude-size relation of Lyman-break galaxy by means of strong lensing, by \citet{Jae++20}.
+It is, however, improper to refer to this effect as strong lensing bias, since the star-galaxy separation is a bias that primarily affects field observations.
+Our analysis showed that the tendency to select compact sources is not a general feature of strong lens samples.
+
+\subsection{Combining lensed quasars with lensed galaxies}
+
+Our experiments were also useful for understanding the possible biases that might incur when combining information from samples of galaxy-galaxy lenses with samples of galaxy-quasar lenses.
+On the one hand the biases in the mass-related quantities ($\asps$ and $\mdmfive$) are very similar in simulations with the same foreground galaxy population and different background sources (see \Fref{fig:quasarbias}).
+On the other hand, these biases are a function of the minimum Einstein radius.
+In order to use a sample of strong lenses as a prior for another sample, then, it is important to make sure that 1) the parent population of foreground galaxies among which lenses are searched for is the same in both surveys, and 2) the two surveys can probe the same distribution in Einstein radius.
+
+The above argument applies to samples of quasars selected regardless of the number of multiple images.
+When dealing with quad lenses, however, the situation is more complicated.
+First of all, because our experiment revealed differences in the bias on the dark matter distribution between quads and the entire sample of lensed quasars.
+Second, because quad lenses tend to have a preferentially higher ellipticity.
+The ellipticity is important in the context of stellar dynamical analyses,
+which are often used in combination with strong lensing to constrain lens mass parameters \citep[see e.g.][]{Yil++20}.
+In order to correctly interpret stellar dynamics data, assumptions on the three-dimensional structure of a lens must be made.
+This is directly related to the projected ellipcitiy: galaxies with an axis ratio close to one tend to be preferentially elongated in the line-of-sight direction, and viceversa.
+When using stellar dynamics-based mass measurements to inform the properties of a sample of quad lenses, therefore, it is important to take into account possible biases due to the different three-dimensional structure of the two samples.
+
+\subsection{The importance of the source population properties}
 
@@
 by selecting only quad lenses).
-This is not a coincidence, but is a consequence of the fact that the dependence of the strong lensing cross-section on the lens parameters is a weak function of source magnitude\footnote{The cross-section itself is a strong function of source magnitude, but the trends between $\crosssect$ and the lens parameters are not.} (see section \ref{ssec:axisymmpoint}).
+This is not a coincidence, but is a consequence of the fact that the dependence of the strong lensing cross-section on the lens parameters is a weak function of source magnitude\footnote{The cross-section itself is a strong function of source magnitude, but the trends between $\crosssect$ and the lens parameters are not.}, especially for magnitudes close to the detection limit (see section \ref{ssec:axisymmpoint}).
 We can then conclude that the details of the properties of the background source population play a secondary role in determining the strong lensing bias.
+
+\subsection{Mitigation strategies}
+
+Although the amplitude of the strong lensing bias in difficult-to-observe mass-related quantities is relatively small, compared to current uncertainties, there might be situations in which one wishes to account for it.
+The formally correct way to do this is by explicitly modelling all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
+Doing so, however, can be difficult in practice.
+Therefore, it can be useful to develop alternative strategies to correct for the bias, or to mitigate its effects.
+
+As we discussed throughout this paper, the amplitude of the strong lensing bias depends on the intrinsic scatter in the mass parameters of the foreground galaxy population.
+One way to minimise the bias, therefore, is to identify scaling relations between observable quantities and mass-related properties that can account for part of the scatter.
+Describing the inner dark matter distribution as a function of stellar mass and half-light radius, as we did in section \ref{ssec:galbias}, is the first step in this direction: the bias in $\mdmfive$ at fixed $\mstar$ and $\reff$ is smaller than the overall shift in the median $\mdmfive$ of the lens population.
+This description could be extended by including the central velocity dispersion as an additional control parameter.
+The velocity dispersion is directly related to the mass distribution of the lens galaxy, and therefore the intrinsic scatter in $\asps$ and $\mdmfive$ at fixed velocity dispersion should be smaller.
+
+One problem with this approach is that the amplitude of the scatter in quantities such as $\asps$ or the dark matter parameters is not known before carrying out a strong lensing analysis, and neither is the resulting strong lensing bias.
+An estimate of the bias, however, could be obtained after constraining the lens population distribution, by creating a simulation of the survey under the assumption that the inferred distribution is a good approximation of the truth.
+This would allow to posterior predict the effects of strong lensing bias.
+One could then adjust the parameters of the simulated galaxy population until the posterior predicted lens population matches the observed one.
+%An estimate of the bias, however, could be obtained after constraining the lens population distribution, by creating a simulation of the survey under the assumption that the inferred distribution is a good approximation of the truth.
+The end product would be a posterior prediction of the properties of a sample of strong lenses.
+This procedure could then be iterated until convergence.
+%If the inferred intrinsic scatter at fixed observable properties is small, so will be the amplitude of the strong lensing bias.
+
+%If one wishes to directly correct for the strong lensing bias but wants to avoid to explicitly model all of the aspects entering the selection process, one possibility is to try and approximate $\pdet$.
+If one wishes to directly correct for the strong lensing bias in a computationally manageable way, one possibility is to try and approximate $\pdet$. 
+In general, $\pdet$ depends nontrivially on the position, redshift and surface brightness distribution of the source, as well as on the lens mass density profile and redshift.
+This requires in principle to explicitly model the distribution of all of these quantities \citep[as done by][]{Son22}.
+However, we can make the assumption that $\pdet$ can be separable in a term that depends purely on the strong lensing cross-section and a term that depends purely on the Einstein radius.
+This is a reasonable assumption, as long as the lens finding procedure does not selectively pick lenses with different image configurations depending on their Einstein radius.
+It would allow us to marginalise over the source position and surface brightness parameters, thus simplifying the problem.
+The dependence of $\pdet$ on $\tein$ could then be described empirically and inferred during the analysis.
+This is essentially the same approach adopted by \citet{Son++19} in the analysis of strong and weak lensing data from the Hyper Suprime-Cam Survey.
+
+%However, we can make the assumption that the dependence of $\pdet$ on source position and surface brightness distribution can be separable from the dependence on the lens mass properties and the lens and source redshift.
+%This is a reasonable assumption, given that the dependence of the strong lensing cross-section on the source parameters is similar in lenses with different mass properties (see section \ref{ssec:axisymmpoint}).
+%This assumption would allow us to marginalise over the source parameters, simplifying the problem. The lens detection probability would then be reduced to the sole $\crosssect$, computed
+
+\subsection{Limitations of our analysis}
 

### 2022-08-21 paper/seleff.tex
finished main text

@@
 In principle, this can be done with a Bayesian hierarchical formalism, provided that the lens detection probability $\pdet$ is known \citep{Son22}.
-For most of the existing strong lensing surveys, however, characterising $\pdet$ is very challenging: the process determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
+For most of the existing strong lensing surveys, however, characterising $\pdet$ is very challenging: the process of determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
 
@@
 \begin{equation}
-\kappa(\boldsymbol\theta) = \dfrac{\Sigma(\boldsymbol\theta)}{\Sigma_{\mathrm{cs}}}.
+\kappa(\boldsymbol\theta) = \dfrac{\Sigma(\boldsymbol\theta)}{\Sigma_{\mathrm{cr}}}.
 \end{equation}
@@
 \begin{equation}
-\Sigma_{cr} = \dfrac{c^2D_s}{4\pi G D_d D_{ds}},
-\end{equation}
-where $c$ is the speed of light, $G$ the gravitational constant, and $D_d$, $D_s$, and $D_{ds}$ are the angular diameter distances between the observer and the lens, the observer and the source, and the lens and the source, respectively.
+\Sigma_{\mathrm{cr}} = \dfrac{c^2D_\mathrm{s}}{4\pi G D_\mathrm{d} D_{\mathrm{ds}}},
+\end{equation}
+where $c$ is the speed of light, $G$ the gravitational constant, and $D_\mathrm{d}$, $D_\mathrm{s}$, and $D_{\mathrm{ds}}$ are the angular diameter distances between the observer and the lens, the observer and the source, and the lens and the source, respectively.
 
@@
 The third column shows the source footprint defined with the procedure described above.
-Coral pixels correspond to the largest footprint that maximizes the number of detected images, while purple pixels belong to the $2\sigma$ detection footprint.
+Pink pixels correspond to the largest footprint that maximizes the number of detected images, while purple pixels belong to the $2\sigma$ detection footprint.
 In the first, second and sixth example, these two footprints coincide.
@@
 if the source is a standard candle or a standard ruler).
-Because of the mass-sheet degeneracy \citep{FGS85}, it is possible to vary the magnification of a lensed image while keeping its observed properties fixed: this could lead to the paradox of two identical lenses that are classified differently on the basis of the underlying magnification.
+Because of the mass-sheet degeneracy \citep{FGS85}, it is possible to vary the magnification of a lensed image while keeping its observed properties fixed: this could lead to the paradox of two identical-looking lenses that are classified differently on the basis of the underlying magnification.
 Although we could still use a magnification-based definition for the sake of carrying out our experiments, it would then be difficult to apply our results to real data.
@@
 As a result, the source plane area that is subject to strong lensing is an even stronger function of these properties, since it scales with the square of the radial caustic radius.
-In order to compute the lensing cross-section, however, we must take into account the magnification of the multiple images: according to our definition of a strong lensing event, given in section \ref{ssec:lensdefpoint}, at least two images must be detected in order for a system to qualify as a strong lens.
+In order to compute the lensing cross-section, however, we must take into account the magnification, because that determines whether multiple images can be detected or not.
+%In order to compute the lensing cross-section, however, we must take into account the magnification of the multiple images: according to our definition of a strong lensing event, given in section \ref{ssec:lensdefpoint}, at least two images must be detected in order for a system to qualify as a strong lens.
 
@@
 This implies that surveys that target different families of sources, with different luminosity functions (for instance, galaxies or quasars), can have different strong lensing biases.
-We explore this possibility in \Sref{sect:lenspop}.
+We explore this possibility in \Sref{sect:results}.
 
@@
 %We randomly drew halo masses from the following distribution:
-Using current weak lensing constraints on the halo mass of elliptical galaxise as a reference \citep{Son++22}, we drew halo masses from the following distribution:
+Using current weak lensing constraints on the halo mass of elliptical galaxies as a reference \citep{Son++22}, we drew halo masses from the following distribution:
 \begin{equation}
@@
 
-The distribution in stellar mass, halo mass and dark matter inner slope of our sample of simulated foreground galaxies depends on parameters describing the intinsic scatter in these properties, namely $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_{\gamma}$.
+The distribution in stellar mass, halo mass and dark matter inner slope of our sample of simulated foreground galaxies depends on parameters describing the intinsic scatter in these properties, namely $\sigma_{\mathrm{sps}}$ and $\sigma_h$.
 Direct observational constraints on these quantities are poor.
@@
 We used Jeans modelling for this purpose.
-We generated samples of early-type galaxies with the recipes described above and with different values of $\sigma_{\mathrm{sps}}$, $\sigma_\mathrm{h}$ and $\sigma_\gamma$.
+We generated samples of early-type galaxies with the recipes described above and with different values of $\sigma_{\mathrm{sps}}$ and $\sigma_\mathrm{h}$.
 Using the spherical Jeans equation under the assumption of isotropic orbits, we predicted the central velocity dispersion of each galaxy in the sample.
 Then, we fitted a fundamental plane relation to this mock sample and measured the predicted scatter in velocity dispersion at fixed $\mobs$ and $\reff$.
-We then varied $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$ to match the observed and the predicted scatter.
+We then varied $\sigma_{\mathrm{sps}}$ and $\sigma_h$ to match the observed and the predicted scatter.
 We did not attempt to match the other parameters of the fundamental plane (i.e.
@@
 
-%After comparing the predicted and observed fundamental plane scatter for different sets of values of $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$, we settled on three 
-%After comparing the predicted and observed fundmantal plane scatter, 
-We settled on three different set of intrinsic scatter parameters, as indicated in \Tref{tab:scatter}. We label them the fiducial, the low-scatter and the high-scatter scenarios.
+We settled on three different sets of intrinsic scatter parameters, as indicated in \Tref{tab:scatter}. We label them the fiducial, the low-scatter and the high-scatter scenarios.
 The high-scatter scenario is ruled out both by the fundamental plane and by weak lensing constraints \citep{Son++22}. 
@@
 This ensures that all of the sources lie behind all of the lenses.
-A similar cut could be applied to a real survey using photometric redshift, to reduce the incidence of false positives (e.g.
+A similar cut could be applied to a real survey using photometric redshifts, to reduce the incidence of false positives (e.g.
 arc-like features physically associated with the lens galaxy) in the lens finding phase.
@@
 
-We describe the population of background quasars with the following double power-law luminosity function in the rest-frame UV absolute magnitude $M$:
+We described the population of background quasars with the following double power-law luminosity function in the rest-frame UV absolute magnitude $M$:
 \begin{equation}
@@
 For the sake of consistency with the population of background extended sources, we limited the redshift distribution of quasars to the range $0.8 < \zqso < 2.5$.
-We then truncated the distribution in $\mqso$ at two magnitudes fainter than the detection limit (which is specified in section \ref{ssub:obs}).
+We then truncated the distribution in $\mqso$ at two magnitudes fainter than the detection limit (which is specified in section \ref{ssec:obs}).
 Finally, we randomly placed quasars in the source plane with a projected number density of $70$~arcmin$^{-2}$.
 This is a much larger number density than observed in the real universe.
-Nevertheless, we are not interested in predicting the absolute number of strong lenses, so this is a viable choice.
+Nevertheless, we are not interested in predicting the absolute number of strong lenses, so this is a legitimate choice.
 The advantage of boosting the number density of quasars is that it allows us to produce a large number of lenses without the need for generating too big a population of foreground galaxies.
@@
 
-Strong lensing tends to preferentially select sources at higher redshift, especially in the more restrictive case with $\theta_{\mathrm{Ein,min}} > 1.0''$.
+Strong lensing tends to preferentially select sources at higher redshift, especially in the more restrictive case with $\tein > 1.0''$.
 This is because, at fixed lens properties, increasing the source redshift lowers the critical surface mass density and, consequently, increases the size of the caustics and the Einstein radius.
@@
 However, as we quantify later in this section, the difference is far from large.
-The reason for this behaviour lies in the fact that, in the detection of both lensed and unlensed sources, the most important quantity is surface brightness, which is preserved by gravitational lensing.
+The reason for this behaviour lies in the fact that, in the detection of both lensed and unlensed sources, the most important quantity is surface brightness, which is preserved by gravitational lensing when the source is larger than the PSF size.
 As we showed in section \ref{ssec:ellext}, the lensing cross-section drops to zero once the surface brightness of the source reaches a value that would make it undetectable in the absence of lensing.
@@
 
-In the fiducial scatter scenario, the one consistent with the observed scatter around the fundamental plane, the bias on $\asps$ varies from $0.03$~dex, when no restrictions on the lens Einstein radius are applied, to $0.08$~dex, corresponding to the case in which only lenses with $\tein > 2.0''$ are selected (see \Fref{fig:lensmassbias}).
+In the fiducial scatter scenario, the one consistent with the observed scatter around the fundamental plane, the bias on $\asps$ varies from $0.03$~dex, when no restrictions on the lens Einstein radius are applied, to $0.09$~dex, corresponding to the extreme case in which only lenses with $\tein > 2.0''$ are selected (see \Fref{fig:lensmassbias}).
+A reasonable value for the minimum Einstein radius in a space-based survey like Euclid or the Chinese Space Station Telescope (CSST) is $\theta_{\mathrm{Ein,min}}=0.5''$.
+In this case, the bias on $\asps$ is slightly smaller than $0.04$~dex.
 The current systematic uncertainty on $\asps$ is $0.2-0.3$~dex: this is roughly the difference in mesaurements of the stellar mass of a galaxy obtained with a Chabrier or a Salpeter IMF.
 Compared with this uncertainty, the amplitude of the strong lensing bias on $\asps$ is small: strong lensing observations can be used directly to discriminate between these two alternative choices of IMF, without the need to correct for selection effects.
-This is especially the case when working with samples of lenses that probe the sub-arcsecond Einstein radius regime.
+%This is especially the case when working with samples of lenses that probe the sub-arcsecond Einstein radius regime.
 Such a goal could be reached with a sample size of a thousand lenses and a statistical study of the kind proposed by \citet{S+C21}.
@@
 However, the comparison of section \ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
-This result suggests that the dependence of the strong lensing bias on the deails of the background source population is very weak, as long as no additional selection on the image configuration is applied (e.g.
+This result suggests that the dependence of the strong lensing bias on the details of the background source population is very weak, as long as no additional selection on the image configuration is applied (e.g.
 by selecting only quad lenses).
@@
 This description could be extended by including the central velocity dispersion as an additional control parameter.
-The velocity dispersion is directly related to the mass distribution of the lens galaxy, and therefore the intrinsic scatter in $\asps$ and $\mdmfive$ at fixed velocity dispersion should be smaller.
-
-One problem with this approach is that the amplitude of the scatter in quantities such as $\asps$ or the dark matter parameters is not known before carrying out a strong lensing analysis, and neither is the resulting strong lensing bias.
-An estimate of the bias, however, could be obtained after constraining the lens population distribution, by creating a simulation of the survey under the assumption that the inferred distribution is a good approximation of the truth.
+The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution.
+The central velocity dispersion, however, is very sensitive to the orbital anisotropy and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
+%intrinsic scatter in $\asps$ at fixed velocity dispersion should be smaller than %and therefore the intrinsic scatter in $\asps$ and $\mdmfive$ at fixed velocity dispersion should be smaller.
+
+%One problem with this approach is that the amplitude of the scatter in quantities such as $\asps$ or the dark matter parameters is not known before carrying out a strong lensing analysis, and neither is the resulting strong lensing bias.
+%An estimate of the bias, however, could be obtained after constraining the lens population distribution, by creating a simulation of the survey under the assumption that the inferred distribution is a good approximation of the truth.
 This would allow to posterior predict the effects of strong lensing bias.
 One could then adjust the parameters of the simulated galaxy population until the posterior predicted lens population matches the observed one.
-%An estimate of the bias, however, could be obtained after constraining the lens population distribution, by creating a simulation of the survey under the assumption that the inferred distribution is a good approximation of the truth.
-The end product would be a posterior prediction of the properties of a sample of strong lenses.
-This procedure could then be iterated until convergence.
-%If the inferred intrinsic scatter at fixed observable properties is small, so will be the amplitude of the strong lensing bias.
-
-%If one wishes to directly correct for the strong lensing bias but wants to avoid to explicitly model all of the aspects entering the selection process, one possibility is to try and approximate $\pdet$.
+
 If one wishes to directly correct for the strong lensing bias in a computationally manageable way, one possibility is to try and approximate $\pdet$. 
@@
 
-%However, we can make the assumption that the dependence of $\pdet$ on source position and surface brightness distribution can be separable from the dependence on the lens mass properties and the lens and source redshift.
-%This is a reasonable assumption, given that the dependence of the strong lensing cross-section on the source parameters is similar in lenses with different mass properties (see section \ref{ssec:axisymmpoint}).
-%This assumption would allow us to marginalise over the source parameters, simplifying the problem. The lens detection probability would then be reduced to the sole $\crosssect$, computed
-
 \subsection{Limitations of our analysis}
 
+The simulations on which our analysis is based as as complex as required by the goal of the analysis itself, which is to estimate the amplitude of the strong lensing bias in a few key quantities.
+We did, however, make some simplifying assumptions.
+One such assumption consisted in neglecting the contribution from line-of-sight structure and the environment to the lensing signal, which typically introduce an external shear and convergence.
+The effect of including external shear in a lens model is similar to that of changing the ellipticity of the lens.
+%Therefore the main result that would be affected by neglecting it is the bias on the axis ratio.
+External convergence mimics the effect of adding or removing a constant sheet of invisible mass, producing an effect similar to varying the distribution of dark matter.
+Typical values of the external shear and external convergence are $|\gamma|<0.1$ and $|\kappa_{\mathrm{ext}}|<0.1$ \citep{Mil++20}: these values are small compared to the typical ellipticities and dark matter fractions of our simulated lenses.
+Therefore, while including them might modify slightly the amplitude of the strong lensing bias in the axis ratio and dark matter distribution parameters, the conclusions our our analysis would not be affected.
+
 %__________________________________________________________________
@@
 
-We conclude.
-
-We also introduced a new definition for a strong lensing event, in the case of an extended source.
+We carried out a thorough investigation of the strong lensing bias in photometrically selected samples of strong lenses.
+We first explored how the strong lensing cross-section varies as a function of lens and source properties, expanding upon the work of \citet{MVK09}.
+% by considering both point-like and extended sources.
+We introduced a new definition for a strong lensing event in the case of an extended source, for this purpose.
+%We also introduced a new definition for a strong lensing event, in the case of an extended source.
 This definition, which is based on the number of detected surface brightness peaks, is robust with respect to the mass-sheet degeneracy, unlike all magnification-based definitions.
+We found that the strong lensing cross-section increases primarily with increasing lens mass, with decreasing half-mass radius, and with decreasing source magnitude.
+
+We then simulated photometric strong lensing surveys with data similar to Euclid Wide, and measured how various properties of the lens galaxies and background sources are biased with respect to their parent distributions.
+We focused primarily on mass-related quantities: most importantly, the stellar population synthesis mismatch parameter $\asps$ and the dark matter distribution.
+We found that strong lenses are biased towards larger values of $\asps$ and of the halo mass $\mhalo$.
+The amplitude of the bias is larger in simulations with a larger intrinsic scatter in the distribution of $\asps$ and $\mhalo$, and increases if lenses with a preferentially large Einstein radius are selected.
+
+We used observational constraints from the fundamental plane to set the values of the intrinsic scatter in our fiducial simulation.
+In that case, assuming a survey that probes the Einstein radius distribution down to $\tein=0.5''$ with relatively high completeness, the bias on $\asps$ is as low as $0.04$~dex (10\%).
+This bias is much smaller than the current systematic uncertainty on the stellar population synthesis-based stellar masses.
+Therefore, strong lensing measurements could be used directly to calibrate stellar mass measurements of massive galaxies to $10\%$ accuracy, without the need to correct for selection effects.
+Under the same assumptions, the strong lensing bias on the average halo mass at fixed stellar mass is $0.04$~dex, while that on the inner dark matter distribution at fixed stellar mass and size is negligible.
+
+Strong lensing selection broadens the magnitude distribution of background sources, compared to the population of objects that are detectable without lensing.
+At the same time, we did not find any evidence for a bias in the size distribution of background sources, at fixed magnitude.
+%This follows from the fact that our selection criterion is based on the surface brightness
+
+We also ran simulations with lensed quasars in place of extended sources, and found that the amplitude of the strong lensing bias in the lens-related parameters has very little sensitivity of the details of the source population.
+This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
+Samples of quad lenses, however, are biased towards galaxies with larger ellipcity, which implies that their three-dimensional structure is on average different from that of the entire sample of lenses.
+This means that particular care must be taken when stellar dynamics measurements obtained on galaxy-galaxy lenses is used to inform the properties of quads.
+
+In conclusion, strong lensing selection introduces unavoidable biases in the properties of the lens galaxy and background source populations.
+Biases that affect observable properties, such as the redshift and the light distribution of the lens, can be easily quantified.
+Biases on mass-related quantities, such as the stellar mass-to-light ratio or the dark matter distribution, are more difficult to measure directly, but can be reduced by modelling how they scale with the observable properties of the lenses, thus minimising their intrinsic scatter.
+
+In the future, it might be desirable to directly model strong lensing selection effects, thus accounting for the strong lensing bias.
+Designing strong lensing surveys with clearly-defined and easily-modellable selection criteria would help greatly in this task.
 

### 2022-08-21 paper/seleff.tex
using observed stellar mass

@@
 
-The second panel of \Fref{fig:lensmassbias} shows the median halo mass of galaxies with a stellar mass of $\log{\mstar}=11.5$.
-We measured this quantity by fitting the following relation to the $\mhalo-\mstar$ distribution,
-\begin{equation}
-\log{\mhalo} \sim \mu_{\mathrm{h},0} + \beta_{\mathrm{h}}(\log{\mstar} - 11.5),
+The second panel of \Fref{fig:lensmassbias} shows the median halo mass of galaxies with an observed stellar mass of $\log{\mobs}=11.4$.
+We measured this quantity by fitting the following relation to the $\mhalo-\mobs$ distribution,
+\begin{equation}
+\log{\mhalo} \sim \mu_{\mathrm{h},0} + \beta_{\mathrm{h}}(\log{\mobs} - 11.4),
 \end{equation}
@@
 Similarly to the $\asps$ case, the strong lenses are biased towards larger values of the halo mass, with the bias being larger for higher-scatter simulations and more restrictive cuts on $\tein$.
-In the fiducial scatter scenario with a minimum Einstein radius of $\theta_{\mathrm{Ein,min}}=1.0''$, the halo masses of lenses are on average $0.1$~dex larger than those of their parent population.
-
-The third panel of \Fref{fig:lensmassbias} shows the median projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$, at fixed stellar mass and half-light radius.
-We obtained this quantity by fitting the following relation to the $\mdmfive-\mstar-\reff$ distribution:
+In the fiducial scatter scenario with a minimum Einstein radius of $\theta_{\mathrm{Ein,min}}=1.0''$, the halo masses of lenses are on average $0.16$~dex larger than those of their parent population.
+
+The third panel of \Fref{fig:lensmassbias} shows the median projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$, at fixed observed stellar mass and half-light radius.
+We obtained this quantity by fitting the following relation to the $\mdmfive-\mobs-\reff$ distribution:
 \begin{equation}\label{eq:dmscaling}
-\log{\mdmfive} \sim \mu_{\mathrm{DM},0} + \beta_{\mathrm{DM}}(\log{\mstar} - 11.5) + \xi_{\mathrm{DM}}(\log{\reff} - 1.2).
+\log{\mdmfive} \sim \mu_{\mathrm{DM},0} + \beta_{\mathrm{DM}}(\log{\mobs} - 11.4) + \xi_{\mathrm{DM}}(\log{\reff} - 1.2).
 \end{equation}
@@
 By capturing this dependence in the model of \Eref{eq:dmscaling}, the residual scatter in $\mdmfive$ around the mean is reduced, and so is the strong lensing bias.
-
-In the fiducial scenario, the bias on $\mdmfive$ is negligible for $\theta_{\mathrm{Ein,min}} < 0.5''$.
-Interestingly, the bias is negative in the low-scatter scenario.
-This is a counter-intuitive result that follows from the fact that the model of \Eref{eq:dmscaling} is not a perfect description of the distribution in $\mdmfive$.
+%
+In the fiducial scenario, the bias on $\mdmfive$ is as small as $0.02$~dex for $\theta_{\mathrm{Ein,min}} < 0.5''$, and is negligible in the low-scatter scenario.
+%Interestingly, the bias is negative in the low-scatter scenario.
+%This is a counter-intuitive result that follows from the fact that the model of \Eref{eq:dmscaling} is not a perfect description of the distribution in $\mdmfive$.
 Nevertheless, the amplitude of the bias is only $0.01$~dex, which for all practical purposes is negligible.
 
-In the fourth panel of \Fref{fig:lensmassbias} we show the average inner dark matter slope at fixed stellar mass and half-light radius, which we measured by fitting the following model
+In the fourth panel of \Fref{fig:lensmassbias} we show the average inner dark matter slope at fixed observed stellar mass and half-light radius, which we measured by fitting the following model
 \begin{equation}\label{eq:dmscaling}
-\gammadm \sim \mu_{\gamma,0} + \beta_{\gamma}(\log{\mstar} - 11.5) + \xi_{\gamma}(\log{\reff} - 1.2),
+\gammadm \sim \mu_{\gamma,0} + \beta_{\gamma}(\log{\mobs} - 11.4) + \xi_{\gamma}(\log{\reff} - 1.2),
 \end{equation}
 and taking the resulting value of $\mu_{\gamma}$.
-The trends in the avereage $\gammadm$ reflect very closely those in $\mdmfive$, but with flipped sign.
-This is because, in our model for the dark matter density profile, inner dark matter mass and density slope are perfectly correlated: both $\mdmfive$ and $\gammadm$ are determined univocally once the halo mass, stellar mass and half-light radius are specified.
+Different simulations show different trends in the average $\gammadm$.
+These are the results of underlying correlations between the dark matter density profile and the stellar and halo mass.
+In our model, galaxies of a given size with a larger stellar mass have a steeper dark matter slope, because the response of the dark matter to the infall of baryons is stronger.
+Vice-versa, galaxies with a larger halo mass have a shallower density slope.
+At fixed $\mobs$ and $\reff$, strong lenses have both a larger stellar mass (because their $\asps$ is larger) and a larger halo mass.
+In the simulation with low scatter, the bias in stellar mass is more important, therefore $\gammadm$ has a positive bias.
+In the simulation with large scatter, the bias in halo mass dominates, and therefore $\gammadm$ is negatively biased.
+
+%The trends in the avereage $\gammadm$ reflect very closely those in $\mdmfive$, but with flipped sign.
+%This is because, in our model for the dark matter density profile, inner dark matter mass and density slope are perfectly correlated: both $\mdmfive$ and $\gammadm$ are determined univocally once the halo mass, stellar mass and half-light radius are specified.
 The fact that the bias is negative is somewhat counter-intuitive: lenses with a stronger response to baryons have both a larger value of $\mdmfive$ and a steeper inner dark matter slope.
-However, $\gammadm$ is also anti-correlated with total halo mass (see $\gammadm-\mhalo$ panel in \Fref{fig:lenspars}): since strong lenses tend to have a preferentially larger halo mass, they also have a shallower dark matter slope.
+%However, $\gammadm$ is also anti-correlated with total halo mass (see $\gammadm-\mhalo$ panel in \Fref{fig:lenspars}): since strong lenses tend to have a preferentially larger halo mass, they also have a shallower dark matter slope.
 
@@
 First panel: median $\log{\asps}$. Dotted lines indicate values of $\asps$ corresponding to a Chabrier and a Salpeter IMF ($\asps=1$ corresponds to a Kroupa IMF).
-Second panel: median $\mhalo$ at $\log{\mstar}=11.5$.
-Third panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
-Fourth panel: median $\gammadm$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+Second panel: median $\mhalo$ at $\log{\mobs}=11.4$.
+Third panel: median $\mdmfive$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
+Fourth panel: median $\gammadm$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
 In each panel, the dashed line indicates the value of the parent population.
@@
 Nevertheless, when focusing on the dark matter content in the inner regions, the amplitude of the strong lensing bias is relatively small, especially when controlling for the stellar distribution of the lens galaxies.
-For instance, at fixed stellar mass and half-light radius, the bias on the projected dark matter mass enclosed within $5$~kpc, $\mdmfive$, is only a few percent in the fiducial scatter scenario with $\theta_{\mathrm{Ein,min}} < 1.0''$.
+For instance, at fixed observed stellar mass and half-light radius, the bias on the projected dark matter mass enclosed within $5$~kpc, $\mdmfive$, is only a few percent in the fiducial scatter scenario with $\theta_{\mathrm{Ein,min}} < 1.0''$.
 This means that strong lenses can be used to understand the inner dark matter distribution of galaxies, as long as the dependence of the dark matter distribution on the properties of the baryonic component is accurately modelled \citep[for example, by following the approach of][]{S+C21}.
@@
 Therefore, strong lensing measurements could be used directly to calibrate stellar mass measurements of massive galaxies to $10\%$ accuracy, without the need to correct for selection effects.
-Under the same assumptions, the strong lensing bias on the average halo mass at fixed stellar mass is $0.04$~dex, while that on the inner dark matter distribution at fixed stellar mass and size is negligible.
+Under the same assumptions, the strong lensing bias on the average halo mass at fixed stellar mass is $0.04$~dex, while that on the inner dark matter distribution at fixed stellar mass and size is $0.02$~dex.
 

### 2022-08-22 paper/seleff.tex
abstract and spell-check

@@
 %\usepackage{subcaption}
+\usepackage{color}
 
@@
 \newcommand{\ale}[1]{\textcolor{red}{\textbf{[Ale: #1]}}}
+\newcommand{\todo}[1]{\textcolor{red}{\textbf{[To Do: #1]}}}
 
@@
     {
-There are selection effects. They are important.
+Strong lenses are a biased subset of the general population of galaxies.
 }
@@
    {
-How important?
+The goal of this work is to quantify how lens galaxies and lensed sources differ from their parent distribution, namely the strong lensing bias.
 } 
@@
    {
-We simulate lenses.
+We first studied how the strong lensing cross-section varies as a function of lens and source properties. 
+Then, we simulated strong lensing surveys with data similar to that expected for Euclid and measured the strong lensing bias in different scenarios.
+We focused particularly on two quantities: the stellar population synthesis mismatch parameter, $\asps$, defined as the ratio between the true stellar mass of a galaxy and the stellar mass obtained from photometry, and the central dark matter mass at fixed stellar mass and size.
 }
@@
    {
-Results.
+Strong lens galaxies are biased towards larger stellar masses, smaller half-mass radii and larger dark matter masses.
+The amplitude of the bias depends on the intrinsic scatter in the mass-related parameters of the galaxy population and on the completeness in Einstein radius of the lens sample.
+For values of the scatter that are consistent with observed scaling relations and a minimum detectable Einstein radius of $0.5''$, the strong lensing bias in $\asps$ is $10\%$, while that in the central dark matter mass is $5\%$.
+The bias has little dependence on the properties of the source population: samples of galaxy-galaxy lenses and galaxy-quasar lenses that probe the same Einstein radius distribution are biased in a very similar way.
+Quadruply imaged quasar lenses, however, are biased towards higher ellipticity galaxies.
 }
@@
    {
-We conclude.
+Given current uncertainties, strong lensing observations can be used directly to improve our current knowledge of the inner structure of galaxies, without the need to correct for selection effects.
 }
@@
 \end{equation}
-where $x$ and $y$ are cartesian axes centred on the lens centre, with $x$ pointing towards the major axis, and where $q$ is the minor-to-major axis ratio.
+where $x$ and $y$ are Cartesian axes centred on the lens centre, with $x$ pointing towards the major axis, and where $q$ is the minor-to-major axis ratio.
 
@@
 The third column shows the source footprint defined with the procedure described above.
-Pink pixels correspond to the largest footprint that maximizes the number of detected images, while purple pixels belong to the $2\sigma$ detection footprint.
+Pink pixels correspond to the largest footprint that maximises the number of detected images, while purple pixels belong to the $2\sigma$ detection footprint.
 In the first, second and sixth example, these two footprints coincide.
@@
 
-We measured $\crosssect$ for lenses with a fixed radial density profile and different ellipticities, with respect to point sources of diferent brightnesses.
+We measured $\crosssect$ for lenses with a fixed radial density profile and different ellipticities, with respect to point sources of different brightnesses.
 In particular, we set $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\tein=\teff$, and set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
@@
 
-Most of the sources that result in detectable lenses produce two detectable images. These are sources that are located in the region enclosed bewteen the radial and the tangential caustic. 
+Most of the sources that result in detectable lenses produce two detectable images. These are sources that are located in the region enclosed between the radial and the tangential caustic. 
 If the source is located within the tangential caustic, however, four detectable images are usually created.
@@
 
-The distribution in stellar mass, halo mass and dark matter inner slope of our sample of simulated foreground galaxies depends on parameters describing the intinsic scatter in these properties, namely $\sigma_{\mathrm{sps}}$ and $\sigma_h$.
+The distribution in stellar mass, halo mass and dark matter inner slope of our sample of simulated foreground galaxies depends on parameters describing the intrinsic scatter in these properties, namely $\sigma_{\mathrm{sps}}$ and $\sigma_h$.
 Direct observational constraints on these quantities are poor.
@@
 
+\todo{Shun-Sheng please fill in}
 %\ale{To be filled by Shun-Sheng}.
@@
 \end{equation}
-and charachteristic magnitude
+and characteristic magnitude
 \begin{equation}
@@
 This is motivated by the fact that the quasar lenses are meant to simulate a sample assembled for the purpose of carrying out time delay measurements, which in turn require combining single-visit detections over many epochs.
+The scenario that we are simulating, then, is that of a lens search in a Euclid-like survey, followed-up with LSST time-domain observations.
 
@@
 More interesting is the relative number of quad lenses with respect to the total, which is about $9\%$. This is a slightly smaller value than the fraction of quads predicted by \citet{O+M10}.
-The reason for this discrepancy lies in the differences bewteen the lens mass models in the two simulations.
-
-\subsection{Einstein radius cuts}\label{ssec:teincuts}
+The reason for this discrepancy lies in the differences between the lens mass models in the two simulations.
+
+%\subsection{Einstein radius cuts}\label{ssec:teincuts}
+\subsection{Completeness in Einstein radius}\label{ssec:teincuts}
 
@@
 Although all of the lenses in these samples are detected, it does not necessarily follow that they would all be included in a strong lensing study.
-There can be a few reasons for escluding certain lenses from a sample. 
+There can be a few reasons for excluding certain lenses from a sample. 
 One is the low accuracy of lens finders: current automated lens finding algorithms tend to produce lens candidates samples with low purity \citep[see e.g.][]{Son++18a,Pet++19,Sav++22}.
@@
 
-Both of these scenarios can result in samples in which lenses with a larger Einstein radius are preferentially selected.
+Both of these scenarios can result in samples that are incomplete below a certain value of the Einstein radius. %in which lenses with a larger Einstein radius are preferentially selected.
 In order to simulate these situations, we apply an additional cut to lens samples, by selecting only lenses with an angular Einstein radius larger than a limiting value $\theta_{\mathrm{Ein,min}}$. 
+We refer to this value as the completeness limit.
 This is equivalent to multiplying $\pdet$ by a step function in the Einstein radius, which is zero for $\tein < \theta_{\mathrm{Ein,min}}$ and one otherwise.
@@
 \Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the parent population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
-A minimum Einstein radius of $0.5''$ is close to what can be currently be achieved via visual inspection of high-resolution space-based images \cite{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
+A completeness limit of $0.5''$ is close to what can be currently be achieved via visual inspection of high-resolution space-based images \cite{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
 \begin{figure*}
@@
 \caption{
-Comparison bewteen the properties of lens samples and the parent population. Distribution in foreground galaxy parameters.
+Comparison between the properties of lens samples and the parent population. Distribution in foreground galaxy parameters.
 Filled contours: distribution of the parent sample.
@@
 \caption{
-Comparison bewteen the properties of lens samples and the parent population. Distribution in background source parameters.
+Comparison between the properties of lens samples and the parent population. Distribution in background source parameters.
 Filled contours: distribution of the parent sample.
@@
 This can be explained as follows.
-For a given value of the observed stellar mass, strong lensing selection favors galaxies with a larger $\asps$ or $\mhalo$, or with a smaller half-light radius.
+For a given value of the observed stellar mass, strong lensing selection favours galaxies with a larger $\asps$ or $\mhalo$, or with a smaller half-light radius.
 Since the galaxy stellar mass function of the parent population is steep, lenses tend to have a relatively small $\mobs$ and large values of $\asps$ or $\mhalo$ for their observed stellar mass (this is shown more clearly in the next section).
@@
 Similarly to the $\asps$ case, the strong lenses are biased towards larger values of the halo mass, with the bias being larger for higher-scatter simulations and more restrictive cuts on $\tein$.
-In the fiducial scatter scenario with a minimum Einstein radius of $\theta_{\mathrm{Ein,min}}=1.0''$, the halo masses of lenses are on average $0.16$~dex larger than those of their parent population.
+In the fiducial scatter scenario with a completeness limit of $\theta_{\mathrm{Ein,min}}=1.0''$, the halo masses of lenses are on average $0.16$~dex larger than those of their parent population.
 
@@
 The first two points are intrinsic to a strong lensing survey and constitute an unavoidable source of bias. 
-The best case scenario occurs when the efficiency of including a detected strong lens in a survey is always one.
+The best case scenario occurs when the efficiency of including a detected strong lens in a survey is always one, and the sample is $100\%$ complete.
 In this case, the third condition does not introduce any further selection and the strong lensing bias is minimised.
@@
 In this case, the bias on $\asps$ is slightly smaller than $0.04$~dex.
-The current systematic uncertainty on $\asps$ is $0.2-0.3$~dex: this is roughly the difference in mesaurements of the stellar mass of a galaxy obtained with a Chabrier or a Salpeter IMF.
+The current systematic uncertainty on $\asps$ is $0.2-0.3$~dex: this is roughly the difference in measurements of the stellar mass of a galaxy obtained with a Chabrier or a Salpeter IMF.
 Compared with this uncertainty, the amplitude of the strong lensing bias on $\asps$ is small: strong lensing observations can be used directly to discriminate between these two alternative choices of IMF, without the need to correct for selection effects.
@@
 The \citet{O+A17} sample was selected primarily via spectroscopy, by looking for signatures of two galaxies at different redshifts in the Sloan Digital Sky Survey \citep[SDSS][]{Yor++00} data.
-SDSS spectra were taken in fibers with a $1.5''$ radius.
-Lensed galaxies that are comparable in size to this scale, or larger, are less likely to be detected, because part of their flux extends outside of the fiber.
+SDSS spectra were taken in fibres with a $1.5''$ radius.
+Lensed galaxies that are comparable in size to this scale, or larger, are less likely to be detected, because part of their flux extends outside of the fibre.
 Compact galaxies with a large overall magnification, instead, are more likely to be detected \citep[see the discussion in section 5.3 of][]{O+A17}.
@@
 On the one hand the biases in the mass-related quantities ($\asps$ and $\mdmfive$) are very similar in simulations with the same foreground galaxy population and different background sources (see \Fref{fig:quasarbias}).
-On the other hand, these biases are a function of the minimum Einstein radius.
+On the other hand, these biases are a function of the completeness limit.
 In order to use a sample of strong lenses as a prior for another sample, then, it is important to make sure that 1) the parent population of foreground galaxies among which lenses are searched for is the same in both surveys, and 2) the two surveys can probe the same distribution in Einstein radius.
@@
 In order to correctly interpret stellar dynamics data, assumptions on the three-dimensional structure of a lens must be made.
-This is directly related to the projected ellipcitiy: galaxies with an axis ratio close to one tend to be preferentially elongated in the line-of-sight direction, and viceversa.
+This is directly related to the projected ellipticity: galaxies with an axis ratio close to one tend to be preferentially elongated in the line-of-sight direction, and vice-versa.
 When using stellar dynamics-based mass measurements to inform the properties of a sample of quad lenses, therefore, it is important to take into account possible biases due to the different three-dimensional structure of the two samples.
@@
 We used observational constraints from the fundamental plane to set the values of the intrinsic scatter in our fiducial simulation.
-In that case, assuming a survey that probes the Einstein radius distribution down to $\tein=0.5''$ with relatively high completeness, the bias on $\asps$ is as low as $0.04$~dex (10\%).
+In that case, assuming a survey that is complete down to $\tein=0.5''$, the bias on $\asps$ is smaller than $0.04$~dex (10\%).
 This bias is much smaller than the current systematic uncertainty on the stellar population synthesis-based stellar masses.
 Therefore, strong lensing measurements could be used directly to calibrate stellar mass measurements of massive galaxies to $10\%$ accuracy, without the need to correct for selection effects.
-Under the same assumptions, the strong lensing bias on the average halo mass at fixed stellar mass is $0.04$~dex, while that on the inner dark matter distribution at fixed stellar mass and size is $0.02$~dex.
+Under the same assumptions, the strong lensing bias on the average halo mass at fixed stellar mass is $0.07$~dex, while that on the inner dark matter distribution at fixed stellar mass and size is $0.02$~dex.
 
@@
 This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
-Samples of quad lenses, however, are biased towards galaxies with larger ellipcity, which implies that their three-dimensional structure is on average different from that of the entire sample of lenses.
+Samples of quad lenses, however, are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is on average different from that of the entire sample of lenses.
 This means that particular care must be taken when stellar dynamics measurements obtained on galaxy-galaxy lenses is used to inform the properties of quads.
@@
 \Fref{fig:qhist} shows a histogram of the observed distribution.
-We fitted this with a beta disribution:
+We fitted this with a beta distribution:
 \begin{equation}\label{eq:betaappendix}

### 2022-08-23 paper/seleff.tex
revisions after Anowar's comments

@@
 \def\tein{\theta_{\mathrm{Ein}}}
+\def\teinmin{\theta_{\mathrm{Ein,min}}}
 \def\teff{\theta_{\mathrm{e}}}
@@
 \def\pdet{{\rm P}_\mathrm{det}}
+\def\psel{{\rm P}_\mathrm{sel}}
+\def\pfind{{\rm P}_\mathrm{find}}
 \def\pdetone{{\rm P}_{\mathrm{det},1}}
@@
 \begin{equation}\label{eq:one}
-\prsl(\psilens,\psisource|S) \propto \prlens(\psilens)\prsource(\psisource)\pdet(\psilens,\psisource|S).
-\end{equation}
-In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\pdet$ describes the probability of detecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
-This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be recognised as a strong lens.
+\prsl(\psilens,\psisource|S) \propto \prlens(\psilens)\prsource(\psisource)\psel(\psilens,\psisource|S).
+\end{equation}
+In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\psel$ describes the probability of selecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
+This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be detected and recognised as a strong lens.
 
@@
 If the main goal of a lensing survey is to characterise the properties of the strong lens population, then it can be accomplished by directly analysing this term.
-For many applications of strong lensing, however, the aim is to constrain the properties of the general galaxy or source population, $\prlens$ and $\prsource$, which are coupled in a nontrivial way via the lens detection probability $\pdet$.
+For many applications of strong lensing, however, the aim is to constrain the properties of the general galaxy or source population, $\prlens$ and $\prsource$, which are coupled in a nontrivial way via the lens selection probability $\psel$.
 In order to obtain an unbiased estimate of either $\prlens$ or $\prsource$, then, it is necessary to invert \Eref{eq:one}.
-In principle, this can be done with a Bayesian hierarchical formalism, provided that the lens detection probability $\pdet$ is known \citep{Son22}.
-For most of the existing strong lensing surveys, however, characterising $\pdet$ is very challenging: the process of determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
-
-The problem of inverting \Eref{eq:one} is a difficult one to tackle exactly.
+In principle, this can be done with a Bayesian hierarchical formalism, provided that the lens selection probability $\psel$ is known \citep{Son22}.
+%For most of the existing strong lensing surveys, however, characterising $\psel$ is very challenging.
+%: the process of determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
+This factor can be written as the following product:
+\begin{equation}\label{eq:two}
+\psel(\psilens,\psisource|S) = \pdet(\psilens,\psisource)\pfind(\psilens,\psisource|S),%\footnote{\citet{Son22} implicitly assumed $\pfind\equiv1$, therefore $\pdet$ and $\psel$ can be used interchangeably in the context of that work},
+\end{equation}
+where $\pdet$ is the probability of detecting a strong lensing event, and $\pfind(\psilens,\psisource|S,\mathrm{det})$ is the probability of correctly classifying it as such\footnote{\citet{Son22} implicitly assumed $\pfind\equiv1$, therefore $\pdet$ and $\psel$ can be used interchangeably in the context of that work.}.
+The detection probability $\pdet$ can be obtained via simulation.
+The main challenge is characterising $\pfind$: in most of the existing strong lensing surveys, the process of determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
+
+For the above reasons, the problem of inverting \Eref{eq:one} is a difficult one to tackle exactly.
 Few studies have attempted to explicitly account for strong lensing selection effects, usually by making ad-hoc simplifying assumptions \citep{Son++15,O+A17,Son++19}.
@@
 their stellar and dark matter mass density profiles); b) the ellipticity of the lenses; c) the size-magnitude distribution of the lensed sources.
-The answer to this question depends on 1) how the lens detection probability $\pdet$ varies as a function of galaxy and source properties, and on 2) the shape of the galaxy and source parameters distribution $\prlens$ and $\prsource$.
-To understand this second point we can imagine the limiting case in which both $\prlens$ and $\prsource$ are Dirac delta functions: in this limit, $\prsl$ simply reduces to the product $\prlens\prsource$ up to a scaling constant, corresponding to a case in which the lensing bias is none.
+The answer to this question depends on 1) how the lens detection probability $\pdet$ varies as a function of galaxy and source properties, 2) the efficiency of a survey at correctly classifying detected strong lenses (i.e.
+$\pfind$), and on 3) the shape of the galaxy and source parameters distribution $\prlens$ and $\prsource$.
+To understand this third point we can imagine the limiting case in which both $\prlens$ and $\prsource$ are Dirac delta functions (i.e.
+all lenses and sources are identical): in this limit, $\prsl$ simply reduces to the product $\prlens\prsource$ up to a scaling constant, corresponding to a case in which the lensing bias is none.
 
 \citet{MVK09} carried out a thorough study of point 1): they quantified how the properties of a lens galaxy determine its probability of creating a lensing event with a point source.
-In this work we revisit the \citet{MVK09} study, we expand it to the extended source case and, most importantly, we address point 2) as well: we simulate large populations of strong lenses using empirical models, simulate lensing surveys with realistic assumptions for the lens probability $\pdet$ and quantify the lensing bias under various scenarios.
+In this work we revisited the \citet{MVK09} study, expanding it to the extended source case.
+We simulated individual lenses and examined how the lens detection probability varies with lens and source properties.
+In addition, we addressed point 2) and 3) as well: we simulated large populations of strong lenses using empirical models, we simulated the lens detection and finding phase, and quantified the lensing bias under various scenarios.
 %We explore how the results change as a function of the scatter in mass parameters at fixed light, of the efficiency of finding lenses with small image separation, and of source type: we simulate both galaxy and quasar sources.
 %We explore how the results change as a function of the scatter in mass parameters at fixed light, and as a function of the efficiency of a lens survey at discovering small image separation lenses.
-We explore how the results change as a function of the efficiency of a lens survey at discovering small image separation lenses, and of the scatter in mass parameters at fixed light.
+We explored how the results change as a function of the efficiency of a lens survey at discovering small image separation lenses, and of the scatter in mass parameters at fixed light.
 
 %In light of the results of our experiments, we discuss to what extent can studies of galaxy structure based on strong lenses be generalised to describe non-lens galaxies.
-Finally, we address the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses.
+Finally, we addressed the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses.
 This last point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
@@
 In \Sref{sect:basics} we introduce the basics of gravitational lensing.
-In \Sref{sect:indlenses} we study individual lens systems, to look for trends in the lensing cross-section with various lens and source properties.
-In \Sref{sect:lenspop} we describe our simulations of lens populations.
-In \Sref{sect:results} we show the results of our analysis on the simulated lens populations.
+In \Sref{sect:indlenses} we study individual lens systems. %, to look for trends in the lensing cross-section with various lens and source properties.
+In \Sref{sect:lenspop} we describe our simulations of lens surveys.
+In \Sref{sect:results} we show the results of our analysis on the simulated lens survey data.
 We discuss the results in \Sref{sect:discuss} and draw conclusions in \Sref{sect:concl}.
@@
 First column: caustics (red curves) and source position (blue circle).
-The radius of the circle corresponds to the radius at which the surface brightness is equal to $2\sigma$ the sky background fluctuation for a single pixel.
+The radius of the circle corresponds to the radius at which the surface brightness is equal to $2\sigma$ the sky background rms fluctuation for a single pixel.
 In other words, the blue circle delimits the area of the source that can be detected.
@@
 In the limit of low density of background sources, which is satisfied in all practical cases, the probability of a lensing event is proportional to $\crosssect$.
+The lensing cross-section defined via \Eref{eq:crosssect} depends solely on the lens detection probability $\pdet$ and does not take into account whether the lens can be correctly classified by a lens finder.
+This separate selection step is captured by the term $\pfind$ in \Eref{eq:two}.
+In this section we consider solely on the detectability of a lens, and therefore focus only on $\crosssect$.
+In \Sref{sect:lenspop}, when considering specific lens survey simulations, we introduce $\pfind$.
 
@@
 In most practical cases this is the second brightest image.
-As \Fref{fig:1dmag} shows, the magnification is very large for sources close to the lens (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
+As \Fref{fig:1dmag} shows, the magnification is very large for sources close to the lens centre (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
 While for the model with $f_{\mathrm{DM}}=0.8$ (purple curve) the magnification is above unity everywhere, other lens models can produce highly de-magnified secondary images for large values of $\beta$.
@@
 
-We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification.
+We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources randomly distributed over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification.
 We then measured the fraction of sources that are strongly lensed according to the criterion of \Eref{eq:pointdet} and multiplied this value by the area over which sources are located.
@@
 We also assumed that the background noise is an uncorrelated Gaussian field.
-We carried out experiments with sources with different values of the total flux, $f$, and half-light radius, $\theta_{\mathrm{e,s}}$.
+We carried out experiments with sources with different values of the total unlensed flux, $f$, and half-light radius, $\theta_{\mathrm{e,s}}$.
 The results are shown in \Fref{fig:extcs}.
-The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius.
+The total flux of the source, indicated in the legend, is measured in units of the sky background rms fluctuation within an area equal to the square of the Einstein radius.
 
@@
 The lens model is fixed to be the same of \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
-Each solid line corresponds to a different value of the total flux of the source. The flux $f$ is expressed in terms of the background noise fluctuation measured over an area equal to $\tein^2$.
+Each solid line corresponds to a different value of the total unlensed flux of the source.
+The flux $f$ is expressed in terms of the background noise rms fluctuation measured over an area equal to $\tein^2$.
 The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
@@
 From a qualitative point of view, this behaviour can be observed also in the absence of lensing: increasing the size of a galaxy while keeping its flux fixed lowers its average surface brightness.
-If the surface brightness drops below the sky fluctuation level, then it becomes very difficult to detect it.
+If the surface brightness drops below the sky rms fluctuation level, then it becomes very difficult to detect it.
 In order to determine whether there are lensing-specific features in the $\crosssect-\theta_{\mathrm{e,s}}$ relation of \Fref{fig:extcs}, we computed, for each source flux, the maximum half-light radius for which it can be detected in the absence of lensing.
@@
 
-In the previous section we showed how the lensing cross-section, which is closely related to the term $\pdet$ in \Eref{eq:one}, varies as a function of lens and source properties.
+In the previous section we showed how the lensing cross-section, which is closely related to the lens detection probability $\pdet$, varies as a function of lens and source properties.
 From here on we focus on the effect that those trends have on populations of lenses.
@@
 %\subsection{Einstein radius cuts}\label{ssec:teincuts}
-\subsection{Completeness in Einstein radius}\label{ssec:teincuts}
+\subsection{Lens finding probability}\label{ssec:teincuts}
 
@@
 Both of these scenarios can result in samples that are incomplete below a certain value of the Einstein radius. %in which lenses with a larger Einstein radius are preferentially selected.
-In order to simulate these situations, we apply an additional cut to lens samples, by selecting only lenses with an angular Einstein radius larger than a limiting value $\theta_{\mathrm{Ein,min}}$. 
-We refer to this value as the completeness limit.
-This is equivalent to multiplying $\pdet$ by a step function in the Einstein radius, which is zero for $\tein < \theta_{\mathrm{Ein,min}}$ and one otherwise.
-%Both for the lensed galaxies and lensed quasars, we apply an additional cut, by selecting only lenses with an angular Einstein radius larger than a limiting value $\theta_{\mathrm{Ein,min}}$.
-%This additional cut is meant to approximate the difficulty in detecting strongly lensed images with small separation, because of the contamination from the light of the lens galaxy.
-We explored scenarios with different values of $\theta_{\mathrm{Ein,min}}$.
+We simulate these situation via the following Einstein radius-dependent lens finding probability:
+\begin{equation}
+\pfind(\tein|S,\mathrm{det}) = \left\{\begin{array}{ll} 1 & \rm{if}\,\tein > \teinmin \\
+0 & \rm{otherwise}\end{array}\right. .
+\end{equation}
+In words, all lenses with Einstein radius larger than $\theta_{\mathrm{Ein,min}}$ are included in the sample, while all those with smaller Einstein radius are excluded.
+We refer to $\teinmin$ as the completeness limit: our simulated lens samples are complete down to $\tein=\teinmin$.
+We explored scenarios with different values of $\teinmin$.
 As the next section shows, the larger the minimum Einstein radius, the higher the strong lensing bias.
@@
 Each one of these conditions introduces a bias with respect to the parent population of foreground galaxies and background sources.
-Together, they define the lens detection probability term $\pdet$ of \Eref{eq:one}.
+Together, they define the lens selection probability term $\psel$ of \Eref{eq:one}.
 %In the best case scenario, the only sources of bias are the first two points.
@@
 
-Although the amplitude of the strong lensing bias in difficult-to-observe mass-related quantities is relatively small, compared to current uncertainties, there might be situations in which one wishes to account for it.
-The formally correct way to do this is by explicitly modelling all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
-Doing so, however, can be difficult in practice.
-Therefore, it can be useful to develop alternative strategies to correct for the bias, or to mitigate its effects.
+%Although the amplitude of the strong lensing bias in difficult-to-observe mass-related quantities is relatively small, compared to current uncertainties, there might be situations in which one wishes to account for it.
+%The formally correct way to do this is by explicitly modelling all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
+%When the entire lens selection procedure is fully automated, then it is possible to use machine-learning to make this problem computationally tractable \citep{Leg++22}. 
+%If human selection is involved, however, it can be a difficult problem to solve in practice.
+%Therefore, it can be useful to develop alternative strategies to correct for the bias, or to mitigate its effects.
 
@@
 This description could be extended by including the central velocity dispersion as an additional control parameter.
-The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution.
+The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution marginalised over the whole population.
 The central velocity dispersion, however, is very sensitive to the orbital anisotropy and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
@@
 
-If one wishes to directly correct for the strong lensing bias in a computationally manageable way, one possibility is to try and approximate $\pdet$. 
-In general, $\pdet$ depends nontrivially on the position, redshift and surface brightness distribution of the source, as well as on the lens mass density profile and redshift.
-This requires in principle to explicitly model the distribution of all of these quantities \citep[as done by][]{Son22}.
-However, we can make the assumption that $\pdet$ can be separable in a term that depends purely on the strong lensing cross-section and a term that depends purely on the Einstein radius.
+If one wishes to directly account for the strong lensing bias, the formally correct procedure is to explicitly model all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
+Although this can be computationally challenging, machine learning can offer an efficient alternative \citep{Leg++22}.
+In order for either of these approaches to work, however, it is essential that the lens selection procedure can be simulated.
+This can be difficult to do if human selection is involved in the definition of the lens sample.
+In that case, a possible alternative is to approximate the lens finding probability $\pfind$. %, which is the component of $\psel$ that is difficult to simulate. 
+For example, we can make the assumption that $\pfind$ depends purely on the Einstein radius.
 This is a reasonable assumption, as long as the lens finding procedure does not selectively pick lenses with different image configurations depending on their Einstein radius.
-It would allow us to marginalise over the source position and surface brightness parameters, thus simplifying the problem.
-The dependence of $\pdet$ on $\tein$ could then be described empirically and inferred during the analysis.
+%It would allow us to marginalise over the source position and surface brightness parameters, thus simplifying the problem.
+The dependence of $\pfind$ on $\tein$ could then be described empirically and inferred during the analysis.
 This is essentially the same approach adopted by \citet{Son++19} in the analysis of strong and weak lensing data from the Hyper Suprime-Cam Survey.
+
+%Although the amplitude of the strong lensing bias in difficult-to-observe mass-related quantities is relatively small, compared to current uncertainties, there might be situations in which one wishes to account for it.
+%The formally correct way to do this is by explicitly modelling all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
+%When the entire lens selection procedure is fully automated, then it is possible to use machine-learning to make this problem computationally tractable \citep{Leg++22}. 
+%If human selection is involved, however, it can be a difficult problem to solve in practice.
+%Therefore, it can be useful to develop alternative strategies to correct for the bias, or to mitigate its effects.
+
+%In general, $\pdet$ depends nontrivially on the position, redshift and surface brightness distribution of the source, as well as on the lens mass density profile and redshift.
+%This requires in principle to explicitly model the distribution of all of these quantities \citep[as done by][ in a simplified case]{Son22}.
+%However, we can make the assumption that $\pdet$ can be separable in a term that depends purely on the strong lensing cross-section and a term that depends purely on the Einstein radius.
+%This is a reasonable assumption, as long as the lens finding procedure does not selectively pick lenses with different image configurations depending on their Einstein radius.
+%It would allow us to marginalise over the source position and surface brightness parameters, thus simplifying the problem.
+%The dependence of $\pdet$ on $\tein$ could then be described empirically and inferred during the analysis.
+%This is essentially the same approach adopted by \citet{Son++19} in the analysis of strong and weak lensing data from the Hyper Suprime-Cam Survey.
 

### 2022-08-23 paper/seleff.tex
fixed typo
added acknowledgements

@@
 This separate selection step is captured by the term $\pfind$ in \Eref{eq:two}.
-In this section we consider solely on the detectability of a lens, and therefore focus only on $\crosssect$.
+In this section we consider exclusively the detectability of a lens, and therefore focus only on $\crosssect$.
 In \Sref{sect:lenspop}, when considering specific lens survey simulations, we introduce $\pfind$.
@@
 
-%\begin{acknowledgements}
-
-%\end{acknowledgements}
+\begin{acknowledgements}
+
+The collaboration leading to this work was initiated at the 2022 Lorentz Center workshop “Bridging gaps between dynamical probes of galaxies”.
+
+\end{acknowledgements}
 

### 2022-09-13 paper/seleff.tex
added background sources description

@@
 
-\todo{Shun-Sheng please fill in}
-%\ale{To be filled by Shun-Sheng}.
+Our background galaxy population is taken from the \textsc{surfs}-based KiDS-Legacy-Like Simulation (SKiLLS) input catalogue, a hybrid simulation catalogue integrating cosmological simulation with high-quality imaging observations (Li et al., in prep).
+The cosmological simulation is obtained from the Synthetic UniveRses For Surveys (\textsc{surfs}) simulations, a set of $N$-body simulations from \citet{Ela++18}.
+The galaxy properties, including the star formation history and the metallicity history, are from an open-source semi-analytic model named \textsc{Shark}\footnote{\url{https://github.com/ICRAR/shark}}~\citep{Lag++18}.
+The original photometry is drawn from a stellar population synthesis technique using stellar synthesis libraries with physically motivated dust attenuation and re-emission models~\citep{Rob++20}.
+Li et al.
+(in prep) further applied an empirical correction to the original synthetic photometry to better agree with the COSMOS2015 observations~\citep{Lai++16}.
+The galaxy morphology is described by a S\'ersic profile with three parameters: the half-light radius in angular units $\theta_{
+mathrm{e,s}}$, the S\'ersic index $n_\mathrm{s}$, and the axis ratio $q_\mathrm{s}$.
+These structural parameters are learned from the imaging data obtained with the Advanced Camera for Surveys (ACS) instrument on the \textit{Hubble} Space Telescope~\citep{Gri++12}.
+We refer to Li et al.
+(in prep) for details on the learning algorithm and validation.
+
+The complete SKiLLS catalogue contains ${\sim}108~{\rm deg}^2$ of galaxies with redshift up to $2.5$ and $r$-band apparent magnitude down to $27$.
 We applied a lower limit to the source redshift, by selecting only sources with $z_s > 0.8$.

### 2022-12-27 paper/seleff.tex
starting to address Ned's comments

@@
 In those cases it is difficult to establish whether the arc is a set of blending images or not, and the decision of including such systems in a strong lens sample is often arbitrary.
-Here we adopt the following definition: an extended source is strongly lensed if the number of surface brightness peaks that are detected is larger than its intrinsic number of peaks in the absence of lensing.
+Here we adopt the following working definition: {\em an extended source is strongly lensed if the number of surface brightness peaks that are detected is larger than its intrinsic number of peaks in the absence of lensing}.
 
@@
 
-Throughout this work we describe the mass distribution of the lens galaxies with a two-component model consisting of a S\'{e}rsic profile, describing the baryonic component of a galaxy, and a generalised Navarro, Frenk \& White (gNFW) profile, describing the dark matter halo.
-We assume that the two components are concentric.
-We also assume that the baryonic component consists entirely of stars.
-This is a good approximation for the majority of the known strong lenses, which are massive elliptical galaxies with little gas.
+%In this work we focus on massive elliptical galaxies as lenses. This is because this class of objects makes up the vast majority of known strong lenses.
+%The type of lenses that we focus on in this work are massive elliptical galaxies: these make up the vast majority of known strong lenses.
+%The lens galaxies that we focus on in this work are massive early-types: these make up the vast majority of known strong lenses.
+In this work we focus on massive early-type galaxies as lenses, as these make up the vast majority of known strong lenses.
+We describe their mass distribution with a model consisting of two concentric components, one describing the baryons and one for the dark matter.
+We assume that the baryonic component consists entirely of stars, thereby neglecting gas, which is known to contribute very little to the mass of early-type galaxies in the inner regions that are probed by strong lensing.
+We then assume that the stars follow a S\'{e}rsic profile, with %with a fixed index of $n=4$ (also known as a de Vaucouleurs profile).
+projected surface mass density given by
+\begin{equation}
+\Sigma(R) = \Sigma_0 \exp{\left\{-b\left(\frac{R}{\reff}\right)^{1/n}\right\}}.
+\end{equation}
+In the above equation,
+\begin{equation}
+\Sigma_0 = \frac{\mstar b^{2n}}{2\pi n \reff^2 \Gamma(2n)},
+\end{equation}
+$\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is given by \citep{C+B99}
+\begin{equation}
+b(n) \approx 2n -\frac13 + \frac{4}{405n} + \frac{46}{25515n^2} + O(n^{-3}).
+\end{equation}
+Throughout this paper we indicate with $\teff$ the angular size of the half-light radius.
+
+Throughout our work we fix the S\'{e}rsic index of the lenses to $n=4$, also known as a de Vaucouleurs model.
+Although early-type galaxies are often described with a free S\'{e}rsic index, a de Vaucouleurs profile is able to reproduce their surface brightness distribution to a few percent in the radial range $1\rm{kpc} < R < 30\rm{kpc}$ \citep[see e.g.][]{Son20}, which is the region that is relevant for strong lensing.
 Finally, we assume that the light distribution of the stellar component follows its mass distribution exactly.
@@
 
-The projected surface mass density of a S\'{e}rsic profile is given by
-\begin{equation}
-\Sigma(R) = \Sigma_0 \exp{\left\{-b\left(\frac{R}{\reff}\right)^{1/4}\right\}},
-\end{equation}
-with
-\begin{equation}
-\Sigma_0 = \frac{\mstar b^{2n}}{2\pi n \reff^2 \Gamma(2n)},
-\end{equation}
-where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is given by \citep{C+B99}
-\begin{equation}
-b(n) \approx 2n -\frac13 + \frac{4}{405n} + \frac{46}{25515n^2} + O(n^{-3}).
-\end{equation}
-Throughout this paper we indicate with $\teff$ the angular size of the half-light radius.
-
-The dark matter component is first defined by means of its three-dimensional distribution, which for a spherically symmetric gNFW profile is
+%Throughout this work we describe the mass distribution of the lens galaxies with a two-component model consisting of a S\'{e}rsic profile, describing the baryonic component of a galaxy, and a generalised Navarro, Frenk \& White (gNFW) profile, describing the dark matter halo.
+%We assume that the two components are concentric.
+%We also assume that the baryonic component consists entirely of stars.
+%This is a good approximation for the majority of the known strong lenses, which are massive elliptical galaxies with little gas.
+%Finally, we assume that the light distribution of the stellar component follows its mass distribution exactly.
+That is, we do not allow for the presence of gradients in the stellar mass-to-light ratio.
+This assumption is probably broken in real galaxies but, as we discuss in \Sref{sect:discuss}, does not impact our investigation.
+
+%The projected surface mass density of a S\'{e}rsic profile is given by
+%\begin{equation}
+%\Sigma(R) = \Sigma_0 \exp{\left\{-b\left(\frac{R}{\reff}\right)^{1/n}\right\}},
+%\end{equation}
+%with
+%\begin{equation}
+%\Sigma_0 = \frac{\mstar b^{2n}}{2\pi n \reff^2 \Gamma(2n)},
+%\end{equation}
+%where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is given by \citep{C+B99}
+%\begin{equation}
+%b(n) \approx 2n -\frac13 + \frac{4}{405n} + \frac{46}{25515n^2} + O(n^{-3}).
+%\end{equation}
+%Throughout this paper we indicate with $\teff$ the angular size of the half-light radius.
+
+We describe the dark matter halo with a generalised Navarro, Frenk \& White (gNFW) profile.
+We first define it by its three-dimensional distribution, which for a spherically symmetric profile is
 \begin{equation}
@@
 \Fref{fig:kappa} shows the dimensionless surface mass density profile of S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
-All of the profiles have a S\'{e}rsic index $n=4$, a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component.
+All of the profiles have %a S\'{e}rsic index $n=4$, 
+a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component (these assumptions are dropped later).
 Two main features emerge from \Fref{fig:kappa}.
@@
 Magnification of the secondary image as a function of the source position.
-The lenses are axisymmetric composite model. Their density profile is plotted in \Fref{fig:kappa}.
+The lenses are axisymmetric composite models. Their density profile is plotted in \Fref{fig:kappa}.
 Vertical dotted lines mark the position of the radial caustic.
@@
 The lens galaxy is a composite model, introduced in section \ref{ssec:profile}, with angular half-light radius equal to the Einstein radius.
-Left panel: cross-section, in units of $\pi\tein^2$, as a function of the inner dark matter slope. The dark matter fraction is fixed at $\fdm=0.5$.
-Right panel: cross-section as a function of the fraction of dark matter within the half-light radius. The inner dark matter slope is fixed at $\gammadm=1.5$.
+The left panel shows the effect of varying the slope $\gammadm$ for fixed $\fdm$; conversely, the right panel shows variations as a function of $\fdm$ with $\gammadm$ fixed.
+%Left panel: cross-section, in units of $\pi\tein^2$, as a function of the inner dark matter slope. The dark matter fraction is fixed at $\fdm=0.5$.
+%Right panel: cross-section as a function of the fraction of dark matter within the half-light radius. The inner dark matter slope is fixed at $\gammadm=1.5$.
 The system is defined as a strong lens if at least two images are detected.
@@
 For instance, at fixed observed stellar mass and half-light radius, the bias on the projected dark matter mass enclosed within $5$~kpc, $\mdmfive$, is only a few percent in the fiducial scatter scenario with $\theta_{\mathrm{Ein,min}} < 1.0''$.
-This means that strong lenses can be used to understand the inner dark matter distribution of galaxies, as long as the dependence of the dark matter distribution on the properties of the baryonic component is accurately modelled \citep[for example, by following the approach of][]{S+C21}.
+This means that strong lenses can indeed be used to understand the inner dark matter distribution of galaxies, as long as the dependence of the dark matter distribution on the properties of the baryonic component is accurately modelled \citep[for example, by following the approach of][]{S+C21}.
 

### 2022-12-28 paper/seleff.tex
Ned and Anowar's comments

@@
 
-Throughout our work we fix the S\'{e}rsic index of the lenses to $n=4$, also known as a de Vaucouleurs model.
-Although early-type galaxies are often described with a free S\'{e}rsic index, a de Vaucouleurs profile is able to reproduce their surface brightness distribution to a few percent in the radial range $1\rm{kpc} < R < 30\rm{kpc}$ \citep[see e.g.][]{Son20}, which is the region that is relevant for strong lensing.
+We fix the S\'{e}rsic index of the lenses to $n=4$, corresponding to a de Vaucouleurs model.
+Although early-type galaxies are often described with a free S\'{e}rsic index, a de Vaucouleurs profile is able to reproduce their surface brightness distribution to a few percent in the radial range $1\rm{kpc} < R < 30\rm{kpc}$ \citep[see e.g.][]{Son20}, which is the region that is most relevant for strong lensing.
 Finally, we assume that the light distribution of the stellar component follows its mass distribution exactly.
 That is, we do not allow for the presence of gradients in the stellar mass-to-light ratio.
-This assumption is probably broken in real galaxies but, as we discuss in \Sref{sect:discuss}, does not impact our investigation.
+This assumption is probably broken in real galaxies but, as we discuss in \Sref{ssec:limitations}, does not impact our investigation.
 
@@
 This result suggests that, to first approximation, lensing does not introduce a strong selection in source size.
-This is expected, since gravitational lensing preserves surface brightness: a source that can be detected in the absence of lensing will produce images with the same surface brightness when lensed, which can be detected as well.
+This follows from the fact that gravitational lensing preserves surface brightness: a source that can be detected in the absence of lensing will produce images with the same surface brightness when lensed, which can be detected as well.
 %When the source is extended, however, 
@@
 %
-The dark matter density profile defined in this way is determined uniquely by the halo mass, the stellar mass and the half-light radius (the more concentrated the stellar distribution, the stronger the halo response, the steeper the dark matter density profile).
+The dark matter density profile defined in this way is determined uniquely by the halo mass, the stellar mass and the half-light radius (the more concentrated the stellar distribution, the stronger the halo response and the steeper the dark matter density profile).
+In principle we could have allowed for additional degrees of freedom, for instance by relaxing the assumption of a fixed initial halo concentration.
+In practice, as we explain in section \ref{ssec:limitations}, our main resultsare not affected by this choice.
 %However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope.
@@
 
-Early-type galaxies lie on the stellar mass fundamental plane, a scaling relation between stellar mass, half-light radius and central velocity dispersion \citep{H+B09}:
+Early-type galaxies lie on the stellar mass fundamental plane, a scaling relation between stellar mass, half-light radius and central velocity dispersion \citep{H+B09,deG++21}:
 \begin{equation}\label{eq:fpcompact}
@@
 \end{equation}
-The existence of a fundamental plane relation is a consequence of the virial theorem: the velocity dispersion of a galaxy is directly related to the 3-dimensional mass distribution in its inner regions, which is typically dominated by the stellar component.
+The existence of a fundamental plane relation is a consequence of the virial theorem: the velocity dispersion of a galaxy in dynamical equilibrium is directly related to the 3-dimensional mass distribution in its inner regions, which is typically dominated by the stellar component.
 At fixed observed stellar mass distribution, however, the central velocity dispersion can vary depending on the stellar population synthesis mismatch parameter, on the dark matter mass and on the dark matter density profile.
@@
 
-For the simulation with extended sources, we produced images with a pixel size of $0.1''$.
-We applied a Moffat PSF with an FWHM of $0.2''$.
+For the simulation with extended sources, we produced images with properties similar to those expected for the Euclid Wide survey 
+citep{Sca++22}.
+We used a pixel size of $0.1''$ and we applied a Moffat PSF with an FWHM of $0.2''$ and a $\beta$ parameter of $5.0$.
 Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
-These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
+%These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
 %
@@
 
-\subsection{Limitations of our analysis}
+\subsection{Limitations of our analysis}\label{ssec:limitations}
 
@@
 Therefore, while including them might modify slightly the amplitude of the strong lensing bias in the axis ratio and dark matter distribution parameters, the conclusions our our analysis would not be affected.
+
+We also assumed that the dark matter density profile of the galaxies is determined uniquely by the halo mass and by the stellar mass distribution (see section \ref{ssub:dmprofile}).
+In reality we expect there to be a range of density profiles, for example as a result of halos having a nonzero scatter in their initial concentration parameter.
+We could in principle carry out an experiment with such an additional source of variation in the dark matter profile.
+Allowing for additional flexibility at fixed halo mass scatter parameter $\sigma_h$ results in a larger scatter in the inner dark matter distribution, which would increase the amplitude of the lensing bias.
+However, the value of $\sigma_h$ that we chose for our fiducial experiment was tuned on the basis of the predicted scatter in velocity dispersion around the fundamental plane.
+That scatter would also be increased by adding flexibility to the dark matter density profile.
+Then, in order to be consistent with the analysis carried out in our paper, the value of $\sigma_h$ would need to be lowered accordingly.
+This, in turn, would reduce the amplitude of the strong lensing bias.
+Since lensing and dynamics are sensitive to the mass distribution at comparable scales (the half-light radius and Einstein radius are similar for most of the lenses), we expect these two effects to cancel out to first approximation, leaving the strong lensing bias roughly unchanged.
 

### 2022-12-29 paper/seleff.tex
revising conclusions

@@
 Strong gravitational lensing is a powerful tool for studying galaxies and cosmology.
-Strong lenses have been used to probe the mass structure of massive galaxies \citep{Aug++10, ORF14, Son++15, Sha++21}, to detect substructure \citep{Veg++12, Hez++16, Nie++20}, to carry out detailed studies of magnified star-forming galaxies \citep{Jon++13}, and to measure the expansion rate of the universe with time delays \citep[see][for a review]{T+M16}.
+Strong lenses have been used to probe the mass structure of massive galaxies \citep{Aug++10, ORF14, Son++15, Sha++21}, to detect substructure \citep{Veg++12, Hez++16, Nie++20}, to carry out detailed studies of magnified star-forming galaxies \citep{Jon++13}, and to measure the expansion rate of the universe with time delays \citep[see][ for a review]{T+M16}.
 
@@
 In order to obtain an unbiased estimate of either $\prlens$ or $\prsource$, then, it is necessary to invert \Eref{eq:one}.
-In principle, this can be done with a Bayesian hierarchical formalism, provided that the lens selection probability $\psel$ is known \citep{Son22}.
+In principle, this can be done with a Bayesian hierarchical formalism \citep{Son22}, but knowledge of the lens selection probability $\psel$ is required.
 %For most of the existing strong lensing surveys, however, characterising $\psel$ is very challenging.
@@
 We simulated individual lenses and examined how the lens detection probability varies with lens and source properties.
-In addition, we addressed point 2) and 3) as well: we simulated large populations of strong lenses using empirical models, we simulated the lens detection and finding phase, and quantified the lensing bias under various scenarios.
+In addition, we addressed points 2) and 3) as well: we simulated large populations of strong lenses using empirical models, we simulated the lens detection and finding phase, and quantified the lensing bias under various scenarios.
 %We explore how the results change as a function of the scatter in mass parameters at fixed light, of the efficiency of finding lenses with small image separation, and of source type: we simulate both galaxy and quasar sources.
@@
 As we showed above, this requires the source to lie within the region enclosed by the radial caustic.
-In order to recognise a strong lens in a real strong lensing survey, however, it is not sufficient for multiple images to exist: they must be detected.
-For this reason, given the detection limit of a survey for a point source, $m_{\mathrm{lim}}$, we define as strong lensing event any lens-source system with at least two images brighter than $m_{\mathrm{lim}}$.
-This is the same definition upon which the \citet{MVK09} work is based \citep{VMK09}.
+In order to recognise a strong lens in a real survey, however, it is not sufficient for multiple images to exist: they must be detected.
+For this reason, given the detection limit for a point source, $m_{\mathrm{lim}}$, we define as strong lensing event any lens-source system with at least two images brighter than $m_{\mathrm{lim}}$.
+This is the same definition used by \citep{VMK09}, upon which the \citet{MVK09} work is based.
 Labelling with $m_2$ the magnitude of the second-brightest image, then, in the absence of photometric noise the lens detection probability $\pdet$ becomes
@@
 That is, we do not allow for the presence of gradients in the stellar mass-to-light ratio.
-This assumption is probably broken in real galaxies but, as we discuss in \Sref{ssec:limitations}, does not impact our investigation.
+In section \ref{ssec:limitations} we discuss qualitatively what the implications of relaxing this assumption would be.
+%This assumption is probably broken in real galaxies but, as we discuss in \Sref{ssec:limitations}, does not impact our investigation.
 
@@
 This result suggests that, to first approximation, lensing does not introduce a strong selection in source size.
-This follows from the fact that gravitational lensing preserves surface brightness: a source that can be detected in the absence of lensing will produce images with the same surface brightness when lensed, which can be detected as well.
+While perhaps surprising, this follows from the fact that gravitational lensing preserves surface brightness: a source that can be detected in the absence of lensing will produce images with the same surface brightness when lensed, which can be detected as well.
 %When the source is extended, however, 
@@
 We set the mean of $\log{\asps}$ to $0.1$, as this is an intermediate value among estimates of $\asps$ from the literature \citep{CvD12, Cap++13, SLC15, Son++15, Son++19}.
-For the scatter $\sigma_\alpha$ we adopted a few different values. We explain in \ref{ssub:scat} how these were chosen.
+We adopted a few different values for the scatter $\sigma_\alpha$. We explain in \ref{ssub:scat} how these were chosen.
 
@@
 In principle we could have allowed for additional degrees of freedom, for instance by relaxing the assumption of a fixed initial halo concentration.
-In practice, as we explain in section \ref{ssec:limitations}, our main resultsare not affected by this choice.
+In practice, as we explain in section \ref{ssec:limitations}, our main results are not affected by this choice.
 %However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope.
@@
 \Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the parent population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
-A completeness limit of $0.5''$ is close to what can be currently be achieved via visual inspection of high-resolution space-based images \cite{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
+A completeness limit of $0.5''$ is close to what can currently be achieved via visual inspection of high-resolution space-based images \cite{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
 \begin{figure*}
@@
 Only galaxies with a large $\mobs$ or a small size can therefore act as strong lenses.
+This is an interesting result, because it suggests that it is in principle possible to use the strong lensing bias on $\mobs$ or $\reff$, which is observable, as a way to constrain the amplitude of the intrinsic scatter in the mass parameters, which is poorly known.
 
@@
 %We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we computed the bias for this sample as well.
-We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we also analysed this subsample on its own.
+We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we also analysed that subsample on its own.
 %\Fref{fig:quasarpars} shows the distribution in a 
@@
 In order for a galaxy-source pair to be included as a strong lens in a survey, three condition must be met.
-First of all, part of the source must be multiply imaged.
+First of all, at least part of the source must be multiply imaged.
 Second, the multiple images must be detectable.
@@
 On the one hand, it preferentially selects brighter sources, because the lensing cross-section increases with increasing source brightness.
-On the other hand, it allows the detection of sources that are intrinsically fainter than the detection limit in the absence of lensing.
-Interestingly, we did not detect any significant bias on the source size, at fixed magnitude.
+On the other hand, it allows for the detection of sources that are intrinsically fainter than the detection limit in the absence of lensing.
+Interestingly, we did not find any significant bias on the source size, at fixed magnitude.
 This result follows from the fact that our lens detection criterion relies on a surface brightness threshold, and surface brightness is preserved by lensing.
@@
 The dependence of $\pfind$ on $\tein$ could then be described empirically and inferred during the analysis.
-This is essentially the same approach adopted by \citet{Son++19} in the analysis of strong and weak lensing data from the Hyper Suprime-Cam Survey.
+This is essentially the approach adopted by \citet{Son++19} in the analysis of strong and weak lensing data from the Hyper Suprime-Cam Survey.
 
@@
 
-We also assumed that the dark matter density profile of the galaxies is determined uniquely by the halo mass and by the stellar mass distribution (see section \ref{ssub:dmprofile}).
-In reality we expect there to be a range of density profiles, for example as a result of halos having a nonzero scatter in their initial concentration parameter.
+We also assumed that the dark matter density profile of the galaxies is completely determined by the halo mass and by the stellar mass distribution (see section \ref{ssub:dmprofile}).
+In reality we expect there to be a range of density profiles, for example as a result of halos having a nonzero scatter in their initial (i.e.
+before baryonic infall) concentration parameter.
 We could in principle carry out an experiment with such an additional source of variation in the dark matter profile.
-Allowing for additional flexibility at fixed halo mass scatter parameter $\sigma_h$ results in a larger scatter in the inner dark matter distribution, which would increase the amplitude of the lensing bias.
+Adding a scatter in concentration while keeping the halo mass scatter parameter $\sigma_h$ fixed results in a larger spread in the inner dark matter distribution, which would increase the amplitude of the lensing bias.
+However, the value of $\sigma_h$ that we chose for our fiducial experiment was tuned on the basis of the predicted scatter in velocity dispersion around the fundamental plane.
+That scatter would also be increased by adding flexibility to the dark matter density profile.
+Then, in order to be consistent with the analysis carried out in this paper, the value of $\sigma_h$ would need to be lowered accordingly.
+This, in turn, would reduce the amplitude of the strong lensing bias and introduce a bias on the halo concentration.
+%Allowing for additional flexibility at fixed halo mass scatter parameter $\sigma_h$ results in a larger scatter in the inner dark matter distribution, which would increase the amplitude of the lensing bias.
 However, the value of $\sigma_h$ that we chose for our fiducial experiment was tuned on the basis of the predicted scatter in velocity dispersion around the fundamental plane.
@@
 This, in turn, would reduce the amplitude of the strong lensing bias.
-Since lensing and dynamics are sensitive to the mass distribution at comparable scales (the half-light radius and Einstein radius are similar for most of the lenses), we expect these two effects to cancel out to first approximation, leaving the strong lensing bias roughly unchanged.
+Since lensing and dynamics are sensitive to the mass distribution at comparable scales (the half-light radius and Einstein radius are similar for most of the lenses), we expect these two effects to cancel out to first approximation.
+The lensing bias on quantities directly related to the inner mass distribution, such as $\asps$ and $\mdmfive$, would be left roughly unchanged.
+%leaving the strong lensing bias on roughly unchanged.
+The main effect of adding scatter to the concentration is to reduce the correlation between the total halo mass and the dark matter mass enclosed within the Einstein radius.
+This would then reduce the amplitude of the strong lensing bias on the halo mass.
+
+Another simplifying assumption that we made was that of adopting a constant mass-to-light ratio for the stellar component.
+Massive galaxies are known to have colour gradients, the main effect of which is to cause the stellar half-mass radius to be smaller than the half-light radius \citep{Szo++13,Sue++19}.
+Since, as shown in \Sref{sect:results}, strong lensing bias tends to select galaxies with a more compact stellar distribution, it would also preferentially select galaxies with a steeper mass-to-light ratio gradient.
+However, testing for the amplitude of the strong lensing bias on the mass-to-light ratio gradient is beyond the scope of this paper.
 
@@
 
-We carried out a thorough investigation of the strong lensing bias in photometrically selected samples of strong lenses.
-We first explored how the strong lensing cross-section varies as a function of lens and source properties, expanding upon the work of \citet{MVK09}.
+Strong lensing is a very active line of research, but a comprehensive understanding of the selection effects associated with it has so far been lacking.
+This work takes a major step towards filling that knowledge gap.
+%After a thorough investigation of the strong lensing bias in photometrically selected samples of strong lenses, we learned the following important lessons. 
+%After a thorough investigation of the strong lensing bias, we learned the following important lessons that apply to photometrically selected samples of strong lenses.
+After a thorough investigation, we learned several lessons regarding the strong lensing bias in photometrically selected lens samples. The following are the most important ones.
+%The thorough investigation of the strong lensing bias that we carried out in this work allowed us to learn several important lessons.
+\begin{enumerate}
+\item The strong lensing cross-section increases primarily with increasing lens mass, with decreasing half-mass radius, and with increasing source brightness.
+However, it has little dependence on the size of the source, if this is smaller than the Einstein radius and if it is detectable in the absence of lensing.
+%source size when this is smaller than the Einstein radius.
+%Sources with a surface brightness that is too low to be detected in the absence of lensing are still undetected when strongly lensed. Similarly, sources that are detectable 
+\item
+
+\end{enumerate}
+
+%We carried out a thorough investigation of the strong lensing bias in photometrically selected samples of strong lenses.
+%We first explored how the strong lensing cross-section varies as a function of lens and source properties, expanding upon the work of \citet{MVK09}.
 % by considering both point-like and extended sources.
-We introduced a new definition for a strong lensing event in the case of an extended source, for this purpose.
-%We also introduced a new definition for a strong lensing event, in the case of an extended source.
-This definition, which is based on the number of detected surface brightness peaks, is robust with respect to the mass-sheet degeneracy, unlike all magnification-based definitions.
-We found that the strong lensing cross-section increases primarily with increasing lens mass, with decreasing half-mass radius, and with decreasing source magnitude.
+%We introduced a new definition for a strong lensing event in the case of an extended source, for this purpose.
+%This definition, which is based on the number of detected surface brightness peaks, is robust with respect to the mass-sheet degeneracy, unlike all magnification-based definitions.
+%We found that the strong lensing cross-section increases primarily with increasing lens mass, with decreasing half-mass radius, and with decreasing source magnitude.
 

### 2022-12-30 paper/seleff.tex
rewritten conclusions

@@
 By capturing this dependence in the model of \Eref{eq:dmscaling}, the residual scatter in $\mdmfive$ around the mean is reduced, and so is the strong lensing bias.
+This, however, is a minor effect: we repeated the analysis while setting the dependence on size to zero and found minimal differences on the derived values of $\mu_{\mathrm{DM},0}$.
 %
@@
 
-%Although the amplitude of the strong lensing bias in difficult-to-observe mass-related quantities is relatively small, compared to current uncertainties, there might be situations in which one wishes to account for it.
-%The formally correct way to do this is by explicitly modelling all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
-%When the entire lens selection procedure is fully automated, then it is possible to use machine-learning to make this problem computationally tractable \citep{Leg++22}. 
-%If human selection is involved, however, it can be a difficult problem to solve in practice.
-%Therefore, it can be useful to develop alternative strategies to correct for the bias, or to mitigate its effects.
-
 As we discussed throughout this paper, the amplitude of the strong lensing bias depends on the intrinsic scatter in the mass parameters of the foreground galaxy population.
 One way to minimise the bias, therefore, is to identify scaling relations between observable quantities and mass-related properties that can account for part of the scatter.
-Describing the inner dark matter distribution as a function of stellar mass and half-light radius, as we did in section \ref{ssec:galbias}, is the first step in this direction: the bias in $\mdmfive$ at fixed $\mstar$ and $\reff$ is smaller than the overall shift in the median $\mdmfive$ of the lens population.
+Describing the inner dark matter distribution as a function of stellar mass and half-light radius, as we did in section \ref{ssec:galbias}, is the first step in this direction: the bias in $\mdmfive$ at fixed $\mstar$ and $\reff$ is smaller than the overall shift in the median $\mdmfive$ of the lens population (albeit by a marginal amount, as we pointed out earlier).
 This description could be extended by including the central velocity dispersion as an additional control parameter.
@@
 The central velocity dispersion, however, is very sensitive to the orbital anisotropy and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
-%intrinsic scatter in $\asps$ at fixed velocity dispersion should be smaller than %and therefore the intrinsic scatter in $\asps$ and $\mdmfive$ at fixed velocity dispersion should be smaller.
-
-%One problem with this approach is that the amplitude of the scatter in quantities such as $\asps$ or the dark matter parameters is not known before carrying out a strong lensing analysis, and neither is the resulting strong lensing bias.
-%An estimate of the bias, however, could be obtained after constraining the lens population distribution, by creating a simulation of the survey under the assumption that the inferred distribution is a good approximation of the truth.
-This would allow to posterior predict the effects of strong lensing bias.
-One could then adjust the parameters of the simulated galaxy population until the posterior predicted lens population matches the observed one.
 
@@
 This can be difficult to do if human selection is involved in the definition of the lens sample.
-In that case, a possible alternative is to approximate the lens finding probability $\pfind$. %, which is the component of $\psel$ that is difficult to simulate. 
+In that case, a possible alternative is to approximate the lens finding probability $\pfind$. 
 For example, we can make the assumption that $\pfind$ depends purely on the Einstein radius.
 This is a reasonable assumption, as long as the lens finding procedure does not selectively pick lenses with different image configurations depending on their Einstein radius.
-%It would allow us to marginalise over the source position and surface brightness parameters, thus simplifying the problem.
 The dependence of $\pfind$ on $\tein$ could then be described empirically and inferred during the analysis.
 This is essentially the approach adopted by \citet{Son++19} in the analysis of strong and weak lensing data from the Hyper Suprime-Cam Survey.
-
-%Although the amplitude of the strong lensing bias in difficult-to-observe mass-related quantities is relatively small, compared to current uncertainties, there might be situations in which one wishes to account for it.
-%The formally correct way to do this is by explicitly modelling all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
-%When the entire lens selection procedure is fully automated, then it is possible to use machine-learning to make this problem computationally tractable \citep{Leg++22}. 
-%If human selection is involved, however, it can be a difficult problem to solve in practice.
-%Therefore, it can be useful to develop alternative strategies to correct for the bias, or to mitigate its effects.
-
-%In general, $\pdet$ depends nontrivially on the position, redshift and surface brightness distribution of the source, as well as on the lens mass density profile and redshift.
-%This requires in principle to explicitly model the distribution of all of these quantities \citep[as done by][ in a simplified case]{Son22}.
-%However, we can make the assumption that $\pdet$ can be separable in a term that depends purely on the strong lensing cross-section and a term that depends purely on the Einstein radius.
-%This is a reasonable assumption, as long as the lens finding procedure does not selectively pick lenses with different image configurations depending on their Einstein radius.
-%It would allow us to marginalise over the source position and surface brightness parameters, thus simplifying the problem.
-%The dependence of $\pdet$ on $\tein$ could then be described empirically and inferred during the analysis.
-%This is essentially the same approach adopted by \citet{Son++19} in the analysis of strong and weak lensing data from the Hyper Suprime-Cam Survey.
 
@@
 This, in turn, would reduce the amplitude of the strong lensing bias and introduce a bias on the halo concentration.
-%Allowing for additional flexibility at fixed halo mass scatter parameter $\sigma_h$ results in a larger scatter in the inner dark matter distribution, which would increase the amplitude of the lensing bias.
-However, the value of $\sigma_h$ that we chose for our fiducial experiment was tuned on the basis of the predicted scatter in velocity dispersion around the fundamental plane.
-That scatter would also be increased by adding flexibility to the dark matter density profile.
-Then, in order to be consistent with the analysis carried out in our paper, the value of $\sigma_h$ would need to be lowered accordingly.
-This, in turn, would reduce the amplitude of the strong lensing bias.
 Since lensing and dynamics are sensitive to the mass distribution at comparable scales (the half-light radius and Einstein radius are similar for most of the lenses), we expect these two effects to cancel out to first approximation.
 The lensing bias on quantities directly related to the inner mass distribution, such as $\asps$ and $\mdmfive$, would be left roughly unchanged.
-%leaving the strong lensing bias on roughly unchanged.
 The main effect of adding scatter to the concentration is to reduce the correlation between the total halo mass and the dark matter mass enclosed within the Einstein radius.
@@
 \item The strong lensing cross-section increases primarily with increasing lens mass, with decreasing half-mass radius, and with increasing source brightness. 
-However, it has little dependence on the size of the source, if this is smaller than the Einstein radius and if it is detectable in the absence of lensing.
-%source size when this is smaller than the Einstein radius.
-%Sources with a surface brightness that is too low to be detected in the absence of lensing are still undetected when strongly lensed. Similarly, sources that are detectable 
+At fixed stellar distribution and fixed dark matter mass enclosed within a given aperture, varying the inner dark matter slope has little impact on the strong lensing cross-section (as long as the aperture within which the dark matter mass is normalised is comparable to the Einstein radius).
+\item The strong lensing cross-section has little dependence on the size of the source, if this is smaller than the Einstein radius and if it is detectable in the absence of lensing.
+Sources with a surface brightness that is too low to be detected are still undetected when strongly lensed. 
+%\item At fixed stellar distribution and fixed dark matter mass enclosed within a given physical aperture, varying the inner dark matter slope has little impact on the strong lensing cross-section.
+\item Lens galaxies tend to be more massive and more compact than their non-lens counterparts. Their redshift distribution is also modified with respect to the general galaxy population.
+At fixed observed stellar mass (i.e.
+inferred by means of stellar population synthesis), lens galaxies have a larger intrinsic stellar mass (i.e.
+a larger stellar population synthesis mismatch parameter $\asps$) and a larger dark matter halo mass.
+At fixed stellar mass and size, lens galaxies are still biased towards a larger dark matter content.
+\item The amplitude of the strong lensing bias depends on how broad is the distribution of the parameters describing the lens population. 
+Important quantities in this regard are the intrinsic scatter in the stellar population synthesis mismatch parameter, $\sigma_{\mathrm{sps}}$, and in the dark matter halo mass, $\sigma_{\mathrm{h}}$.
+Increasing the values of $\sigma_{\mathrm{sps}}$ and $\sigma_\mathrm{h}$ results in a stronger bias on $\asps$ and on the dark matter mass, and a weaker bias on the observed stellar mass and half-light radius.
+This implies that, in principle, we could constrain the intrinsic scatter parameters, which are currently poorly known, by measuring the amplitude of the strong lensing bias on $\mobs$ and $\reff$, which is easily observable.
+\item The strong lensing bias varies depending on the completeness of the lens sample as a function of Einstein radius.
+Surveys that can discover lenses with a smaller Einstein radius have a smaller associated strong lensing bias in all quantities.
 \item 
-
-\end{enumerate}
-
-%We carried out a thorough investigation of the strong lensing bias in photometrically selected samples of strong lenses.
-%We first explored how the strong lensing cross-section varies as a function of lens and source properties, expanding upon the work of \citet{MVK09}.
-% by considering both point-like and extended sources.
-%We introduced a new definition for a strong lensing event in the case of an extended source, for this purpose.
-%This definition, which is based on the number of detected surface brightness peaks, is robust with respect to the mass-sheet degeneracy, unlike all magnification-based definitions.
-%We found that the strong lensing cross-section increases primarily with increasing lens mass, with decreasing half-mass radius, and with decreasing source magnitude.
-
-We then simulated photometric strong lensing surveys with data similar to Euclid Wide, and measured how various properties of the lens galaxies and background sources are biased with respect to their parent distributions.
-We focused primarily on mass-related quantities: most importantly, the stellar population synthesis mismatch parameter $\asps$ and the dark matter distribution.
-We found that strong lenses are biased towards larger values of $\asps$ and of the halo mass $\mhalo$.
-The amplitude of the bias is larger in simulations with a larger intrinsic scatter in the distribution of $\asps$ and $\mhalo$, and increases if lenses with a preferentially large Einstein radius are selected.
-
-We used observational constraints from the fundamental plane to set the values of the intrinsic scatter in our fiducial simulation.
-In that case, assuming a survey that is complete down to $\tein=0.5''$, the bias on $\asps$ is smaller than $0.04$~dex (10\%).
+%We used observational constraints from the fundamental plane to set the values of the intrinsic scatter in our fiducial simulation.
+Under reasonable assumptions on the intrinsic scatter parameters, for a Euclid-like survey that is complete down to $\tein=0.5''$ the bias on $\asps$ is smaller than $0.04$~dex (10\%).
 This bias is much smaller than the current systematic uncertainty on the stellar population synthesis-based stellar masses.
@@
 Under the same assumptions, the strong lensing bias on the average halo mass at fixed stellar mass is $0.07$~dex, while that on the inner dark matter distribution at fixed stellar mass and size is $0.02$~dex.
-
-Strong lensing selection broadens the magnitude distribution of background sources, compared to the population of objects that are detectable without lensing.
+%
+\item Strong lensing selection broadens the magnitude distribution of background sources, compared to the population of objects that are detectable without lensing.
 At the same time, we did not find any evidence for a bias in the size distribution of background sources, at fixed magnitude.
-%This follows from the fact that our selection criterion is based on the surface brightness
-
-We also ran simulations with lensed quasars in place of extended sources, and found that the amplitude of the strong lensing bias in the lens-related parameters has very little sensitivity of the details of the source population.
+
+\item Simulations with lensed quasars in place of extended sources showed that the amplitude of the strong lensing bias in the lens-related parameters has very little sensitivity on the details of the source population.
 This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
-Samples of quad lenses, however, are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is on average different from that of the entire sample of lenses.
+\item Samples of quad lenses are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is also biased.
 This means that particular care must be taken when stellar dynamics measurements obtained on galaxy-galaxy lenses is used to inform the properties of quads.
+\end{enumerate}
 
@@
 Biases that affect observable properties, such as the redshift and the light distribution of the lens, can be easily quantified.
-Biases on mass-related quantities, such as the stellar mass-to-light ratio or the dark matter distribution, are more difficult to measure directly, but can be reduced by modelling how they scale with the observable properties of the lenses, thus minimising their intrinsic scatter.
-
-In the future, it might be desirable to directly model strong lensing selection effects, thus accounting for the strong lensing bias.
+Biases on mass-related quantities, such as the stellar mass-to-light ratio or the dark matter distribution, are more difficult to measure directly and must be modelled by taking into account selection effects.
+%, but can be reduced by modelling how they scale with the observable properties of the lenses, thus minimising their intrinsic scatter.
+%In the future, it might be desirable to directly model strong lensing selection effects, thus accounting for the strong lensing bias.
 Designing strong lensing surveys with clearly-defined and easily-modellable selection criteria would help greatly in this task.
+A peak-based definition of a strong lensing event such as that introduced in section \ref{ssec:lensdefext} could be used for this purpose.
 

### 2023-01-12 paper/seleff.tex
going back to A&A cause it's cheaper

@@
 
-   \author{Alessandro Sonnenfeld\inst{1}\and
-           Others
-          }
-
-   \institute{Leiden Observatory, Leiden University, Niels Bohrweg 2, 2333 CA Leiden, the Netherlands\\
-              \email{sonnenfeld@strw.leidenuniv.nl}
+   \author{Alessandro Sonnenfeld\inst{\ref{sjtu},\ref{leiden}}\and
+           Shun-Sheng Li\inst{\ref{leiden}}\and
+           Giulia Despali\inst{\ref{ita}}\and
+           Anowar Shajib\inst{\ref{chicago},\ref{kicp}}\and
+           Edward N. Taylor\inst{\ref{swin}}
+          }
+
+   \institute{
+Department of Astronomy, School of Physics and Astronomy, Shanghai Jiao Tong University, Shanghai 200240, China\\
+              \email{sonnenfeld@sjtu.edu.cn}\label{sjtu}
+\and
+Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\label{leiden}\and
+Institut für Theoretische Astrophysik, Zentrum für Astronomie, Heidelberg Universität, Albert-Ueberle-Str. 2, 69120, Heidelberg, Germany\label{ita}\and
+Department of Astronomy \& Astrophysics, University of Chicago,
+Chicago, IL 60637, USA\label{chicago}\and
+Kavli Institute for Cosmological Physics, University of Chicago,
+Chicago, IL 60637, USA\label{kicp}\and
+Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn 3122, Australia\label{swin}\and
              }
@@
 Low scatter & 0.05 & 0.10 & 0.021 \\
-High scatter & 0.10 & 0.30 & 0.433 \\
+High scatter & 0.10 & 0.30 & 0.043 \\
 \end{tabular}
@@
 In order for either of these approaches to work, however, it is essential that the lens selection procedure can be simulated.
-This can be difficult to do if human selection is involved in the definition of the lens sample.
+This, in turn, requires having an objective definition of a strong lensing event. A peak-based definition such as that introduced in section \ref{ssec:lensdefext} could be used for this purpose.
+
+Nevertheless, it can still be difficult to fully forward model strong lensing selection if visual inspection by humans is involved in the definition of the lens sample.
 In that case, a possible alternative is to approximate the lens finding probability $\pfind$. 
@@
 
-The simulations on which our analysis is based as as complex as required by the goal of the analysis itself, which is to estimate the amplitude of the strong lensing bias in a few key quantities.
+The simulations on which our analysis is based are as complex as required by the goal of the analysis itself, which is to estimate the amplitude of the strong lensing bias in a few key quantities.
 We did, however, make some simplifying assumptions.
@@
 Designing strong lensing surveys with clearly-defined and easily-modellable selection criteria would help greatly in this task.
-A peak-based definition of a strong lensing event such as that introduced in section \ref{ssec:lensdefext} could be used for this purpose.
+%A peak-based definition of a strong lensing event such as that introduced in section \ref{ssec:lensdefext} could be used for this purpose.
 

### 2023-01-30 single_lens/elliptical_ext/get_largesource_crosssect.py
making it strong_lensing_tools-friendly

@@ -1,9 +1,4 @@
 import numpy as np
-from astropy.io import fits as pyfits
-from skimage import measure
-from scipy.special import gamma as gfunc
-from scipy.signal import convolve2d
-from scipy.optimize import brentq
 from sl_profiles import sersic
 from lensdet import detect_lens
 import h5py

### 2023-01-30 single_lens/elliptical_ext/plot_example_detections.py

@@ -1,7 +1,6 @@
 import numpy as np
 import glafic
 from sl_profiles import sersic
-from scipy.special import gamma as gfunc
 from scipy.signal import convolve2d
 import pylab
 from astropy.io import fits as pyfits

### 2023-01-30 single_lens/elliptical_ext/plot_largesource_cs.py

@@ -1,9 +1,6 @@
 import numpy as np
 import pylab
 import h5py
-from astropy.io import fits as pyfits
-from scipy.interpolate import splrep, splev
-from plotters import probcontour
 from labellines import labelLine, labelLines
 from matplotlib.ticker import MultipleLocator
 from matplotlib import rc

### 2023-01-30 paper/seleff.tex
minor changes

@@
 \usepackage{hyperref}
+\usepackage{siunitx}
 % To add links in your PDF file, use the package "hyperref"
@@
            Giulia Despali\inst{\ref{ita}}\and
-           Anowar Shajib\inst{\ref{chicago},\ref{kicp}}\and
+           Anowar Shajib\inst{\ref{chicago},\ref{kicp},\ref{einstein}}\and
            Edward N. Taylor\inst{\ref{swin}}
@@
 Chicago, IL 60637, USA\label{kicp}\and
+NHFP Einstein Fellow\label{einstein}\and
 Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn 3122, Australia\label{swin}\and
@@
 
-The Python code used for the simulation and analysis of the lens sample can be found in a dedicated section of a GitHub repository\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools}}.
+The Python code used for the simulation and analysis of the lens sample can be found in a dedicated section of a GitHub repository\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/papers/selection_effects}}.
 
@@
 A very important quantity for describing the strength of a strong lens is the Einstein radius, $\tein$.
-For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:lenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
+For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:1dlenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
 The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form here have infinite magnification in the tangential direction.
@@
 Lines of different colour correspond to sources of different intrinsic magnitude.
+The dashed blue line, corresponding to the brightest source magnitude, overlaps completely with the dashed orange line.
+The dashed purple line is zero: very faint sources cannot produce any detectable quads.
 %The lens consists of a galaxy at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
@@
 Moreover, as we discussed in section \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
-For this reason, we split our analysis in two parts: first we explore the small source size regime, then we consider cases in which the source size is comparable to or bigger than the lens caustics.
+For this reason, we split our analysis into two parts.
+First, we explore the small source size regime, where the source size is comparable to or smaller than the lens caustics.
+Then, we consider cases in which the source size is bigger than the lens caustics, which we refer to as the large source regime.
 
@@
 Strong lensing cross-section of an extended source.
-Small source size regime.
+Small source size regime ($\theta_{\mathrm{e,s}} < \tein$).
 The cross-section is plotted as a function of the ratio between the half-light radius of the source and the Einstein radius of the lens.
@@
 \caption{
-Strong lensing cross-section of an extended source. Large source size regime.
+Strong lensing cross-section of an extended source. Large source size regime ($\theta_{\mathrm{e,s}} \gtrsim \tein$).
 The cross-section is plotted as a function of the ratio between the Einstein radius of the lens and the half-light radius of the source.
@@
 We set the mean of $\log{\asps}$ to $0.1$, as this is an intermediate value among estimates of $\asps$ from the literature \citep{CvD12, Cap++13, SLC15, Son++15, Son++19}.
-We adopted a few different values for the scatter $\sigma_\alpha$. We explain in \ref{ssub:scat} how these were chosen.
+We adopted a few different values for the scatter $\sigma_\alpha$. We explain in Section \ref{ssub:scat} how these were chosen.
 
@@
 The galaxy morphology is described by a S\'ersic profile with three parameters: the half-light radius in angular units $\theta_{
-mathrm{e,s}}$, the S\'ersic index $n_\mathrm{s}$, and the axis ratio $q_\mathrm{s}$.
+\mathrm{e,s}}$, the S\'ersic index $n_\mathrm{s}$, and the axis ratio $q_\mathrm{s}$.
 These structural parameters are learned from the imaging data obtained with the Advanced Camera for Surveys (ACS) instrument on the \textit{Hubble} Space Telescope~\citep{Gri++12}.
@@
 For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering $1000$ square degrees.
-The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around $300000$.
+The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around \num{300000}.
 %Then we generated a population of sources behind the lenses.
@@
 
-For the simulation with extended sources, we produced images with properties similar to those expected for the Euclid Wide survey 
-citep{Sca++22}.
+For the simulation with extended sources, we produced images with properties similar to those expected for the Euclid Wide survey \citep{Sca++22}.
 We used a pixel size of $0.1''$ and we applied a Moffat PSF with an FWHM of $0.2''$ and a $\beta$ parameter of $5.0$.
@@
 For the sake of conciseness, we focus instead on the quantities that we consider most important.
-Nevertheless, the output of our simulations is available online\footnote{\url{https://github.com/astrosonnen/sl_selection_effects}}.
+Nevertheless, the output of our simulations is available online\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/papers/selection_effects}}.
 We encourage readers who are interested in studying aspects of the strong lensing bias that are not covered in this section to download our data and analyse them directly.
@@
 
-In order for a galaxy-source pair to be included as a strong lens in a survey, three condition must be met.
+In order for a galaxy-source pair to be included as a strong lens in a survey, three conditions must be met.
 First of all, at least part of the source must be multiply imaged.
@@
 This description could be extended by including the central velocity dispersion as an additional control parameter. 
+%Galaxies selected in velocity dispersion are similar to existing samples of lenses in terms of other observable quantities
 The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution marginalised over the whole population.
-The central velocity dispersion, however, is very sensitive to the orbital anisotropy and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
+The central velocity dispersion, however, is very sensitive to the orbital anisotropy, to the three-dimensional structure, and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
 
@@
 Typical values of the external shear and external convergence are $|\gamma|<0.1$ and $|\kappa_{\mathrm{ext}}|<0.1$ \citep{Mil++20}: these values are small compared to the typical ellipticities and dark matter fractions of our simulated lenses.
-Therefore, while including them might modify slightly the amplitude of the strong lensing bias in the axis ratio and dark matter distribution parameters, the conclusions our our analysis would not be affected.
+Therefore, while including them might modify slightly the amplitude of the strong lensing bias in the axis ratio and dark matter distribution parameters, the conclusions of our analysis would not be affected.
 
@@
 The collaboration leading to this work was initiated at the 2022 Lorentz Center workshop “Bridging gaps between dynamical probes of galaxies”.
+AS and SSL are supported by NOVA, the Netherlands Research School for Astronomy.
+Support for this work was provided by NASA through the NASA Hubble Fellowship grant HST-HF2-51492 awarded to AJS by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS5-26555.
 

### 2023-01-30 paper/seleff.tex
final edits

@@
            Giulia Despali\inst{\ref{ita}}\and
-           Anowar Shajib\inst{\ref{chicago},\ref{kicp},\ref{einstein}}\and
+           Anowar J. Shajib\inst{\ref{chicago},\ref{kicp},\ref{einstein}}\and
            Edward N. Taylor\inst{\ref{swin}}
@@
 NHFP Einstein Fellow\label{einstein}\and
-Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn 3122, Australia\label{swin}\and
+Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn 3122, Australia\label{swin}
              }
@@
 This last point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
-Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{B+T21}, but any difference between the two lens classes can in principle introduce biases, if not corrected for.
+Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{B+T21}, but any difference between the two lens classes can introduce biases, if not corrected for.
 With this study we aim to quantify this bias.
@@
 For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:1dlenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
-The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form here have infinite magnification in the tangential direction.
+The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form there have infinite magnification in the tangential direction.
 %
@@
 The outer curve is the radial caustic, while the inner diamond is the tangential caustic.
-Sources located outside the caustic are not strongly lensed.
+Point sources located outside the caustic are not strongly lensed.
 Sources that lie in the region enclosed between the two caustics produce three images, while sources inside the tangential caustic produce five images.
@@
 \caption{
-Criterion used to classify lensed images of extended sources. Four examples.
+Criterion used to classify lensed images of extended sources. Six examples.
 First column: caustics (red curves) and source position (blue circle).
@@
 
-In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the source surface brightness distribution, with its radial profile, shape and orientation, and the point spread function (PSF).
+In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the source surface brightness distribution, with its radial profile, shape and orientation, and the PSF.
 Moreover, as we discussed in section \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
@@
 We simulated a large number of images of extended sources with {\sc Glafic} and measured the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
-For the small source size experiment, we used pixels with a size equal to $1/20\tein$ and convolved the images with a Moffat PSF with a FWHM of two pixels.
+For the small source size experiment, we used pixels with a size equal to $1/20\tein$ and convolved the images with a Moffat PSF with a FWHM of two pixels and a $\beta$ parameter of $5.0$.
 We also assumed that the background noise is an uncorrelated Gaussian field.
@@
 
-Our background galaxy population is taken from the \textsc{surfs}-based KiDS-Legacy-Like Simulation (SKiLLS) input catalogue, a hybrid simulation catalogue integrating cosmological simulation with high-quality imaging observations (Li et al., in prep).
+Our background galaxy population is taken from the \textsc{surfs}-based KiDS-Legacy-Like Simulation (SKiLLS) input catalogue, a hybrid simulation catalogue integrating cosmological simulation with high-quality imaging observations \citep{Li++23}.
 The cosmological simulation is obtained from the Synthetic UniveRses For Surveys (\textsc{surfs}) simulations, a set of $N$-body simulations from \citet{Ela++18}.
@@
 The original photometry is drawn from a stellar population synthesis technique using stellar synthesis libraries with physically motivated dust attenuation and re-emission models~\citep{Rob++20}.
-Li et al.
-(in prep) further applied an empirical correction to the original synthetic photometry to better agree with the COSMOS2015 observations~\citep{Lai++16}.
+\citet{Li++23} further applied an empirical correction to the original synthetic photometry to better agree with the COSMOS2015 observations~\citep{Lai++16}.
 The galaxy morphology is described by a S\'ersic profile with three parameters: the half-light radius in angular units $\theta_{
@@
 These structural parameters are learned from the imaging data obtained with the Advanced Camera for Surveys (ACS) instrument on the \textit{Hubble} Space Telescope~\citep{Gri++12}.
-We refer to Li et al.
-(in prep) for details on the learning algorithm and validation.
+We refer to \citet{Li++23} for details on the learning algorithm and validation.
 
@@
 Finally, we randomly placed quasars in the source plane with a projected number density of $70$~arcmin$^{-2}$.
-This is a much larger number density than observed in the real universe.
-Nevertheless, we are not interested in predicting the absolute number of strong lenses, so this is a legitimate choice.
+This is a much larger number density than observed in the real universe, but we are allowed to do so because we are not interested in predicting the absolute number of strong lenses.
 The advantage of boosting the number density of quasars is that it allows us to produce a large number of lenses without the need for generating too big a population of foreground galaxies.
@@
 
-For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering $1000$ square degrees.
+For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering \num{1000} square degrees.
 The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around \num{300000}.
@@
 When working with large samples of lenses, obtaining spectroscopic measurements is not a viable option, and photometric redshifts are a necessity.
-Measuring the photometric redshift of a strongly lensed source, however, is challenging, especially when the Einstein radius is small and the source light is blended with the light from the lens \citep{Lan++22}.
+Measuring the photometric redshift of a strongly lensed source, however, is challenging, especially when the Einstein radius is small and the source light is blended with the light from the lens \citep{Lan++23}.
 
 Both of these scenarios can result in samples that are incomplete below a certain value of the Einstein radius. %in which lenses with a larger Einstein radius are preferentially selected.
-We simulate these situation via the following Einstein radius-dependent lens finding probability:
+We simulate this situation via the following Einstein radius-dependent lens finding probability:
 \begin{equation}
@@
 \Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the parent population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
-A completeness limit of $0.5''$ is close to what can currently be achieved via visual inspection of high-resolution space-based images \cite{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
+A completeness limit of $0.5''$ is close to what can currently be achieved via visual inspection of high-resolution space-based images \citep{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
 \begin{figure*}
@@
 Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
-Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
+Red dashed lines: distribution of the lenses with Einstein radius larger than $1.0''$.
 \label{fig:lenspars}
@@
 Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
-Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
+Red dashed lines: distribution of the lenses with Einstein radius larger than $1.0''$.
 \label{fig:sourcepars}
@@
 \end{equation}
-and taking the resulting value of $\mu_{\gamma}$.
+and taking the resulting value of $\mu_{\gamma,0}$.
 Different simulations show different trends in the average $\gammadm$.
@@
 First panel: median $\log{\asps}$.
-Second panel: median $\mhalo$ at $\log{\mstar}=11.5$.
-Third panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+Second panel: median $\mhalo$ at $\log{\mobs}=11.4$.
+Third panel: median $\mdmfive$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
 Fourth panel: median lens axis ratio.
@@
 The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution marginalised over the whole population.
-The central velocity dispersion, however, is very sensitive to the orbital anisotropy, to the three-dimensional structure, and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
+The central velocity dispersion, however, is very sensitive to the orbital anisotropy, to the three-dimensional structure, and to gradients in stellar mass-to-light ratio, which are not well known.
+Therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
 

### 2023-01-30 lens_population/make_lens_cornerplot.py

@@ -49,7 +49,7 @@ major_step = [0.2, 0.5, 0.1, 0.5, 0.5, 2, 0.2]
 minor_step = [0.05, 0.1, 0.02, 0.1, 0.1, 0.5, 0.05]
 
 fig = pylab.figure()
-pylab.subplots_adjust(left=0.08, right=0.99, bottom=0.12, top=0.99, hspace=0.1, wspace=0.1)
+pylab.subplots_adjust(left=0.10, right=0.99, bottom=0.12, top=0.99, hspace=0.1, wspace=0.1)
 #pylab.figtext(0.45, 0.95, 'Extended model', fontsize=fsize+3, backgroundcolor=(0., 1., 0.))
 
 for i in range(npars):

### 2023-01-30 single_lens/elliptical_ext/plot_cs_vs_re.py

@@ -17,7 +17,7 @@ ylim = (0.085, 1.5)
 logscale = True
 
 if logscale:
-    leftm = 0.14
+    leftm = 0.15
 else:
     leftm = 0.08
 

### 2023-01-30 paper/seleff.tex
final edits

@@
 
-%The luminosity and size distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies.
-%For instance, lensing magnification allows the detection of fainter objects with respect to the field, and, in case lenses are selected on the basis of the magnification of the source, galaxies that are very extended with respect to the Einstein radius are less likely to be classified as strongly lensed.
-This is because their magnification averaged over their size tends to be low \citep[see the discussion in section 5.3 of][]{O+A17}.
-
 In general, the probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
@@
 In principle, this can be done with a Bayesian hierarchical formalism \citep{Son22}, but knowledge of the lens selection probability $\psel$ is required.
-%For most of the existing strong lensing surveys, however, characterising $\psel$ is very challenging.
-%: the process of determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
 This factor can be written as the following product:
 \begin{equation}\label{eq:two}
-\psel(\psilens,\psisource|S) = \pdet(\psilens,\psisource)\pfind(\psilens,\psisource|S),%\footnote{\citet{Son22} implicitly assumed $\pfind\equiv1$, therefore $\pdet$ and $\psel$ can be used interchangeably in the context of that work},
+\psel(\psilens,\psisource|S) = \pdet(\psilens,\psisource)\pfind(\psilens,\psisource|S),
 \end{equation}
@@
 Few studies have attempted to explicitly account for strong lensing selection effects, usually by making ad-hoc simplifying assumptions \citep{Son++15,O+A17,Son++19}.
-%The problem of the invertibility of \Eref{eq:one} can be simplified by adopting lens selection criteria that can be easily forward-modelled: for instance by building samples that are complete above a well-defined observational cut \citep{Son22}.
-%The problem can be simplified by adopting lens selection criteria that can be easily forward-modelled: for instance by building samples that are complete above a well-defined observational cut \citep{Son22}.
 Whether it is necessary to invert \Eref{eq:one}, however, depends on the severity of the strong lensing bias that needs to be corrected and on the accuracy requirements on the key quantities of interest.
@@
 In this paper we aim to quantify the strength of the strong lensing bias on a series of foreground galaxy and background source parameters. 
-%, under realistic assumptions for the lens detection probability $\pdet$.
 In particular, we aim to determine how strong lenses differ from the parent population of foreground galaxies and background sources in terms of a) the radial mass structure of the lenses (i.e.
@@
 In addition, we addressed points 2) and 3) as well: we simulated large populations of strong lenses using empirical models, we simulated the lens detection and finding phase, and quantified the lensing bias under various scenarios.
-%We explore how the results change as a function of the scatter in mass parameters at fixed light, of the efficiency of finding lenses with small image separation, and of source type: we simulate both galaxy and quasar sources.
-%We explore how the results change as a function of the scatter in mass parameters at fixed light, and as a function of the efficiency of a lens survey at discovering small image separation lenses.
 We explored how the results change as a function of the efficiency of a lens survey at discovering small image separation lenses, and of the scatter in mass parameters at fixed light.
 
-%In light of the results of our experiments, we discuss to what extent can studies of galaxy structure based on strong lenses be generalised to describe non-lens galaxies.
 Finally, we addressed the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses.
@@
 In \Sref{sect:basics} we introduce the basics of gravitational lensing.
-In \Sref{sect:indlenses} we study individual lens systems. %, to look for trends in the lensing cross-section with various lens and source properties.
+In \Sref{sect:indlenses} we study individual lens systems. 
 In \Sref{sect:lenspop} we describe our simulations of lens surveys.
@@
 \end{equation}
-%which is the ratio between the projected surface mass density of the lens and the critical surface mass density for lensing.
 The latter quantity is defined as
-%The quantity $\Sigma_{\matrhrm{cs}}$ is the critical surface mass density for lensing, defined as
 \begin{equation}
@@
 and the expression for the deflection angle reduces to
-%In the special case of an axisymmetric lens, the deflection angle, calculated along an arbitrary axis passing through the lens centre, reduces to
 \begin{equation}
@@
 \end{equation}
-%Based on \Eref{eq:criteq},
-%we generalise the definition of $\tein$ to the non-axisymmetric case as the circularised radius of the isodensity contour that encloses an average surface mass density equal to $\Sigma_{\mathrm{cr}}$.
 
 Axisymmetric lenses of the kind considered in this work 
-%Axisymmetric lenses with a density profile of the kind considered in section \ref{ssec:profile} 
 typically produce either one or three images of a point source.
 \Fref{fig:scheme} helps to visualise this property.
-%\Fref{fig:scheme}, which shows the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, helps to visualise this property. 
 Plotted in \Fref{fig:scheme} is the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, for a few lens models.
-%for the same lens models considered in \Fref{fig:kappa}.
 According to \Eref{eq:1dlenseq}, images form at the locations where this quantity equals the position of the source. 
@@
 As can be seen from \Fref{fig:scheme}, a source at this location is mapped to the stationary point of the $\theta - \alpha(\theta)$ curve. 
-%this value corresponds to the stationary point of the $\theta - \alpha(\theta)$ curve. 
 That location on the image plane is known as the radial critical curve. 
@@
 \end{figure}
-%The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the magnification of an image in the radial direction. 
-%This can be understood by noting that, according to \Eref{eq:1dlenseq}, the derivative of $\theta - \alpha(\theta)$ is equal to $d\beta/d\theta$, which quantifies the ratio between an infinitesimal displacement in the radial direction in the source plane and the corresponding displacement in the image plane.
-%Stationary points of the $\theta - \alpha(\theta)$ curve, therefore, are points of infinite magnification in the radial direction.
-For a given axisymmetric lens, these points describe a circle known as the radial critical curve.
-%The corresponding curve in the source plane, the radius of which is given by the value of $\theta - \alpha(\theta)$ at the stationary point, is known as the radial caustic.
-%As can be seen in \Fref{fig:scheme}, sources that lie within the radial caustic are strongly lensed into three images, while sources that lie outside of it produce only one image.
-
-%Strong lenses also have a tangential critical curve, where the magnification along the tangential direction is formally infinite.
-For an axisymmetric lens, the tangential critical curve is a circle with radius equal to the Einstein radius, and the corresponding tangential caustic collapses to the $\beta=0$ point in the source plane.
 
@@
 The outermost curves are radial caustics, while the inner ones are tangential ones.
-%While in the axisymmetric case these are a circle and a point respectively (blue curves), in the elliptical case they are transformed into a 
 The most striking difference with respect to the axisymmetric case (blue curves in \Fref{fig:ellcaust}) is the fact that the tangential caustic is transformed from a point into a diamond-like curve.
@@
 In samples of lenses defined via visual inspection this process is typically done implicitly, by identifying multiply imaged blobs that belong to the same source element.
-%Applying our lens definition criterion to a real lens requires showing that the observed number of surface brightness peaks can be reproduced with a lens model in which the source has a smaller number of peaks.
 
@@
 Our definition of a strong lensing event, instead, is robust with respect to the mass-sheet degeneracy.
-%Because of the mass-sheet degeneracy
 
@@
 
-%In this work we focus on massive elliptical galaxies as lenses. This is because this class of objects makes up the vast majority of known strong lenses.
-%The type of lenses that we focus on in this work are massive elliptical galaxies: these make up the vast majority of known strong lenses.
-%The lens galaxies that we focus on in this work are massive early-types: these make up the vast majority of known strong lenses.
 In this work we focus on massive early-type galaxies as lenses, as these make up the vast majority of known strong lenses.
@@
 We assume that the baryonic component consists entirely of stars, thereby neglecting gas, which is known to contribute very little to the mass of early-type galaxies in the inner regions that are probed by strong lensing.
-We then assume that the stars follow a S\'{e}rsic profile, with %with a fixed index of $n=4$ (also known as a de Vaucouleurs profile).
+We then assume that the stars follow a S\'{e}rsic profile, with 
 projected surface mass density given by
@@
 In section \ref{ssec:limitations} we discuss qualitatively what the implications of relaxing this assumption would be.
-%This assumption is probably broken in real galaxies but, as we discuss in \Sref{ssec:limitations}, does not impact our investigation.
-
-%Throughout this work we describe the mass distribution of the lens galaxies with a two-component model consisting of a S\'{e}rsic profile, describing the baryonic component of a galaxy, and a generalised Navarro, Frenk \& White (gNFW) profile, describing the dark matter halo.
-%We assume that the two components are concentric.
-%We also assume that the baryonic component consists entirely of stars.
-%This is a good approximation for the majority of the known strong lenses, which are massive elliptical galaxies with little gas.
-%Finally, we assume that the light distribution of the stellar component follows its mass distribution exactly.
-That is, we do not allow for the presence of gradients in the stellar mass-to-light ratio.
-This assumption is probably broken in real galaxies but, as we discuss in \Sref{sect:discuss}, does not impact our investigation.
-
-%The projected surface mass density of a S\'{e}rsic profile is given by
-%\begin{equation}
-%\Sigma(R) = \Sigma_0 \exp{\left\{-b\left(\frac{R}{\reff}\right)^{1/n}\right\}},
-%\end{equation}
-%with
-%\begin{equation}
-%\Sigma_0 = \frac{\mstar b^{2n}}{2\pi n \reff^2 \Gamma(2n)},
-%\end{equation}
-%where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is given by \citep{C+B99}
-%\begin{equation}
-%b(n) \approx 2n -\frac13 + \frac{4}{405n} + \frac{46}{25515n^2} + O(n^{-3}).
-%\end{equation}
-%Throughout this paper we indicate with $\teff$ the angular size of the half-light radius.
 
@@
 \Fref{fig:kappa} shows the dimensionless surface mass density profile of S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
-All of the profiles have %a S\'{e}rsic index $n=4$, 
+All of the profiles have 
 a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component (these assumptions are dropped later).
@@
 In order to compute the lensing cross-section, however, we must take into account the magnification, because that determines whether multiple images can be detected or not.
-%In order to compute the lensing cross-section, however, we must take into account the magnification of the multiple images: according to our definition of a strong lensing event, given in section \ref{ssec:lensdefpoint}, at least two images must be detected in order for a system to qualify as a strong lens.
 
@@
 Depending on the intrinsic brightness of the lensed source, these images may or may not be detected.
-%This has two implications for the lensing cross-section.
-First, a larger caustic size does not necessarily translate into a larger value of $\crosssect$.
-Second, the cross-section depends on the intrinsic brightness of the lensed source: an extremely bright point source can be detected regardless of its position in the source plane, but a source with a finite brightness can go undetected depending on its value of $\beta$.
 
@@
 The left panel shows the effect of varying the slope $\gammadm$ for fixed $\fdm$; conversely, the right panel shows variations as a function of $\fdm$ with $\gammadm$ fixed.
-%Left panel: cross-section, in units of $\pi\tein^2$, as a function of the inner dark matter slope. The dark matter fraction is fixed at $\fdm=0.5$.
-%Right panel: cross-section as a function of the fraction of dark matter within the half-light radius. The inner dark matter slope is fixed at $\gammadm=1.5$.
 The system is defined as a strong lens if at least two images are detected.
@@
 This is achieved by adjusting other properties of the lens, such as the total mass of the baryonic or the dark matter component (see \Fref{fig:kappa}). 
-%This means that other properties of the lens, such as the total mass of the baryonic component or the dark matter component, are adjusted to maintain a constant $\tein$ (see \Fref{fig:kappa}). 
 In practice, when varying one ingredient of the lens density profile, the Einstein radius varies in response.
@@
 \Fref{fig:phys_cs} shows $\crosssect$ as a function of stellar mass, half-light radius, inner dark matter slope, and projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$.
-%The reference lens is a galaxy at $z=0.3$ with $\log{\mstar}=11.5$, $\reff=7\,{\rm kpc}$, $\gammadm=1.5$, $\log{\mfive}=11.0$, and a source at $z=1.5$.
 
@@
 The dashed purple line is zero: very faint sources cannot produce any detectable quads.
-%The lens consists of a galaxy at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
 The parameters of the lens density profile are $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\teff=\tein$.
@@
 The horizontal dotted line indicates the cross-section for a very bright point source (i.e. the area enclosed within the caustics).
-%Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
 \label{fig:extcs}
@@
 \end{figure}
-
-%We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
-%We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_{\mathrm{e,s}}$.
-%The results are shown in \Fref{fig:extcs}.
-%The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius.
-We assumed that the background noise is an uncorrelated Gaussian field, for this purpose.
 
@@
 While perhaps surprising, this follows from the fact that gravitational lensing preserves surface brightness: a source that can be detected in the absence of lensing will produce images with the same surface brightness when lensed, which can be detected as well.
-%When the source is extended, however, 
 In order to classify a source as strongly lensed, however, we require that multiple images are observed. 
@@
 This lowers the signal-to-noise ratio of the multiple images compared to the point source case.
-%The result is the decrease in $\crosssect$ at values of the source size smaller than the 
 The result is that $\crosssect$ starts to decrease with increasing source size at values of $\theta_{\mathrm{e,s}}$ that are smaller than the no-lensing detection limit, as observed in \Fref{fig:extcs}.
 
-%At the brightest flux explored in the experiment (red line in \Fref{fig:extcs}), $\crosssect$ increases with increasing source size, before dropping to zero.
 At the brightest flux explored in the experiment (green line in \Fref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (black dotted line in \Fref{fig:extcs}).
@@
 
-%As we explained in \Sref{sect:intro}, a key ingredient determining the amplitude of the strong lensing bias is the scatter in the mass properties of the foreground galaxy population.
-%Although we do not have robust measurements of the intrinsic scatter in properties such as the stellar mass-to-light ratio or the dark matter halo mass, we set upper limits on their amplitude using the
-%Our simulations explore different amplitudes of the scatter. Scatter
-
 \subsection{Foreground galaxies}\label{ssec:lenses}
@@
 We chose to focus on early-type galaxies because many strong lensing surveys have preferentially targeted this class of objects in their lens-finding phase \citep[e.g.][]{Bol++06,Gav++12,Son++18a}.
-%This, in turn, was motivated by two points:
 This, in turn, had a dual motivation:
@@
 
-%We randomly assign a value of $\asps$ to each galaxy in the sample, by drawing it from the following distribution:
 For each galaxy in the sample, we randomly drew a value of $\log{\asps}$ from the following distribution:
@@
 We defined this as the mass enclosed within a spherical shell with average density equal to $200$ times the critical density of the Universe.
-%We randomly drew halo masses from the following distribution:
 Using current weak lensing constraints on the halo mass of elliptical galaxies as a reference \citep{Son++22}, we drew halo masses from the following distribution:
@@
 which is a Gaussian in $\log{\mhalo}$ with a mean that scales with stellar mass and scatter $\sigma_h$.
-%The values of the coefficients describing the mean of the distribution were chosen to be close to current weak lensing constraints on the halo mass of elliptical galaxies \citep{Son++22}.
-%The intrinsic scatter in halo mass is more uncertain from an observational point of view, therefore we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
 The intrinsic scatter in halo mass is not well constrained observationally, therefore we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
@@
 We assumed the dark matter distribution to be initially described by an NFW profile with a concentration\footnote{The concentration is the ratio between the halo virial radius and the scale radius $r_s$.} of five, then used the prescription of \citet{Cau++20} to model the response of the halo to the infall of baryons.
-% on the basis of the present stellar mass distribution.
 In the \citet{Cau++20} model the halo response is approximated with an analytical function that depends on the present stellar mass distribution, and typically results in a more concentrated and steeper density profile compared to the original NFW model.
@@
 In practice, as we explain in section \ref{ssec:limitations}, our main results are not affected by this choice.
-%However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope.
-Therefore, we applied a random scatter on $\gammadm$ of amplitude $\sigma_{\gamma}$.
-We explored different values for $\sigma_{\gamma}$, as explained in the next section.
 
@@
 Nevertheless, we use it in our experiment in order to obtain a more conservative upper limit on the amplitude of the strong lensing bias.
-%A model with the same intrinsic scatter parameters and a more realistic 3-dimensional and orbital structure would likely predict a fundamental plane scatter that is at least as large as the observed one.
 \begin{table}
@@
 arc-like features physically associated with the lens galaxy) in the lens finding phase.
-%This selection is meant to simulate a cut on the redshift (possibly photometric) of the arcs, which would help reduce the incidence of false positives during the lens finding phase.
 The resulting number density of sources is $70$~arcmin$^{-2}$.
@@
 Finally, we randomly placed quasars in the source plane with a projected number density of $70$~arcmin$^{-2}$.
-This is a much larger number density than observed in the real universe, but we are allowed to do so because we are not interested in predicting the absolute number of strong lenses.
+This is a much larger number density than observed in the real universe, but we are allowed to do so because we are not interested in predicting the absolute number of strong lenses, so this is a legitimate choice.
 The advantage of boosting the number density of quasars is that it allows us to produce a large number of lenses without the need for generating too big a population of foreground galaxies.
@@
 The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around \num{300000}.
-%Then we generated a population of sources behind the lenses.
-%We approximated the foreground galaxies as isolated: for each source, we only modelled the contribution to the lensing field generated by a single lens.
 We approximated the foreground galaxies as isolated: when determining whether a galaxy acts as a strong lens, we only modelled the contribution to the lensing signal from the galaxy itself, and neglected that of the environment.
@@
 Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
-%These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
-%
 We then applied the peak detection-based lens selection criterion introduced in section \ref{ssec:lensdefext} to find the strong lenses.
@@
 
-%\subsection{Einstein radius cuts}\label{ssec:teincuts}
 \subsection{Lens finding probability}\label{ssec:teincuts}
@@
 
-Both of these scenarios can result in samples that are incomplete below a certain value of the Einstein radius. %in which lenses with a larger Einstein radius are preferentially selected.
+Both of these scenarios can result in samples that are incomplete below a certain value of the Einstein radius. 
 We simulate this situation via the following Einstein radius-dependent lens finding probability:
@@
 Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
-Red dashed lines: distribution of the lenses with Einstein radius larger than $1.0''$.
+Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
 \label{fig:lenspars}
@@
 The most striking difference between the samples is in the stellar mass: strong lensing selects preferentially galaxies with larger values of $\mobs$.
-%, which is much larger in the lenses than in the general population. 
 Lenses tend to also have a larger halo mass, a smaller half-light radius at fixed stellar mass, and a larger stellar population synthesis mismatch parameter.
 The distribution in ellipticity and inner dark matter slope of the lenses instead look very similar to that of the parent population.
-%This outcome was expected, given the results of the analysis of \Sref{sect:indlenses}.
 Additionally, we can see that the amplitude of the strong lensing bias appears to be always larger in the lens sample with the more restrictive selection on Einstein radius.
@@
 Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
-Red dashed lines: distribution of the lenses with Einstein radius larger than $1.0''$.
+Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
 \label{fig:sourcepars}
@@
 For this reason, also the distribution in half-light radius is very similar between the lensed sources and the detectable source population.
-%We can also see a bias in favour of sources with a very low S\'{e}rsic index.
-This is a consequence of the selection in magnitude: our source sample shows a strong correlation between $\msource$ and $n$: the brightest sources tend to have the smallest values of the S\'{e}rsic index.
 
@@
 The horizontal dashed line in each panel shows the value of the parent population: the larger the distance between the curve of a simulation and this line, the higher the strong lensing bias.
-
-%\Fref{fig:lensobsbias} shows the median of various lens galaxy observable quantities as a function of the minimum Einstein radius, for the three lens population simulations
 
@@
 In the fiducial scenario, the bias on $\mdmfive$ is as small as $0.02$~dex for $\theta_{\mathrm{Ein,min}} < 0.5''$, and is negligible in the low-scatter scenario.
-%Interestingly, the bias is negative in the low-scatter scenario.
-%This is a counter-intuitive result that follows from the fact that the model of \Eref{eq:dmscaling} is not a perfect description of the distribution in $\mdmfive$.
-Nevertheless, the amplitude of the bias is only $0.01$~dex, which for all practical purposes is negligible.
 
@@
 In the simulation with large scatter, the bias in halo mass dominates, and therefore $\gammadm$ is negatively biased.
-
-%The trends in the avereage $\gammadm$ reflect very closely those in $\mdmfive$, but with flipped sign.
-%This is because, in our model for the dark matter density profile, inner dark matter mass and density slope are perfectly correlated: both $\mdmfive$ and $\gammadm$ are determined univocally once the halo mass, stellar mass and half-light radius are specified.
-The fact that the bias is negative is somewhat counter-intuitive: lenses with a stronger response to baryons have both a larger value of $\mdmfive$ and a steeper inner dark matter slope.
-%However, $\gammadm$ is also anti-correlated with total halo mass (see $\gammadm-\mhalo$ panel in \Fref{fig:lenspars}): since strong lenses tend to have a preferentially larger halo mass, they also have a shallower dark matter slope.
 
@@
 For this reason, we only ran simulations with lensed quasars in the fiducial scatter scenario and compared the results with those from the extended source simulation.
-%We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we computed the bias for this sample as well.
 We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we also analysed that subsample on its own.
-%\Fref{fig:quasarpars} shows the distribution in a 
-%\Fref{fig:quasarbias} shows the strong lensing bias obtained with the lensed quasars simulation, compared with that
 
@@
 This trend was expected, given the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
-%As we discuss in \Sref{sect:discuss}, the difference in ellipticity between quad lenses and 
 Quad lenses tend to also have a slightly larger halo mass at fixed stellar mass, for $\theta_{\mathrm{Ein},min}<1''$.
@@
 Bias in the population of quasar lenses (all and quads only), compared to the extended source simulation.
-%Bias on lens mass properties as a function of the minimum Einstein radius.
 First panel: median $\log{\asps}$.
-Second panel: median $\mhalo$ at $\log{\mobs}=11.4$.
-Third panel: median $\mdmfive$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
+Second panel: median $\mhalo$ at $\log{\mstar}=11.5$.
+Third panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
 Fourth panel: median lens axis ratio.
@@
 \end{figure}
-
-%As a last step, we also analysed the bias in the source-related parameters of the lensed quasars simulation.
-%\Fref{fig:quasar_cp} shows the distribution in source redshift and magnitude of the lensed quasars, in the $\theta_{\mathrm{Ein,min}}>0.5''$ case, compared to that of the parent and detectable populations.
 
@@
 Together, they define the lens selection probability term $\psel$ of \Eref{eq:one}.
-%In the best case scenario, the only sources of bias are the first two points.
 The first two points are intrinsic to a strong lensing survey and constitute an unavoidable source of bias. 
@@
 Compared with this uncertainty, the amplitude of the strong lensing bias on $\asps$ is small: strong lensing observations can be used directly to discriminate between these two alternative choices of IMF, without the need to correct for selection effects.
-%This is especially the case when working with samples of lenses that probe the sub-arcsecond Einstein radius regime.
 Such a goal could be reached with a sample size of a thousand lenses and a statistical study of the kind proposed by \citet{S+C21}.
@@
 However, it appears to be in contradiction with the work of \citet{O+A17}, who argued that their strong lens sample selected preferentially compact sources.
-%However, \citet{O+A17} showed that their strong lens sample selected preferentially compact sources. 
 The origin of this discrepancy probably lies in the differences between the criteria used to define a strong lens in the two studies.
@@
 Compact galaxies with a large overall magnification, instead, are more likely to be detected \citep[see the discussion in section 5.3 of][]{O+A17}.
-
-%Our analysis based on photometric detection showed that the tendency to select compact sources is not a general feature of strong lens samples.
+We refer the reader to \citet{Arn++12} for a thorough study of selection effects associated with spectroscopy.
+
 There is also an indirect way in which strong lensing could preferentially probe more compact sources than possible in the field.
@@
 This description could be extended by including the central velocity dispersion as an additional control parameter. 
-%Galaxies selected in velocity dispersion are similar to existing samples of lenses in terms of other observable quantities
 The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution marginalised over the whole population.
@@
 The effect of including external shear in a lens model is similar to that of changing the ellipticity of the lens. 
-%Therefore the main result that would be affected by neglecting it is the bias on the axis ratio.
 External convergence mimics the effect of adding or removing a constant sheet of invisible mass, producing an effect similar to varying the distribution of dark matter.
@@
 This work takes a major step towards filling that knowledge gap.
-%After a thorough investigation of the strong lensing bias in photometrically selected samples of strong lenses, we learned the following important lessons. 
-%After a thorough investigation of the strong lensing bias, we learned the following important lessons that apply to photometrically selected samples of strong lenses.
 After a thorough investigation, we learned several lessons regarding the strong lensing bias in photometrically selected lens samples. The following are the most important ones.
-%The thorough investigation of the strong lensing bias that we carried out in this work allowed us to learn several important lessons.
 \begin{enumerate}
@@
 Sources with a surface brightness that is too low to be detected are still undetected when strongly lensed. 
-%\item At fixed stellar distribution and fixed dark matter mass enclosed within a given physical aperture, varying the inner dark matter slope has little impact on the strong lensing cross-section.
 \item Lens galaxies tend to be more massive and more compact than their non-lens counterparts. Their redshift distribution is also modified with respect to the general galaxy population.
@@
 \item 
-%We used observational constraints from the fundamental plane to set the values of the intrinsic scatter in our fiducial simulation.
 Under reasonable assumptions on the intrinsic scatter parameters, for a Euclid-like survey that is complete down to $\tein=0.5''$ the bias on $\asps$ is smaller than $0.04$~dex (10\%).
@@
 Biases on mass-related quantities, such as the stellar mass-to-light ratio or the dark matter distribution, are more difficult to measure directly and must be modelled by taking into account selection effects.
-%, but can be reduced by modelling how they scale with the observable properties of the lenses, thus minimising their intrinsic scatter.
-%In the future, it might be desirable to directly model strong lensing selection effects, thus accounting for the strong lensing bias.
 Designing strong lensing surveys with clearly-defined and easily-modellable selection criteria would help greatly in this task.
-%A peak-based definition of a strong lensing event such as that introduced in section \ref{ssec:lensdefext} could be used for this purpose.
 
@@
 These cuts produced a sample of $8078$ galaxies.
-%We focused on the $r-$band de Vaucouleurs profile measurements obtained by \citet{Mee++15}, for these objects.
-%We then took measurements of their total flux, half-light radius and axis ratio from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}. Using the 
-%We then took measurements of the total flux, half-light radius and axis ratio obtained by \citet{Mee++15} from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}.
-Using the
-%For each object, we took the measurement of its de Vaucouleurs model-based $r-band$ total flux, half-light radius and axis ratio, as provided by \citet{Mee++15}.
 We then focused on the $r-$band de Vaucouleurs model-based photometric measurements of \citet{Mee++15}.
@@
 
-%We did not attempt to match the other parameters of the fundamental plane (i.e.
-the constant of proportionality of \Eref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.
-
 \end{document}

### 2023-05-26 paper/seleff.tex
rewritten intro, magnification-dependent selection

@@
            Giulia Despali\inst{\ref{ita}}\and
+           Raphael Gavazzi\inst{\ref{marseille}}\and
            Anowar J. Shajib\inst{\ref{chicago},\ref{kicp},\ref{einstein}}\and
@@
 Institut für Theoretische Astrophysik, Zentrum für Astronomie, Heidelberg Universität, Albert-Ueberle-Str. 2, 69120, Heidelberg, Germany\label{ita}\and
+CNRS and CNES, Laboratoire d'Astrophysique de Marseille, Aix-Marseille Universit\'{e} 38, Rue Fr\'{e}d\'{e}ric Joliot-Curie, F-13388 Marseille, France\label{marseille}\and
 Department of Astronomy \& Astrophysics, University of Chicago,
@@
 Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
-A necessary condition for a galaxy to act as a strong lens with respect to a given source is that its projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
+In order for a lens-source pair to be included in a strong lensing survey, the following conditions must be met:
+\begin{enumerate}
+\item The lens must produce multiple images of the source
+\item The multiple images must be detected and resolved
+\item The system must be recognized as a lens by the survey.
+\end{enumerate}
+Condition 1 poses constraints on the mass distribution of the lens, as well as on the geometry of the system. 
+First of all, not all lenses are capable of producing multiple images, but only those whose surface mass density $\Sigma(\boldsymbol\theta)$ is larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
 This condition excludes objects with a very diffuse mass distribution from the population of lenses.
-In general, galaxies with a higher concentration of mass are more likely to be strong lenses, and are therefore over-represented in lens samples.
-The distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies: for instance, lensing magnification allows the detection of fainter objects with respect to the field.
-
-In general, the probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
+Second, objects with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples.
+Third, at fixed lens properties, sources that are located further away along the line of sight are more likely to be lensed.
+
+%First, the projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}, a condition that excludes objects with a very diffuse mass distribution from the population of lenses.
+%Second, the source must be sufficiently close in projection to the lens center, in a way that depends on the lens mass distribution.
+Galaxies with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples.
+Third, at fixed lens properties, sources that
+
+Condition 2 sets limits on the observable properties of the sources and, again, the lens mass distribution, in a way that depends on the survey characteristics.
+On the one hand, brighter sources are more likely to be detected when lensed. On the other hand, lensing magnification allows the detection of fainter objects with respect to the field.
+However, if the angular separation between the multiple images is too small, the system may not be classified as a lens.
+Since the image separation increases with lens mass, this means that the lens sample is biased against low-mass objects.
+
+Condition 3 can introduce additional biases, depending on the efficiency of the lens finding process on which the survey is based.
+Most commonly, lenses with a smaller image separation are more difficult to identify even if detected and resolved, because of contamination from the lens galaxy.
+This introduces an additional bias towards lenses with a larger or more concentrated mass.
+
+%A necessary condition for a galaxy to act as a strong lens with respect to a given source is that its projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
+%This condition excludes objects with a very diffuse mass distribution from the population of lenses.
+%In general, galaxies with a higher concentration of mass are more likely to be strong lenses, and are therefore over-represented in lens samples.
+%The distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies: for instance, lensing magnification allows the detection of fainter objects with respect to the field.
+
+The cumulative effect of these conditions can be summarised as follows.
+%In general, 
+The probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
 \begin{equation}\label{eq:one}

### 2023-05-26 lens_population/minmu_extended_sources.py

@@ -1,0 +1,334 @@
+import numpy as np
+import os
+import glafic
+import h5py
+from simpars import *
+from scipy.signal import convolve2d
+from astropy.io import fits as pyfits
+from scipy.interpolate import splev
+from scipy.optimize import brentq
+from sl_profiles import sersic, gnfw, deVaucouleurs as deV
+import sl_cosmology
+from sl_cosmology import G, M_Sun, Mpc, c
+from scipy.special import gamma as gfunc
+from lensdet import detect_lens
+import sys
+
+
+modelname = sys.argv[1]
+sourcecat = pyfits.open('/Users/alessandro/catalogs/skills_sourceonly_zcut.fits')[1].data
+pop = h5py.File('%s_galaxies.hdf5'%modelname, 'r+')
+
+psf = pyfits.open('psf.fits')[0].data
+
+modeldir = 'extended_sims/%s/'%modelname
+if not os.path.isdir(modeldir):
+    os.system('mkdir %s'%modeldir)
+
+nsamp = pop.attrs['nsamp']
+islens_samp = np.zeros(nsamp, dtype=bool)
+tein_zs_samp = np.zeros(nsamp)
+rein_zs_samp = np.zeros(nsamp)
+
+f = open('%s_sources.cat'%modelname, 'r')
+sourcelines = f.readlines()[1:]
+f.close()
+
+# primary parameters
+omegaM = 0.3
+omegaL = 0.7
+weos = -1.
+hubble = 0.7
+prefix = 'tmp'
+xmin = -5.
+ymin = -5.
+xmax = 5.
+ymax = 5.
+pix_ext = pix_arcsec
+pix_poi = 0.1
+maxlev = 5
+
+glafic.init(omegaM, omegaL, weos, hubble, prefix, xmin, ymin, xmax, ymax, pix_ext, pix_poi, maxlev, verb = 0)
+glafic.set_secondary('flag_hodensity 2')
+glafic.set_secondary('nfw_users 1')
+glafic.set_secondary('hodensity 200')
+
+glafic.startup_setnum(2, 1, 0)
+glafic.set_lens(1, 'gnfw', 0.3, 1e13, 0.0, 0.0, 0., 90.0, 10., 1.5)
+glafic.set_lens(2, 'sers', 0.3, 1e11, 0.0, 0.0, 0., 90.0, 1., 4.)
[…]

### 2023-05-26 lens_population/simpars.py

@@ -18,6 +18,7 @@ pix_arcsec = 0.1
 nsigma_pixdet = 2.
 sky_rms = 0.00433
 npix_min = 1
+minmu = 4. # minimum magnification (only applied to alternative selection sim)
 
 # stellar mass function
 phi_muz = 1.009e-3

### 2023-05-31 paper/seleff.tex
minmu

@@
 
+{\bf
 Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
@@
 The cumulative effect of these conditions can be summarised as follows.
+}
 %In general, 
@@
 However, testing for the amplitude of the strong lensing bias on the mass-to-light ratio gradient is beyond the scope of this paper.
+
+\subsection{Alternative lens definitions}
+
+{\bf
+Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in section \ref{ssec:lensdefext}.
+However, many previous studies aimed at simulating populations of lenses adopted a magnification-based definition, in which a lens-source system is classified as a strong lens if the magnification of the source is larger than a minimum value \citep[see e.g.][]{Col15, Rob++20}.
+In order to check whether our results are robust to the choice of definition of a strong lensing event, we re-analysed the fiducial sample with this alternative definition.
+%repeated the analysis of the fiducial sample with extended sources.
+In particular, we defined as strong lenses all systems with a detected background source and $\mu_{\mathrm{tot}} > 3$, where $\mu_{\mathrm{tot}}$ is the total magnification (obtained by integrating the flux of all of the lensed images and dividing it by the intrinsic source flux).
+The total number of strong lenses increases by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
+\Fref{fig:minmubias} shows the strong lensing bias on a few selected quantities (stellar population synthesis mismatch parameter, projected dark matter mass within $5$~kpc, source magnitude and source half-light radius).
+
+}
+
+%However, many previous simulation studies adopted criteria based on magnification \citep[see e.g.][]{Col15, Rob++20}.
+
+However, in the literature it is common to find alternative criteria based on magnification \citep[see e.g.][]{Col15,
+
+it is common to find 
+In the 
+Other simulation works, however, use a
+
+Our analysis was based on the definition of a strong lens given in sections \ref{ssec:lensdefpoint} and \ref{ssec:lensdefext}. 
+These are not the only possible definitions.
+In the case of an extended source, in particular, it is common to find in the literature magnification-based
 

### 2023-05-31 lens_population/plot_minmu_merged_bias.py

@@ -1,0 +1,171 @@
+import numpy as np
+import pylab
+import h5py
+from simpars import lmobs_piv, lreff_piv
+from scipy.optimize import leastsq
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 20
+
+smag_min = 24.8
+smag_max = 25.2
+
+lpopnames = ['fiducial_1000sqdeg_lenses.hdf5', 'fiducial_1000sqdeg_minmulenses.hdf5']
+labels = ['Fiducial', '$\mu_{\mathrm{tot}} > 3$']
+lensdefs = ['islens', 'minmulens']
+teindefs = ['tein_zs', 'minmutein_zs']
+nsims = len(lpopnames)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+colors = [colseq[0], colseq[5]]
+linestyles = ['-', '--']
+
+fig, ax = pylab.subplots(4, 1, figsize=(6, 13))
+
+pylab.subplots_adjust(left=0.2, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
+
+ntein = 21
+tein_arr = np.linspace(0., 2., ntein)
+
+detectable = h5py.File('detectable_sources.hdf5', 'r')
+
+zmed_det = np.median(detectable['zs'][()])
+smagmed_det = np.median(detectable['smag'][()])
+
+smagbin_det = (detectable['smag'][()] > smag_min) & (detectable['smag'][()] < smag_max)
+
+nsermed_det = np.median(detectable['nser'][smagbin_det])
+sreffmed_det = np.median(detectable['sreff'][smagbin_det])
+
+qmed_det = np.median(detectable['sq'][()])
+
+galpop = h5py.File('fiducial_1000sqdeg_galaxies.hdf5', 'r')
+laspsmed_gal = np.median(galpop['lasps'][()])
+
+for n in range(nsims):
+
+    laspsmed_arr = np.zeros(ntein)
+    laspserr_arr = np.zeros(ntein)
+
+    lmdm5med_arr = np.zeros(ntein)
+    lmdm5err_arr = np.zeros(ntein)
+
+    def mdm5fitfunc(p):
[…]

### 2023-05-31 lens_population/simpars.py

@@ -18,7 +18,7 @@ pix_arcsec = 0.1
 nsigma_pixdet = 2.
 sky_rms = 0.00433
 npix_min = 1
-minmu = 4. # minimum magnification (only applied to alternative selection sim)
+minmu = 3. # minimum magnification (only applied to alternative selection sim)
 
 # stellar mass function
 phi_muz = 1.009e-3

### 2023-06-12 paper/seleff.tex

@@
 Strictly speaking, the results shown in \Sref{sect:results} apply only to samples of lenses and background sources with the same properties as our simulations.
+{\bf Real lens samples can have more complex source distributions.
+For example, many existing lens-finding algorithms are tuned to identify sources with different colours from the lenses, introducing a selection that can modify the source distribution in redshift-magnitude space.
+}
 However, the comparison of section \ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
@@
 In particular, we defined as strong lenses all systems with a detected background source and $\mu_{\mathrm{tot}} > 3$, where $\mu_{\mathrm{tot}}$ is the total magnification (obtained by integrating the flux of all of the lensed images and dividing it by the intrinsic source flux).
-The total number of strong lenses increases by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
+The total number of strong lenses increased by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
 \Fref{fig:minmubias} shows the strong lensing bias on a few selected quantities (stellar population synthesis mismatch parameter, projected dark matter mass within $5$~kpc, source magnitude and source half-light radius).
-
-}
-
-%However, many previous simulation studies adopted criteria based on magnification \citep[see e.g.][]{Col15, Rob++20}.
-
-However, in the literature it is common to find alternative criteria based on magnification \citep[see e.g.][]{Col15,
-
-it is common to find 
-In the 
-Other simulation works, however, use a
-
-Our analysis was based on the definition of a strong lens given in sections \ref{ssec:lensdefpoint} and \ref{ssec:lensdefext}. 
-These are not the only possible definitions.
-In the case of an extended source, in particular, it is common to find in the literature magnification-based
+Besides the change in the source magnitude (third panel), we see very little differences in the bias with respect to the peak-based definition of the fiducial analysis.
+We then conclude that the exact definition of a strong lensing event does not have a big impact on the strong lensing bias of lens-related properties.
+We stress out, however, that magnification is not a directly-observable quantity, and therefore we advise against applying magnification-based lens definitions to define lens samples, for the reasons discussed in section \ref{ssec:lensdefext}.
+\begin{figure}
+\includegraphics[width=\columnwidth]{minmu_bias.eps}
+\caption{
+Strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves).
+First panel: median $\log{\asps}$.
+Second panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+Third panel: median source magnitude.
+Fourth panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
+}
+\label{fig:minmubias}
+\end{figure}
+}
 

### 2023-06-18 paper/reply_to_referee.txt
revised version

We thank the referee for the useful suggestions. We have addressed the points raised in the report, as explained below.

- "The authors have a novel definition of a lens, and a toy definition of lens finding. The results are likely very sensitive to these definitions (indeed their conclusion about source size biases is demonstrably at odds with other authors because they have different definitions of a detectable lens). The authors should investigate which of their results are robust to different definitions of a lens and a discovered lens, and which findings are not robust e.g. if they impose a minimum magnification cut, or only choose objects with a centroid within the caustic."

Reply - We added a subsection in the Discussion (6.6) to check how the key results vary with a magnification-based lens definition.

- "I also wonder how much source colour plays a role - we often think of lenses as red and lensed sources as blue. Given their source catalogue has colours, the authors might consider investigating if colour based finders can introduce a potential selection bias too."

Reply - A selection on colour is equivalent to a change in the properties of the underlying source population. Given the small differences in the strong lensing bias between galaxy-galaxy and galaxy-quasar lenses, which have different source redshift distributions, we expect the effect of colour selection to be similarly small.

- "This work doesn't seem to be a huge step forward compared to Collett & Cunnington 2016 (https://ui.adsabs.harvard.edu/abs/2016MNRAS.462.3255C/abstract). The authors don't seem to be aware of that paper and I think they need to rework their lensed quasar section substantially to compare their result to Collett & Cunnington, in a similar way to how they compare with Mandelbaum 2008 for galaxy-galaxy lenses."

Reply - We thank the referee for pointing out this important work that we had missed. Our paper, however, addresses a different question from the ones covered by Collett & Cunnington 2016: how galaxy-quasar lenses are different from galaxy-galaxy lenses. We hope that the revised text make this point more clear.

- "It would be good if this section had a more concrete application to time-delay cosmography. The main potential H0 bias in applying galaxy-galaxy results to lensed quasars is likely to be differences in the total density slope of the populations. A figure comparing these would be very useful. Ideally, the authors would also seek to quantify if the H0 shifts seen by Birrer 2020 (https://ui.adsabs.harvard.edu/abs/2020A%26A...643A.165B/abstract) are consistent with differences in the selection function of the SLACS and TDCOSMO lenses."

Reply - We are not convinced that comparing the total density slopes would help much with understanding the impact on H0: the slopes of time-delay lenses are typically measured with very high precision, so there are not many concrete cases in which having a prior on the density slope would help with the inference of H0. What currently dominates the uncertainty on H0 is the deviation from a pure power-law density profile, which Birrer et al. 2020 described by a mass-sheet transformation parameter. We added a subsection (5.2.2) to quantify the bias on lambda_mst, which can be directly translated into a bias on H0.

- "More context should be given to the size of the biases detected - are they likely to substantially change current or near-future conclusions about astrophysics/cosmology?"

Reply - We're a bit unsure about how to address this point. We believe that the quantitative statements in the conclusions (particularly bullet point 6 and 8) already give an answer to this question.

- "I found the introduction to get very technical, very quickly. Equation 1 isn't really an equation, more an illustration of what goes into the discovered lens population. I think a bit more qualitative text about lensing selection biases would be helpful before jumping in to equation 1."

Reply - We restructured the introduction accordingly.

### 2023-06-18 paper/seleff.tex
revised version

@@
 
+\def\lmst{\lambda_{\mathrm{mst}}}
+
 \def\Sref#1{Section~\ref{#1}\xspace}
@@
 The bias has little dependence on the properties of the source population: samples of galaxy-galaxy lenses and galaxy-quasar lenses that probe the same Einstein radius distribution are biased in a very similar way.
-Quadruply imaged quasar lenses, however, are biased towards higher ellipticity galaxies.
+%Quadruply imaged quasar lenses, however, are biased towards higher ellipticity galaxies.
+%{\bf
+%The bias on time-delay measurements of $H_0$ due to the use of prior information from galaxy-galaxy lenses is on the order of $1\%$.
+%}
 }
@@
 Given current uncertainties, strong lensing observations can be used directly to improve our current knowledge of the inner structure of galaxies, without the need to correct for selection effects.
+{\bf Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.}
+%{\bf Galaxy-galaxy lenses can be used to inform the properties of quasar lenses for the purpose of measuring $H_0$ from time delays
 }
@@
 
-Finally, we addressed the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses.
-This last point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
-Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{B+T21}, but any difference between the two lens classes can introduce biases, if not corrected for.
+{\bf Finally, we considered galaxy-quasar lenses. 
+The strong lensing bias of lensed quasars was studied by \citep{C+C16}, in the context of power-law lens mass models with external convergence.
+Here we addressed the question of how different are galaxy-quasar lenses from galaxy-galaxy lens samples. 
+}
+%Finally, we addressed the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses. 
+This %last% 
+point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
+Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{Bir++20,B+T21}, but any difference between the two lens classes can introduce biases, if not corrected for.
 With this study we aim to quantify this bias.
@@
 
+\subsubsection{Bias in lens parameters}
+
 The main goal of the experiment with lensed quasars is to check whether there are any differences in the strong lensing bias with respect to the extended source case, at fixed properties of the foreground galaxy population.
@@
 
-\Fref{fig:quasarbias} shows the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
+{\bf The first four rows of \Fref{fig:quasarbias} show} the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
 The biases of the population of lensed quasars (red curves) are very similar to that of lensed galaxies (blue curves), especially for values of $\theta_{\mathrm{Ein},min}<1''$, 
 When considering only quad lenses, however, there are some differences, the most remarkable of which is the bias in the axis ratio: quad lenses tend to be on average galaxies with a higher ellipticity.
-This trend was expected, given the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
+This is a well-known effect \citep[see e.g.][]{KKS97} which can be explained in the context of the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
+%This trend was expected, given the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
 Quad lenses tend to also have a slightly larger halo mass at fixed stellar mass, for $\theta_{\mathrm{Ein},min}<1''$.
@@
 Fourth panel: median lens axis ratio.
+Fifth panel: median mass-sheet transformation parameter, defined in Appendix~\ref{sect:appendixc}.
 In each panel, the dashed line indicates the value of the parent population.
@@
 \end{figure}
+
+\subsubsection{Impact on $H_0$}
+
+{\bf
+A measurement of the time delay between the images of a lensed quasar can be converted into an estimate of the expansion rate of the Universe, $H_0$.
+%, provided that the mass distribution of the lens galaxy (and of matter along the line-of-sight) is accurately known.
+In general, though, the value of $H_0$ is degenerate with the mass distribution of the lens (and of matter along the line-of-sight), which can be difficult to constrain on an individual lens basis.
+One possible strategy to break this denegeracy is to use information gathered from a larger set of lenses as a prior for the time-delay lenses.
+This was the strategy adopted by \citet{Bir++20} in the most recent measurement of $H_0$ by the TDCOSMO collaboration: they combined a larger sample of extended source lenses with a smaller set of time-delay lenses.
+In this section we want to quantify the possible bias on $H_0$ that can result from this approach.
+
+The lens mass model used by \citet{Bir++20} consists in a power-law density profile modified by a mass-sheet transformation. The corresponding projected density profile is
+\begin{equation}\label{eq:mstprofile}
+\kappa(\theta) \propto \lmst\theta^{1-\gamma} + (1 - \lmst),
+\end{equation}
+where $\lmst$ is called the mass-sheet transformation parameter.
+Strong lensing data can constrain very well the power-law index $\gamma$, but not at all $\lmst$.
+%, on which the inference on the Hubble constant depends.
+Given the time delay, the inferred value of the Hubble constant $H_0$ scales linearly with $1/\lmst$.
+\citet{Bir++20} measured the population distribution of $\lmst$ on the sample of extended source lenses, by combining strong lensing and stellar kinematics data, and then used it as a prior in the time-delay analysis.
+Although the mass model used in our work is different from that of \Eref{eq:mstprofile}, we can estimate the bias associated with this approach by computing the value of $\lmst$ for each lens and then comparing the strong lensing bias on $\lmst$ of extended source and quasar lenses.
+Appendix~\ref{sect:appendixc} explains how we obtained $\lmst$ for each lens.
+The strong lensing bias on the median value of $\lmst$ is shown in the fifth panel of \Fref{fig:quasarbias}\footnote{Since $\lmst$ is only defined for strong lenses, the median value of $\lmst$ of the general population of galaxies is not defined.}.
+
+At fixed minimum Einstein radius, there is a small difference between the median $\lmst$ in extended source and quasar lenses, typically of order 1\%.
+The associated bias on $H_0$ is, therefore, equally small.
+The value of $\lmst$, however, increases with increasing minimum $\tein$. This implies that if the two samples probe different ranges in Einstein radius, the bias on $H_0$ can be larger than that.
+The bias can also be larger if the properties of the general population of galaxies (as opposed to the strong lenses) are used to inform the structure of time-delay lenses \citep[see][]{C+C16}.
+This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in \Fref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines)
+}
 
@@
 This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
+{\bf
+The associated bias on the inference of $H_0$ is on the order of $1\%$.
+}
 \item Samples of quad lenses are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is also biased.
@@
 
+\section{Mass-sheet transformation parameter}\label{sect:appendixc}
+
+In the time-delay study of \citet{Bir++20}, lenses are described with a mass density profile given by \Eref{eq:mstprofile}.
+\citet{Bir++20} defined the profile by first fitting a pure power-law mass model to the strong lensing data, thus constraining the Einstein radius and power-law index $\gamma$, and then applying a mass-sheet transformation of the following kind:
+\begin{equation}\label{eq:mst}
+\kappa(\boldsymbol\theta) \rightarrow \lmst\kappa(\boldsymbol\theta) + 1 - \lmst.
+\end{equation}
+%\citet{Bir++20} fixed the slope $\gamma$ of the power-law component to the value obtained by fitting a pure power-law mass model to the strong lensing data.
+The mass-sheet transformation parameter $\lmst$ can be interpreted as a quantity describing a departure from a pure power-law profile.
+One possible way of estimating $\lmst$ for our lenses is to fit a power-law to the simulated images and then optimize for the value of $\lmst$ that best matches the true density profile.
+That, however, would be too time-consuming.
+Instead, we emulate this process as follows.
+
+\citet{Son18} showed that, when the main arc and the counter-image of an extended source are well resolved, their relative widths constrain the following combination of derivatives of the lens potential at the Einstein radius:
+\begin{equation}\label{eq:invariant}
+\frac{\psi'''}{1 - \psi''}.
+\end{equation}
+In other words, the above quantity can be measured robustly (i.e. in a model-independent way).
+When fitting a pure power-law lens model to lensing data, then, the value of $\gamma$ that we obtain is the one that reproduces the true $\psi'''/(1-\psi'')$\footnote{Strictly speaking, this is only true in the limit of well-resolved arcs and nearly axisymmetric lenses}.
+This is given by
+\begin{equation}
+\gamma = 2 + \tein \frac{\psi'''}{1-\psi''}
+\end{equation}
+Hence, given a lens in our sample, we first define a pure power-law model with the same Einstein radius as the true one and with $\gamma$ given by the equation above. 
+Then, given a pure power-law model, we find $\lmst$ such that the first three derivatives of the lens potential match the true values.
+In particular, since a mass-sheet transformation changes the second derivative of the potential as
+\begin{equation}
+\psi'' \rightarrow \lmst\psi'' + 1 - \lmst,
+\end{equation}
+we set
+\begin{equation}
+\lmst = \frac{1 - \psi''_{\mathrm{PL}}}{1 - \psi''},
+\end{equation}
+where $\psi''_{\mathrm{PL}}$ is the value of $\psi''$ of the pure power-law lens model.
+
 \end{document}

### 2023-06-26 paper/reply_to_referee.txt
reply to referee

@@
 
-Reply - We added a subsection in the Discussion (6.6) to check how the key results vary with a magnification-based lens definition.
+Reply - We added a subsection in the Discussion (6.7) to check how the key results vary with a magnification-based lens definition.
 
@@
 Given the small differences in the strong lensing bias between galaxy-galaxy and galaxy-quasar lenses, which have different source redshift distributions, we expect the effect of colour selection to be similarly small.
+We added a comment on this point in section 6.3.
 
@@
 Our paper, however, addresses a different question from the ones covered by Collett & Cunnington 2016: how galaxy-quasar lenses are different from galaxy-galaxy lenses.
-We hope that the revised text make this point more clear.
+We hope that the revised text makes this point more clear.
 
@@
 Reply - We restructured the introduction accordingly.
+
+Finally, we added a section to the Discussion to address the issue of lens vs. source selection (section 6.6).

### 2023-06-26 paper/seleff.tex
adding lens-source selection
reply to referee

@@
 However, testing for the amplitude of the strong lensing bias on the mass-to-light ratio gradient is beyond the scope of this paper.
+
+\subsection{Lens vs. source selection}
+
+{\bf
+The simulations described in \Sref{sect:sims}, on which our experiment is based, are examples of so-called lens-selected surveys:
+the lens samples are built by first defining a population of possible lens galaxies and then searching for strong lenses among them.
+Photometric lens searches are typically carried out in this way \citep[see e.g.][]{Gav++12,Son++18,Pet++19}.
+An alternative approach is to build a source-selected sample, by first searching for signatures of strong lensing within a population of possible background sources and then obtaining information on the lens galaxies \citep[see e.g.][]{Hez++13}.
+In principle, these two approaches lead to different strong lensing biases. %Therefore, strictly speaking, our results only apply to a lens-selected sample.
+%For example, from \Fref{fig:lenspars} we can see that the distribution in redshift and stellar mass of the lenses extends beyond the bounds applied to the population of foreground galaxies: in principle, a source-based selection method
+For example, from \Fref{fig:lenspars} we can see that the stellar mass distribution of the lenses extends below the minimum mass cut that we applied to the population of foregrounds galaxies: a source-based search that is able detect those lenses would result in a different sample and a different associated strong lensing bias.
+However, when considering only lenses with $\tein > 1''$, lens-selection and source-selection yield the same sample (see red histogram in \Fref{fig:lenspars}): in the limit of large Einstein radius, the two operations (identifying suitable lens galaxies and detecting a strongly lensed source) commute.
+
+%In general, speaking of a lens- or source-selected sample is improper.
+In general, there is no such thing as a purely lens- or source-selected sample. 
+A lens-based search is always also source-based, because a detection of the lensed source is required to identify a lens.
+A source-based search is also lens-based, because the lens properties determine important observational features of the lensed source, such as the number of images, the image separation and the total flux.
+}
 

### 2023-06-28 paper/seleff.tex
fixed refs and added Dogruel

@@
 The bias can also be larger if the properties of the general population of galaxies (as opposed to the strong lenses) are used to inform the structure of time-delay lenses \citep[see][]{C+C16}.
-This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in \Fref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines)
+This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in \Fref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines).
 }
@@
 {\bf
-The simulations described in \Sref{sect:sims}, on which our experiment is based, are examples of so-called lens-selected surveys:
+The simulations described in \Sref{sect:lenspop}, on which our experiment is based, are examples of so-called lens-selected surveys:
 the lens samples are built by first defining a population of possible lens galaxies and then searching for strong lenses among them.
-Photometric lens searches are typically carried out in this way \citep[see e.g.][]{Gav++12,Son++18,Pet++19}.
+Photometric lens searches are typically carried out in this way \citep[see e.g.][]{Gav++12,Son++18a,Pet++19}.
 An alternative approach is to build a source-selected sample, by first searching for signatures of strong lensing within a population of possible background sources and then obtaining information on the lens galaxies \citep[see e.g.][]{Hez++13}.
@@
 In principle we should also account for observational uncertainties on the stellar mass measurement, as they too contribute to the inferred scatter in $\sigma_{\mathrm{ap}}$.
-In practice, however, it is difficult to estimate observational uncertainties on stellar population synthesis measurements.
+In practice, however, it is difficult to estimate observational uncertainties on stellar population synthesis measurements \citep[but see][]{Dog++23}.
 For this reason we chose not to propagate uncertainties on $\mobs$. As a result, the inferred scatter parameter $\sigma_{\sigma}$ is slightly overestimated.

### 2023-07-03 paper/seleff.tex
added sentence on color selection
resubmitted version

@@
            Shun-Sheng Li\inst{\ref{leiden}}\and
-           Giulia Despali\inst{\ref{ita}}\and
+           Giulia Despali\inst{\ref{bologna}}\and
            Raphael Gavazzi\inst{\ref{marseille}}\and
@@
 Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\label{leiden}\and
-Institut für Theoretische Astrophysik, Zentrum für Astronomie, Heidelberg Universität, Albert-Ueberle-Str. 2, 69120, Heidelberg, Germany\label{ita}\and
+Alma Mater Studiorum - Università di Bologna, Dipartimento di Fisica e Astronomia "Augusto Righi", Via Gobetti 93/2, Bologna, Italy\label{bologna}\and
 CNRS and CNES, Laboratoire d'Astrophysique de Marseille, Aix-Marseille Universit\'{e} 38, Rue Fr\'{e}d\'{e}ric Joliot-Curie, F-13388 Marseille, France\label{marseille}\and
@@
 We can then conclude that the details of the properties of the background source population play a secondary role in determining the strong lensing bias.
+{\bf 
+A selection aimed at specific kinds of sources, such as blue star-forming galaxies or quasars, results primarily in different numbers of lenses being discovered, but does not change how biased the lenses are with respect to the general galaxy population.
+}
 
@@
 If one wishes to directly account for the strong lensing bias, the formally correct procedure is to explicitly model all of the selection steps in a Bayesian hierarchical formalism, as explained by \citet{Son22}.
-Although this can be computationally challenging, machine learning can offer an efficient alternative \citep{Leg++22}.
+Although this can be computationally challenging, machine learning can offer an efficient alternative \citep{Leg++23}.
 In order for either of these approaches to work, however, it is essential that the lens selection procedure can be simulated.
@@
 Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in section \ref{ssec:lensdefext}.
-However, many previous studies aimed at simulating populations of lenses adopted a magnification-based definition, in which a lens-source system is classified as a strong lens if the magnification of the source is larger than a minimum value \citep[see e.g.][]{Col15, Rob++20}.
+However, many previous studies aimed at simulating populations of lenses adopted a magnification-based definition, in which a lens-source system is classified as a strong lens if the magnification of the source is larger than a minimum value \citep[see e.g.][]{Col15, Robertson++20}.
 In order to check whether our results are robust to the choice of definition of a strong lensing event, we re-analysed the fiducial sample with this alternative definition.
@@
 
+{\bf
 In the time-delay study of \citet{Bir++20}, lenses are described with a mass density profile given by \Eref{eq:mstprofile}.
@@
 where $\psi''_{\mathrm{PL}}$ is the value of $\psi''$ of the pure power-law lens model.
+}
 

### 2023-07-13 paper/seleff.tex
resubmitted version

@@
 The bias has little dependence on the properties of the source population: samples of galaxy-galaxy lenses and galaxy-quasar lenses that probe the same Einstein radius distribution are biased in a very similar way.
-%Quadruply imaged quasar lenses, however, are biased towards higher ellipticity galaxies.
-%{\bf
-%The bias on time-delay measurements of $H_0$ due to the use of prior information from galaxy-galaxy lenses is on the order of $1\%$.
-%}
 }
@@
 Given current uncertainties, strong lensing observations can be used directly to improve our current knowledge of the inner structure of galaxies, without the need to correct for selection effects.
-{\bf Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.}
-%{\bf Galaxy-galaxy lenses can be used to inform the properties of quasar lenses for the purpose of measuring $H_0$ from time delays
+bf Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.
 }
@@
 
-{\bf
 Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
@@
 
-%First, the projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}, a condition that excludes objects with a very diffuse mass distribution from the population of lenses.
-%Second, the source must be sufficiently close in projection to the lens center, in a way that depends on the lens mass distribution.
-Galaxies with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples.
-Third, at fixed lens properties, sources that
-
 Condition 2 sets limits on the observable properties of the sources and, again, the lens mass distribution, in a way that depends on the survey characteristics.
@@
 
-%A necessary condition for a galaxy to act as a strong lens with respect to a given source is that its projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
-%This condition excludes objects with a very diffuse mass distribution from the population of lenses.
-%In general, galaxies with a higher concentration of mass are more likely to be strong lenses, and are therefore over-represented in lens samples.
-%The distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies: for instance, lensing magnification allows the detection of fainter objects with respect to the field.
-
 The cumulative effect of these conditions can be summarised as follows.
-}
-%In general, 
 The probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
@@
 
-{\bf Finally, we considered galaxy-quasar lenses. 
+Finally, we considered galaxy-quasar lenses. 
 The strong lensing bias of lensed quasars was studied by \citep{C+C16}, in the context of power-law lens mass models with external convergence.
 Here we addressed the question of how different are galaxy-quasar lenses from galaxy-galaxy lens samples. 
-}
-%Finally, we addressed the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses. 
-This %last% 
-point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
+This point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
 Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{Bir++20,B+T21}, but any difference between the two lens classes can introduce biases, if not corrected for.
@@
 
-%\subsubsection{Lensed galaxies}
-
 For the simulation with extended sources, we produced images with properties similar to those expected for the Euclid Wide survey \citep{Sca++22}.
@@
 
-{\bf The first four rows of \Fref{fig:quasarbias} show} the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
+The first four rows of \Fref{fig:quasarbias} show the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
 The biases of the population of lensed quasars (red curves) are very similar to that of lensed galaxies (blue curves), especially for values of $\theta_{\mathrm{Ein},min}<1''$, 
@@
 This is a well-known effect \citep[see e.g.][]{KKS97} which can be explained in the context of the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
-%This trend was expected, given the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
 Quad lenses tend to also have a slightly larger halo mass at fixed stellar mass, for $\theta_{\mathrm{Ein},min}<1''$.
@@
 
-{\bf
 A measurement of the time delay between the images of a lensed quasar can be converted into an estimate of the expansion rate of the Universe, $H_0$. 
-%, provided that the mass distribution of the lens galaxy (and of matter along the line-of-sight) is accurately known.
 In general, though, the value of $H_0$ is degenerate with the mass distribution of the lens (and of matter along the line-of-sight), which can be difficult to constrain on an individual lens basis.
@@
 Strong lensing data can constrain very well the power-law index $\gamma$, but not at all $\lmst$.
-%, on which the inference on the Hubble constant depends.
 Given the time delay, the inferred value of the Hubble constant $H_0$ scales linearly with $1/\lmst$.
@@
 This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in \Fref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines).
-}
 
@@
 Strictly speaking, the results shown in \Sref{sect:results} apply only to samples of lenses and background sources with the same properties as our simulations.
-{\bf Real lens samples can have more complex source distributions.
+Real lens samples can have more complex source distributions.
 For example, many existing lens-finding algorithms are tuned to identify sources with different colours from the lenses, introducing a selection that can modify the source distribution in redshift-magnitude space.
-}
 However, the comparison of section \ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
@@
 We can then conclude that the details of the properties of the background source population play a secondary role in determining the strong lensing bias.
-{\bf 
 A selection aimed at specific kinds of sources, such as blue star-forming galaxies or quasars, results primarily in different numbers of lenses being discovered, but does not change how biased the lenses are with respect to the general galaxy population.
-}
 
@@
 
-{\bf
 The simulations described in \Sref{sect:lenspop}, on which our experiment is based, are examples of so-called lens-selected surveys:
@@
 An alternative approach is to build a source-selected sample, by first searching for signatures of strong lensing within a population of possible background sources and then obtaining information on the lens galaxies \citep[see e.g.][]{Hez++13}.
-In principle, these two approaches lead to different strong lensing biases. %Therefore, strictly speaking, our results only apply to a lens-selected sample.
-%For example, from \Fref{fig:lenspars} we can see that the distribution in redshift and stellar mass of the lenses extends beyond the bounds applied to the population of foreground galaxies: in principle, a source-based selection method
+In principle, these two approaches lead to different strong lensing biases. 
 For example, from \Fref{fig:lenspars} we can see that the stellar mass distribution of the lenses extends below the minimum mass cut that we applied to the population of foregrounds galaxies: a source-based search that is able detect those lenses would result in a different sample and a different associated strong lensing bias.
@@
 
-%In general, speaking of a lens- or source-selected sample is improper.
 In general, there is no such thing as a purely lens- or source-selected sample. 
@@
 A source-based search is also lens-based, because the lens properties determine important observational features of the lensed source, such as the number of images, the image separation and the total flux.
-}
 
@@
 
-{\bf
 Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in section \ref{ssec:lensdefext}.
@@
 In order to check whether our results are robust to the choice of definition of a strong lensing event, we re-analysed the fiducial sample with this alternative definition. 
-%repeated the analysis of the fiducial sample with extended sources.
 In particular, we defined as strong lenses all systems with a detected background source and $\mu_{\mathrm{tot}} > 3$, where $\mu_{\mathrm{tot}}$ is the total magnification (obtained by integrating the flux of all of the lensed images and dividing it by the intrinsic source flux).
 The total number of strong lenses increased by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
-\Fref{fig:minmubias} shows the strong lensing bias on a few selected quantities (stellar population synthesis mismatch parameter, projected dark matter mass within $5$~kpc, source magnitude and source half-light radius).
-Besides the change in the source magnitude (third panel), we see very little differences in the bias with respect to the peak-based definition of the fiducial analysis.
+\Fref{fig:minmulbias} and \Fref{fig:minmusbias} show the strong lensing bias on a series of lens- and source-related properties.
+Besides the change in the source redshift and magnitude (first two panels in \Fref{fig:minmusbias}), we see very little differences in the bias with respect to the peak-based definition of the fiducial analysis.
 We then conclude that the exact definition of a strong lensing event does not have a big impact on the strong lensing bias of lens-related properties.
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{minmu_bias.eps}
-\caption{
-Strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves).
+\includegraphics[width=\columnwidth]{minmu_lens_bias.eps}
+\caption{
+Lens strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves).
 First panel: median $\log{\asps}$.
 Second panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
-Third panel: median source magnitude.
-Fourth panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
-}
-\label{fig:minmubias}
+Third panel: median $\gammadm$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
+Fourth panel: median axis ratio.
+Fifth panel: median mass-sheet transformation parameter, defined in Appendix~\ref{sect:appendixc}.
+In each panel, the dashed line indicates the value of the parent population.
+}
+\label{fig:minmulbias}
 \end{figure}
-}
-
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{minmu_source_bias.eps}
+\caption{
+Source strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves). 
+First panel: median source redshift.
+Second panel: median source magnitude.
+Third panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
+The dashed line indicates the median value of the population of sources detectable in the absence of lensing.
+}
+\label{fig:minmusbias}
+\end{figure}
 %__________________________________________________________________
@@
 This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
-{\bf
 The associated bias on the inference of $H_0$ is on the order of $1\%$.
-}
 \item Samples of quad lenses are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is also biased.
@@
 
-{\bf
 In the time-delay study of \citet{Bir++20}, lenses are described with a mass density profile given by \Eref{eq:mstprofile}.
@@
 \end{equation}
-%\citet{Bir++20} fixed the slope $\gamma$ of the power-law component to the value obtained by fitting a pure power-law mass model to the strong lensing data.
 The mass-sheet transformation parameter $\lmst$ can be interpreted as a quantity describing a departure from a pure power-law profile.
@@
 where $\psi''_{\mathrm{PL}}$ is the value of $\psi''$ of the pure power-law lens model.
-}
 

### 2023-07-13 lens_population/plot_minmu_lens_bias.py

@@ -1,0 +1,195 @@
+import numpy as np
+import pylab
+import h5py
+from simpars import lmobs_piv, lreff_piv
+from scipy.optimize import leastsq
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 20
+
+smag_min = 24.8
+smag_max = 25.2
+
+lpopnames = ['fiducial_1000sqdeg_lenses.hdf5', 'fiducial_1000sqdeg_minmulenses.hdf5']
+labels = ['Fiducial', '$\mu_{\mathrm{tot}} > 3$']
+lensdefs = ['islens', 'minmulens']
+teindefs = ['tein_zs', 'minmutein_zs']
+nsims = len(lpopnames)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+colors = [colseq[0], colseq[5]]
+linestyles = ['-', '--']
+
+fig, ax = pylab.subplots(5, 1, figsize=(6, 14))
+
+pylab.subplots_adjust(left=0.23, right=1.00, bottom=0.05, top=1., wspace=0., hspace=0.)
+
+ntein = 21
+tein_arr = np.linspace(0., 2., ntein)
+
+detectable = h5py.File('detectable_sources.hdf5', 'r')
+
+zmed_det = np.median(detectable['zs'][()])
+smagmed_det = np.median(detectable['smag'][()])
+
+smagbin_det = (detectable['smag'][()] > smag_min) & (detectable['smag'][()] < smag_max)
+
+nsermed_det = np.median(detectable['nser'][smagbin_det])
+sreffmed_det = np.median(detectable['sreff'][smagbin_det])
+
+qmed_det = np.median(detectable['sq'][()])
+
+galpop = h5py.File('fiducial_1000sqdeg_galaxies.hdf5', 'r')
+laspsmed_gal = np.median(galpop['lasps'][()])
+qmed_gal = np.median(galpop['q'][()])
+
+for n in range(nsims):
+
+    lenspop = h5py.File(lpopnames[n], 'r')
+
+    laspsmed_arr = np.zeros(ntein)
+    laspserr_arr = np.zeros(ntein)
+
+    lmdm5med_arr = np.zeros(ntein)
+    lmdm5err_arr = np.zeros(ntein)
+
[…]

### 2023-07-13 lens_population/plot_minmu_source_bias.py

@@ -1,0 +1,135 @@
+import numpy as np
+import pylab
+import h5py
+from simpars import lmobs_piv, lreff_piv
+from scipy.optimize import leastsq
+from matplotlib.ticker import MultipleLocator
+from matplotlib.colors import ListedColormap
+from matplotlib import cm
+from matplotlib import rc
+rc('text', usetex=True)
+
+
+fsize = 20
+
+smag_min = 24.8
+smag_max = 25.2
+
+lpopnames = ['fiducial_1000sqdeg_lenses.hdf5', 'fiducial_1000sqdeg_minmulenses.hdf5']
+labels = ['Fiducial', '$\mu_{\mathrm{tot}} > 3$']
+lensdefs = ['islens', 'minmulens']
+teindefs = ['tein_zs', 'minmutein_zs']
+nsims = len(lpopnames)
+
+colseq = pylab.rcParams['axes.prop_cycle'].by_key()['color']
+colors = [colseq[0], colseq[5]]
+linestyles = ['-', '--']
+
+fig, ax = pylab.subplots(3, 1, figsize=(6, 10))
+
+pylab.subplots_adjust(left=0.22, right=1.00, bottom=0.07, top=1., wspace=0., hspace=0.)
+
+ntein = 21
+tein_arr = np.linspace(0., 2., ntein)
+
+detectable = h5py.File('detectable_sources.hdf5', 'r')
+
+zmed_det = np.median(detectable['zs'][()])
+smagmed_det = np.median(detectable['smag'][()])
+
+smagbin_det = (detectable['smag'][()] > smag_min) & (detectable['smag'][()] < smag_max)
+
+nsermed_det = np.median(detectable['nser'][smagbin_det])
+sreffmed_det = np.median(detectable['sreff'][smagbin_det])
+
+qmed_det = np.median(detectable['sq'][()])
+
+galpop = h5py.File('fiducial_1000sqdeg_galaxies.hdf5', 'r')
+laspsmed_gal = np.median(galpop['lasps'][()])
+
+for n in range(nsims):
+
+    laspsmed_arr = np.zeros(ntein)
+    laspserr_arr = np.zeros(ntein)
+
+    lmdm5med_arr = np.zeros(ntein)
+    lmdm5err_arr = np.zeros(ntein)
+
+    def mdm5fitfunc(p):
[…]

### 2023-07-25 paper/seleff.tex
accepted version

@@
 Given current uncertainties, strong lensing observations can be used directly to improve our current knowledge of the inner structure of galaxies, without the need to correct for selection effects.
-bf Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.
+Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.
 }
@@
 \item The multiple images must be detected and resolved
-\item The system must be recognized as a lens by the survey.
+\item The system must be recognised as a lens by the survey.
 \end{enumerate}
@@
 In general, though, the value of $H_0$ is degenerate with the mass distribution of the lens (and of matter along the line-of-sight), which can be difficult to constrain on an individual lens basis.
-One possible strategy to break this denegeracy is to use information gathered from a larger set of lenses as a prior for the time-delay lenses.
+One possible strategy to break this degeneracy is to use information gathered from a larger set of lenses as a prior for the time-delay lenses.
 This was the strategy adopted by \citet{Bir++20} in the most recent measurement of $H_0$ by the TDCOSMO collaboration: they combined a larger sample of extended source lenses with a smaller set of time-delay lenses.
@@
 The mass-sheet transformation parameter $\lmst$ can be interpreted as a quantity describing a departure from a pure power-law profile.
-One possible way of estimating $\lmst$ for our lenses is to fit a power-law to the simulated images and then optimize for the value of $\lmst$ that best matches the true density profile.
+One possible way of estimating $\lmst$ for our lenses is to fit a power-law to the simulated images and then optimise for the value of $\lmst$ that best matches the true density profile.
 That, however, would be too time-consuming.

### 2024-03-10 paper/seleff.tex
italics Euclid

@@
 We first studied how the strong lensing cross-section varies as a function of lens and source properties. 
-Then, we simulated strong lensing surveys with data similar to that expected for Euclid and measured the strong lensing bias in different scenarios.
+Then, we simulated strong lensing surveys with data similar to that expected for \textit{Euclid} and measured the strong lensing bias in different scenarios.
 We focused particularly on two quantities: the stellar population synthesis mismatch parameter, $\asps$, defined as the ratio between the true stellar mass of a galaxy and the stellar mass obtained from photometry, and the central dark matter mass at fixed stellar mass and size.
@@
    {
-Strong lens galaxies are biased towards larger stellar masses, smaller half-mass radii and larger dark matter masses.
+Strong lens galaxies are biased towards higher stellar masses, smaller half-mass radii, and higher dark matter masses.
 The amplitude of the bias depends on the intrinsic scatter in the mass-related parameters of the galaxy population and on the completeness in Einstein radius of the lens sample.
@@
 Given current uncertainties, strong lensing observations can be used directly to improve our current knowledge of the inner structure of galaxies, without the need to correct for selection effects.
-Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.
+Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radii.
 }
@@
 Strong gravitational lensing is a powerful tool for studying galaxies and cosmology.
-Strong lenses have been used to probe the mass structure of massive galaxies \citep{Aug++10, ORF14, Son++15, Sha++21}, to detect substructure \citep{Veg++12, Hez++16, Nie++20}, to carry out detailed studies of magnified star-forming galaxies \citep{Jon++13}, and to measure the expansion rate of the universe with time delays \citep[see][ for a review]{T+M16}.
+Strong lenses have been used to probe the mass structure of massive galaxies \citep{Aug++10, ORF14, Son++15, Sha++21}, to detect substructure \citep{Veg++12, Hez++16, Nie++20}, to carry out detailed studies of magnified star-forming galaxies \citep{Jon++13}, and to measure the expansion rate of the Universe with time delays \citep[see][ for a review]{T+M16}.
 
 Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
-In order for a lens-source pair to be included in a strong lensing survey, the following conditions must be met:
-\begin{enumerate}
-\item The lens must produce multiple images of the source
-\item The multiple images must be detected and resolved
-\item The system must be recognised as a lens by the survey.
-\end{enumerate}
+In order for a lens-source pair to be included in a strong lensing survey, the following conditions must be met: (1) the lens must produce multiple images of the source, (2) the multiple images must be detected and resolved, and (3) the system must be recognised as a lens by the survey.
+
 Condition 1 poses constraints on the mass distribution of the lens, as well as on the geometry of the system. 
-First of all, not all lenses are capable of producing multiple images, but only those whose surface mass density $\Sigma(\boldsymbol\theta)$ is larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
+First of all, not all lenses are capable of producing multiple images, only those whose surface mass density, $\Sigma(\boldsymbol\theta)$, is higher than the critical surface mass density for lensing, $\Sigma_{\mathrm{cr}}$, at at least one position, $\boldsymbol\theta$ \citep{SEF92}.
 This condition excludes objects with a very diffuse mass distribution from the population of lenses.
-Second, objects with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples.
-Third, at fixed lens properties, sources that are located further away along the line of sight are more likely to be lensed.
+Second, objects with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged and are therefore over-represented in lens samples.
+Third, with fixed lens properties, sources that are located farther away along the line of sight are more likely to be lensed.
 
@@
 Most commonly, lenses with a smaller image separation are more difficult to identify even if detected and resolved, because of contamination from the lens galaxy.
-This introduces an additional bias towards lenses with a larger or more concentrated mass.
+This introduces an additional bias towards lenses with a higher or more concentrated mass.
 
 The cumulative effect of these conditions can be summarised as follows.
-The probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
+The probability distribution, $\prsl$, of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}
 \begin{equation}\label{eq:one}
@@
 \end{equation}
-In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\psel$ describes the probability of selecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
-This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be detected and recognised as a strong lens.
-
-The left-hand side of \Eref{eq:one} is directly accessible from strong lensing observations.
+In this equation, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\psel$ describes the probability of selecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion, $S$, used to define a lens.
+This last factor takes both physical and survey selection effects into account, that is, whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and whether such an image can be detected and recognised as a strong lens.
+
+The information needed for the left-hand side of Eq.
+\ref{eq:one} is directly accessible from strong lensing observations.
 If the main goal of a lensing survey is to characterise the properties of the strong lens population, then it can be accomplished by directly analysing this term.
-For many applications of strong lensing, however, the aim is to constrain the properties of the general galaxy or source population, $\prlens$ and $\prsource$, which are coupled in a nontrivial way via the lens selection probability $\psel$.
-In order to obtain an unbiased estimate of either $\prlens$ or $\prsource$, then, it is necessary to invert \Eref{eq:one}.
-In principle, this can be done with a Bayesian hierarchical formalism \citep{Son22}, but knowledge of the lens selection probability $\psel$ is required.
+For many applications of strong lensing, however, the aim is to constrain the properties of the general galaxy or source population, $\prlens$ and $\prsource$, which are coupled in a non-trivial way via the lens selection probability, $\psel$.
+In order to obtain an unbiased estimate of either $\prlens$ or $\prsource$, it is necessary to invert Eq. \ref{eq:one}.
+In principle, this can be done with a Bayesian hierarchical formalism \citep{Son22}, but knowledge of the lens selection probability, $\psel$, is required.
 This factor can be written as the following product:
@@
 \end{equation}
-where $\pdet$ is the probability of detecting a strong lensing event, and $\pfind(\psilens,\psisource|S,\mathrm{det})$ is the probability of correctly classifying it as such\footnote{\citet{Son22} implicitly assumed $\pfind\equiv1$, therefore $\pdet$ and $\psel$ can be used interchangeably in the context of that work.}.
-The detection probability $\pdet$ can be obtained via simulation.
-The main challenge is characterising $\pfind$: in most of the existing strong lensing surveys, the process of determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a nontrivial visual selection step.
-
-For the above reasons, the problem of inverting \Eref{eq:one} is a difficult one to tackle exactly.
-Few studies have attempted to explicitly account for strong lensing selection effects, usually by making ad-hoc simplifying assumptions \citep{Son++15,O+A17,Son++19}.
-Whether it is necessary to invert \Eref{eq:one}, however, depends on the severity of the strong lensing bias that needs to be corrected and on the accuracy requirements on the key quantities of interest.
+where $\pdet$ is the probability of detecting a strong lensing event, and $\pfind(\psilens,\psisource|S,\mathrm{det})$ is the probability of correctly classifying it as such\footnote{\citet{Son22} implicitly assumed $\pfind\equiv1$; therefore, $\pdet$ and $\psel$ can be used interchangeably in the context of that work.}.
+The detection probability, $\pdet$, can be obtained via simulations.
+The main challenge is characterising $\pfind$: in most of the existing strong lensing surveys, the process of determining whether a system is included in a strong lens sample is typically a combination of several cuts, usually involving a non-trivial visual selection step.
+
+For the above reasons, the problem of inverting Eq. \ref{eq:one} is a difficult one to tackle exactly.
+A few studies have attempted to explicitly account for strong lensing selection effects, usually by making ad hoc simplifying assumptions \citep{Son++15,O+A17,Son++19}.
+Whether it is necessary to invert Eq.
+\ref{eq:one}, however, depends on the severity of the strong lensing bias that needs to be corrected and on the accuracy requirements for the key quantities of interest.
 
 In this paper we aim to quantify the strength of the strong lensing bias on a series of foreground galaxy and background source parameters. 
-In particular, we aim to determine how strong lenses differ from the parent population of foreground galaxies and background sources in terms of a) the radial mass structure of the lenses (i.e.
-their stellar and dark matter mass density profiles); b) the ellipticity of the lenses; c) the size-magnitude distribution of the lensed sources.
-The answer to this question depends on 1) how the lens detection probability $\pdet$ varies as a function of galaxy and source properties, 2) the efficiency of a survey at correctly classifying detected strong lenses (i.e.
-$\pfind$), and on 3) the shape of the galaxy and source parameters distribution $\prlens$ and $\prsource$.
+In particular, we aim to determine how strong lenses differ from the parent population of foreground galaxies and background sources in terms of (a) the radial mass structure of the lenses (i.e.
+their stellar and dark matter mass density profiles), (b) the ellipticity of the lenses, and (c) the size-magnitude distribution of the lensed sources.
+This determination depends on (1) how the lens detection probability, $\pdet$, varies as a function of galaxy and source properties, (2) the efficiency of a survey in correctly classifying detected strong lenses (i.e.
+$\pfind$), and (3) the shape of the galaxy and source parameter distribution, $\prlens$ and $\prsource$.
 To understand this third point we can imagine the limiting case in which both $\prlens$ and $\prsource$ are Dirac delta functions (i.e.
-all lenses and sources are identical): in this limit, $\prsl$ simply reduces to the product $\prlens\prsource$ up to a scaling constant, corresponding to a case in which the lensing bias is none.
-
-\citet{MVK09} carried out a thorough study of point 1): they quantified how the properties of a lens galaxy determine its probability of creating a lensing event with a point source.
-In this work we revisited the \citet{MVK09} study, expanding it to the extended source case.
+all lenses and sources are identical): in this limit, $\prsl$ simply reduces to the product $\prlens\prsource$ up to a scaling constant; this corresponds to a case in which there is no lensing bias.
+
+\citet{MVK09} carried out a thorough study of point (1): they quantified how the properties of a lens galaxy determine its probability of creating a lensing event with a point source.
+In this work we revisit the \citet{MVK09} study, expanding it to the extended source case.
 We simulated individual lenses and examined how the lens detection probability varies with lens and source properties.
-In addition, we addressed points 2) and 3) as well: we simulated large populations of strong lenses using empirical models, we simulated the lens detection and finding phase, and quantified the lensing bias under various scenarios.
-We explored how the results change as a function of the efficiency of a lens survey at discovering small image separation lenses, and of the scatter in mass parameters at fixed light.
+In addition, we address points (2) and (3): we simulated large populations of strong lenses using empirical models, we simulated the lens detection and finding phase, and we quantified the lensing bias under various scenarios.
+We explored how the results change as a function of the efficiency of a lens survey in discovering small image separation lenses, and as a function of the scatter in mass parameters at fixed light.
 
 Finally, we considered galaxy-quasar lenses. 
-The strong lensing bias of lensed quasars was studied by \citep{C+C16}, in the context of power-law lens mass models with external convergence.
-Here we addressed the question of how different are galaxy-quasar lenses from galaxy-galaxy lens samples. 
-This point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
-Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{Bir++20,B+T21}, but any difference between the two lens classes can introduce biases, if not corrected for.
+The strong lensing bias of lensed quasars was studied by \cite{C+C16} in the context of power-law lens mass models with external convergence.
+Here we address the question of how different galaxy-quasar lenses are from galaxy-galaxy lens samples. 
+This point is relevant for time-delay cosmography studies, in which measurements of the time delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe.
+Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{Bir++20,B+T21}, but any difference between the two lens classes can introduce biases if not corrected for.
 With this study we aim to quantify this bias.
 
-The structure of this work is the following.
-In \Sref{sect:basics} we introduce the basics of gravitational lensing.
-In \Sref{sect:indlenses} we study individual lens systems. 
-In \Sref{sect:lenspop} we describe our simulations of lens surveys.
-In \Sref{sect:results} we show the results of our analysis on the simulated lens survey data.
-We discuss the results in \Sref{sect:discuss} and draw conclusions in \Sref{sect:concl}.
-
-The Python code used for the simulation and analysis of the lens sample can be found in a dedicated section of a GitHub repository\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/papers/selection_effects}}.
+The structure of this paper is as follows.
+In Sect. \ref{sect:basics} we introduce the basics of gravitational lensing.
+In Sect. \ref{sect:indlenses} we study individual lens systems. 
+In Sect. \ref{sect:lenspop} we describe our simulations of lens surveys.
+In Sect. \ref{sect:results} we show the results of our analysis of the simulated lens survey data.
+We discuss the results in Sect. \ref{sect:discuss} and draw conclusions in Sect. \ref{sect:concl}.
+
+The Python code used for the simulation and analysis of the lens sample can be found in a dedicated section of a GitHub repository\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/tree/main/papers/selection_effects}}.
 
@@
 
-The lensing properties of an object with respect to a source depend solely on its dimensionless surface mass density distribution $\kappa(\boldsymbol\theta)$ (also referred to as the convergence). 
+The lensing properties of an object with respect to a source depend solely on its dimensionless surface mass density distribution, $\kappa(\boldsymbol\theta)$ (also referred to as the convergence). 
 This is the ratio between the surface mass density and the critical surface mass density for lensing:
@@
 In the special case of an axisymmetric lens we can simplify the notation by considering a single coordinate axis with origin at the centre of the lens.
-We label the components of the image position, source position and deflection angle along this axis as $\theta$, $\beta$ and $\alpha$, respectively.
+We label the components of the image position, source position and deflection angle along this axis as $\theta$, $\beta$, and $\alpha$, respectively.
 The lens equation for an axisymmetric lens then becomes
@@
 A very important quantity for describing the strength of a strong lens is the Einstein radius, $\tein$.
-For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:1dlenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
+For an axisymmetric lens, this is the radius corresponding to the solution of Eq.
+\ref{eq:1dlenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
 The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form there have infinite magnification in the tangential direction.
@@
 typically produce either one or three images of a point source.
-\Fref{fig:scheme} helps to visualise this property.
+\Fref{fig:scheme} helps visualise this property.
 Plotted in \Fref{fig:scheme} is the quantity $\theta - \alpha(\theta)$ as a function of the position in the image plane $\theta$, for a few lens models.
-According to \Eref{eq:1dlenseq}, images form at the locations where this quantity equals the position of the source. 
+According to Eq. \ref{eq:1dlenseq}, images form at the locations where this quantity equals the position of the source. 
 Therefore, given a source position $\beta$, the number of images and their location can be determined by drawing a horizontal line at the value $\beta$ on the vertical axis, and finding the points where this line intersects the $\theta - \alpha(\theta)$ curve.
@@
 Images that form there have a formally infinite magnification in the radial direction\footnote{The slope of the $\theta - \alpha(\theta)$ curve is the inverse of the radial magnification.
-This can be understood by taking the derivative of \Eref{eq:1dlenseq} with respect to $\theta$.}.
+This can be understood by taking the derivative of Eq.
+\ref{eq:1dlenseq} with respect to $\theta$.}.
 
@@
 \caption{
-The lens equation of axisymmetric lenses. Solid lines: right-hand side of \Eref{eq:1dlenseq} for five lenses with different density profiles. 
+Lens equation of axisymmetric lenses. Solid lines represent the right-hand side of Eq. \ref{eq:1dlenseq} for five lenses with different density profiles. 
 Given a source at position $\beta$, its lensed images form at the values of $\theta$ where the solid line intersects a horizontal line located at $\beta$ on the vertical axis. 
-Dotted lines: positions of the radial caustics. Sources located within the radial caustic of a given lens produce three images.
-The lens models used in this simulations consist of a stellar component and a dark matter halo, as described in section \ref{ssec:profile}. Their density profile is plotted in \Fref{fig:kappa}.
+Dotted lines represent positions of the radial caustics. Sources located within the radial caustic of a given lens produce three images.
+The lens models used in this simulation consist of a stellar component and a dark matter halo, as described in Sect. \ref{ssec:profile}. Their density profile is plotted in Fig. \ref{fig:kappa}.
 \label{fig:scheme}
@@
 Point sources located outside the caustic are not strongly lensed.
-Sources that lie in the region enclosed between the two caustics produce three images, while sources inside the tangential caustic produce five images.
-The lens model adopted for this experiment consists of a stellar component and a dark matter halo, as described in \ref{ssec:profile}. The two components have the same ellipticity.
+Sources that lie in the region enclosed by the two caustics produce three images, while sources inside the tangential caustic produce five images.
+The lens model adopted for this experiment consists of a stellar component and a dark matter halo, as described in Sect. \ref{ssec:profile}. The two components have the same ellipticity.
 \label{fig:ellcaust}
@@
 
-\subsection{Lensing event definition: point sources}\label{ssec:lensdefpoint}
+\subsection{Lensing event definition: Point sources}\label{ssec:lensdefpoint}
 
@@
 
-\subsection{Lensing event definition: extended sources}\label{ssec:lensdefext}
+\subsection{Lensing event definition: Extended sources}\label{ssec:lensdefext}
 
@@
 In those cases it is difficult to establish whether the arc is a set of blending images or not, and the decision of including such systems in a strong lens sample is often arbitrary.
-Here we adopt the following working definition: {\em an extended source is strongly lensed if the number of surface brightness peaks that are detected is larger than its intrinsic number of peaks in the absence of lensing}.
+Here we adopt the following working definition: an extended source is strongly lensed if the number of surface brightness peaks that are detected is larger than its intrinsic number of peaks in the absence of lensing.
 
@@
 We point out that, although our lens definition relies on peak detections, it is not necessary for the peak of the source surface brightness to lie within the caustics in order for the source to be strongly lensed.
-This can be observed in the fourth example of \Fref{fig:lensdef}: although the source centroid lies outside of the caustic (as shown in the first column), the outskirts of the source overlap with the lens, and therefore multiple images are produced.
+This can be observed in the fourth example of Fig.
+\ref{fig:lensdef}: although the source centroid lies outside of the caustic (as shown in the first column), the outskirts of the source overlap with the lens, and therefore multiple images are produced.
 
@@
 \caption{
-Criterion used to classify lensed images of extended sources. Six examples.
-First column: caustics (red curves) and source position (blue circle).
+Criterion used to classify lensed images of extended sources. Six examples are shown.
+First column: Caustics (red curves) and source position (blue circle).
 The radius of the circle corresponds to the radius at which the surface brightness is equal to $2\sigma$ the sky background rms fluctuation for a single pixel.
 In other words, the blue circle delimits the area of the source that can be detected.
-Second column: mock image of the lensed source, with added noise.
-Third column: footprint of the source. The purple footprint is obtained with a $2\sigma$ detection criterion. The coral region is the largest footprint with the highest number of detected images.
+Second column: Mock image of the lensed source, with added noise.
+Third column: Footprint of the source. The purple footprint is obtained with a $2\sigma$ detection criterion. The coral region is the largest footprint with the highest number of detected images.
 \label{fig:lensdef}
@@
 Most, if not all, of the galaxy-scale strong lenses known belong to these two categories.
-However, there is a third regime in strong lensing, corresponding to the case in which the source size is much bigger than the caustics, such as in the fifth example of \Fref{fig:lensdef}.
+However, there is a third regime in strong lensing, corresponding to the case in which the source size is much bigger than the caustics, such as in the fifth example of Fig. \ref{fig:lensdef}.
 In this regime, the overall size and total flux of the source are roughly preserved, and the lens produces only a relatively minor perturbation on a localised region of the image.
@@
 In order to do so, it is useful to introduce the concept of strong lensing cross-section.
-Given a foreground galaxy with parameters $\psilens$, a background source with parameters $\psisource$, and a criterion $S$ to define a strong lensing event, the strong lensing cross-section is defined as \citep{Son22}:
+Given a foreground galaxy with parameters $\psilens$, a background source with parameters $\psisource$, and a criterion $S$ to define a strong lensing event, the strong lensing cross-section is defined as \citep{Son22}
 \begin{equation}\label{eq:crosssect}
@@
 In the limit of low density of background sources, which is satisfied in all practical cases, the probability of a lensing event is proportional to $\crosssect$.
-The lensing cross-section defined via \Eref{eq:crosssect} depends solely on the lens detection probability $\pdet$ and does not take into account whether the lens can be correctly classified by a lens finder.
-This separate selection step is captured by the term $\pfind$ in \Eref{eq:two}.
+The lensing cross-section defined via Eq.
+\ref{eq:crosssect} depends solely on the lens detection probability $\pdet$ and does not take into account whether the lens can be correctly classified by a lens finder.
+This separate selection step is captured by the term $\pfind$ in Eq. \ref{eq:two}.
 In this section we consider exclusively the detectability of a lens, and therefore focus only on $\crosssect$.
-In \Sref{sect:lenspop}, when considering specific lens survey simulations, we introduce $\pfind$.
-
-We compute $\crosssect$ in a series of different scenarios of increasing complexity.
-The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments. We describe this in section \ref{ssec:profile}.
-In section \ref{ssec:axisymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
-In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.
+In Sect.
+\ref{sect:lenspop}, when considering specific lens survey simulations, we introduce $\pfind$.
+
+We computed $\crosssect$ in a series of different scenarios of increasing complexity.
+The model family adopted to describe the radial density profile of the lenses was the same in all of our experiments. We describe this in Sect. \ref{ssec:profile}.
+In Sect. \ref{ssec:axisymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
+In Sect. \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in Sect. \ref{ssec:ellext} we replace point sources with extended sources.
 
@@
 That is, we do not allow for the presence of gradients in the stellar mass-to-light ratio.
-In section \ref{ssec:limitations} we discuss qualitatively what the implications of relaxing this assumption would be.
-
-We describe the dark matter halo with a generalised Navarro, Frenk \& White (gNFW) profile.
+In Sect.
+\ref{ssec:limitations} we discuss qualitatively what the implications of relaxing this assumption would be.
+
+We describe the dark matter halo with a generalised Navarro, Frenk, and White (gNFW) profile.
 We first define it by its three-dimensional distribution, which for a spherically symmetric profile is
@@
 
-\Fref{fig:kappa} shows the dimensionless surface mass density profile of S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
+Figure \ref{fig:kappa} shows the dimensionless surface mass density profile of S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
 All of the profiles have 
 a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component (these assumptions are dropped later).
-Two main features emerge from \Fref{fig:kappa}.
+Two main features emerge from Fig.
+\ref{fig:kappa}.
 First, the baryons generally dominate the total density in the inner regions $(\theta < \tein)$, while the dark matter is the main component at large radii.
 Second, models with different dark matter fractions and inner dark matter slopes can conspire to produce very similar total density profiles. 
-This is the case, for example, of the ($\fdm=0.5,\gammadm=1.5$) and the ($\fdm=0.2,\gammadm=2.0$) models (green and blue lines in \Fref{fig:kappa}).
+This is the case, for example, of the ($\fdm=0.5,\gammadm=1.5$) and the ($\fdm=0.2,\gammadm=2.0$) models (green and blue lines in Fig. \ref{fig:kappa}).
 
@@
 Dimensionless surface mass density profile of S\'{e}rsic + gNFW composite models.
-The S\'{e}rsic index of the baryonic component is fixed to $n=4$ and the scale radius of the dark matter component is fixed to ten times the half-light radius.
+The S\'{e}rsic index of the baryonic component is fixed to $n=4,$ and the scale radius of the dark matter component is fixed to ten times the half-light radius.
 All of the profiles are normalised in such a way that the Einstein radius is equal to the half-light radius.
-Solid lines: total density profile. Dotted lines: dark matter density profile. Dashed lines: baryonic density profile.
-The blue, orange and green dashed lines are identical, as they correspond to profiles with the same fraction of baryonic mass within the half-light radius.
+Solid lines show the total density profile, dotted lines the dark matter density profile, and dashed lines the baryonic density profile.
+The dashed blue, orange, and green lines are identical, as they correspond to profiles with the same fraction of baryonic mass within the half-light radius.
 \label{fig:kappa}
@@
 
-Axisymmetric lenses with a density profile of the kind introduced in section \ref{ssec:profile} can produce either one or three images of a point source.
-This can be seen in \Fref{fig:scheme}, which shows the lens equation for various values of the dark matter fraction and the inner dark matter slope.
+Axisymmetric lenses with a density profile of the kind introduced in Sect. \ref{ssec:profile} can produce either one or three images of a point source.
+This can be seen in Fig. \ref{fig:scheme}, which shows the lens equation for various values of the dark matter fraction and the inner dark matter slope.
 All of the lenses shown in this figure have the same Einstein radius, which is equal in size to the half-mass radius of the stellar component.
-The radius of the radial caustic, marked by the dotted lines in \Fref{fig:scheme}, is a strong function of the lens properties: it is largest in lenses with a smaller dark matter fraction or steeper dark matter slopes.
+The radius of the radial caustic, marked by the dotted lines in Fig.
+\ref{fig:scheme}, is a strong function of the lens properties: it is largest in lenses with a smaller dark matter fraction or steeper dark matter slopes.
 As a result, the source plane area that is subject to strong lensing is an even stronger function of these properties, since it scales with the square of the radial caustic radius.
-In order to compute the lensing cross-section, however, we must take into account the magnification, because that determines whether multiple images can be detected or not.
-
-\Fref{fig:1dmag} shows the magnification of the secondary image as a function of source position.
+In order to compute the lensing cross-section, however, we must take the magnification into account  because it determines whether multiple images can be detected or not.
+
+Figure \ref{fig:1dmag} shows the magnification of the secondary image as a function of source position.
 The secondary image is the one located in the region between the radial and tangential critical curves, opposite to the source with respect to the lens centre.
 In most practical cases this is the second brightest image.
-As \Fref{fig:1dmag} shows, the magnification is very large for sources close to the lens centre (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
+As Fig.
+\ref{fig:1dmag} shows, the magnification is very large for sources close to the lens centre (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
 While for the model with $f_{\mathrm{DM}}=0.8$ (purple curve) the magnification is above unity everywhere, other lens models can produce highly de-magnified secondary images for large values of $\beta$.
@@
 Magnification of the secondary image as a function of the source position.
-The lenses are axisymmetric composite models. Their density profile is plotted in \Fref{fig:kappa}.
+The lenses are axisymmetric composite models. Their density profile is plotted in Fig. \ref{fig:kappa}.
 Vertical dotted lines mark the position of the radial caustic.
@@
 
-Using the definition of \Eref{eq:crosssect}, we computed the lensing cross-section of a set of axisymmetric lenses, with respect to point sources with different brightnesses.
+Using the definition of Eq. \ref{eq:crosssect}, we computed the lensing cross-section of a set of axisymmetric lenses, with respect to point sources with different brightnesses.
 In particular, we considered model lenses with fixed Einstein radius, with angular half-mass radius $\teff$ fixed to $\tein$, and varying values of the dark matter fraction and dark matter inner slope.
-The results are shown in \Fref{fig:fixedap_cs}.
+The results are shown in Fig. \ref{fig:fixedap_cs}.
 Each line corresponds to a source with a given intrinsic (i.e. unlensed) magnitude, $m_s$. The difference between this magnitude and the limiting magnitude of the survey is indicated as follows:
@@
 
-We can see a clear trend between $\Delta m$ and the lensing cross-section, in both panels of \Fref{fig:fixedap_cs}: $\crosssect$ is larger for brighter sources.
+We can see a clear trend between $\Delta m$ and the lensing cross-section, in both panels of Fig. \ref{fig:fixedap_cs}: $\crosssect$ is larger for brighter sources.
 The trend saturates below a certain $\Delta m$, for sufficiently small values of $\gammadm$ or for sufficiently large values of $\fdm$.
@@
 Strong lensing cross-section of an axisymmetric lens and a point source.
-The lens galaxy is a composite model, introduced in section \ref{ssec:profile}, with angular half-light radius equal to the Einstein radius.
+The lens galaxy is a composite model, introduced in Sect. \ref{ssec:profile}, with an angular half-light radius equal to the Einstein radius.
 The left panel shows the effect of varying the slope $\gammadm$ for fixed $\fdm$; conversely, the right panel shows variations as a function of $\fdm$ with $\gammadm$ fixed.
 The system is defined as a strong lens if at least two images are detected.
-Different lines correspond to the difference $\Delta m$ between the source magnitude and the survey detection limit for a point source.
+Different lines correspond to the difference, $\Delta m$, between the source magnitude and the survey detection limit for a point source.
 \label{fig:fixedap_cs}
@@
 At fixed source brightness, trends with the dark matter inner slope or dark matter fraction are generally weak.
-The lines of \Fref{fig:fixedap_cs}, however, have been computed by keeping the Einstein radius fixed while varying $\gammadm$ or $\fdm$.
-This is achieved by adjusting other properties of the lens, such as the total mass of the baryonic or the dark matter component (see \Fref{fig:kappa}). 
+The lines of Fig. \ref{fig:fixedap_cs}, however, have been computed by keeping the Einstein radius fixed while varying $\gammadm$ or $\fdm$.
+This is achieved by adjusting other properties of the lens, such as the total mass of the baryonic or the dark matter component (see Fig. \ref{fig:kappa}). 
 In practice, when varying one ingredient of the lens density profile, the Einstein radius varies in response.
 To get a complete view of how the lensing cross-section depends on different lens properties, we also computed how $\crosssect$ responds in absolute terms by varying one lens parameter at a time.
-\Fref{fig:phys_cs} shows $\crosssect$ as a function of stellar mass, half-light radius, inner dark matter slope, and projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$.
+Figure \ref{fig:phys_cs} shows $\crosssect$ as a function of stellar mass, half-light radius, inner dark matter slope, and projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$.
 
@@
 In each panel, only one property of the lens is varied, as indicated on the label of the horizontal axis.
-Each curve corresponds to a different value of $\Delta m$, in accordance with \Fref{fig:fixedap_cs}.
+Each curve corresponds to a different value of $\Delta m$, in accordance with Fig. \ref{fig:fixedap_cs}.
 The dashed line in each panel shows the quantity $\pi\tein^2$.
@@
 
-\Fref{fig:fixedap_cs} and \Fref{fig:phys_cs} show that the trends between lens properties and the strong lensing cross-section can be different for sources with different brightnesses.
+Figures \ref{fig:fixedap_cs} and \ref{fig:phys_cs} show that the trends between lens properties and the strong lensing cross-section can be different for sources with different brightnesses.
 The net effect in a strong lensing survey is an average over the source population, weighted by the source luminosity function. 
 This implies that surveys that target different families of sources, with different luminosity functions (for instance, galaxies or quasars), can have different strong lensing biases.
-We explore this possibility in \Sref{sect:results}.
+We explore this possibility in Sect.
+\ref{sect:results}.
 
@@
 In particular, we set $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\tein=\teff$, and set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
-This is the lens model used to produce the caustics plot of \Fref{fig:ellcaust}.
-\Fref{fig:ellcaust} suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
-Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.
+This is the lens model used to produce the caustics plot of Fig. \ref{fig:ellcaust}.
+Figure \ref{fig:ellcaust} suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
+Therefore, we expect the strong lensing cross-section to be a weak function of ellipticity.
 
 We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources randomly distributed over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification.
-We then measured the fraction of sources that are strongly lensed according to the criterion of \Eref{eq:pointdet} and multiplied this value by the area over which sources are located.
-The resulting $\crosssect$ is shown in \Fref{fig:ellpoint_cs} as a function of the minor-to-major axis ratio $q$.
+We then measured the fraction of sources that are strongly lensed according to the criterion of Eq.
+\ref{eq:pointdet} and multiplied this value by the area over which sources are located.
+The resulting $\crosssect$ is shown in Fig. \ref{fig:ellpoint_cs} as a function of the minor-to-major axis ratio, $q$.
 
@@
 Point source strong lensing cross-section as a function of lens axis ratio.
-Solid lines: cross-section based on the lens event definition of \Eref{eq:pointdet}.
-Dashed lines: cross-section based on the detection of four images (quad cross-section).
-Lines of different colour correspond to sources of different intrinsic magnitude.
+Solid lines show the cross-section based on the lens event definition of Eq. \ref{eq:pointdet} and dashed lines the cross-section based on the detection of four images (quad cross-section).
+Lines of different colours correspond to sources of different intrinsic magnitudes.
 The dashed blue line, corresponding to the brightest source magnitude, overlaps completely with the dashed orange line.
 The dashed purple line is zero: very faint sources cannot produce any detectable quads.
-The parameters of the lens density profile are $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\teff=\tein$.
+The parameters of the lens density profile are $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, and $\teff=\tein$.
 The baryonic and dark matter components have the same ellipticity and direction of the major axis.
@@
 For this reason we also computed an alternative lensing cross-section, in which the definition of a lensing event requires the detection of four images, instead of two.
-This is plotted in \Fref{fig:ellpoint_cs} with dashed lines.
+This is plotted in Fig. \ref{fig:ellpoint_cs} with dashed lines.
 The cross-section for quads is a strong function of lens ellipticity, for bright sources.
-This is a consequence of the fact that the area enclosed within the tangential caustic, which is where a source needs to be in order to produce four images, increases with increased lens ellipticity, as \Fref{fig:ellcaust} shows.
+This is a consequence of the fact that the area enclosed within the tangential caustic, which is where a source needs to be in order to produce four images, increases with increased lens ellipticity, as Fig.
+\ref{fig:ellcaust} shows.
 For sources that are intrinsically fainter than the detection limit, however, the quad cross-section is extremely small, regardless of ellipticity.
@@
 In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the source surface brightness distribution, with its radial profile, shape and orientation, and the PSF.
-Moreover, as we discussed in section \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
+Moreover, as we discussed in Sect. \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
 For this reason, we split our analysis into two parts.
@@
 a S\'{e}rsic model with $n=1$).
-We then took a lens model with the same parameters used in section \ref{ssec:ellpoint} and a minor-to-major axis ratio of $q=0.7$.
-We simulated a large number of images of extended sources with {\sc Glafic} and measured the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
-For the small source size experiment, we used pixels with a size equal to $1/20\tein$ and convolved the images with a Moffat PSF with a FWHM of two pixels and a $\beta$ parameter of $5.0$.
+We then took a lens model with the same parameters used in Sect. \ref{ssec:ellpoint} and a minor-to-major axis ratio of $q=0.7$.
+We simulated a large number of images of extended sources with {\sc Glafic} and measured the fraction of them that results in a strong lens, according to the definition of Sect. \ref{ssec:lensdefext}.
+For the small source size experiment, we used pixels with a size equal to $1/20\tein$ and convolved the images with a Moffat PSF with a full width at half maximum (FWHM) of two pixels and a $\beta$ parameter of $5.0$.
 We also assumed that the background noise is an uncorrelated Gaussian field.
 We carried out experiments with sources with different values of the total unlensed flux, $f$, and half-light radius, $\theta_{\mathrm{e,s}}$.
-The results are shown in \Fref{fig:extcs}.
+The results are shown in Fig. \ref{fig:extcs}.
 The total flux of the source, indicated in the legend, is measured in units of the sky background rms fluctuation within an area equal to the square of the Einstein radius.
@@
 \caption{
-Strong lensing cross-section of an extended source.
-Small source size regime ($\theta_{\mathrm{e,s}} < \tein$).
+Strong lensing cross-section of an extended source: small source size regime ($\theta_{\mathrm{e,s}} < \tein$).
 The cross-section is plotted as a function of the ratio between the half-light radius of the source and the Einstein radius of the lens.
-The lens model is fixed to be the same of \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
+The lens model is fixed to be the same as in Fig. \ref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
 Each solid line corresponds to a different value of the total unlensed flux of the source.
-The flux $f$ is expressed in terms of the background noise rms fluctuation measured over an area equal to $\tein^2$.
+The flux, $f$, is expressed in terms of the background noise rms fluctuation measured over an area equal to $\tein^2$.
 The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
@@
 If the surface brightness drops below the sky rms fluctuation level, then it becomes very difficult to detect it.
-In order to determine whether there are lensing-specific features in the $\crosssect-\theta_{\mathrm{e,s}}$ relation of \Fref{fig:extcs}, we computed, for each source flux, the maximum half-light radius for which it can be detected in the absence of lensing.
-We used the same criterion as that of section \ref{ssec:lensdefext} to define a detection: we defined the source footprint as the ensemble of pixels that are $2\sigma$ above the background and required the total signal-to-noise ratio within the footprint to be larger than ten.
-The resulting limiting sizes are shown as vertical lines in \Fref{fig:extcs}.
+In order to determine whether there are lensing-specific features in the $\crosssect-\theta_{\mathrm{e,s}}$ relation of Fig.
+\ref{fig:extcs}, we computed, for each source flux, the maximum half-light radius for which it can be detected in the absence of lensing.
+We used the same criterion as that of Sect.
+\ref{ssec:lensdefext} to define a detection: we defined the source footprint as the ensemble of pixels that are $2\sigma$ above the background and required the total signal-to-noise ratio within the footprint to be larger than ten.
+The resulting limiting sizes are shown as vertical lines in Fig.
+\ref{fig:extcs}.
 For each given total flux, the non-lensing size limit is similar to the value of $\theta_{\mathrm{e,s}}$ at which the lensing cross-section drops.
 
-This result suggests that, to first approximation, lensing does not introduce a strong selection in source size.
+This result suggests that, to the first approximation, lensing does not introduce a strong selection in source size.
 While perhaps surprising, this follows from the fact that gravitational lensing preserves surface brightness: a source that can be detected in the absence of lensing will produce images with the same surface brightness when lensed, which can be detected as well.
@@
 This lowers the signal-to-noise ratio of the multiple images compared to the point source case.
-The result is that $\crosssect$ starts to decrease with increasing source size at values of $\theta_{\mathrm{e,s}}$ that are smaller than the no-lensing detection limit, as observed in \Fref{fig:extcs}.
-
-At the brightest flux explored in the experiment (green line in \Fref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (black dotted line in \Fref{fig:extcs}).
+The result is that $\crosssect$ starts to decrease with increasing source size at values of $\theta_{\mathrm{e,s}}$ that are smaller than the no-lensing detection limit, as observed in Fig.
+\ref{fig:extcs}.
+
+At the brightest flux explored in the experiment (green line in Fig.
+\ref{fig:extcs}), the lensing cross-section at small source sizes is larger than the area enclosed within the radial caustic (black dotted line in Fig.
+\ref{fig:extcs}).
 This is because, when the source is very bright, it can give rise to multiple images even while its centroid lies outside of the radial caustic, as long as its surface brightness distribution extends into it.
@@
 Then, starting from the lens model used in the previous section, we progressively increased the critical surface mass density to reduce the Einstein radius down to values comparable to the pixel size.
-\Fref{fig:largesourcecs} shows the resulting strong lensing cross-section as a function of the ratio between lens Einstein radius and source half-light radius.
+Figure \ref{fig:largesourcecs} shows the resulting strong lensing cross-section as a function of the ratio between lens Einstein radius and source half-light radius.
 For values of the Einstein radius close to the size of the PSF, the lensing cross-section (blue solid line) is very small: this is because the perturbations caused by lensing are not well resolved.
@@
 \caption{
-Strong lensing cross-section of an extended source. Large source size regime ($\theta_{\mathrm{e,s}} \gtrsim \tein$).
+Strong lensing cross-section of an extended source: large source size regime ($\theta_{\mathrm{e,s}} \gtrsim \tein$).
 The cross-section is plotted as a function of the ratio between the Einstein radius of the lens and the half-light radius of the source.
@@
 The source half-light radius measures $20$ pixels, and the PSF has an FWHM of two pixels.
-The lens model is varied: starting from the model of \Fref{fig:extcs}, the critical surface mass density is increased to produce progressively smaller values of $\tein$.
-Solid line: lensing cross-section, in units of the footprint area of the source in the absence of lensing, as a function of the ratio between the lens Einstein radius and the source half-light radius.
-Dashed line: area enclosed within the lens caustics.
-Dotted line: value of the cross-section corresponding to the area of the source.
-Dash-dotted line: FWHM of the PSF.
+The lens model is varied: starting from the model of Fig. \ref{fig:extcs}, the critical surface mass density is increased to produce progressively smaller values of $\tein$.
+The solid line shows the lensing cross-section, in units of the footprint area of the source in the absence of lensing, as a function of the ratio between the lens Einstein radius and the source half-light radius.
+The dashed line delimits the area enclosed within the lens caustics, and the dotted line shows the value of the cross-section corresponding to the area of the source.
+The dash-dotted line is the FWHM of the PSF.
 \label{fig:largesourcecs}
@@
 
-\section{Lens populations simulations}\label{sect:lenspop}
+\section{Lens population simulations}\label{sect:lenspop}
 
@@
 Our simulations are based on empirical models, in which existing observations of the baryonic component of galaxies are complemented with a set of assumptions on the mass distribution of the lenses.
-In section \ref{ssec:lenses} we explain how we built our foreground galaxy sample,
-while in section \ref{ssec:sources} we describe the simulation of the background source population.
-In section \ref{ssec:obs} we describe how our mock observations of lenses are generated.
-In section \ref{ssec:teincuts} we apply a further selection step, based on the angular size of the Einstein radius.
+In Sect. \ref{ssec:lenses} we explain how we built our foreground galaxy sample,
+while in Sect. \ref{ssec:sources} we describe the simulation of the background source population.
+In Sect. \ref{ssec:obs} we describe how our mock observations of lenses are generated.
+In Sect. \ref{ssec:teincuts} we apply a further selection step, based on the angular size of the Einstein radius.
 
@@
 
-We describe lenses with elliptical versions of the two-component model introduced in section \ref{ssec:profile}.
+We describe lenses with elliptical versions of the two-component model introduced in Sect. \ref{ssec:profile}.
 In the following sections we explain how their parameters are generated.
@@
 We drew stellar masses from the stellar mass function of quiescent galaxies measured by \citet{Muz++13}.
-In particular, we used the following comoving number density distribution
+In particular, we used the following comoving number density distribution:
 \begin{equation}
@@
 We set the mean of $\log{\asps}$ to $0.1$, as this is an intermediate value among estimates of $\asps$ from the literature \citep{CvD12, Cap++13, SLC15, Son++15, Son++19}.
-We adopted a few different values for the scatter $\sigma_\alpha$. We explain in Section \ref{ssub:scat} how these were chosen.
+We adopted a few different values for the scatter $\sigma_\alpha$. We explain in Sect. \ref{ssub:scat} how these were chosen.
 
@@
 The choice for these distributions was motivated by observations of a sample of early-type galaxies.
-In Appendix~\ref{sect:appendixa} we explain how this sample was defined and how the coefficients of \Eref{eq:masssize} and \Eref{eq:qdist} were determined.
+In Appendix~\ref{sect:appendixa} we explain how this sample was defined and how the coefficients of Eq.
+\ref{eq:masssize} and Eq.
+\ref{eq:qdist} were determined.
 
@@
 
-We modelled the dark matter distribution of each lens galaxy with an elliptical gNFW halo (see section \ref{ssec:profile} for its definition).
+We modelled the dark matter distribution of each lens galaxy with an elliptical gNFW halo (see Sect. \ref{ssec:profile} for its definition).
 The parameters of the radial density profile were assigned as follows.
@@
 which is a Gaussian in $\log{\mhalo}$ with a mean that scales with stellar mass and scatter $\sigma_h$.
-The intrinsic scatter in halo mass is not well constrained observationally, therefore we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
+The intrinsic scatter in halo mass is not well constrained observationally; therefore, we ran simulations with different values of $\sigma_h$, as for the stellar population synthesis mismatch parameter.
 
 To determine the density profile of the halo we relied on a theoretically motivated model that takes into account the effect of baryons on the dark matter.
-We assumed the dark matter distribution to be initially described by an NFW profile with a concentration\footnote{The concentration is the ratio between the halo virial radius and the scale radius $r_s$.} of five, then used the prescription of \citet{Cau++20} to model the response of the halo to the infall of baryons.
+We assumed the dark matter distribution to be initially described by an NFW profile with a concentration\footnote{The concentration is the ratio between the halo virial radius and the scale radius, $r_s$.} of five, then used the prescription of \citet{Cau++20} to model the response of the halo to the infall of baryons.
 In the \citet{Cau++20} model the halo response is approximated with an analytical function that depends on the present stellar mass distribution, and typically results in a more concentrated and steeper density profile compared to the original NFW model.
@@
 In principle we could have allowed for additional degrees of freedom, for instance by relaxing the assumption of a fixed initial halo concentration.
-In practice, as we explain in section \ref{ssec:limitations}, our main results are not affected by this choice.
-
-Given the radial profile of the dark matter halo, we obtained an elliptical version of it by applying a transformation of the kind of \Eref{eq:elltransf} to its projected surface mass density.
+In practice, as we explain in Sect. \ref{ssec:limitations}, our main results are not affected by this choice.
+
+Given the radial profile of the dark matter halo, we obtained an elliptical version of it by applying a transformation of the kind of Eq. \ref{eq:elltransf} to its projected surface mass density.
 We assumed that the axis ratio and orientation of the halo is the same as that of the stellar component.
@@
 \end{equation}
-The existence of a fundamental plane relation is a consequence of the virial theorem: the velocity dispersion of a galaxy in dynamical equilibrium is directly related to the 3-dimensional mass distribution in its inner regions, which is typically dominated by the stellar component.
+The existence of a fundamental plane relation is a consequence of the virial theorem: the velocity dispersion of a galaxy in dynamical equilibrium is directly related to the three-dimensional mass distribution in its inner regions, which is typically dominated by the stellar component.
 At fixed observed stellar mass distribution, however, the central velocity dispersion can vary depending on the stellar population synthesis mismatch parameter, on the dark matter mass and on the dark matter density profile.
@@
 We did not attempt to match the other parameters of the fundamental plane (i.e.
-the constant of proportionality of \Eref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.
+the constant of proportionality of Eq.
+\ref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.
 The details of this procedure are given in Appendix~\ref{sect:appendixb}.
@@
 \begin{table}
-\caption{Intrinsic scatter scenarios}
+\caption{Intrinsic scatter scenarios.}
 \label{tab:scatter}
@@
 The resulting number density of sources is $70$~arcmin$^{-2}$.
-We approximated their spatial distribution as uniform in the sky, that is we neglected clustering of the sources.
+We approximated their spatial distribution as uniform in the sky, that is, we neglected clustering of the sources.
 
@@
 
-We described the population of background quasars with the following double power-law luminosity function in the rest-frame UV absolute magnitude $M$:
+We described the population of background quasars with the following double power-law luminosity function in the rest-frame UV absolute magnitude, $M$:
 \begin{equation}
@@
 \begin{equation}
-\log{\Phi^*} = -6.0991 + 0.0209\zqso + 0.0171\zqso^2,
+\log{\Phi^*} = -6.0991 + 0.0209\zqso + 0.0171\zqso^2
 \end{equation}
@@
 For the sake of consistency with the population of background extended sources, we limited the redshift distribution of quasars to the range $0.8 < \zqso < 2.5$.
-We then truncated the distribution in $\mqso$ at two magnitudes fainter than the detection limit (which is specified in section \ref{ssec:obs}).
+We then truncated the distribution in $\mqso$ at two magnitudes fainter than the detection limit (which is specified in Sect. \ref{ssec:obs}).
 Finally, we randomly placed quasars in the source plane with a projected number density of $70$~arcmin$^{-2}$.
@@
 For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering \num{1000} square degrees.
-The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around \num{300000}.
+The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in Sect. \ref{ssub:mstarz}, is around \num{300000}.
 We approximated the foreground galaxies as isolated: when determining whether a galaxy acts as a strong lens, we only modelled the contribution to the lensing signal from the galaxy itself, and neglected that of the environment.
-We discuss the possible implications of this approximation in \Sref{sect:discuss}.
+We discuss the possible implications of this approximation in Sect. \ref{sect:discuss}.
 For each lens, we determined its caustics relative to the highest source redshift, $\zsource=2.5$.
@@
 
-For the simulation with extended sources, we produced images with properties similar to those expected for the Euclid Wide survey \citep{Sca++22}.
+For the simulation with extended sources, we produced images with properties similar to those expected for the \textit{Euclid} Wide survey \citep{Sca++22}.
 We used a pixel size of $0.1''$ and we applied a Moffat PSF with an FWHM of $0.2''$ and a $\beta$ parameter of $5.0$.
 Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
-We then applied the peak detection-based lens selection criterion introduced in section \ref{ssec:lensdefext} to find the strong lenses.
+We then applied the peak detection-based lens selection criterion introduced in Sect. \ref{ssec:lensdefext} to find the strong lenses.
 
 The fiducial scatter simulation with extended background sources produced a sample of $2113$ lenses, corresponding to a number density of $2.1$~deg$^{-2}$.
-This is about a factor of five smaller than the number density predicted by \citet{Col15} for the Euclid survey.
+This is about a factor of five smaller than the number density predicted by \citet{Col15} for the \textit{Euclid} survey.
 This is a result of differences in the description of the source population, in the criteria used to define a strong lensing event, and in the redshift and stellar mass cuts that we applied to define the foreground galaxy population.
@@
 For the lensed quasars we did not simulate pixel-level data, but simply computed the observed magnitudes of the multiple images.
-Following the definition of section \ref{ssec:lensdefpoint}, we included in the sample of strong lenses only systems with at least two images brighter than a limiting magnitude $m_{\mathrm{lim}}$.
-We set $m_{\mathrm{lim}}=23.3$, which corresponds to the $10\sigma$ detection limit of the LSST in a single visit \citep{O+M10}.
-This is motivated by the fact that the quasar lenses are meant to simulate a sample assembled for the purpose of carrying out time delay measurements, which in turn require combining single-visit detections over many epochs.
-The scenario that we are simulating, then, is that of a lens search in a Euclid-like survey, followed-up with LSST time-domain observations.
+Following the definition of Sect. \ref{ssec:lensdefpoint}, we included in the sample of strong lenses only systems with at least two images brighter than a limiting magnitude $m_{\mathrm{lim}}$.
+We set $m_{\mathrm{lim}}=23.3$, which corresponds to the $10\sigma$ detection limit of the Legacy Survey of Space and Time (LSST) in a single visit \citep{O+M10}.
+This is motivated by the fact that the quasar lenses are meant to simulate a sample assembled for the purpose of carrying out time-delay measurements, which in turn require combining single-visit detections over many epochs.
+The scenario that we are simulating, then, is that of a lens search in a \textit{Euclid}-like survey, followed-up with LSST time-domain observations.
 
@@
 
-\Fref{fig:teinhist} shows the distribution in Einstein radius of the simulated lens samples.
+Figure \ref{fig:teinhist} shows the distribution in Einstein radius of the simulated lens samples.
 In all cases, the distribution peaks at $\tein\approx0.7''$.
@@
 Einstein radius distribution of the simulated lens samples.
-These are: galaxy-galaxy lenses in the fiducial, high scatter and low scatter scenarios; galaxy-quasar lenses with fiducial scatter, considering all lenses or only lenses that produce four images.
+These are: galaxy-galaxy lenses in the fiducial, high scatter, and low scatter scenarios and galaxy-quasar lenses with fiducial scatter, considering all lenses or only lenses that produce four images.
 \label{fig:teinhist}
@@
 \end{equation}
-In words, all lenses with Einstein radius larger than $\theta_{\mathrm{Ein,min}}$ are included in the sample, while all those with smaller Einstein radius are excluded.
+In words, all lenses with Einstein radii larger than $\theta_{\mathrm{Ein,min}}$ are included in the sample, while all those with smaller Einstein radius are excluded.
 We refer to $\teinmin$ as the completeness limit: our simulated lens samples are complete down to $\tein=\teinmin$.
@@
 In this section we present the results of the lens population simulations. 
-Section \ref{ssec:galbias} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:quasarbias} focuses on the population of galaxy-quasar lenses.
+Section \ref{ssec:galbias} shows the results of the galaxy-galaxy lens experiment, while Sect. \ref{ssec:quasarbias} focuses on the population of galaxy-quasar lenses.
 Given the number of parameters that are needed to describe our model, providing a complete characterisation of the strong lensing bias is a problem with relatively high dimensionality, and is beyond the scope of this paper.
 For the sake of conciseness, we focus instead on the quantities that we consider most important.
-Nevertheless, the output of our simulations is available online\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/papers/selection_effects}}.
+Nevertheless, the output of our simulations is available online\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/tree/main/papers/selection_effects}}.
 We encourage readers who are interested in studying aspects of the strong lensing bias that are not covered in this section to download our data and analyse them directly.
@@
 
-\Fref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the parent population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
+Figure \ref{fig:lenspars} shows the distribution in the parameters of the foreground galaxies of the lens systems, compared to those of the parent population, for the fiducial scatter scenario with two different values of the minimum Einstein radius: $0.5''$ and $1.0''$.
 A completeness limit of $0.5''$ is close to what can currently be achieved via visual inspection of high-resolution space-based images \citep{Gar++22}, while the $1.0''$ limit can be seen as a more conservative case.
@@
 \caption{
-Comparison between the properties of lens samples and the parent population. Distribution in foreground galaxy parameters.
-Filled contours: distribution of the parent sample.
-Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
-Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
+Comparison between the properties of lens samples and the parent population: distribution in foreground galaxy parameters.
+Filled contours represent the distribution of the parent sample, solid green lines the distribution of the lenses with Einstein radii larger than $0.5''$, and solid red lines the distribution of the lenses with Einstein radii larger than $1.0''$.
 \label{fig:lenspars}
@@
 
-\Fref{fig:sourcepars} shows the distribution in parameters describing the background source population, for the same simulations of \Fref{fig:lenspars}. 
-Additionally, \Fref{fig:sourcepars} shows the subset of the parent population that consists of detectable sources.
+Figure \ref{fig:sourcepars} shows the distribution in parameters describing the background source population, for the same simulations of Fig. \ref{fig:lenspars}. 
+Additionally, Fig. \ref{fig:sourcepars} shows the subset of the parent population that consists of detectable sources.
 These are background galaxies that, in the absence of lensing, can be detected according to the same criterion used for the lensed sources (i.e.
@@
 \caption{
-Comparison between the properties of lens samples and the parent population. Distribution in background source parameters.
-Filled contours: distribution of the parent sample.
-Black solid lines: distribution of the detectable sources.
-Green solid lines: distribution of the lenses with Einstein radius larger than $0.5''$.
-Red solid lines: distribution of the lenses with Einstein radius larger than $1.0''$.
+Comparison between the properties of lens samples and the parent population: distribution in background source parameters.
+Filled contours show the distribution of the parent sample, solid black lines the distribution of the detectable sources, solid green lines the distribution of the lenses with Einstein radii larger than $0.5''$, and solid red lines the distribution of the lenses with Einstein radii larger than $1.0''$.
 \label{fig:sourcepars}
@@
 The reason for this behaviour lies in the fact that, in the detection of both lensed and unlensed sources, the most important quantity is surface brightness, which is preserved by gravitational lensing when the source is larger than the PSF size.
-As we showed in section \ref{ssec:ellext}, the lensing cross-section drops to zero once the surface brightness of the source reaches a value that would make it undetectable in the absence of lensing.
+As we showed in Sect. \ref{ssec:ellext}, the lensing cross-section drops to zero once the surface brightness of the source reaches a value that would make it undetectable in the absence of lensing.
 For this reason, also the distribution in half-light radius is very similar between the lensed sources and the detectable source population.
 
-Qualitatively, the results shown in \Fref{fig:lenspars} and \Fref{fig:sourcepars} match our expectations from \Sref{sect:indlenses}.
+Qualitatively, the results shown in Figs. \ref{fig:lenspars} and \ref{fig:sourcepars} match our expectations from Sect. \ref{sect:indlenses}.
 In the rest of this section we quantify the lensing bias.
@@
 
-\Fref{fig:lensobsbias} shows the median redshift, median $\mobs$, median size for a given $\mobs$, and median axis ratio of various lens population simulations, as a function of the minimum Einstein radius.
+Figure \ref{fig:lensobsbias} shows the median redshift, median $\mobs$, median size for a given $\mobs$, and median axis ratio of various lens population simulations, as a function of the minimum Einstein radius.
 We defined the median size at fixed $\mobs$ by fitting the following mass-size relation to the $\reff-\mobs$ distribution:
@@
 \end{equation}
-The quantity shown in the third panel of \Fref{fig:lensobsbias} is the parameter $\mu_{R,0}$, which is the average $\log{\reff}$ at an observed stellar mass of $\log{\mobs}=11.4$.
+The quantity shown in the third panel of Fig. \ref{fig:lensobsbias} is the parameter $\mu_{R,0}$, which is the average $\log{\reff}$ at an observed stellar mass of $\log{\mobs}=11.4$.
 The horizontal dashed line in each panel shows the value of the parent population: the larger the distance between the curve of a simulation and this line, the higher the strong lensing bias.
@@
 Bias on lens observable properties as a function of the minimum Einstein radius.
-First panel: median redshift.
-Second panel: median observed stellar mass.
-Third panel: median half-light radius at an observed mass of $\log{\mobs}=11.4$.
-Fourth panel: median axis ratio.
-In each panel, the black dashed line indicates the value of the parent population.
+First panel: Median redshift.
+Second panel: Median observed stellar mass.
+Third panel: Median half-light radius at an observed mass of $\log{\mobs}=11.4$.
+Fourth panel: Median axis ratio.
+In each panel, the dashed black line indicates the value of the parent population.
 Error bars indicate the standard deviation of the mean of the lens sample.
@@
 
-As was already visible in \Fref{fig:lenspars}, there is a clear bias towards lower redshift, higher stellar mass and smaller sizes, with the bias becoming stronger for more restrictive cuts in Einstein radius.
+As was already visible in Fig.
+\ref{fig:lenspars}, there is a clear bias towards lower redshift, higher stellar mass and smaller sizes, with the bias becoming stronger for more restrictive cuts in Einstein radius.
 Biases in stellar mass and size are stronger for the simulation with low scatter in $\asps$ and $\mhalo$.
@@
 
-\Fref{fig:lensmassbias} shows the median of the distribution in various mass-related quantities, as a function of the minimum Einstein radius.
+Figure \ref{fig:lensmassbias} shows the median of the distribution in various mass-related quantities, as a function of the minimum Einstein radius.
 The top panel is the median stellar population synthesis mismatch parameter.
@@
 
-The second panel of \Fref{fig:lensmassbias} shows the median halo mass of galaxies with an observed stellar mass of $\log{\mobs}=11.4$.
+The second panel of Fig. \ref{fig:lensmassbias} shows the median halo mass of galaxies with an observed stellar mass of $\log{\mobs}=11.4$.
 We measured this quantity by fitting the following relation to the $\mhalo-\mobs$ distribution,
@@
 
-The third panel of \Fref{fig:lensmassbias} shows the median projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$, at fixed observed stellar mass and half-light radius.
+The third panel of Fig. \ref{fig:lensmassbias} shows the median projected dark matter mass enclosed within an aperture of $5$~kpc, $\mdmfive$, at fixed observed stellar mass and half-light radius.
 We obtained this quantity by fitting the following relation to the $\mdmfive-\mobs-\reff$ distribution:
@@
 \end{equation}
-\Fref{fig:lensmassbias} shows the value of $\mu_{\mathrm{DM},0}$.
+Figure \ref{fig:lensmassbias} shows the value of $\mu_{\mathrm{DM},0}$.
 The bias on this quantity is qualitatively similar to that on the total halo mass, but much smaller in amplitude.
@@
 Second, while $\mu_{\mathrm{h},0}$ is the halo mass at fixed stellar mass, $\mu_{\mathrm{DM},0}$ is measured at fixed half-light radius as well.
-The density profile of the dark matter halos in our simulation have a dependence on galaxy size: the response of dark matter to baryons is stronger for more concentrated stellar distributions (see section \ref{ssub:dmprofile}).
-By capturing this dependence in the model of \Eref{eq:dmscaling}, the residual scatter in $\mdmfive$ around the mean is reduced, and so is the strong lensing bias.
+The density profile of the dark matter halos in our simulation have a dependence on galaxy size: the response of dark matter to baryons is stronger for more concentrated stellar distributions (see Sect.
+\ref{ssub:dmprofile}).
+By capturing this dependence in the model of Eq. \ref{eq:dmscaling}, the residual scatter in $\mdmfive$ around the mean is reduced, and so is the strong lensing bias.
 This, however, is a minor effect: we repeated the analysis while setting the dependence on size to zero and found minimal differences on the derived values of $\mu_{\mathrm{DM},0}$.
@@
 
-In the fourth panel of \Fref{fig:lensmassbias} we show the average inner dark matter slope at fixed observed stellar mass and half-light radius, which we measured by fitting the following model
+In the fourth panel of Fig. \ref{fig:lensmassbias} we show the average inner dark matter slope at fixed observed stellar mass and half-light radius, which we measured by fitting the following model
 \begin{equation}\label{eq:dmscaling}
@@
 In our model, galaxies of a given size with a larger stellar mass have a steeper dark matter slope, because the response of the dark matter to the infall of baryons is stronger.
-Vice-versa, galaxies with a larger halo mass have a shallower density slope.
+Vice versa, galaxies with a larger halo mass have a shallower density slope.
 At fixed $\mobs$ and $\reff$, strong lenses have both a larger stellar mass (because their $\asps$ is larger) and a larger halo mass.
-In the simulation with low scatter, the bias in stellar mass is more important, therefore $\gammadm$ has a positive bias.
+In the simulation with low scatter, the bias in stellar mass is more important; therefore, $\gammadm$ has a positive bias.
 In the simulation with large scatter, the bias in halo mass dominates, and therefore $\gammadm$ is negatively biased.
@@
 Bias on lens mass properties as a function of the minimum Einstein radius.
-First panel: median $\log{\asps}$. Dotted lines indicate values of $\asps$ corresponding to a Chabrier and a Salpeter IMF ($\asps=1$ corresponds to a Kroupa IMF).
-Second panel: median $\mhalo$ at $\log{\mobs}=11.4$.
-Third panel: median $\mdmfive$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
-Fourth panel: median $\gammadm$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
+First panel: Median $\log{\asps}$. Dotted lines indicate values of $\asps$ corresponding to a Chabrier and a Salpeter IMF ($\asps=1$ corresponds to a Kroupa IMF).
+Second panel: Median $\mhalo$ at $\log{\mobs}=11.4$.
+Third panel: Median $\mdmfive$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
+Fourth panel: Median $\gammadm$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
 In each panel, the dashed line indicates the value of the parent population.
@@
 
-\Fref{fig:sourcebias} shows the lensing bias in source-related parameters, with respect to the population of sources that are detectable without lensing.
-As previously seen in \Fref{fig:sourcepars}, lensed sources are biased towards higher redshift (top panel of \Fref{fig:sourcebias}), with the trend being larger for more restrictive cuts on the Einstein radius.
+Figure \ref{fig:sourcebias} shows the lensing bias in source-related parameters, with respect to the population of sources that are detectable without lensing.
+As previously seen in Fig.
+\ref{fig:sourcepars}, lensed sources are biased towards higher redshift (top panel of Fig.
+\ref{fig:sourcebias}), with the trend being larger for more restrictive cuts on the Einstein radius.
 The second panel shows the bias in the source magnitude.
@@
 Naively one might have expected the opposite trend, as strong lensing magnification allows the detection of sources that are intrinsically fainter than the detection limit.
-However, as the analysis of \Sref{sect:indlenses} clearly shows, the lensing cross-section is always larger for brighter sources, and that explains the sign of the bias.
+However, as the analysis of Sect. \ref{sect:indlenses} clearly shows, the lensing cross-section is always larger for brighter sources, and that explains the sign of the bias.
 At the same time, the median does not capture the whole picture of the bias in $\msource$.
-For instance, in the fiducial simulation with no cut on the Einstein radius, the $90\%$-ile of the $\msource$ distribution is $26.05$, while that of the population of detectable sources is $0.14$~mag brighter: indeed, strong lensing allows the detection of fainter sources.
-
-The third and fourth panels of \Fref{fig:sourcepars} show the bias in half-light radius and S\'{e}rsic index, respectively, for sources within a magnitude bin centred on $\msource=25$ and with a $0.4$~mag width.
+For instance, in the fiducial simulation with no cut on the Einstein radius, the $90$th percentile of the $\msource$ distribution is $26.05$, while that of the population of detectable sources is $0.14$~mag brighter: indeed, strong lensing allows the detection of fainter sources.
+
+The third and fourth panels of Fig.
+\ref{fig:sourcepars} show the bias in half-light radius and S\'{e}rsic index, respectively, for sources within a magnitude bin centred on $\msource=25$ and with a $0.4$~mag width.
 We detected no clear sign of bias.
-Finally, the fifth panel shows the bias in source axis ratio. Also in this case no obvious sign of bias was detected, except in the largest values of $\theta_{\mathrm{Ein,min}}$.
+Finally, the fifth panel shows the bias in source axis ratio. Also in this, case no obvious sign of bias was detected, except in the largest values of $\theta_{\mathrm{Ein,min}}$.
 
@@
 Bias on source properties as a function of the minimum Einstein radius.
-First panel: median source redshift.
-Second panel: median source magnitude.
-Third panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
-Fourth panel: median S\'{e}rsic index in the magnitude bin $24.8 < \msource < 25.2$.
-Fifth panel: median axis ratio.
+First panel: Median source redshift.
+Second panel: Median source magnitude.
+Third panel: Median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
+Fourth panel: Median S\'{e}rsic index in the magnitude bin $24.8 < \msource < 25.2$.
+Fifth panel: Median axis ratio.
 In each panel, the dashed line indicates the value of the population of detectable sources.
-Error bars indicate the standard deviation of the mean of either the full sample (first, second and fifth panel) or the bin (third and fourth panel).
+Error bars indicate the standard deviation of the mean of either the full sample (first, second, and fifth panel) or the bin (third and fourth panel).
 \label{fig:sourcebias}
@@
 For this reason, we only ran simulations with lensed quasars in the fiducial scatter scenario and compared the results with those from the extended source simulation.
-We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses, therefore we also analysed that subsample on its own.
-
-The first four rows of \Fref{fig:quasarbias} show the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
+We are also interested in understanding how the subset of quad lenses differs from the entire population of quasar lenses; therefore, we also analysed that subsample on its own.
+
+The first four rows of Fig.
+\ref{fig:quasarbias} show the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
 The biases of the population of lensed quasars (red curves) are very similar to that of lensed galaxies (blue curves), especially for values of $\theta_{\mathrm{Ein},min}<1''$, 
 When considering only quad lenses, however, there are some differences, the most remarkable of which is the bias in the axis ratio: quad lenses tend to be on average galaxies with a higher ellipticity.
-This is a well-known effect \citep[see e.g.][]{KKS97} which can be explained in the context of the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
+This is a well-known effect \citep[see e.g.][]{KKS97} that can be explained in the context of the results of Sect.
+\ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
 Quad lenses tend to also have a slightly larger halo mass at fixed stellar mass, for $\theta_{\mathrm{Ein},min}<1''$.
@@
 Bias in the population of quasar lenses (all and quads only), compared to the extended source simulation.
-First panel: median $\log{\asps}$.
-Second panel: median $\mhalo$ at $\log{\mstar}=11.5$.
-Third panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
-Fourth panel: median lens axis ratio.
-Fifth panel: median mass-sheet transformation parameter, defined in Appendix~\ref{sect:appendixc}.
+First panel: Median $\log{\asps}$.
+Second panel: Median $\mhalo$ at $\log{\mstar}=11.5$.
+Third panel: Median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+Fourth panel: Median lens axis ratio.
+Fifth panel: Median mass-sheet transformation parameter, defined in Appendix~\ref{sect:appendixc}.
 In each panel, the dashed line indicates the value of the parent population.
@@
 A measurement of the time delay between the images of a lensed quasar can be converted into an estimate of the expansion rate of the Universe, $H_0$. 
-In general, though, the value of $H_0$ is degenerate with the mass distribution of the lens (and of matter along the line-of-sight), which can be difficult to constrain on an individual lens basis.
+In general, though, the value of $H_0$ is degenerate with the mass distribution of the lens (and of matter along the line of sight), which can be difficult to constrain on an individual lens basis.
 One possible strategy to break this degeneracy is to use information gathered from a larger set of lenses as a prior for the time-delay lenses.
@@
 where $\lmst$ is called the mass-sheet transformation parameter.
-Strong lensing data can constrain very well the power-law index $\gamma$, but not at all $\lmst$.
+Strong lensing data can constrain  the power-law index $\gamma$ very well, but not at all $\lmst$.
 Given the time delay, the inferred value of the Hubble constant $H_0$ scales linearly with $1/\lmst$.
 \citet{Bir++20} measured the population distribution of $\lmst$ on the sample of extended source lenses, by combining strong lensing and stellar kinematics data, and then used it as a prior in the time-delay analysis.
-Although the mass model used in our work is different from that of \Eref{eq:mstprofile}, we can estimate the bias associated with this approach by computing the value of $\lmst$ for each lens and then comparing the strong lensing bias on $\lmst$ of extended source and quasar lenses.
+Although the mass model used in our work is different from that of Eq.
+\ref{eq:mstprofile}, we can estimate the bias associated with this approach by computing the value of $\lmst$ for each lens and then comparing the strong lensing bias on $\lmst$ of extended source and quasar lenses.
 Appendix~\ref{sect:appendixc} explains how we obtained $\lmst$ for each lens.
-The strong lensing bias on the median value of $\lmst$ is shown in the fifth panel of \Fref{fig:quasarbias}\footnote{Since $\lmst$ is only defined for strong lenses, the median value of $\lmst$ of the general population of galaxies is not defined.}.
+The strong lensing bias on the median value of $\lmst$ is shown in the fifth panel of Fig.
+\ref{fig:quasarbias}\footnote{Since $\lmst$ is only defined for strong lenses, the median value of $\lmst$ of the general population of galaxies is not defined.}.
 
@@
 The bias can also be larger if the properties of the general population of galaxies (as opposed to the strong lenses) are used to inform the structure of time-delay lenses \citep[see][]{C+C16}.
-This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in \Fref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines).
+This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in Fig.
+\ref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines).
 
@@
 Each one of these conditions introduces a bias with respect to the parent population of foreground galaxies and background sources.
-Together, they define the lens selection probability term $\psel$ of \Eref{eq:one}.
+Together, they define the lens selection probability term $\psel$ of Eq.
+\ref{eq:one}.
 The first two points are intrinsic to a strong lensing survey and constitute an unavoidable source of bias. 
@@
 In this case, the third condition does not introduce any further selection and the strong lensing bias is minimised.
-We explored this scenario in \Sref{sect:results} when setting the minimum Einstein radius to zero.
-If only lenses with Einstein radius larger than a given threshold are selected, however, the bias generally increases.
+We explore this scenario in Sect. \ref{sect:results} when setting the minimum Einstein radius to zero.
+If only lenses with Einstein radii larger than a given threshold are selected, however, the bias generally increases.
 
@@
 
-In the fiducial scatter scenario, the one consistent with the observed scatter around the fundamental plane, the bias on $\asps$ varies from $0.03$~dex, when no restrictions on the lens Einstein radius are applied, to $0.09$~dex, corresponding to the extreme case in which only lenses with $\tein > 2.0''$ are selected (see \Fref{fig:lensmassbias}).
-A reasonable value for the minimum Einstein radius in a space-based survey like Euclid or the Chinese Space Station Telescope (CSST) is $\theta_{\mathrm{Ein,min}}=0.5''$.
+In the fiducial scatter scenario, the one consistent with the observed scatter around the fundamental plane, the bias on $\asps$ varies from $0.03$~dex, when no restrictions on the lens Einstein radius are applied, to $0.09$~dex, corresponding to the extreme case in which only lenses with $\tein > 2.0''$ are selected (see Fig.
+\ref{fig:lensmassbias}).
+A reasonable value for the minimum Einstein radius in a space-based survey like \textit{Euclid} or the Chinese Space Station Telescope is $\theta_{\mathrm{Ein,min}}=0.5''$.
 In this case, the bias on $\asps$ is slightly smaller than $0.04$~dex.
@@
 The origin of this discrepancy probably lies in the differences between the criteria used to define a strong lens in the two studies.
-The \citet{O+A17} sample was selected primarily via spectroscopy, by looking for signatures of two galaxies at different redshifts in the Sloan Digital Sky Survey \citep[SDSS][]{Yor++00} data.
+The \citet{O+A17} sample was selected primarily via spectroscopy, by looking for signatures of two galaxies at different redshifts in the Sloan Digital Sky Survey \citep[SDSS;][]{Yor++00} data.
 SDSS spectra were taken in fibres with a $1.5''$ radius.
 Lensed galaxies that are comparable in size to this scale, or larger, are less likely to be detected, because part of their flux extends outside of the fibre.
-Compact galaxies with a large overall magnification, instead, are more likely to be detected \citep[see the discussion in section 5.3 of][]{O+A17}.
+Compact galaxies with a large overall magnification, instead, are more likely to be detected \citep[see the discussion in Sect.
+5.3 of][]{O+A17}.
 We refer the reader to \citet{Arn++12} for a thorough study of selection effects associated with spectroscopy.
@@
 Our experiments were also useful for understanding the possible biases that might incur when combining information from samples of galaxy-galaxy lenses with samples of galaxy-quasar lenses.
-On the one hand the biases in the mass-related quantities ($\asps$ and $\mdmfive$) are very similar in simulations with the same foreground galaxy population and different background sources (see \Fref{fig:quasarbias}).
+On the one hand the biases in the mass-related quantities ($\asps$ and $\mdmfive$) are very similar in simulations with the same foreground galaxy population and different background sources (see Fig.
+\ref{fig:quasarbias}).
 On the other hand, these biases are a function of the completeness limit.
@@
 In order to correctly interpret stellar dynamics data, assumptions on the three-dimensional structure of a lens must be made.
-This is directly related to the projected ellipticity: galaxies with an axis ratio close to one tend to be preferentially elongated in the line-of-sight direction, and vice-versa.
-When using stellar dynamics-based mass measurements to inform the properties of a sample of quad lenses, therefore, it is important to take into account possible biases due to the different three-dimensional structure of the two samples.
+This is directly related to the projected ellipticity: galaxies with an axis ratio close to one tend to be preferentially elongated in the line-of-sight direction, and vice versa.
+When using stellar dynamics-based mass measurements to inform the properties of a sample of quad lenses, it is therefore important to take into account possible biases due to the different three-dimensional structure of the two samples.
 
@@
 
-Strictly speaking, the results shown in \Sref{sect:results} apply only to samples of lenses and background sources with the same properties as our simulations.
+Strictly speaking, the results shown in Sect. \ref{sect:results} apply only to samples of lenses and background sources with the same properties as our simulations.
 Real lens samples can have more complex source distributions.
 For example, many existing lens-finding algorithms are tuned to identify sources with different colours from the lenses, introducing a selection that can modify the source distribution in redshift-magnitude space.
-However, the comparison of section \ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
+However, the comparison of Sect.
+\ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
 This result suggests that the dependence of the strong lensing bias on the details of the background source population is very weak, as long as no additional selection on the image configuration is applied (e.g.
 by selecting only quad lenses).
-This is not a coincidence, but is a consequence of the fact that the dependence of the strong lensing cross-section on the lens parameters is a weak function of source magnitude\footnote{The cross-section itself is a strong function of source magnitude, but the trends between $\crosssect$ and the lens parameters are not.}, especially for magnitudes close to the detection limit (see section \ref{ssec:axisymmpoint}).
+This is not a coincidence, but is a consequence of the fact that the dependence of the strong lensing cross-section on the lens parameters is a weak function of source magnitude\footnote{The cross-section itself is a strong function of source magnitude, but the trends between $\crosssect$ and the lens parameters are not.}, especially for magnitudes close to the detection limit (see Sect.
+\ref{ssec:axisymmpoint}).
 We can then conclude that the details of the properties of the background source population play a secondary role in determining the strong lensing bias.
@@
 One way to minimise the bias, therefore, is to identify scaling relations between observable quantities and mass-related properties that can account for part of the scatter.
-Describing the inner dark matter distribution as a function of stellar mass and half-light radius, as we did in section \ref{ssec:galbias}, is the first step in this direction: the bias in $\mdmfive$ at fixed $\mstar$ and $\reff$ is smaller than the overall shift in the median $\mdmfive$ of the lens population (albeit by a marginal amount, as we pointed out earlier).
+Describing the inner dark matter distribution as a function of stellar mass and half-light radius, as we did in Sect.
+\ref{ssec:galbias}, is the first step in this direction: the bias in $\mdmfive$ at fixed $\mstar$ and $\reff$ is smaller than the overall shift in the median $\mdmfive$ of the lens population (albeit by a marginal amount, as we pointed out earlier).
 This description could be extended by including the central velocity dispersion as an additional control parameter. 
@@
 The central velocity dispersion, however, is very sensitive to the orbital anisotropy, to the three-dimensional structure, and to gradients in stellar mass-to-light ratio, which are not well known.
-Therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
+Therefore, it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.
 
@@
 In order for either of these approaches to work, however, it is essential that the lens selection procedure can be simulated.
-This, in turn, requires having an objective definition of a strong lensing event. A peak-based definition such as that introduced in section \ref{ssec:lensdefext} could be used for this purpose.
+This, in turn, requires having an objective definition of a strong lensing event. A peak-based definition such as that introduced in Sect. \ref{ssec:lensdefext} could be used for this purpose.
 
@@
 
-We also assumed that the dark matter density profile of the galaxies is completely determined by the halo mass and by the stellar mass distribution (see section \ref{ssub:dmprofile}).
-In reality we expect there to be a range of density profiles, for example as a result of halos having a nonzero scatter in their initial (i.e.
+We also assumed that the dark matter density profile of the galaxies is completely determined by the halo mass and by the stellar mass distribution (see Sect.
+\ref{ssub:dmprofile}).
+In reality we expect there to be a range of density profiles, for example as a result of halos having a non-zero scatter in their initial (i.e.
 before baryonic infall) concentration parameter.
@@
 Massive galaxies are known to have colour gradients, the main effect of which is to cause the stellar half-mass radius to be smaller than the half-light radius \citep{Szo++13,Sue++19}.
-Since, as shown in \Sref{sect:results}, strong lensing bias tends to select galaxies with a more compact stellar distribution, it would also preferentially select galaxies with a steeper mass-to-light ratio gradient.
+Since, as shown in Sect.
+\ref{sect:results}, strong lensing bias tends to select galaxies with a more compact stellar distribution, it would also preferentially select galaxies with a steeper mass-to-light ratio gradient.
 However, testing for the amplitude of the strong lensing bias on the mass-to-light ratio gradient is beyond the scope of this paper.
 
-\subsection{Lens vs. source selection}
-
-The simulations described in \Sref{sect:lenspop}, on which our experiment is based, are examples of so-called lens-selected surveys:
+\subsection{Lens versus source selection}
+
+The simulations described in Sect. \ref{sect:lenspop}, on which our experiment is based, are examples of so-called lens-selected surveys:
 the lens samples are built by first defining a population of possible lens galaxies and then searching for strong lenses among them.
@@
 In principle, these two approaches lead to different strong lensing biases. 
-For example, from \Fref{fig:lenspars} we can see that the stellar mass distribution of the lenses extends below the minimum mass cut that we applied to the population of foregrounds galaxies: a source-based search that is able detect those lenses would result in a different sample and a different associated strong lensing bias.
-However, when considering only lenses with $\tein > 1''$, lens-selection and source-selection yield the same sample (see red histogram in \Fref{fig:lenspars}): in the limit of large Einstein radius, the two operations (identifying suitable lens galaxies and detecting a strongly lensed source) commute.
+For example, from Fig.
+\ref{fig:lenspars} we can see that the stellar mass distribution of the lenses extends below the minimum mass cut that we applied to the population of foreground galaxies: a source-based search that is able detect those lenses would result in a different sample and a different associated strong lensing bias.
+However, when considering only lenses with $\tein > 1''$, lens-selection and source-selection yield the same sample (see the red histogram in Fig.
+\ref{fig:lenspars}): in the limit of large Einstein radius, the two operations (identifying suitable lens galaxies and detecting a strongly lensed source) commute.
 
@@
 
-Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in section \ref{ssec:lensdefext}.
+Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in Sect. \ref{ssec:lensdefext}.
 However, many previous studies aimed at simulating populations of lenses adopted a magnification-based definition, in which a lens-source system is classified as a strong lens if the magnification of the source is larger than a minimum value \citep[see e.g.][]{Col15, Robertson++20}.
@@
 The total number of strong lenses increased by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
-\Fref{fig:minmulbias} and \Fref{fig:minmusbias} show the strong lensing bias on a series of lens- and source-related properties.
-Besides the change in the source redshift and magnitude (first two panels in \Fref{fig:minmusbias}), we see very little differences in the bias with respect to the peak-based definition of the fiducial analysis.
+Figures \ref{fig:minmulbias} and  \ref{fig:minmusbias} show the strong lensing bias on a series of lens- and source-related properties.
+Besides the change in the source redshift and magnitude (first two panels in Fig.
+\ref{fig:minmusbias}), we see very little differences in the bias with respect to the peak-based definition of the fiducial analysis.
 We then conclude that the exact definition of a strong lensing event does not have a big impact on the strong lensing bias of lens-related properties.
-We stress out, however, that magnification is not a directly-observable quantity, and therefore we advise against applying magnification-based lens definitions to define lens samples, for the reasons discussed in section \ref{ssec:lensdefext}.
+We stress, however, that magnification is not a directly observable quantity, and therefore we advise against applying magnification-based lens definitions to define lens samples, for the reasons discussed in Sect.
+\ref{ssec:lensdefext}.
 \begin{figure}
@@
 Lens strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves).
-First panel: median $\log{\asps}$.
-Second panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
-Third panel: median $\gammadm$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
-Fourth panel: median axis ratio.
-Fifth panel: median mass-sheet transformation parameter, defined in Appendix~\ref{sect:appendixc}.
+First panel: Median $\log{\asps}$.
+Second panel: Median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
+Third panel: Median $\gammadm$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
+Fourth panel: Median axis ratio.
+Fifth panel: Median mass-sheet transformation parameter, defined in Appendix~\ref{sect:appendixc}.
 In each panel, the dashed line indicates the value of the parent population.
@@
 Source strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves). 
-First panel: median source redshift.
-Second panel: median source magnitude.
-Third panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
+First panel: Median source redshift.
+Second panel: Median source magnitude.
+Third panel: Median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
 The dashed line indicates the median value of the population of sources detectable in the absence of lensing.
@@
 This work takes a major step towards filling that knowledge gap.
-After a thorough investigation, we learned several lessons regarding the strong lensing bias in photometrically selected lens samples. The following are the most important ones.
+After a thorough investigation, we have learned several lessons regarding the strong lensing bias in photometrically selected lens samples. The following are the most important ones.
 \begin{enumerate}
@@
 At fixed stellar distribution and fixed dark matter mass enclosed within a given aperture, varying the inner dark matter slope has little impact on the strong lensing cross-section (as long as the aperture within which the dark matter mass is normalised is comparable to the Einstein radius).
-\item The strong lensing cross-section has little dependence on the size of the source, if this is smaller than the Einstein radius and if it is detectable in the absence of lensing.
+\item The strong lensing cross-section has little dependence on the size of the source if the former is smaller than the Einstein radius and is detectable in the absence of lensing.
 Sources with a surface brightness that is too low to be detected are still undetected when strongly lensed. 
-\item Lens galaxies tend to be more massive and more compact than their non-lens counterparts. Their redshift distribution is also modified with respect to the general galaxy population.
+\item Lens galaxies tend to be more massive and more compact than their non-lens counterparts. Their redshift distribution also differs from that of the general galaxy population.
 At fixed observed stellar mass (i.e.
-inferred by means of stellar population synthesis), lens galaxies have a larger intrinsic stellar mass (i.e.
-a larger stellar population synthesis mismatch parameter $\asps$) and a larger dark matter halo mass.
+inferred by means of stellar population synthesis), lens galaxies have a higher intrinsic stellar mass (i.e.
+a larger stellar population synthesis mismatch parameter, $\asps$) and a higher dark matter halo mass.
 At fixed stellar mass and size, lens galaxies are still biased towards a larger dark matter content.
-\item The amplitude of the strong lensing bias depends on how broad is the distribution of the parameters describing the lens population. 
+\item The amplitude of the strong lensing bias depends on how broad the distribution of the parameters describing the lens population is. 
 Important quantities in this regard are the intrinsic scatter in the stellar population synthesis mismatch parameter, $\sigma_{\mathrm{sps}}$, and in the dark matter halo mass, $\sigma_{\mathrm{h}}$.
@@
 This implies that, in principle, we could constrain the intrinsic scatter parameters, which are currently poorly known, by measuring the amplitude of the strong lensing bias on $\mobs$ and $\reff$, which is easily observable.
-\item The strong lensing bias varies depending on the completeness of the lens sample as a function of Einstein radius.
+\item The strong lensing bias varies depending on the completeness of the lens sample as a function of the Einstein radius.
 Surveys that can discover lenses with a smaller Einstein radius have a smaller associated strong lensing bias in all quantities.
 \item 
-Under reasonable assumptions on the intrinsic scatter parameters, for a Euclid-like survey that is complete down to $\tein=0.5''$ the bias on $\asps$ is smaller than $0.04$~dex (10\%).
+Under reasonable assumptions regarding the intrinsic scatter parameters, for a \textit{Euclid}-like survey that is complete down to $\tein=0.5''$ the bias on $\asps$ is smaller than $0.04$~dex (10\%).
 This bias is much smaller than the current systematic uncertainty on the stellar population synthesis-based stellar masses.
@@
 \item Strong lensing selection broadens the magnitude distribution of background sources, compared to the population of objects that are detectable without lensing.
-At the same time, we did not find any evidence for a bias in the size distribution of background sources, at fixed magnitude.
-
-\item Simulations with lensed quasars in place of extended sources showed that the amplitude of the strong lensing bias in the lens-related parameters has very little sensitivity on the details of the source population.
-This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
-The associated bias on the inference of $H_0$ is on the order of $1\%$.
-\item Samples of quad lenses are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is also biased.
-This means that particular care must be taken when stellar dynamics measurements obtained on galaxy-galaxy lenses is used to inform the properties of quads.
+At the same time, we did not find any evidence for a bias in the size distribution of background sources at fixed magnitude.
+
+\item Simulations with lensed quasars in place of extended sources showed that the amplitude of the strong lensing bias in the lens-related parameters is not very sensitive to the details of the source population.
+This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples probe the same range in Einstein radius and lens observable properties.
+The associated bias on the inference of $H_0$ is of the order of $1\%$.
+\item Samples of quad lenses are biased towards galaxies with larger ellipticities, which implies that their three-dimensional structure is also biased.
+This means that particular care must be taken when stellar dynamics measurements obtained on galaxy-galaxy lenses are used to inform the properties of quads.
 \end{enumerate}
@@
 Biases that affect observable properties, such as the redshift and the light distribution of the lens, can be easily quantified.
-Biases on mass-related quantities, such as the stellar mass-to-light ratio or the dark matter distribution, are more difficult to measure directly and must be modelled by taking into account selection effects.
-Designing strong lensing surveys with clearly-defined and easily-modellable selection criteria would help greatly in this task.
+Biases on mass-related quantities, such as the stellar mass-to-light ratio or the dark matter distribution, are more difficult to measure directly and must be modelled taking selection effects into account.
+Designing strong lensing surveys with clearly defined and easily modellable selection criteria would help greatly in this task.
 
@@
 
-To assign half-light radii and ellipticities to the simulated lenses we relied on observations of a sample of early-type galaxies selected from the Sloan Digital Sky Survey \citep[SDSS][]{Yor++00}.
+To assign half-light radii and ellipticities to the simulated lenses we relied on observations of a sample of early-type galaxies selected from the SDSS \citep[][]{Yor++00}.
 This sample was selected as follows.
@@
 We then proceeded to fit for the axis ratio distribution of the same sample of galaxies.
-\Fref{fig:qhist} shows a histogram of the observed distribution.
+Figure \ref{fig:qhist} shows a histogram of the observed distribution.
 We fitted this with a beta distribution:
@@
 Distribution in axis ratio of a sample of early-type galaxies.
-The sample consists of $8078$ galaxies from the SDSS, selected by means of cuts in redshift, colour and S\'{e}rsic index as explained in the text.
+The sample consists of $8078$ galaxies from the SDSS, selected by means of cuts in redshift, colour, and S\'{e}rsic index, as explained in the text.
 Measurements of the axis ratio are taken from the de Vaucouleurs model fits of \citet{Mee++15}.
-The model curve is a beta distribution, \Eref{eq:betaappendix}, with $\alpha=6.28$ and $\beta=2.05$.
+The model curve is a beta distribution (Eq. \ref{eq:betaappendix}) with $\alpha=6.28$ and $\beta=2.05$.
 \label{fig:qhist}
@@
 \end{equation}
-We accounted for observational uncertainties on $\sigma_{\mathrm{ap}}$ when doing the fit, therefore the parameter $\sigma_{\sigma}$ describes the intrinsic scatter in the logarithm of the velocity dispersion, deconvolved from the observational scatter.
+We accounted for observational uncertainties on $\sigma_{\mathrm{ap}}$ when doing the fit; therefore, the parameter $\sigma_{\sigma}$ describes the intrinsic scatter in the logarithm of the velocity dispersion, deconvolved from the observational scatter.
 In principle we should also account for observational uncertainties on the stellar mass measurement, as they too contribute to the inferred scatter in $\sigma_{\mathrm{ap}}$.
@@
 
-We then generated samples of $z=0.2$ galaxies from the model of section \ref{ssec:lenses} and used the spherical Jeans equation to predict their central velocity dispersion.
+We then generated samples of $z=0.2$ galaxies from the model of Sect. \ref{ssec:lenses} and used the spherical Jeans equation to predict their central velocity dispersion.
 The spherical Jeans equation is \citep{B+T87}
@@
 \end{equation}
-where $\rho_*(r)$ is the 3-dimensional distribution of dynamical tracers (i.e.
+where $\rho_*(r)$ is the three-dimensional distribution of dynamical tracers (i.e.
 the stars), $\sigma_r$ the radial component of the velocity dispersion, $\beta(r)$ the orbital anisotropy parameter, and $M(r)$ the mass enclosed within a spherical shell of radius $r$.
-We assumed isotropic orbits ($\beta=0$), then integrated \Eref{eq:jeans} to obtain the seeing-convolved, surface brightness-weighted line-of-sight velocity dispersion within the SDSS spectroscopic aperture\footnote{We used Python code available at \url{https://github.com/astrosonnen/spherical_jeans}, for this purpose.}.
-Finally, we fitted the fundamental plane relation of \Eref{eq:fpfull} to the observed stellar mass, half-light radius and central velocity dispersion of the mock sample.
+We assumed isotropic orbits ($\beta=0$), then integrated Eq.
+\ref{eq:jeans} to obtain the seeing-convolved, surface brightness-weighted line-of-sight velocity dispersion within the SDSS spectroscopic aperture\footnote{We used the Python code available at \url{https://github.com/astrosonnen/spherical_jeans} for this purpose.}.
+Finally, we fitted the fundamental plane relation of Eq. \ref{eq:fpfull} to the observed stellar mass, half-light radius and central velocity dispersion of the mock sample.
 We repeated this procedure for different values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$, $\sigma_\mathrm{h}$ and $\sigma_\gamma$, then settled on the three scenarios listed in \Tref{tab:scatter}.
@@
 
-In the time-delay study of \citet{Bir++20}, lenses are described with a mass density profile given by \Eref{eq:mstprofile}.
+In the time-delay study of \citet{Bir++20}, lenses are described with a mass density profile given by Eq. \ref{eq:mstprofile}.
 \citet{Bir++20} defined the profile by first fitting a pure power-law mass model to the strong lensing data, thus constraining the Einstein radius and power-law index $\gamma$, and then applying a mass-sheet transformation of the following kind:
@@
 \gamma = 2 + \tein \frac{\psi'''}{1-\psi''}
-\end{equation}
+.\end{equation}
 Hence, given a lens in our sample, we first define a pure power-law model with the same Einstein radius as the true one and with $\gamma$ given by the equation above. 
